### 2020-06-04 lmm-genetic-review.tex
Initial Overleaf Import

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{comment}
\usepackage{float}
\usepackage{tikz}
\allowdisplaybreaks
%\usepackage{biblatex}
%\AtEveryCitekey{%
%  \clearfield{note}%
%}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\usetikzlibrary{arrows}
\providecommand{\note}[1]{\textcolor{red}{#1}}
\providecommand{\anna}[1]{\textcolor{blue}{#1}}
\newcommand{\bbeta}{\boldsymbol{\beta}}
\newcommand{\bbetaT}{\widetilde{\boldsymbol{\beta}}}
\newcommand{\bbetaHat}{\widehat{\boldsymbol{\beta}}}
\newcommand{\bgamma}{\boldsymbol{\gamma}}
\newcommand{\beps}{\boldsymbol{\varepsilon}}
\newcommand{\bepsT}{\widetilde{\boldsymbol{\varepsilon}}}
\newcommand{\bX}{\textbf{X}}
\newcommand{\bXT}{\widetilde{\textbf{X}}}
\newcommand{\bZ}{\textbf{Z}}
\newcommand{\bZT}{\widetilde{\textbf{Z}}}
\newcommand{\by}{\textbf{y}}
\newcommand{\byT}{\widetilde{\textbf{y}}}
\newcommand{\bD}{\textbf{D}}
\newcommand{\bK}{\textbf{K}}
\newcommand{\bI}{\textbf{I}}
\newcommand{\bV}{\textbf{V}}
\newcommand{\bU}{\textbf{U}}
\newcommand{\bR}{\textbf{R}}
\newcommand{\bW}{\textbf{W}} 
\newcommand{\sigmaee}{\sigma^2_e}
\newcommand{\sigmagg}{\sigma^2_g}

\title{Review Paper Draft}
\author{Anna Reisetter}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\note{Needs rewriting?}
\anna{Yes, probably, but I'd like to get a better sense of how the rest of the paper shakes out before trying to summarize it.}
Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to confounding due to unobserved factors. This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for. Linear mixed models (LMMs) are an attractive method to correct for unobserved confounding. These methods simultaneously correct for relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data. However, population structure itself does not confound the phenotype-genotype relationship, rather differential environmental exposures among subgroups do. Population structure may or may not serve as a good proxy for such differences. Given these subtle distinctions in confounding sources, the ability of LMMs to accurately estimate fixed, sparse genetic effects has not been well studied. Considering this, we evaluate the performance of penalized LMMs in terms of MSE and false positive rates in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.
\end{abstract}

\input{main}

\newpage

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

### 2020-06-04 main.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

An objective of many genetic studies is to accurately identify variants associated with a particular trait or disease, and to estimate the effects of those variants. Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a disease, and identifying potential drug targets. However, this effort may be hindered by several factors.

Genetic variants are often correlated with each other, meaning that univariate testing of genetic variants may result in false positives and biased effect estimates. This occurs when the variant being tested is correlated with other variants, so that the effects of these correlated variants are erroneously attributed to the variant in question. One method for analyzing data in the presence of LD is pruning, where only uncorrelated SNPs are used in modeling or testing procedures. Pruning limits analysis to SNPs which are representative of a particular region, and may result in improved predictive accuracy. However, pruning is limited in that it can only identify SNPs in LD with the true causal SNP. In addition, it has been shown that a worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \cite{prive2018efficient}.

An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously. However, such methods rely on data where the number of observations is less than the number of of features. This is often infeasible in genetic studies where the number of features frequently exceeds the number of observations. Penalized regression methods make such problems tractable, by constraining the solution space of coefficient estimates.

The Lasso \cite{tibshirani1996regression} is a widely used penalized regression method. It relies on an assumption of sparsity, which allows for computational speedup while performing coefficient estimation and variable selection simultaneously. These features make the Lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.

Lasso penalized regression models assume independent observations and uncorrelated errors. In the presence of population structure, some samples are more correlated than others, violating the independent errors assumption. If left unaccounted for, this presence of such dependency can result in the identification of spurious associations. Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \cite{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, there remain open and relevant problems in analyzing structured genetic data \cite{lawson2019population, barton2019population}.

One reason that population stratification warrants such concern in genetic studies is that due to ethnic and geographic segregation, population stratification tends to be associated with differing environmental exposures and cultural practices.  In other words, random genetic variation is likely to be associated with non-genetic factors that also influence the trait of interest, thereby confounding the genotype-phenotype relationship and leading to spurious associations and biased estimates of SNP effects. Biased estimates of this nature severely hinder prediction and understanding differences among populations \cite{barton2019population}. Although bias at individual loci may be small, this bias may become magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \cite{barton2019population}.

In order to clarify and emphasize the importance of this subtle distinction, we present a detailed review of relevant concepts and methods. The remainder of this paper is organized as follows. In section \ref{structureSources} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper. We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \cite{Sillanpaeae2011, sul2018population} \anna{other papers? NB: re-read Sul, 2018 paper - a lot of similar review concepts}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures. In so doing, we refine the definition of confounding due to population structure in terms of a non-genetic mechanism. In section \ref{methodsHistorical} we briefly review historical methods of correcting for population stratification and relatedness. In section \ref{methodsFixedRandom} we review methods of correcting for population stratification as a fixed versus random effect. These correspond to PCA adjusted lasso penalized regression and lasso penalized multivariate LMMs, respectively, the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present. In addition, we review the statistical details of the consequences of population structure and heterogeneous environmental exposures for these methods. In sections \ref{simulation} and \ref{results}, we illustrate the concepts described in section \ref{methodsFixedRandom} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity. These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized LMMs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------%
\subsection{Structure in genetic data}

Population structure is a common phenomenon in genetic data. Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations. For example, European American \cite{campbell2005demonstrating}, Han Chinese \cite{xu2009genomic, chen2009genetic}, and most recently, cohorts within the UK Biobank data \cite{haworth2019apparent}.

Relatedness describes how genetically similar individuals are, and can refer to both ancient and recent relatedness.  Recent relatedness is characterized by small clusters of high levels of similarity and is indicative of familial or cryptic relatedness. Ancient relatedness, or common ancestry, is generally characterized by large clusters of low levels of similarity. The presence of multiple distinct ancestry groups is known as population stratification.

Pedigree-based methods may be used to explicitly model recent relatedness if familial relationships are known. However, in this review we focus on methods to account for unobserved relatedness. Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are be present. Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations. This phenomenon is commonly described as \textbf{confounding due to population stratification}. However, the mechanism of this phenomenon warrants some discussion. In particular, it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense, rather there must exist some non-genetic mechanism by which population stratification affects both the phenotype and genotype \cite{barton2019population, vilhjalmsson2012nature}. \\

\anna{DAG here with only population stratification, observed snps, and phenotype ... maybe an arrow from population stratification to phenotype with a question mark (?)}

%------------------------------%
\subsection{Motivating example}

As an example, consider a genetic study to assess genetic variants associated with lung cancer in a sample comprised of subjects from two distinct subpopulationsk A and B. Assume the major allele of SNP $X$ is present with higher frequency in subpopulation A compared to subpopulation B, but it has no direct effect on lung cancer. Now suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality. A GWAS analysis of data from these subpopulations would find SNP $X$ to be significantly associated with lung cancer. This spurious association is not identified because of the presence of population stratification itself, but because that stratification is related to a causal environmental exposure. Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be found to be associated with the phenotype.\\

\anna{description of DAG below, in reference to previous one}

The mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env}. The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction. Otherwise stated, we do not assume that environmental factors influence population-specific allele frequencies, or vice versa, only that the two may be related. Additionally, we assume that the environmental exposure has no effect on on observed SNPs.

\begin{figure}[H]
\centering
\begin{tikzpicture}[%
>=latex',
circ/.style={draw, shape=circle, node distance=1cm, line width=1pt}]%Define the arrow type and style for circled nodes
\draw[dashed] (0,1) node[left] (P) {Population Stratification} -- (2,1) node[right] (E) {Environment}; 
\draw[->] (0,0) node[left] (X) {Observed SNPs} -- (2,0) node[right] (Y) {Phenotype}; %Line between X and Y
\draw[->] (P)   -- (X) ; 
\draw[->] (E)   -- (Y) ; 
% \filldraw[color=red!60, fill=none, very thick] (0,1) ellipse (4.5 and 0.35);
\end{tikzpicture}
\end{figure}

\anna{An additional consideration in genetic studies that we briefly review  is that of genetic background. Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed. As \cite{vilhjalmsson2012nature} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding. Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype. }

\anna{
This work will focus on environmental confounding, however, we wish to acknowledge and describe genetic background as a potential confounder for the sake of completeness. 
}

%------------------------------%
\subsection{Correcting for population structure}

Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts. It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs.

The genomic control method is a univariate testing procedure which modifies the Cochran-Armitage test for trend by a correction factor. This correction attempts to quantify and remove the effect of population structure on the distribution of the chi-square statistic using markers unrelated to the phenotype of interest and independent of the marker being tested for association \cite{devlin1999genomic, bacanu2000power, wang2009testing}. The genomic control method retains the disadvantages of univariate SNP testing and its power to detect true associations is likely to suffer in the presence of linkage disequilibrium.

The structured association method attempts to infer subject membership within a discrete number of nonoverlapping subpopulations. Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \cite{pritchard1999use, pritchard2000association}. By partitioning the overall sample size in this manner, the structured association method is particularly vulnerable to a loss of power. Additionally, it relies on the assumption that these subpopulations are distinct and nonoveralapping. This assumption may not accurately reflect the characteristics of a particular sample, such as one consisting of admixed populations, or recent relatedness. Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall somewhere between these two extremes. By clustering those individuals within in one subpopulation or another we lose information about more high dimensional relatedness.

Another approach is to create an approximation of population structure using surrogate variables, and to adjust for these as an additional model covariates. Principal component-adjustment (PC-adjustment) methods use the first several principal components of the design matrix as these surrogate variables \cite{price2006principal}. This approach may be used in univariate or multivariate frameworks, and allows for analysis on the entirety of a structured sample.

Like PC-adjustment methods, LMMs rely on a pair-wise similarity matrix to estimate and subsequently correct for relatedness among subjects. However, the term LMM has been used to describe a number of distinct procedures for analyzing structured genetic data which differ in their objectives and methods. There are two main objectives of LMMs used in a structured genetic context: estimating narrow-sense heritability, and estimating individual SNP effects.

Among LMM methods that aim to estimate heritability, the most well-known example is genome-wide complex trait analysis (GCTA) \cite{yang2011gcta}. In this framework, all observed SNPs are treated as random effects. These effects are integrated out in order to estimate a genetic variance component, which in turn, is used to quantify the total narrow-sense heritability of a trait \cite{yang2010common}. \anna{We will briefly return to a discussion of GCTA and related methods later, though techniques for estimating heritability are not the focus of this paper.}

Within the category of LMMs that aim to estimate SNP effects are univariate and multivariate approaches. Univariate approaches attempt to estimate univariate SNP effects on phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure using a genetic similarity matrix \cite{yu2006unified, kang2010variance, kang2008efficient}. However, as with all univariate testing approaches univariate LMM implementations may still produce spurious associations and biased effect estimates in the presence of LD.

Multivariate mixed model approaches reduce these problems by assessing the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while controlling for environmental confounding  \cite{Rakitsch2012, bhatnagar2019simultaneous}. This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes. These SNPs will be included in the model as both fixed and random effects, while SNPs with smaller effect sizes are modeled only as part of the random effect in order to control for environmental confounding. \\

\anna{say more about logistic?}\\
The remainder of this review will focus on the performance of multivariate Lasso-penalized methods for the analysis of continuous, normally distributed outcomes. Specifically, PC-adjustment and LMM methods, which we will refer to as PC-Lasso and LMM-Lasso, respectively. We consider two similar implementations of the LMM-Lasso method, one developed by \cite{Rakitsch2012} and a more recent implementation from \cite{bhatnagar2019simultaneous}. Subsequent references to LMM-Lasso apply to both implementations, unless otherwise stated. Where necessary, we will differentiate between these implementations as LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, respectively.

Using simulations, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding. \\

\anna{Do we want to include the sign rate stuff?}

Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \cite{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\

\anna{find that 2020 paper that simulates the data sort of similarly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------%
\subsection{Model}

To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$. We denote the standardized version of this genotype matrix as $\bX$. We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,

\begin{equation}
    \label{eqn:model}
    \by = \bX\bbeta + \bZ\bgamma + \beps
\end{equation}

where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that indicates population structure, $\bgamma$ is an $q \times 1$ vector of subpopulation-specific environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.

Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and relative relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}. Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:

\begin{align}
    \label{eqn:k}
    \bK &= \frac{1}{p} \bX \bX^T \nonumber \\
                   &= \frac{1}{p} \bU \bD^2 \bU^T \nonumber \\
                   &= \frac{1}{p} \bU \bD (\bU \bD)^T \nonumber \\
                   &\propto \bR \bR^T
\end{align}

where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the eigenvalues of $\bX$ (equivalently, the singular values of $\bK$.) The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.

%------------------------------%
\subsection{PC-Lasso}
PC-Lasso uses the first $k$ principal components of $\bX$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates. It has been shown \cite{hoffman2013correcting} that this corresponds to the regression model

\begin{equation}
    \label{eqn:pca_reg}
    \by = \bX\bbeta + \bU_{1:k}\boldsymbol{\alpha} + \beps 
\end{equation}

where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \ref{eqn:model}. In a penalized regression framework, this model yields the objective function

\begin{equation}
    \label{eqn:pca_obj}
    Q(\boldsymbol{\beta, \alpha| X, y}) = \frac{1}{2n} ||\boldsymbol{y} - \boldsymbol{X \beta} - \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \boldsymbol{\beta} ||_1
\end{equation}

where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients. \anna{do I need to define the ell2 norm?} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.

%which requires estimation of $p + k + 1$ parameters. Determining the appropriate number of principal components to include is a non-trivial task. However, often in practice the first 10 principal components are used. \anna{and so that's what we do in all our analyses.}

%------------------------------%
\subsection{LMM-Lasso}
LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$. Often $\bZ$ is not observed directly, and a common convention is to replace $\bZ$ with $\bX$, yielding the random effect $\bX\bgamma \sim N(\boldsymbol{0}, \sigmagg \textbf{K})$. \anna{Equivalently, $\bX\bgamma$ may be denoted $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.} This results in the following LMM regression model

\begin{equation}
    \label{eqn:lmm_reg}
    \by = \bX\bbeta + \bX\bgamma+ \beps
\end{equation}

where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\bX \bbeta, \bV := \sigmagg \textbf{K} + \sigmaee \textbf{I})$. LMM-Lasso then performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that

\begin{equation}
\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}}\bX \bbeta, \bI).
\end{equation}

\anna{(However, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...)}

Using the spectral decomposition of $\bK$ described in equation \ref{eqn:k}, we can gain further insight into the mechanism of this transformation. Note that $\bV^{-1}$ can be expressed using the following matrix factorizations

\begin{align*}
    \bV^{-1} &= (\sigmagg \bK + \sigmaee \bI)^{-1}\\
    &=(\sigmagg \boldsymbol{U D^2 U}^T + \sigmaee \bU \bU^T)^{-1}\\
    &= (\bU (\sigmagg \boldsymbol{D}^2 + \sigmaee \bI) \bU^T)^{-1}\\
    &=\bU (\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1} \bU^T.
\end{align*}

\noindent By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$, $\byT = \bU^T \by$, and $\bW = \text{diag}\{w_1, ... w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1}$. Together, this yields the objective function

\begin{equation}
\label{eqn:lmmObj}
%Q(\boldsymbol{\beta| \widetilde{X}, \widetilde{y}, w}) = ||\Sigma_{i=1}^n w_i ({\widetilde{y}_i} - {\boldsymbol{\widetilde{x}}_i^T \boldsymbol{\beta}})||_2^2 + \lambda || \boldsymbol{\beta} ||_1
Q(\bbeta | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
\end{equation}

\anna{
Modeling population structure using a random effect, as opposed to fixed effects as in PC-Lasso, requires the estimation of substantially fewer parameters ($p + 2$ for LMM-Lasso-Rakitsch, $p + 3$ for LMM-Lasso-ggmix).
}

%------------------------------%
\subsubsection{LMM-Lasso variants}

We have so far used the term $\bW$ to describe the general LMM-Lasso method. However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.

LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, in which the variance components and corresponding weights are estimated once under the assumption of a null model. A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \anna{glmnet citation} or \texttt{ncvreg} \anna{ncvreg citation}. LMM-Lasso-ggmix \cite{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm. This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \cite{bhatnagar2019simultaneous}. \anna{We postpone our discussion of the differences in performance of each method to the results section.}

%------------------------------%
\subsubsection{\anna{Intuition from GLS (?) p-values and SEs in response to rotation?}}

%------------------------------%
\subsection{Equivalence of PC-Lasso and LMM-Lasso}

Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \ref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects. \cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \ref{eqn:k}.

Thus, whether modeled as a fixed or random effect, the underlying effect of $\bZ \bgamma$ on phenotype remains the same, but its treatment as a fixed or random effect has important implications in terms of model fitting and estimation. Including principal components as fixed effects requires estimation of a greater number of parameters, which may result in a loss of power \cite{zhang2015principal}. Additionally, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \cite{patterson2006population, zhao2018practical}. It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \cite{zhang2015principal}. However, in practice it is often recommended that the ten largest principal components be included \cite{zhao2018practical}. This is the implementation of PC-Lasso we will use throughout our simulations.

%------------------------------%
\subsection{Simulating environmental confounding}
Based on the LMM regression model \ref{eqn:lmm_reg}, many simulation studies of the performance of LMMs \cite{Rakitsch2012, bhatnagar2019simultaneous} \anna{cite more} simulate the quantity $\bX\bgamma$ as a random draw from a $N(\boldsymbol{0}, \sigmagg \bK)$. However, simulating environmental confounding according to this framework, where $\bgamma$ is a random quantity, "washes out" bias in $\hat{\bbeta}$ while inflating the variance. This is due to the fact that the sign of any element of the vector $\bX\bgamma$ will change from one simulation to the next. We can see this by considering the expectation and variance of low-dimensional OLS estimates of $\hat{\bbeta}$ based on model \ref{eqn:lmm_reg}:

\begin{align*}
    \mathbb{E}(\hat{\bbeta}) &= (\bX^T \bX)^{-1} \bX^T \mathbb{E}(\by)\\
    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{E}(\bX\bbeta + \bX\bgamma + \beps)\\
    &\quad\\
    \mathbb{V}(\hat{\bbeta}) &= (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\by)\\
    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps)\\
    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^T \bX)^{-1}
\end{align*}

Under the assumption that $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$, $\mathbb{E}(\hat{\bbeta}) = \bbeta, \mathbb{V}(\hat{\bbeta}) = \frac{\sigmagg}{p}\bI + \sigmaee (\bX^T \bX)^{-1}$. Considered in the context of our air quality example, such an assumption would be akin to subpopulation A being exposed to good air quality in one simulation, and bad air quality in the next. Instead, considering $\bgamma$ as a fixed quantity is more consistent with environmental heterogeneity and allows us to explore the bias in $\hat{\bbeta}$. Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\hat{\bbeta}) = \bbeta + \bgamma, \mathbb{V}(\hat{\bbeta}) = \sigmaee (\bX^T \bX)^{-1}$. Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework.

Although we derive these quantities under model \ref{eqn:lmm_reg}, where $\bZ = \bX$, as our assumed data-generating model, these results also hold for the more general model \ref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of $\bX$. Assume $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ where $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$. Then,

\begin{align*}
    \by &= \bX\bbeta + \bZ\bgamma + \beps\\
    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps\\
    &=  \bX \bbeta' + \beps'
\end{align*}

where $\boldsymbol{\tau}$ quantifies bias in the estimated $\boldsymbol{\hat{\beta}'}$ and $\boldsymbol{\psi}$ is absorbed by the error term.

%------------------------------%
\subsection{Simulation study}

%------------------------------%
\subsubsection{Data generation}
We generate pseudophenotypes using three types of $\bX$ matrices:
\begin{enumerate}
    \item Simulated genetic data from 4 independent subpopulations \cite{ochoa2019simulate}
    \item Simulateed genetic data from 4 admixed subpopulations \cite{ochoa2019simulate}
    \item Real genetic data from 4 subpopulations (African, African American, European, Japanese) \cite{efron2016computer}
\end{enumerate}

%------------------------------%
\subsubsection{Pseudophenotype generation}
\begin{itemize}
    \item We generate pseudophenotypes as follows
    \begin{equation}
        \label{eqn:sim}
        \by = a \bX \bbeta + b \bZ \bgamma + c \beps \notag
    \end{equation}
    \item Where $a, b, c$ are scaling factors such that the variance of $\by$ is equal to 1, and is partitioned as follows:
    \begin{equation}
        \label{eqn:parition_y}
        \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi) \notag
    \end{equation}
    \item $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability
    \item $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise
    \item We note that this is similar to the simulation technique employed by THAT 2020 PAPER FIND IT....
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------%
\subsection{Re-estimation of $\eta$ may, in some cases, slightly improve our estimate of $\hat{\eta}$, but does not improve the estimation of $\beta$, and takes a long time}
Put this at the end of results
%------------------------------%

\anna{We have been unable to find a scenario where ggmix outperforms the LMM-Lasso method implemented by \cite{Rakitsch2012}. Ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the matrix K. Although it is primarily intended for use in high dimensional settings, where it performs on par with LMM-Lasso, this is also the scenario where it is ~100 times slower than LMM-Lasso. }\\

\anna{ggmix vs lmmlasso eta-hat plot} - I have this, make it pretty\\

\anna{ggmix vs lmmlasso MSE plot}- I have this, make it pretty\\

\anna{ggmix vs lmmlasso prc plot (?)}\\

\anna{ggmix vs. lmmlasso time plot}- I have this, make it pretty\\

%------------------------------%
\subsection{LMM-Lasso outperforms Lasso - \anna{keep low-dimensional results? Only 'real data' for now... could use VanderWas to generate a $y$...}}
%------------------------------%
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.7]{figures/fig1a.png}
    \caption{Low-dimensional setting ($n=200, p=100$)}
    \label{fig:ld}
\end{figure}

%------------------------------%
\subsection{PC-Lasso and LMM-Lasso outperform Lasso}
%------------------------------%
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.7]{figures/fig1b.png}
    \caption{High dimensional setting ($n=200, p=1000$)}
    \label{fig:hd}
\end{figure}

\anna{add high-dimensional real data case using vanderwas?}

%------------------------------%
\subsection{Population structure}
%------------------------------%
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.7]{figures/kinship1}
    \caption{A few large subpopulations}
    \label{fig:big_blocks}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.7]{figures/kinship2}
    \caption{Many small subpopulations}
    \label{fig:small_blocks}
\end{figure}

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when simulated subpopulations are small}
%------------------------------%
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.7]{figures/fig2a}
    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy.}
    \label{fig:big_vs_small}
\end{figure}

%------------------------------%
\subsection{$\gamma$ structure}
%------------------------------%
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.7]{figures/gamma_structure.png}
    \caption{$\bgamma$ structures}
    \label{fig:gamma_structures}
\end{figure}

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when $\gamma$ is heterogeneous}
%------------------------------%
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/fig2b.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
    \label{fig:my_label}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion/conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{itemize}
    \item LMM-Lasso outperforms PC-Lasso when (1) subpopulations are small and (2) when confounding effects are nonlinear and heterogeneous
    \item LMM-Lasso \cite{Rakitsch2012} and ggmix  \cite{bhatnagar2019simultaneous} perform similarly in many scenarios
    \item However, ggmix is time-consuming and may not be practical for more realistic genetic data sets
    \item Both PC-Lasso and LMM-Lasso methods suffer in estimation accuracy due to their bias from $\lambda$
    \item Bias-reduction techniques for penalized LMMs may be an attractive avenue for future work
\end{itemize}

### 2020-06-05 main.tex
Section 3 updates

@@
 LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, in which the variance components and corresponding weights are estimated once under the assumption of a null model.
-A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \anna{glmnet citation} or \texttt{ncvreg} \anna{ncvreg citation}.
+A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.
 LMM-Lasso-ggmix \cite{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm.
-This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \cite{bhatnagar2019simultaneous}.
+This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \cite{ggmix, bhatnagar2019simultaneous}.
 \anna{We postpone our discussion of the differences in performance of each method to the results section.}
@@
 \subsection{Simulating environmental confounding}
-Based on the LMM regression model \ref{eqn:lmm_reg}, many simulation studies of the performance of LMMs \cite{Rakitsch2012, bhatnagar2019simultaneous} \anna{cite more} simulate the quantity $\bX\bgamma$ as a random draw from a $N(\boldsymbol{0}, \sigmagg \bK)$.
+Based on the LMM regression model \ref{eqn:lmm_reg}, numerous simulation studies of the performance of LMMs \cite{Rakitsch2012, bhatnagar2019simultaneous} simulate the quantity $\bX\bgamma$ as a random draw from a $N(\boldsymbol{0}, \sigmagg \bK)$.
 However, simulating environmental confounding according to this framework, where $\bgamma$ is a random quantity, "washes out" bias in $\hat{\bbeta}$ while inflating the variance.
@@
 Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\hat{\bbeta}) = \bbeta + \bgamma, \mathbb{V}(\hat{\bbeta}) = \sigmaee (\bX^T \bX)^{-1}$.
-Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework.
 
@@
 where $\boldsymbol{\tau}$ quantifies bias in the estimated $\boldsymbol{\hat{\beta}'}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
+Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
 
-%------------------------------%
-\subsection{Simulation study}
-
-%------------------------------%
-\subsubsection{Data generation}
-We generate pseudophenotypes using three types of $\bX$ matrices:
-\begin{enumerate}
-    \item Simulated genetic data from 4 independent subpopulations \cite{ochoa2019simulate}
-    \item Simulateed genetic data from 4 admixed subpopulations \cite{ochoa2019simulate}
-    \item Real genetic data from 4 subpopulations (African, African American, European, Japanese) \cite{efron2016computer}
-\end{enumerate}
-
-%------------------------------%
-\subsubsection{Pseudophenotype generation}
-\begin{itemize}
-    \item We generate pseudophenotypes as follows
 \begin{equation}
     \label{eqn:sim}
-        \by = a \bX \bbeta + b \bZ \bgamma + c \beps \notag
+    \by = a \bX \bbeta + b \bZ \bgamma + c \beps 
 \end{equation}
-    \item Where $a, b, c$ are scaling factors such that the variance of $\by$ is equal to 1, and is partitioned as follows:
+
+where $\by$ is a pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
+
 \begin{equation}
     \label{eqn:parition_y}
-        \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi) \notag
+    \mathbb{V}(\by) = \eta + (1 - \eta) \xi + (1-\eta)(1-\xi).
 \end{equation}
-    \item $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability
-    \item $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise
-    \item We note that this is similar to the simulation technique employed by THAT 2020 PAPER FIND IT....
-\end{itemize}
+
+where $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability \anna{also SNR...} and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
+\anna{We note that this is similar to the simulation technique employed by THAT 2020 PAPER FIND IT...}
+
+We generate these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian, African American, and Hispanic populations.
+\anna{cite Vanderwas data somehow, add n = ???
+for all three groups).} For each $\bX$ matrix, $\bZ$ is a corresponding incidence matrix of 1s and 0s identifying subpopulation membership.
+Several structures of the fixed $\bgamma$ were considered.
+\anna{We give more details about this in the results section.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%------------------------------%
-\subsection{Re-estimation of $\eta$ may, in some cases, slightly improve our estimate of $\hat{\eta}$, but does not improve the estimation of $\beta$, and takes a long time}
-Put this at the end of results
-%------------------------------%
-
-\anna{We have been unable to find a scenario where ggmix outperforms the LMM-Lasso method implemented by \cite{Rakitsch2012}.
-Ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the matrix K.
-Although it is primarily intended for use in high dimensional settings, where it performs on par with LMM-Lasso, this is also the scenario where it is ~100 times slower than LMM-Lasso.
-}\\
-
-\anna{ggmix vs lmmlasso eta-hat plot} - I have this, make it pretty\\
-
-\anna{ggmix vs lmmlasso MSE plot}- I have this, make it pretty\\
-
-\anna{ggmix vs lmmlasso prc plot (?)}\\
-
-\anna{ggmix vs. lmmlasso time plot}- I have this, make it pretty\\
 
@@
 
+%------------------------------%
+\subsection{Re-estimation of $\eta$ may, in some cases, slightly improve our estimate of $\hat{\eta}$, but does not improve the estimation of $\beta$, and takes a long time}
+Put this at the end of results
+%------------------------------%
+
+\anna{We have been unable to find a scenario where ggmix outperforms the LMM-Lasso method implemented by \cite{Rakitsch2012}.
+Ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the matrix K.
+Although it is primarily intended for use in high dimensional settings, where it performs on par with LMM-Lasso, this is also the scenario where it is ~100 times slower than LMM-Lasso.
+}\\
+
+\anna{ggmix vs lmmlasso eta-hat plot} - I have this, make it pretty\\
+
+\anna{ggmix vs lmmlasso MSE plot}- I have this, make it pretty\\
+
+\anna{ggmix vs lmmlasso prc plot (?)}\\
+
+\anna{ggmix vs. lmmlasso time plot}- I have this, make it pretty\\
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-06 main.tex
Section 4 updates

@@
 \end{tikzpicture}
+\caption{Directed acyclic graph (DAG) of confounding due to population structure}
+\label{fig:ps_env}
 \end{figure}
@@
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-
 LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, in which the variance components and corresponding weights are estimated once under the assumption of a null model.
@@
 
-We generate these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian, African American, and Hispanic populations.
-\anna{cite Vanderwas data somehow, add n = ???
-for all three groups).} For each $\bX$ matrix, $\bZ$ is a corresponding incidence matrix of 1s and 0s identifying subpopulation membership.
-Several structures of the fixed $\bgamma$ were considered.
-\anna{We give more details about this in the results section.}
+We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian (\anna{n = ?}), African American (\anna{n = ?}), and Hispanic (\anna{n = ?}).
+\anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership.
+500 simulations were performed for each data type and parameter combination.
 
@@
 %------------------------------%
-\subsection{LMM-Lasso outperforms Lasso - \anna{keep low-dimensional results? Only 'real data' for now... could use VanderWas to generate a $y$...}}
-%------------------------------%
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.7]{figures/fig1a.png}
-    \caption{Low-dimensional setting ($n=200, p=100$)}
-    \label{fig:ld}
-\end{figure}
-
-%------------------------------%
 \subsection{PC-Lasso and LMM-Lasso outperform Lasso}
 %------------------------------%
+Firstly, these penalized regression methods were compared in the presence of varying levels of environmental confounding.
+Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.
+Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
+The top row of figure \ref{fig:mse} shows that when $\xi = 0$, and no environmental confounding is present, lasso and LMM-Lasso perform comparably.
+PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that the top 10 principal components do not explain any variability in the outcome, yet are included in the final model, since they are unpenalized.
+The bottom row of figure \ref{fig:mse} illustrates the resulting MSE when $\xi = 0.8$ and thus 40\% of the variability of the phenotype is attributable to environmental confounding.
+In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
+This improvement is more substantial in the independent subpoplulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.
+\anna{need details about the real data once it's run}.\\
+
+\anna{Do I need to discuss the large subpops, heterogeneous gamma here?} \\
+
+\anna{add high-dimensional real data case using vanderwas? add dashed line to legend = null model mse}\\
 \begin{figure}[H]
@@
     \includegraphics[scale = 0.7]{figures/fig1b.png}
-    \caption{High dimensional setting ($n=200, p=1000$)}
-    \label{fig:hd}
-\end{figure}
-
-\anna{add high-dimensional real data case using vanderwas?}
-
-%------------------------------%
-\subsection{Population structure}
-%------------------------------%
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.7]{figures/kinship1}
-    \caption{A few large subpopulations}
-    \label{fig:big_blocks}
-\end{figure}
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.7]{figures/kinship2}
-    \caption{Many small subpopulations}
-    \label{fig:small_blocks}
+    \caption{MSE for different data types in the presence and absence of environmental confounding.}
+    \label{fig:mse}
 \end{figure}
@@
 %------------------------------%
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations \anna{with moderate relatedness (?)} to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air qualitty example, and (2) a large number of small, closely related subpopulations intended to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.
+Figure \ref{fig:kinship} illustrates the simulated kinship structures in these scenarios, \anna{as well as that present in the empirical SNP data}.
+We note that we show results for simulated data generated such that subpopulations were of equal size.
+Those of unequal size were also considered, but did not result in substantively different results (not shown - could be in supplement).
+
+\begin{figure}[H]
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=.7\linewidth]{figures/kinship1}  
+  \caption{Independent subpopulations simulated kinship; a small number of large groups}
+  \label{fig:large_groups}
+\end{subfigure}
+\newline
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.7\linewidth]{figures/kinship2}  
+  \caption{Independent subpopulations simulated kinship; a large number of small groups}
+  \label{fig:small_groups}
+\end{subfigure}
+\newline
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.7\linewidth]{figures/vanderwas_kinship.pdf}  
+  \caption{\anna{Vanderbilt data genome-wide} kinship}
+  \label{fig:vandy_groups}
+\end{subfigure}
+\caption{\anna{obviously these figures need to be tweaked:sizes, uniform scales} \\ \anna{Should kinship in the first figure be lower - more moderate?}}
+\label{fig:kinship}
+\end{figure}
+
+Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
+\anna{Due to constraints within the BNPSD package and the nature of the admixed population model they use (???), only the independent subpopulations method was able to simulate a large number of small subpopulations.
+For this reason, the presented simulated results all correspond to this simulation method from the bnpsd package.} Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
+The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.
+In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line.
+However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations.
+IN the presence of many small subpopulations, however, the entire box and error bars remain above the 0 line, indicating the clear superior estimation accuracy of LMM-Lasso in this context.
+\cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
+As such, it makes sense that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the large subpopulation setting, but fall short in the presence of many small subgroups.
+\anna{Additionally, its performance in this context may also me augmented by the fact that the subpopulations are more discretized  easily distinguishable based on their observed genetic data in the independent subpopulation simulated data.
+Although small subpopulations are unable to be simulated in the BNPSD admixed subpopulation setting, the corresponding results for the large subgroups in the admixed simulated data show much more comparable results between PC-Lasso and LMM-Lasso.
+- do this}.
+\anna{Vanderwas?}
+
+\anna{Should Vanderwas simulated data be here?}
 \begin{figure}[H]
@@
 %------------------------------%
-\subsection{$\gamma$ structure}
+\subsection{LMM-Lasso outperforms PC-Lasso when $\gamma$ is heterogeneous}
 %------------------------------%
+
+In addition to considering the effect of population structure, or forms of $\bZ$, on these methods' estimation accuracy, we also evaluated the effect of varying forms of $\bgamma$, the subpopulation-specific effect sizes.
+Our initial thinking was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that do not lend themselves to linear approximations.
+To that end we investigated four distinct $\bgamma$ structures: linear homogeneous, linear heterogeneous, nonlinear homogeneous, and nonlinear heterogeneous, \anna{names should change...
+half and half, not exponential is the better nonlinear setting}.
+In our naming convention, linear/nonlinear refers to whether the coefficients follow a linear pattern versus a nonlinear form, while homogeneous/heterogeneous indicates whether coefficients have the same, or differing signs, respectively.
+For example, (1, -2, 3, -4) is an example of a linear-heterogeneous structure, whereas (2, 2, 1, 1) is a nonlinear-homogeneous form.
+Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures.
+\\
+
+\anna{put half and half in this plot instead of exponential? In addition?}
 \begin{figure}[H]
     \centering
-    \includegraphics[scale = 0.7]{figures/gamma_structure.png}
+    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
     \caption{$\bgamma$ structures}
@@
 
-%------------------------------%
-\subsection{LMM-Lasso outperforms PC-Lasso when $\gamma$ is heterogeneous}
-%------------------------------%
+We thus evaluate PC-Lasso and LMM-Lasso in the presence of large and small subpopulations as well as distinct coefficient structures.
+Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) for all simulations, broken down by both subpopulation and coefficient structure.
+The right-hand panels of \ref{fig:big_small_gamma} are consistent with figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of small subpopulations, regardless of coefficient linearity or homogeneity.
+The panel of figure \ref{fig:big_small_gamma} corresponding to large subpopulations with heterogeneous coefficients also appears to be a setting where LMM-Lasso consistently outperforms PC-Lasso, regardless of coefficient linearity.
+Only in the setting of large supopulations with homogeneous coefficients does PC-Lasso outperform LMM-Lasso.
+In line with our initial intuition, PC-Lasso performs best within this setting when coefficients are linear in nature.
+
 \begin{figure}[H]
@@
     \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
-    \label{fig:my_label}
+    \label{fig:big_small_gamma}
 \end{figure}
@@
 \subsection{Re-estimation of $\eta$ may, in some cases, slightly improve our estimate of $\hat{\eta}$, but does not improve the estimation of $\beta$, and takes a long time}
-Put this at the end of results
 %------------------------------%
 
-\anna{We have been unable to find a scenario where ggmix outperforms the LMM-Lasso method implemented by \cite{Rakitsch2012}.
-Ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the matrix K.
-Although it is primarily intended for use in high dimensional settings, where it performs on par with LMM-Lasso, this is also the scenario where it is ~100 times slower than LMM-Lasso.
-}\\
+In section 3 briefly mentioned the existence of to LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+The LMM-Lasso results shown so far are those of LMM-Lasso-Rakitsch, though they are representative of LMM-Lasso-ggmix.
+The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods which aim to correct for population structure and environmental heterogeneity.
+However, we also want to explain why we used LMM-Lasso-Rakitsch, and provide guidance for potential users of either method.
+
+LMM-Lasso-Rakitsch estimates the variance components once, under the assumption of a null model.
+LMM-Lasso-ggmix seeks to improve upon this by re-estimating the variance components and $\bbeta$ iteratively.
+Although this underlying idea seems reasonable, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$ or $\eta$.
+
+\anna{ggmix vs lmmlasso MSE plot}- I have this, make it pretty\\
 
@@
 
-\anna{ggmix vs lmmlasso MSE plot}- I have this, make it pretty\\
-
-\anna{ggmix vs lmmlasso prc plot (?)}\\
-
-\anna{ggmix vs. lmmlasso time plot}- I have this, make it pretty\\
+In addition, we have found that LMM-Lasso-ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the RRM.
+Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners.
+However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, \anna{as shown in two previous mse figs}, this is also a setting in which the computational time required may be prohibitive for practical applications.
+Figure \anna{time fig} shows a ratio of the computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, where the horizontal dashed line at 1 indicates both methods require the same computing time.
+In high dimenions, where LMM-Lasso-ggmix performs well, it is between \anna{something and something else} times slower than LMM-Lasso-Rakitsch.
+Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method.
+We have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM - or something more fun}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.
 

### 2020-06-09 lmm-genetic-review.tex
Section 4 updates
Updated abstract and discussion.

@@
 \usepackage{tikz}
+\usepackage{subcaption}
+\usepackage{url}
 \allowdisplaybreaks
@@
 \begin{abstract}
-\note{Needs rewriting?}
-\anna{Yes, probably, but I'd like to get a better sense of how the rest of the paper shakes out before trying to summarize it.}
 Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to confounding due to unobserved factors.
 This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
-Linear mixed models (LMMs) are an attractive method to correct for unobserved confounding.
-These methods simultaneously correct for relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data.
-However, population structure itself does not confound the phenotype-genotype relationship, rather differential environmental exposures among subgroups do.
-Population structure may or may not serve as a good proxy for such differences.
-Given these subtle distinctions in confounding sources, the ability of LMMs to accurately estimate fixed, sparse genetic effects has not been well studied.
-Considering this, we evaluate the performance of penalized LMMs in terms of MSE and false positive rates in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.
+Penalized linear mixed models (LMMs) are an attractive method to correct for unobserved confounding.
+These methods correct for varying levels of relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data.
+Despite extensive discussion and simulation studies of penalized LMMs in the literature, the non-genetic mechanism by which population stratification confounds the phenotype-genotype relationship is rarely discussed or treated explicitly.
+As such, the ability of LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
+Considering this, we evaluate the performance of several penalized regression methods in terms of MSE in presence of a number of confounding structures.
+% in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.
 \end{abstract}

### 2020-06-09 main.tex
Updated abstract and discussion.

@@
 The remainder of this paper is organized as follows.
-In section \ref{structureSources} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
-We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \cite{Sillanpaeae2011, sul2018population} \anna{other papers?
-NB: re-read Sul, 2018 paper - a lot of similar review concepts}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures.
+In section 2 we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
+We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \cite{Sillanpaeae2011, sul2018population}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures.
 In so doing, we refine the definition of confounding due to population structure in terms of a non-genetic mechanism.
-In section \ref{methodsHistorical} we briefly review historical methods of correcting for population stratification and relatedness.
-In section \ref{methodsFixedRandom} we review methods of correcting for population stratification as a fixed versus random effect.
-These correspond to PCA adjusted lasso penalized regression and lasso penalized multivariate LMMs, respectively, the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present.
-In addition, we review the statistical details of the consequences of population structure and heterogeneous environmental exposures for these methods.
-In sections \ref{simulation} and \ref{results}, we illustrate the concepts described in section \ref{methodsFixedRandom} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
-These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized LMMs.
+In addition, we briefly review historical methods of correcting for population stratification and relatedness.
+In section 3 we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-Lasso) and lasso penalized multivariate LMMs (LMM-Lasso).
+We review the statistical details of these methods as well as the consequences of the ways in which they attempt to correct for population structure and environmental heterogeneity.
+In section 4, we illustrate the concepts described in section 3 via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
+These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods.
+In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
 
@@
 
-\anna{DAG here with only population stratification, observed snps, and phenotype ... maybe an arrow from population stratification to phenotype with a question mark (?)}
+% \anna{DAG here with only population stratification, observed snps, and phenotype ... maybe an arrow from population stratification to phenotype with a question mark (?)}
 
@@
 
-\anna{description of DAG below, in reference to previous one}
+% \anna{description of DAG below, in reference to previous one}
 
@@
 
-\anna{An additional consideration in genetic studies that we briefly review  is that of genetic background.
+An additional consideration in genetic studies that we will briefly mention is that of genetic background.
 Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
@@
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
-}
-
-\anna{
-This work will focus on environmental confounding, however, we wish to acknowledge and describe genetic background as a potential confounder for the sake of completeness. 
-}
+While we acknowledge the challenges presented by genetic background, this work will focus on correction methods for confounding due to environmental heterogeneity, for which genetic similarity may serve as a proxy.
 
@@
 These effects are integrated out in order to estimate a genetic variance component, which in turn, is used to quantify the total narrow-sense heritability of a trait \cite{yang2010common}.
-\anna{We will briefly return to a discussion of GCTA and related methods later, though techniques for estimating heritability are not the focus of this paper.}
 
@@
 These SNPs will be included in the model as both fixed and random effects, while SNPs with smaller effect sizes are modeled only as part of the random effect in order to control for environmental confounding.
-\\
 
-\anna{say more about logistic?}\\
 The remainder of this review will focus on the performance of multivariate Lasso-penalized methods for the analysis of continuous, normally distributed outcomes.
@@
 Using simulations, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
-\\
-
-\anna{Do we want to include the sign rate stuff?}
 
 Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \cite{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\
-
-\anna{find that 2020 paper that simulates the data sort of similarly}
 
@@
 \anna{do I need to define the ell2 norm?} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
-
-%which requires estimation of $p + k + 1$ parameters.
-Determining the appropriate number of principal components to include is a non-trivial task.
-However, often in practice the first 10 principal components are used.
-\anna{and so that's what we do in all our analyses.}
+Modeling unobserved confounding using fixed effects requires the estimation of $p + k + 1$ parameters.
 
@@
 \subsection{LMM-Lasso}
-LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$.
+LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$.
 Often $\bZ$ is not observed directly, and a common convention is to replace $\bZ$ with $\bX$, yielding the random effect $\bX\bgamma \sim N(\boldsymbol{0}, \sigmagg \textbf{K})$.
-\anna{Equivalently, $\bX\bgamma$ may be denoted $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.} This results in the following LMM regression model
+In the literature \cite{Rakitsch2012}, these unobserved confounding effects may equivalently denoted as $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.
+This results in the following LMM regression model
 
@@
 
-\anna{(However, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...)}
+\anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
 
@@
     \bV^{-1} &= (\sigmagg \bK + \sigmaee \bI)^{-1}\\
-    &=(\sigmagg \boldsymbol{U D^2 U}^T + \sigmaee \bU \bU^T)^{-1}\\
-    &= (\bU (\sigmagg \boldsymbol{D}^2 + \sigmaee \bI) \bU^T)^{-1}\\
-    &=\bU (\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1} \bU^T.
+    &=(\sigmagg \bU \bD^2 \bU^T + \sigmaee \bU \bU^T)^{-1}\\
+    &= (\bU (\sigmagg \bD^2 + \sigmaee \bI) \bU^T)^{-1}\\
+    &=\bU (\sigmagg \bD^2 + \sigmaee \bI)^{-1} \bU^T.
 \end{align*}
@@
 \begin{equation}
-\label{eqn:lmmObj}
-%Q(\boldsymbol{\beta| \widetilde{X}, \widetilde{y}, w}) = ||\Sigma_{i=1}^n w_i ({\widetilde{y}_i} - {\boldsymbol{\widetilde{x}}_i^T \boldsymbol{\beta}})||_2^2 + \lambda || \boldsymbol{\beta} ||_1
+\label{eqn:lmm_obj}
 Q(\bbeta | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
@@
 
-\anna{
-Modeling population structure using a random effect, as opposed to fixed effects as in PC-Lasso, requires the estimation of substantially fewer parameters ($p + 2$ for LMM-Lasso-Rakitsch, $p + 3$ for LMM-Lasso-ggmix).
-}
+An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of substantially fewer parameters: $p + 2$ for LMM-Lasso-Rakitsch, and $p + 3$ for LMM-Lasso-ggmix.
 
@@
 This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \cite{ggmix, bhatnagar2019simultaneous}.
-\anna{We postpone our discussion of the differences in performance of each method to the results section.}
+We postpone further discussion of the differences between these methods to the results section.
 
 %------------------------------%
-\subsubsection{\anna{Intuition from GLS (?) p-values and SEs in response to rotation?}}
+% \subsubsection{\anna{Intuition from GLS (?) p-values and SEs in response to rotation?}}
 
@@
 
-where $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability \anna{also SNR...} and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-\anna{We note that this is similar to the simulation technique employed by THAT 2020 PAPER FIND IT...}
+where $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
+Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
 
-We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian (\anna{n = ?}), African American (\anna{n = ?}), and Hispanic (\anna{n = ?}).
+We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
 \anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership.
@@
 In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
-This improvement is more substantial in the independent subpoplulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.
-\anna{need details about the real data once it's run}.\\
+This improvement is more substantial in the independent subpopulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.
 
-\anna{Do I need to discuss the large subpops, heterogeneous gamma here?} \\
+\anna{
+\begin{itemize}
+    \item Need details about the real data once it's run
+    \item Update figure with Vanderwas-generated simulation data
+    \item Add dashed line to legend (dashed = null model mse)
+    \item Should I mention subpopulation sizes (a few large ones) or gamma structure (half-and-half-heterogeneous) here? Those are covered in more detail later.
+\end{itemize}
+}
 
-\anna{add high-dimensional real data case using vanderwas? add dashed line to legend = null model mse}\\
 \begin{figure}[H]
     \centering
-    \includegraphics[scale = 0.7]{figures/fig1b.png}
+    \includegraphics[scale = 0.9]{figures/fig1b.png}
     \caption{MSE for different data types in the presence and absence of environmental confounding.}
@@
 %------------------------------%
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations \anna{with moderate relatedness (?)} to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air qualitty example, and (2) a large number of small, closely related subpopulations intended to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.
-Figure \ref{fig:kinship} illustrates the simulated kinship structures in these scenarios, \anna{as well as that present in the empirical SNP data}.
-We note that we show results for simulated data generated such that subpopulations were of equal size.
-Those of unequal size were also considered, but did not result in substantively different results (not shown - could be in supplement).
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.
+Figure \ref{fig:kinship} illustrates the simulated kinship structures in these scenarios, as well as that present in the empirical SNP data.
+We note that we show results for simulated data generated such that subpopulations are of equal size, while the empirical SNP data illustrates a scenario of unequal subpopulation sizes.
+
+\anna{
+\begin{itemize}
+    \item "Spend" this many figures on kinship?
+\end{itemize}
+}
 
@@
 \end{subfigure}
-\caption{\anna{obviously these figures need to be tweaked:sizes, uniform scales} \\ \anna{Should kinship in the first figure be lower - more moderate?}}
+\caption{\anna{obviously these figures need to be tweaked:sizes, uniform scales}}
 \label{fig:kinship}
@@
 Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
-\anna{Due to constraints within the BNPSD package and the nature of the admixed population model they use (???), only the independent subpopulations method was able to simulate a large number of small subpopulations.
-For this reason, the presented simulated results all correspond to this simulation method from the bnpsd package.} Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
 The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.
@@
 However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations.
-IN the presence of many small subpopulations, however, the entire box and error bars remain above the 0 line, indicating the clear superior estimation accuracy of LMM-Lasso in this context.
+In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
 \cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
-As such, it makes sense that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the large subpopulation setting, but fall short in the presence of many small subgroups.
-\anna{Additionally, its performance in this context may also me augmented by the fact that the subpopulations are more discretized  easily distinguishable based on their observed genetic data in the independent subpopulation simulated data.
-Although small subpopulations are unable to be simulated in the BNPSD admixed subpopulation setting, the corresponding results for the large subgroups in the admixed simulated data show much more comparable results between PC-Lasso and LMM-Lasso.
-- do this}.
-\anna{Vanderwas?}
+As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
+Additionally, its performance in this context may also be augmented by the fact that the subpopulations are more discretized  easily distinguishable based on their observed genetic data in the independent subpopulation simulated data.
+The BNPSD 1-dimensional admixture simulation method is unable to accommodate a large number of small subpopulations.
+\anna{However, the results for the few large subgroups in the 1-dimensional admixture simulated data show more comparable results between PC-Lasso and LMM-Lasso.
+} \anna{(I still need to run these sims and add the results to the figure below)}
 
-\anna{Should Vanderwas simulated data be here?}
+\anna{
+\begin{itemize}
+    \item Vanderwas data to show results for subpopulations of unequal size
+    \item results for BNPSD 1d admixture data with large subgroups
+\end{itemize}
+}
+
 \begin{figure}[H]
@@
 Our initial thinking was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that do not lend themselves to linear approximations.
-To that end we investigated four distinct $\bgamma$ structures: linear homogeneous, linear heterogeneous, nonlinear homogeneous, and nonlinear heterogeneous, \anna{names should change...
-half and half, not exponential is the better nonlinear setting}.
+To that end we investigated four distinct $\bgamma$ structures: linear homogeneous, linear heterogeneous, nonlinear homogeneous, and nonlinear heterogeneous, \anna{(names probably need to change.
+We've moved to half-and-half, not exponential, as a more illustrative nonlinear setting)}.
 In our naming convention, linear/nonlinear refers to whether the coefficients follow a linear pattern versus a nonlinear form, while homogeneous/heterogeneous indicates whether coefficients have the same, or differing signs, respectively.
@@
 
-\anna{put half and half in this plot instead of exponential? In addition?}
+\anna{
+\begin{itemize}
+    \item Replace exponential with half-and-half
+\end{itemize}
+}
+
 \begin{figure}[H]
@@
 The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods which aim to correct for population structure and environmental heterogeneity.
-However, we also want to explain why we used LMM-Lasso-Rakitsch, and provide guidance for potential users of either method.
+However, we also wish to briefly discuss the key differences between methods, and provide guidance for potential users of either method.
 
@@
 LMM-Lasso-ggmix seeks to improve upon this by re-estimating the variance components and $\bbeta$ iteratively.
-Although this underlying idea seems reasonable, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$ or $\eta$.
-
-\anna{ggmix vs lmmlasso MSE plot}- I have this, make it pretty\\
-
-\anna{ggmix vs lmmlasso eta-hat plot} - I have this, make it pretty\\
+Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$ or $\eta$.
+\anna{
+\begin{itemize}
+    \item ggmix vs lmmlasso MSE plot will go here once run and pretty
+    \item ggmix vs lmmlasso eta-hat plot will go here once run and pretty
+\end{itemize}
+}
 
@@
 Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners.
-However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, \anna{as shown in two previous mse figs}, this is also a setting in which the computational time required may be prohibitive for practical applications.
-Figure \anna{time fig} shows a ratio of the computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, where the horizontal dashed line at 1 indicates both methods require the same computing time.
-In high dimenions, where LMM-Lasso-ggmix performs well, it is between \anna{something and something else} times slower than LMM-Lasso-Rakitsch.
-Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method.
-We have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM - or something more fun}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.
+However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in \anna{figure xxx}, this is also a setting in which the computational time required may be prohibitive for practical applications.
+Figure \anna{xxx} shows a ratio of the computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, where the horizontal dashed line at 1 indicates both methods require the same computing time.
+In high dimenions, where LMM-Lasso-ggmix performs well, it is \anna{approximately X} times slower than LMM-Lasso-Rakitsch.
+Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{itemize}
-    \item LMM-Lasso outperforms PC-Lasso when (1) subpopulations are small and (2) when confounding effects are nonlinear and heterogeneous
-    \item LMM-Lasso \cite{Rakitsch2012} and ggmix  \cite{bhatnagar2019simultaneous} perform similarly in many scenarios
-    \item However, ggmix is time-consuming and may not be practical for more realistic genetic data sets
-    \item Both PC-Lasso and LMM-Lasso methods suffer in estimation accuracy due to their bias from $\lambda$
-    \item Bias-reduction techniques for penalized LMMs may be an attractive avenue for future work
-\end{itemize}
+In this work, we have reviewed the concept of population stratification and defined explicitly the non-genetic, environmental mechanism by which it may contribute to confounding the genotype-phenotype relationship.
+We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
+In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.
+
+We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively.
+We have also detailed the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity.
+In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
+We then compare the estimation accuracy of a naive lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
+
+Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the naive lasso whenever environmental confounding effects are present.
+However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the naive lasso.
+This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
+Additionally, we find that LMM-Lasso outperforms PC-Lasso when (1) small genetic subpopulations are present which correspond to environmental heterogeneity and (2) when confounding effects are nonlinear and heterogeneous in nature.
+In the presence of large subpopulations which correspond to homogeneous, linear environmental effects, PC-Lasso and LMM-Lasso perform comparably in estimation accuracy.
+While this scenario may be plausible, given that LMM-Lasso is more robust to subpopulation sizes and nonlinear effects, it may present a 'safer choice' for practitioners seeking to obtain unbiased estimates when the exact nature of population structure and corresponding environmental heterogeneity are unknown.
+
+We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
+Although we found that both methods perform similarly in terms of estimation accuracy in many scenarios, LMM-Lasso-ggmix has several distinct disadvantages.
+Firstly, it suffers from instability in low dimensions.
+Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix takes approximately \anna{x} times longer due to its iterative estimation procedure.
+Thirdly, while the additional time required of LMM-Lasso-ggmix could be worthwhile if it resulted in a more accurate estimation of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are superior to those of LMM-Lasso-Rakitsch.
+We have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM (final name tbd)}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.
+
+Even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
+Because of this, bias-reduction techniques for penalized LMMs may be an attractive avenue for future research.

### 2020-06-17 main.tex
Updating figures

@@
 
+Overview of simulation... briefly define gamma structure, subpops size... more details in supplement...
+
 %------------------------------%
@@
     \item Need details about the real data once it's run
-    \item Update figure with Vanderwas-generated simulation data
-    \item Add dashed line to legend (dashed = null model mse)
-    \item Should I mention subpopulation sizes (a few large ones) or gamma structure (half-and-half-heterogeneous) here? Those are covered in more detail later.
+    \item Should I mention subpopulation sizes (a few large ones) or gamma structure (half-and-half-heterogeneous) here? No - do a brief summary where noted above
 \end{itemize}
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/fig1b.png}
+    \includegraphics[scale = 1.1]{figures/beta_mse.png}
     \caption{MSE for different data types in the presence and absence of environmental confounding.}
@@
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
-Additionally, its performance in this context may also be augmented by the fact that the subpopulations are more discretized  easily distinguishable based on their observed genetic data in the independent subpopulation simulated data.
-The BNPSD 1-dimensional admixture simulation method is unable to accommodate a large number of small subpopulations.
-\anna{However, the results for the few large subgroups in the 1-dimensional admixture simulated data show more comparable results between PC-Lasso and LMM-Lasso.
-} \anna{(I still need to run these sims and add the results to the figure below)}
 
@@
 \begin{itemize}
-    \item Vanderwas data to show results for subpopulations of unequal size
-    \item results for BNPSD 1d admixture data with large subgroups
+    \item I ran this but think it just makes things more confusing - go back to only independent subpopulations data - for both (?) 
 \end{itemize}
@@
     \centering
-    \includegraphics[scale = 0.7]{figures/fig2a}
+    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
+     \includegraphics[scale = 0.9]{figures/fig2a.png}
     \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy.}
@@
 Our initial thinking was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that do not lend themselves to linear approximations.
-To that end we investigated four distinct $\bgamma$ structures: linear homogeneous, linear heterogeneous, nonlinear homogeneous, and nonlinear heterogeneous, \anna{(names probably need to change.
-We've moved to half-and-half, not exponential, as a more illustrative nonlinear setting)}.
+To that end we investigated six distinct $\bgamma$ structures: linear homogeneous, linear heterogeneous, nonlinear homogeneous, nonlinear heterogeneous, exponential homogeneous, and exponential heterogeneous.
 In our naming convention, linear/nonlinear refers to whether the coefficients follow a linear pattern versus a nonlinear form, while homogeneous/heterogeneous indicates whether coefficients have the same, or differing signs, respectively.
@@
 \\
-
-\anna{
-\begin{itemize}
-    \item Replace exponential with half-and-half
-\end{itemize}
-}
 
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/fig2b.png}
+    % \includegraphics[scale = 0.9]{figures/fig2b.png}
+    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
     \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
@@
 %------------------------------%
-\subsection{Re-estimation of $\eta$ may, in some cases, slightly improve our estimate of $\hat{\eta}$, but does not improve the estimation of $\beta$, and takes a long time}
+\subsection{Different implementations of LMM-Lasso}
 %------------------------------%
@@
 \begin{itemize}
-    \item ggmix vs lmmlasso MSE plot will go here once run and pretty
-    \item ggmix vs lmmlasso eta-hat plot will go here once run and pretty
+    \item ggmix vs lmmlasso MSE plot will go here once run and pretty: options
 \end{itemize}
 }
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 0.9]{figures/eta_hat.png}
+    \includegraphics[scale = 0.9]{figures/beta_hat.png}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/eta_beta_hat.png}
+\end{figure}
+
+\newpage
+\anna{The error bars are indicating differences across different values of xi and simulated data-types (indepent and 1d linear admixture - this may be confusing/something to change or describe explicitly.
+}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 0.8]{figures/time_ratio.png}
+    % \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
+    \label{fig:time}
+\end{figure}
 
@@
 However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in \anna{figure xxx}, this is also a setting in which the computational time required may be prohibitive for practical applications.
-Figure \anna{xxx} shows a ratio of the computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, where the horizontal dashed line at 1 indicates both methods require the same computing time.
-In high dimenions, where LMM-Lasso-ggmix performs well, it is \anna{approximately X} times slower than LMM-Lasso-Rakitsch.
+Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.
+The horizontal dashed line at 1 indicates both methods require the same computing time.
+Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes approximately 71 to 9 times as long as LMM-Lasso-Rakitsch.
 Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.

### 2020-06-18 lmm-genetic-review.tex
Edits to Introduction

@@
 \usetikzlibrary{arrows}
-\providecommand{\note}[1]{\textcolor{red}{#1}}
+\providecommand{\pb}[1]{\textcolor{red}{#1}}
 \providecommand{\anna}[1]{\textcolor{blue}{#1}}
@@
 \begin{abstract}
-Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to confounding due to unobserved factors.
+Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to confounding due to unobserved factors \pb{``Unobserved factors'' is a bit vague -- think we need to work ``environment'' into this first sentence}.
 This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
@@
 These methods correct for varying levels of relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data.
-Despite extensive discussion and simulation studies of penalized LMMs in the literature, the non-genetic mechanism by which population stratification confounds the phenotype-genotype relationship is rarely discussed or treated explicitly.
+Despite extensive discussion and simulation studies of penalized LMMs in the literature \pb{is the literature on penalized LMMs really ``extensive''?
+I would say that the literature on penalized regression is extensive, and on LMMs is extensive, but not actually that much on penalized LMMs -- hence the need for this review}, the non-genetic mechanism by which population stratification confounds the phenotype-genotype relationship is rarely discussed or treated explicitly.
 As such, the ability of LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.

### 2020-06-18 main.tex
Experimental push
Edits to Introduction

@@
 This occurs when the variant being tested is correlated with other variants, so that the effects of these correlated variants are erroneously attributed to the variant in question.
-One method for analyzing data in the presence of LD is pruning, where only uncorrelated SNPs are used in modeling or testing procedures.
-Pruning limits analysis to SNPs which are representative of a particular region, and may result in improved predictive accuracy.
-However, pruning is limited in that it can only identify SNPs in LD with the true causal SNP.
-In addition, it has been shown that a worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \cite{prive2018efficient}.
+One method for analyzing data in the presence of LD \pb{LD needs to be defined} is pruning, where only uncorrelated SNPs are used in modeling or testing procedures.
+Pruning limits analysis to SNPs which are representative of a particular region, and may result in improved predictive accuracy \pb{is predictive accuracy the main reason for pruning?
+Why does it improve prediction?}.
+However, pruning is limited in that it can only identify SNPs in LD with the true causal SNP \pb{I don't understand this sentence; wouldn't this limitation apply to all methods?}.
+In addition, it has been shown that in the worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \cite{prive2018efficient}.
 
 An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
-However, such methods rely on data where the number of observations is less than the number of of features.
-This is often infeasible in genetic studies where the number of features frequently exceeds the number of observations.
-Penalized regression methods make such problems tractable, by constraining the solution space of coefficient estimates.
+However, such methods rely on data where the number of observations is less than the number of features.
+This is often infeasible in genetic studies where the number of features typically exceeds the number of observations.
+Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
 
 The Lasso \cite{tibshirani1996regression} is a widely used penalized regression method.
-It relies on an assumption of sparsity, which allows for computational speedup while performing coefficient estimation and variable selection simultaneously.
+It relies on an assumption of sparsity, which means that coefficient estimation and variable selection can be accomplished simultaneously, and also provides considerable computational advantages, meaning that it scales up well to very large data sets.
 These features make the Lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
 
-Lasso penalized regression models assume independent observations and uncorrelated errors.
+Lasso penalized regression models typically assume independent observations and uncorrelated errors.
 In the presence of population structure, some samples are more correlated than others, violating the independent errors assumption.
-If left unaccounted for, this presence of such dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \cite{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, there remain open and relevant problems in analyzing structured genetic data \cite{lawson2019population, barton2019population}.
+If unaccounted for, this dependency can result in the identification of spurious associations.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \cite{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, there remain open and relevant problems \pb{this is vague} in analyzing structured genetic data \cite{lawson2019population, barton2019population}.
 
@@
 Using the spectral decomposition of $\bK$ described in equation \ref{eqn:k}, we can gain further insight into the mechanism of this transformation.
-Note that $\bV^{-1}$ can be expressed using the following matrix factorizations
+Note that $\bV^{-1}$ can be expressed using the following matrix factorization \pb{ONLY WORKS IF $\bX$ IS FULL RANK}
 

### 2020-06-18 main.tex
Changing citation style; edits for 2.1 and 2.2

@@
 The presence of multiple distinct ancestry groups is known as population stratification.
+\pb{Should some of these terms, like cryptic relatedness, be bolded?}
 
@@
 However, in this review we focus on methods to account for unobserved relatedness.
-Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are be present.
+Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are present.
 Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations.
@@
 
-As an example, consider a genetic study to assess genetic variants associated with lung cancer in a sample comprised of subjects from two distinct subpopulationsk A and B.
+As an example, consider a genetic study to assess genetic variants associated with lung cancer in a sample comprised of subjects from two distinct subpopulations A and B.
 Assume the major allele of SNP $X$ is present with higher frequency in subpopulation A compared to subpopulation B, but it has no direct effect on lung cancer.
 Now suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality.
-A GWAS analysis of data from these subpopulations would find SNP $X$ to be significantly associated with lung cancer.
+Consequently, a GWAS analysis of data from these subpopulations would be predisposed to find SNP $X$ significantly associated with lung cancer.
 This spurious association is not identified because of the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
@@
 Additionally, we assume that the environmental exposure has no effect on on observed SNPs.
+\pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.}
 
@@
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
-While we acknowledge the challenges presented by genetic background, this work will focus on correction methods for confounding due to environmental heterogeneity, for which genetic similarity may serve as a proxy.
+While we acknowledge the challenges presented by genetic background, this work will focus on correction methods for confounding due to environmental heterogeneity, for which genetic similarity may serve as a proxy.\pb{Overall comment about this paragraph: I feel like we *are* addressing genetic background, by conducting a genomewide multivariate analysis.
+I think its worth pointing out that this is not the only way to deal with genetic background, but it's not as though we're not addressing it at all.}
 

### 2020-06-18 main.tex
Edits to 2.3

@@
 In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
+\pb{don't hard-code ``section 3''; use ref and label instead}
 
@@
 Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts.
-It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs.
+It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs \pb{Shouldn't this be: between environment and SNPs?}.
 
-The genomic control method is a univariate testing procedure which modifies the Cochran-Armitage test for trend by a correction factor.
+The genomic control method is a univariate testing procedure which modifies the Cochran-Armitage test for trend by a correction factor \pb{genomic control is not specific to trend tests; it can be used for any type of test statistic inflation}.
 This correction attempts to quantify and remove the effect of population structure on the distribution of the chi-square statistic using markers unrelated to the phenotype of interest and independent of the marker being tested for association \cite{devlin1999genomic, bacanu2000power, wang2009testing}.
-The genomic control method retains the disadvantages of univariate SNP testing and its power to detect true associations is likely to suffer in the presence of linkage disequilibrium.
+The genomic control method retains the disadvantages of univariate SNP testing and its power to detect true associations is likely to suffer in the presence of linkage disequilibrium \pb{reference?}.
+\pb{This paragraph is a little too negative about genomic control...needs to be softened slightly}
 
-The structured association method attempts to infer subject membership within a discrete number of nonoverlapping subpopulations.
+Stratified analysis attempts to infer subject membership within a discrete number of nonoverlapping subpopulations.
 Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \cite{pritchard1999use, pritchard2000association}.
@@
 This assumption may not accurately reflect the characteristics of a particular sample, such as one consisting of admixed populations, or recent relatedness.
-Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall somewhere between these two extremes.
-By clustering those individuals within in one subpopulation or another we lose information about more high dimensional relatedness.
+Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall somewhere between these two extremes \pb{seems useful to bring in idea of a gradient, or continuously varying relatedness, here}.
+By clustering those individuals within in one subpopulation or another we lose information about more high dimensional relatedness \pb{I think ``high dimensional relatedness'' is confusing/ambiguous here.
+What do you mean by it?}.
 
@@
 
-Like PC-adjustment methods, LMMs rely on a pair-wise similarity matrix to estimate and subsequently correct for relatedness among subjects.
+Like PC-adjustment methods, LMMs rely on a pair-wise similarity matrix \pb{as we discussed, this is a bit odd to say here} to estimate and subsequently correct for relatedness among subjects.
 However, the term LMM has been used to describe a number of distinct procedures for analyzing structured genetic data which differ in their objectives and methods.
@@
 
-Within the category of LMMs that aim to estimate SNP effects are univariate and multivariate approaches.
-Univariate approaches attempt to estimate univariate SNP effects on phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure using a genetic similarity matrix \cite{yu2006unified, kang2010variance, kang2008efficient}.
+LMMs can also be used to estimate SNP effects, and this can be done in either a univariate or multivariate manner.
+Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure using a genetic similarity matrix \cite{yu2006unified, kang2010variance, kang2008efficient}.
 However, as with all univariate testing approaches univariate LMM implementations may still produce spurious associations and biased effect estimates in the presence of LD.
@@
 
-Using simulations, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
-
+In Section~\ref{Sec:res}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
 Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \cite{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\
@@
 
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
+To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?}.
 We denote the standardized version of this genotype matrix as $\bX$.
@@
 \section{Results}
+\label{Sec:res}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-18 main.tex
Edits to 3.1-3.4

@@
 
+\pb{Seems like we should explicitly include an intercept in all these models?}
+
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?}.
-We denote the standardized version of this genotype matrix as $\bX$.
+We denote the standardized version \pb{need more explicit details on standardization} of this genotype matrix as $\bX$.
 We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
@@
 where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that indicates population structure, $\bgamma$ is an $q \times 1$ vector of subpopulation-specific environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+\pb{This is a problem throughout, but ``where'' shouldn't be indented here.
+Either don't have a space between the equation and the next block of text, or use noindent}
 
-Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and relative relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
+Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and relative relationship matrix (RRM) \pb{RRM is the REALIZED relationship matrix, not relative}, $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
 Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
@@
 
-where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the eigenvalues of $\bX$ (equivalently, the singular values of $\bK$.) The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
+where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the eigenvalues of $\bX$ \pb{only square matrices have eigenvalues} (equivalently, the singular values of $\bK$.) The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
 
@@
 where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients.
-\anna{do I need to define the ell2 norm?} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
+\anna{do I need to define the ell2 norm?} \pb{YES} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
 Modeling unobserved confounding using fixed effects requires the estimation of $p + k + 1$ parameters.
@@
 \subsection{LMM-Lasso}
-LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$.
-Often $\bZ$ is not observed directly, and a common convention is to replace $\bZ$ with $\bX$, yielding the random effect $\bX\bgamma \sim N(\boldsymbol{0}, \sigmagg \textbf{K})$.
-In the literature \cite{Rakitsch2012}, these unobserved confounding effects may equivalently denoted as $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.
-This results in the following LMM regression model
+LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$ \pb{too many sub/superscripts for an in-line fraction}.
+Typically, $\bZ$ is not observed directly, and a common convention \citep[e.g.,][]{Rakitsch2012} is to replace $\bZ$ with $\bX$.
+Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.
+This results in the LMM regression model
 
@@
 
-where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\bX \bbeta, \bV := \sigmagg \textbf{K} + \sigmaee \textbf{I})$.
+\pb{need a comma after this equation, just like you would if it were text; this shows up several other places as well}
+
+where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
 LMM-Lasso then performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
@@
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, in which the variance components and corresponding weights are estimated once under the assumption of a null model.
+LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, in which the variance components and corresponding weights are estimated once under the assumption of a null model \pb{I think you should be more specific here about the ``variance components'', use the $\sigma_e^2$ notation you just introduced, etc.}.
 A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.

### 2020-06-23 lmm-genetic-review.tex
Changing citation style; edits for 2.1 and 2.2
Edits to 3.1-3.4
Updating abstract

--- 2020-06-09 lmm-genetic-review.tex
@@
 \allowdisplaybreaks
+\usepackage{natbib}
 %\usepackage{biblatex}
@@
 \newcommand{\bV}{\textbf{V}}
+\newcommand{\bu}{\textbf{u}}
 \newcommand{\bU}{\textbf{U}}
@@
 \begin{abstract}
-Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to confounding due to unobserved factors.
-This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
+Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to \anna{unobserved confounding factors, such as environmental heterogeneity.} This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
 Penalized linear mixed models (LMMs) are an attractive method to correct for unobserved confounding.
 These methods correct for varying levels of relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data.
-Despite extensive discussion and simulation studies of penalized LMMs in the literature, the non-genetic mechanism by which population stratification confounds the phenotype-genotype relationship is rarely discussed or treated explicitly.
-As such, the ability of LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
-Considering this, we evaluate the performance of several penalized regression methods in terms of MSE in presence of a number of confounding structures.
+\anna{Despite extensive literature on penalized regression, and LMMs, there is a need for a statistically rigorous review of penalized LMMs.
+Specifically, the ability of LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
+Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.} Considering this, we evaluate the performance of several penalized regression methods in terms of MSE in the presence of a number of confounding structures.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.
@@
 
-\bibliographystyle{abbrv}
+\bibliographystyle{ims}
 \bibliography{references}

### 2020-06-23 main.tex
Fixing overleaf warnings
Section 3 punctuation/indent fixes

@@
 
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?}.
-We denote the standardized version \pb{need more explicit details on standardization} of this genotype matrix as $\bX$.
+To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?} \anna{In order to show the equivalence of $u \sim N(0, \sigmagg \bK) \text{ and } Z\gamma \sim N(0, \sigmagg \bK)$ I think so, but maybe I can point out that this is the assumption the equivalence is arising from, but we don't' have to have discrete suppopulations in practice...}.
+We denote the standardized version \pb{need more explicit details on standardization} \anna{get details from Hayes 2009} of this genotype matrix as $\bX$.
 We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
-
 \begin{equation}
@@
 \end{equation}
+where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that indicates population structure, $\bgamma$ is an $q \times 1$ vector of subpopulation-specific environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
 
-where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that indicates population structure, $\bgamma$ is an $q \times 1$ vector of subpopulation-specific environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
-\pb{This is a problem throughout, but ``where'' shouldn't be indented here.
-Either don't have a space between the equation and the next block of text, or use noindent}
-
-Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and relative relationship matrix (RRM) \pb{RRM is the REALIZED relationship matrix, not relative}, $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
+Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
 Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
-
 \begin{align}
@@
                    &= \frac{1}{p} \bU \bD (\bU \bD)^T \nonumber \\
-                   &\propto \bR \bR^T
+                   &\propto \bR \bR^T,
 \end{align}
-
 where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the eigenvalues of $\bX$ \pb{only square matrices have eigenvalues} (equivalently, the singular values of $\bK$.) The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
@@
 It has been shown \cite{hoffman2013correcting} that this corresponds to the regression model
-
 \begin{equation}
@@
 \end{equation}
-
 where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \ref{eqn:model}.
 In a penalized regression framework, this model yields the objective function
-
 \begin{equation}
@@
 \end{equation}
-
 where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients.
@@
 This results in the LMM regression model
-
 \begin{equation}
     \label{eqn:lmm_reg}
-    \by = \bX\bbeta + \bX\bgamma+ \beps
+    \by = \bX\bbeta + \bX\bgamma+ \beps,
 \end{equation}
-
-\pb{need a comma after this equation, just like you would if it were text; this shows up several other places as well}
-
 where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
 LMM-Lasso then performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
-
 \begin{equation}
@@
 \end{equation}
-
 \anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
-
 Using the spectral decomposition of $\bK$ described in equation \ref{eqn:k}, we can gain further insight into the mechanism of this transformation.
 Note that $\bV^{-1}$ can be expressed using the following matrix factorization \pb{ONLY WORKS IF $\bX$ IS FULL RANK}
-
 \begin{align*}
@@
 \end{align*}
-
-\noindent By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$, $\byT = \bU^T \by$, and $\bW = \text{diag}\{w_1, ...
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$, $\byT = \bU^T \by$, and $\bW = \text{diag}\{w_1, ...
 w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1}$.
 Together, this yields the objective function
-
 \begin{equation}
@@
 \end{equation}
-
 An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of substantially fewer parameters: $p + 2$ for LMM-Lasso-Rakitsch, and $p + 3$ for LMM-Lasso-ggmix.
@@
 We can see this by considering the expectation and variance of low-dimensional OLS estimates of $\hat{\bbeta}$ based on model \ref{eqn:lmm_reg}:
-
 \begin{align*}
@@
     &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps)\\
-    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^T \bX)^{-1}
+    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^T \bX)^{-1}.
 \end{align*}
-
 Under the assumption that $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$, $\mathbb{E}(\hat{\bbeta}) = \bbeta, \mathbb{V}(\hat{\bbeta}) = \frac{\sigmagg}{p}\bI + \sigmaee (\bX^T \bX)^{-1}$.
@@
 Then,
-
 \begin{align*}
@@
 \end{align*}
-
 where $\boldsymbol{\tau}$ quantifies bias in the estimated $\boldsymbol{\hat{\beta}'}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
 Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
-
 \begin{equation}
     \label{eqn:sim}
-    \by = a \bX \bbeta + b \bZ \bgamma + c \beps 
+    \by = a \bX \bbeta + b \bZ \bgamma + c \beps,
 \end{equation}
-
 where $\by$ is a pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
-
 \begin{equation}
@@
 \end{equation}
-
-where $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
+Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
 Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
@@
 Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$ or $\eta$.
-\anna{
-\begin{itemize}
-    \item ggmix vs lmmlasso MSE plot will go here once run and pretty: options
-\end{itemize}
-}
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.9]{figures/eta_hat.png}
-    \includegraphics[scale = 0.9]{figures/beta_hat.png}
-\end{figure}
-
+\anna{eta part needs to be converted to mse}
 \begin{figure}[H]
@@
     \includegraphics[scale = 0.8]{figures/time_ratio.png}
-    % \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
+    \caption{Caption goes here}
     \label{fig:time}

### 2020-06-23 main.tex
Section 1 edits - in progress

@@
 This occurs when the variant being tested is correlated with other variants, so that the effects of these correlated variants are erroneously attributed to the variant in question.
-One method for analyzing data in the presence of LD \pb{LD needs to be defined} is pruning, where only uncorrelated SNPs are used in modeling or testing procedures.
-Pruning limits analysis to SNPs which are representative of a particular region, and may result in improved predictive accuracy \pb{is predictive accuracy the main reason for pruning?
-Why does it improve prediction?}.
-However, pruning is limited in that it can only identify SNPs in LD with the true causal SNP \pb{I don't understand this sentence; wouldn't this limitation apply to all methods?}.
+\anna{One method for analyzing data in the presence of correlated loci, or linkage disequilibrium (LD), is LD-pruning.
+In LD-pruning methods, SNPs are sequentially thinned, so that only SNPs correlated with each other below some desired threshold are used in modeling and testing procedures.} LD-pruning limits analysis to SNPs which are representative of a particular region.
+This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci can be identified and their effects estimated.
 In addition, it has been shown that in the worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \cite{prive2018efficient}.
+However, LD-pruning does allow for the use of methods which assume uncorrelated loci such as \anna{BLAHHHHH: \url{https://www.nature.com/articles/srep27644/tables/1}}.
+In addition, LD-pruning may result in improved predictive accuracy \anna{because why}, which can be of particular benefit to polygenic risk score methods.
 
@@
 The remainder of this paper is organized as follows.
-In section 2 we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
+In section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
 We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \cite{Sillanpaeae2011, sul2018population}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures.
@@
 In addition, we briefly review historical methods of correcting for population stratification and relatedness.
-In section 3 we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-Lasso) and lasso penalized multivariate LMMs (LMM-Lasso).
+In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-Lasso) and lasso penalized multivariate LMMs (LMM-Lasso).
 We review the statistical details of these methods as well as the consequences of the ways in which they attempt to correct for population structure and environmental heterogeneity.
-In section 4, we illustrate the concepts described in section 3 via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
+In section \ref{sec:results}, we illustrate the concepts described in section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
 These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods.
 In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
-\pb{don't hard-code ``section 3''; use ref and label instead}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}
+\section{Background} \label{sec:background}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In Section~\ref{Sec:res}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
+In section \ref{sec:results}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
 Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \cite{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
+\section{Methods} \label{sec:methods}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{Sec:res}
+\section{Results} \label{sec:results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion/conclusion}
+\section{Discussion} \label{sec:discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-23 main.tex
Section 3 small tweaks - in progress

@@
 
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?} \anna{In order to show the equivalence of $u \sim N(0, \sigmagg \bK) \text{ and } Z\gamma \sim N(0, \sigmagg \bK)$ I think so, but maybe I can point out that this is the assumption the equivalence is arising from, but we don't' have to have discrete suppopulations in practice...}.
-We denote the standardized version \pb{need more explicit details on standardization} \anna{get details from Hayes 2009} of this genotype matrix as $\bX$.
+To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?} \anna{In order to show the equivalence of $u \sim N(0, \sigmagg \bK) \text{ and } Z\gamma \sim N(0, \sigmagg \bK)$ I think so, but maybe I can point out that this is the assumption the equivalence is arising from, but we don't' have to have discrete suppopulations in practice...\textbf{do this}}.
+We denote the standardized version \pb{need more explicit details on standardization} \anna{\textbf{do this} get details from Hayes 2009} of this genotype matrix as $\bX$.
 We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
@@
 Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
-Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
+Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM: \anna{does this sentence need to change?}
 \begin{align}
@@
 \end{align}
-where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the eigenvalues of $\bX$ \pb{only square matrices have eigenvalues} (equivalently, the singular values of $\bK$.) The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
+where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the singular values of $\bK$.
+The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
 
@@
 \end{equation}
-where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients.
-\anna{do I need to define the ell2 norm?} \pb{YES} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
-Modeling unobserved confounding using fixed effects requires the estimation of $p + k + 1$ parameters.
+where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, \anna{and $||.||_2$ the $\ell_2$ norm.} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
+Modeling unobserved confounding using fixed effects in this manner thus requires the estimation of $p + k + 1$ parameters.
 
@@
 \subsection{LMM-Lasso}
-LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$ \pb{too many sub/superscripts for an in-line fraction}.
-Typically, $\bZ$ is not observed directly, and a common convention \citep[e.g.,][]{Rakitsch2012} is to replace $\bZ$ with $\bX$.
+LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
+$$\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p).$$  Typically, $\bZ$ is not observed directly, and a common convention \citep[e.g.,][]{Rakitsch2012} is to replace $\bZ$ with $\bX$.
 Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.
@@
 Using the spectral decomposition of $\bK$ described in equation \ref{eqn:k}, we can gain further insight into the mechanism of this transformation.
-Note that $\bV^{-1}$ can be expressed using the following matrix factorization \pb{ONLY WORKS IF $\bX$ IS FULL RANK}
+Note that $\bV^{-1}$ can be expressed using the following matrix factorization \pb{ONLY WORKS IF $\bX$ IS FULL RANK} \anna{I understand this on paper, but why does it work numerically?
+Even if $n > p$, svd(K) or eigen(K) gives a U that is $n \times n$...
+what are the vectors corresponding to the 0-eigenvalues?
+Toy example...}
 \begin{align*}
@@
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, in which the variance components and corresponding weights are estimated once under the assumption of a null model \pb{I think you should be more specific here about the ``variance components'', use the $\sigma_e^2$ notation you just introduced, etc.}.
-A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.
+LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses \anna{a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$.
+These variance components and the corresponding weights are estimated once under the assumption of a null model.} A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.
 LMM-Lasso-ggmix \cite{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm.

### 2020-06-23 main.tex
Edits to 3.4 and 3.5

@@
 
+\pb{I don't like calling these two methods ``equivalent''.
+It immediately begs the question: if they're equivalent, why is one of the main aims of this paper to compare them?
+I propose renaming this section ``Relationship between PC-Lasso and LMM-Lasso'' and then talking about both their similarities AND their differences.}
+
 Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \ref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
 \cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \ref{eqn:k}.
+\pb{use \textbackslash eqref, not \textbackslash ref, for equations}
 
@@
 It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \cite{zhang2015principal}.
-However, in practice it is often recommended that the ten largest principal components be included \cite{zhao2018practical}.
-This is the implementation of PC-Lasso we will use throughout our simulations.
+Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}, so this is the way we implement PC-Lasso throughout Section~\ref{sec:results}.
 
@@
 \subsection{Simulating environmental confounding}
-Based on the LMM regression model \ref{eqn:lmm_reg}, numerous simulation studies of the performance of LMMs \cite{Rakitsch2012, bhatnagar2019simultaneous} simulate the quantity $\bX\bgamma$ as a random draw from a $N(\boldsymbol{0}, \sigmagg \bK)$.
-However, simulating environmental confounding according to this framework, where $\bgamma$ is a random quantity, "washes out" bias in $\hat{\bbeta}$ while inflating the variance.
-This is due to the fact that the sign of any element of the vector $\bX\bgamma$ will change from one simulation to the next.
-We can see this by considering the expectation and variance of low-dimensional OLS estimates of $\hat{\bbeta}$ based on model \ref{eqn:lmm_reg}:
+
+In carrying out simulation studies of the LMM regression model \ref{eqn:lmm_reg}, a common practice \citep{Rakitsch2012, bhatnagar2019simultaneous} is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis.
+We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method.
+Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
+To see why this is the case, consider the OLS \pb{have to define OLS somewhere, if it's not already defined} estimates of $\hat{\bbeta}$ based on model \ref{eqn:lmm_reg}:
 \begin{align*}
@@
 \end{align*}
-Under the assumption that $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$, $\mathbb{E}(\hat{\bbeta}) = \bbeta, \mathbb{V}(\hat{\bbeta}) = \frac{\sigmagg}{p}\bI + \sigmaee (\bX^T \bX)^{-1}$.
-Considered in the context of our air quality example, such an assumption would be akin to subpopulation A being exposed to good air quality in one simulation, and bad air quality in the next.
-Instead, considering $\bgamma$ as a fixed quantity is more consistent with environmental heterogeneity and allows us to explore the bias in $\hat{\bbeta}$.
+Under the assumption that $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$, $\mathbb{E}(\hat{\bbeta}) = \bbeta, \mathbb{V}(\hat{\bbeta}) = \frac{\sigmagg}{p}\bI + \sigmaee (\bX^T \bX)^{-1}$ \pb{this is not a sentence}.
+To return to our earlier example involving confounding due to air quality, randomly drawing a new value of $\bX\bgamma$ for each replication would mean that subpopulation A may be exposed to good air quality in one replication, and bad air quality in the next.
+Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from fixed populations A and B?
+Those repeated samples would show a systematic bias due to the difference in air quality, but this is masked if we generate new values of $\bX\bgamma$ with each replication.
+Considering $\bgamma$ as a fixed quantity is more consistent with the usual meaning of environmental confounding.
 Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\hat{\bbeta}) = \bbeta + \bgamma, \mathbb{V}(\hat{\bbeta}) = \sigmaee (\bX^T \bX)^{-1}$.
 
-Although we derive these quantities under model \ref{eqn:lmm_reg}, where $\bZ = \bX$, as our assumed data-generating model, these results also hold for the more general model \ref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of $\bX$.
-Assume $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ where $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
+Although we derive these quantities under model \ref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \ref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of $\bX$.
+Assume \pb{NO ASSUMPTION IS NECESSARY HERE!} $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ where $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ \pb{need to say projection matrix, null space, column space}.
 Then,
@@
 where $\boldsymbol{\tau}$ quantifies bias in the estimated $\boldsymbol{\hat{\beta}'}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
+In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{tau}$.
+
+\pb{should this paragraph move to next section?}
 Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
@@
 \end{equation}
-where $\by$ is a pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
+where $\by$ represents the phenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
 \begin{equation}
@@
 Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
+\pb{So, for example, $\eta=0.5$ and $xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
 
 We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
-\anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership.
+\anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership \pb{Is this true for the linear mixing case as well?}.
 500 simulations were performed for each data type and parameter combination.

### 2020-06-25 main.tex
Figure updates - in progress

@@
 It immediately begs the question: if they're equivalent, why is one of the main aims of this paper to compare them?
-I propose renaming this section ``Relationship between PC-Lasso and LMM-Lasso'' and then talking about both their similarities AND their differences.}
+I propose renaming this section ``Relationship between PC-Lasso and LMM-Lasso'' and then talking about both their similarities AND their differences.} \anna{sounds good}
 
 Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \ref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
-\cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \ref{eqn:k}.
-\pb{use \textbackslash eqref, not \textbackslash ref, for equations}
+\cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}.
+\pb{use \textbackslash eqref, not \textbackslash ref, for equations} \anna{ok, but why?}
 
@@
 \pb{should this paragraph move to next section?}
-Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
+\anna{yes} Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
 \begin{equation}
@@
     \item Need details about the real data once it's run
-    \item Should I mention subpopulation sizes (a few large ones) or gamma structure (half-and-half-heterogeneous) here? No - do a brief summary where noted above
+    \item Need to fix facet labels
 \end{itemize}
@@
     \includegraphics[scale = 1.1]{figures/beta_mse.png}
-    \caption{MSE for different data types in the presence and absence of environmental confounding.}
+    \caption{MSE for different data types in the presence and absence of environmental confounding. \anna{$\gamma$ structure is half-and-half-decreasing-heterogeneous, subpop size is large}}
     \label{fig:mse}
@@
 
-\anna{
-\begin{itemize}
-    \item I ran this but think it just makes things more confusing - go back to only independent subpopulations data - for both (?) 
-\end{itemize}
-}
+\anna{admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g.
+4 environmental exposures.
+the small subpopulations scenario more closely mimics gradations of environmental effects - what is usually simulated with a random draw from the MVN distn...
+we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}
 
@@
     % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
-     \includegraphics[scale = 0.9]{figures/fig2a.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy.}
+     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy:  \anna{$\bgamma$ structure is nonlinear (aka half and half) and heterogeneous, $\eta= \xi = 0.5$}}
     \label{fig:big_vs_small}
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures: \anna{structureX is 1d linear admixture, $\eta= \xi = 0.5$}}
     \label{fig:big_small_gamma}
@@
 
-\anna{eta part needs to be converted to mse}
+\anna{Fix factor label greek letters - R update broke some things}
 \begin{figure}[H]
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
+     \caption{Caption goes here \anna{$\xi = 0.5$, structureX = 1d linear, structureGamma is half and half heterogeneous}}
+    \label{fig:eta_beta_mse}
 \end{figure}
-
-\newpage
-\anna{The error bars are indicating differences across different values of xi and simulated data-types (indepent and 1d linear admixture - this may be confusing/something to change or describe explicitly.
-}
 
@@
     \includegraphics[scale = 0.8]{figures/time_ratio.png}
-    \caption{Caption goes here}
+    \caption{Caption goes here \anna{$\eta = \xi = 0.5$, structureX = 1d linear, structureGamma is half and half heterogeneous}}
     \label{fig:time}
@@
 Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners.
-However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in \anna{figure xxx}, this is also a setting in which the computational time required may be prohibitive for practical applications.
+However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications.
 Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.

### 2020-06-25 main.tex
Edits to Section 4 and 5

@@
 \subsubsection{LMM-Lasso variants}
+\label{sec:rak-ggmix}
 
@@
 Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-\pb{So, for example, $\eta=0.5$ and $xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
+\pb{So, for example, $\eta=0.5$ and $\xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
 
@@
 %------------------------------%
-\subsection{PC-Lasso and LMM-Lasso outperform Lasso}
+\subsection{PC-Lasso and LMM-Lasso outperform Lasso in the presence of environmental confounding}
 %------------------------------%
-Firstly, these penalized regression methods were compared in the presence of varying levels of environmental confounding.
+
+We begin by comparing the ordinary lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
 Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
-The top row of figure \ref{fig:mse} shows that when $\xi = 0$, and no environmental confounding is present, lasso and LMM-Lasso perform comparably.
-PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that the top 10 principal components do not explain any variability in the outcome, yet are included in the final model, since they are unpenalized.
-The bottom row of figure \ref{fig:mse} illustrates the resulting MSE when $\xi = 0.8$ and thus 40\% of the variability of the phenotype is attributable to environmental confounding.
+The top row of figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably.
+PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome.
+The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$ (i.e., 40\% of the variability in the phenotype is attributable to environmental confounding).
 In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
 This improvement is more substantial in the independent subpopulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.
+
+\pb{Need to describe how we select $\lambda$}
 
@@
 
+\pb{Just FYI, it's hard to tell bias and var apart in black and white}
+
 %------------------------------%
@@
 %------------------------------%
+
+\pb{I find ``small'' subpopulations unclear; how do you feel about ``fine'' and ``coarse'' structure?}
+
 Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.
@@
 We note that we show results for simulated data generated such that subpopulations are of equal size, while the empirical SNP data illustrates a scenario of unequal subpopulation sizes.
+
+\pb{should describe our empirical structure here -- is it fine or coarse?}
 
@@
 }
+
+\pb{these are nice figures, but to me, better for supplement}
 
@@
 
-Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
+Phenotypes for these simulations were again simulated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
 Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
@@
 In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
-\cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
+This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
@@
 %------------------------------%
+
+\pb{We should discuss this section; to me, this section comes across as overly complicated and the point is not really coming across}
 
@@
 
-In section 3 briefly mentioned the existence of to LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-The LMM-Lasso results shown so far are those of LMM-Lasso-Rakitsch, though they are representative of LMM-Lasso-ggmix.
-The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods which aim to correct for population structure and environmental heterogeneity.
-However, we also wish to briefly discuss the key differences between methods, and provide guidance for potential users of either method.
+Section~\ref{sec:rak-ggmix} briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+As we will see, the two generally perform quite similarly.
+Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of either LMM-Lasso method.
+
+The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.
+However, given the generally strong performance of LMM-Lasso, we also wish to briefly discuss the key differences between these two ways of implementing it, and provide guidance for potential users.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this work, we have reviewed the concept of population stratification and defined explicitly the non-genetic, environmental mechanism by which it may contribute to confounding the genotype-phenotype relationship.
+In this work, we have reviewed the concept of population stratification and defined explicitly the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
@@
 
+\pb{Introduce penalizedLMM here, or in next paper?}
+
 Even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.

### 2020-06-25 main.tex
Figures: tweaks/notes to self

--- 2020-06-25 main.tex
@@
 Otherwise stated, we do not assume that environmental factors influence population-specific allele frequencies, or vice versa, only that the two may be related.
-Additionally, we assume that the environmental exposure has no effect on on observed SNPs.
-\pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.}
+Additionally, we assume that the environmental exposure has no \anna{direct} effect on on observed SNPs.
+\pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.} \anna{we assume the environment is correlated with population structure, and thus genetic information may serve as a proxy for differential environmental exposures, which also affect phenotype}
 
@@
 It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs \pb{Shouldn't this be: between environment and SNPs?}.
+\anna{no...}
 
@@
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
+\\
 
-\anna{admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g.
-4 environmental exposures.
-the small subpopulations scenario more closely mimics gradations of environmental effects - what is usually simulated with a random draw from the MVN distn...
-we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}
+\anna{
+admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g.
+4 environmental exposures; it is a better proxy for environmental effects.\\
+}
+
+\anna{
+the small subpopulations scenario more closely mimics a gradient of environmental effects - what is usually simulated with a random draw from the MVN distn... \\
+}
+
+\anna{
+we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}\\
 

### 2020-06-25 main.tex

--- 2020-06-25 main.tex
@@
 
-Phenotypes for these simulations were again simulated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
+Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
 Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
@@
 In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
-This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
+\cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
+\\
 
-\anna{admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g.
-4 environmental exposures.
-the small subpopulations scenario more closely mimics gradations of environmental effects - what is usually simulated with a random draw from the MVN distn...
-we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}
+\anna{
+admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g.
+4 environmental exposures; it is a better proxy for environmental effects.\\
+}
+
+\anna{
+the small subpopulations scenario more closely mimics a gradient of environmental effects - what is usually simulated with a random draw from the MVN distn... \\
+}
+
+\anna{
+we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}\\
 

### 2020-06-25 main.tex
Section 2: notes/responses
2.2 edits

--- 2020-06-25 main.tex
@@
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
-Otherwise stated, we do not assume that environmental factors influence population-specific allele frequencies, or vice versa, only that the two may be related.
-Additionally, we assume that the environmental exposure has no effect on on observed SNPs.
+Otherwise stated, we do not assume that environmental factors \anna{directly} influence population-specific allele frequencies, or vice versa, only that the two may be related.
 \pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.}

### 2020-06-25 main.tex
Meeting notes/comments

@@
 Otherwise stated, we do not assume that environmental factors \anna{directly} influence population-specific allele frequencies, or vice versa, only that the two may be related.
-\pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.}
+\anna{often end up related for complex reasons...
+the environment and genetics have become associated over time for complicated reasons that we aren't going to get into...
+air quality isn't changing people's genetics, genetics aren't causing people to move to areas of better/worse air quality}
 
@@
 Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts.
-It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs \pb{Shouldn't this be: between environment and SNPs?}.
-\anna{no...}
+It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs.
 
@@
 \pb{Seems like we should explicitly include an intercept in all these models?}
+
+\anna{typically Z is an incidence matrix... subpops q ... but not the only possibility}
 
@@
 what are the vectors corresponding to the 0-eigenvalues?
-Toy example...}
+Toy example...
+PROVIDED U IS N X N}
 \begin{align*}
@@
 \pb{We should discuss this section; to me, this section comes across as overly complicated and the point is not really coming across}
+\anna{homogeneous reflects groups that are genetically similar have more similar environmental exposures, heterogeneous says this is not the case.
+For example...
+negative lung exposures that are geographic make more sense as a homogeneous exposure but something more behavior-based like smoking behaviors are more plausibly be heterogenous: e.g.
+smoking in China vs.
+France - similar behavior, different genetically}
 
@@
 
+\anna{replace this with inedpendent subpopulations}
 \begin{figure}[H]

### 2020-06-29 main.tex
Figures: updating with figures for independent subpopulations, xi = 0.8. Labels still need work.

@@
 Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.
-Figure \ref{fig:kinship} illustrates the simulated kinship structures in these scenarios, as well as that present in the empirical SNP data.
+
+\anna{the following sentence needs to change now that kinship figs are gone}
+Figure \anna{KINSHIP FIGS IN SUPP NOT HERE} illustrates the simulated kinship structures in these scenarios, as well as that present in the empirical SNP data.
 We note that we show results for simulated data generated such that subpopulations are of equal size, while the empirical SNP data illustrates a scenario of unequal subpopulation sizes.
@@
 \pb{should describe our empirical structure here -- is it fine or coarse?}
-
-\anna{
-\begin{itemize}
-    \item "Spend" this many figures on kinship?
-\end{itemize}
-}
-
-\pb{these are nice figures, but to me, better for supplement}
-
-\begin{figure}[H]
-\begin{subfigure}{.8\textwidth}
-  \centering
-  % include first image
-  \includegraphics[width=.7\linewidth]{figures/kinship1}  
-  \caption{Independent subpopulations simulated kinship; a small number of large groups}
-  \label{fig:large_groups}
-\end{subfigure}
-\newline
-\begin{subfigure}{.8\textwidth}
-  \centering
-  % include second image
-  \includegraphics[width=.7\linewidth]{figures/kinship2}  
-  \caption{Independent subpopulations simulated kinship; a large number of small groups}
-  \label{fig:small_groups}
-\end{subfigure}
-\newline
-\begin{subfigure}{.8\textwidth}
-  \centering
-  % include second image
-  \includegraphics[width=.7\linewidth]{figures/vanderwas_kinship.pdf}  
-  \caption{\anna{Vanderbilt data genome-wide} kinship}
-  \label{fig:vandy_groups}
-\end{subfigure}
-\caption{\anna{obviously these figures need to be tweaked:sizes, uniform scales}}
-\label{fig:kinship}
-\end{figure}
 
@@
      \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy:  \anna{$\bgamma$ structure is nonlinear (aka half and half) and heterogeneous, $\eta= \xi = 0.5$}}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy:  \anna{structureX is independent subpops, $\bgamma$ structure is dichotomous and heterogeneous, $\eta= 0.5, \xi$ is varying}}
     \label{fig:big_vs_small}
@@
 
-\anna{replace this with inedpendent subpopulations}
 \begin{figure}[H]
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures: \anna{structureX is 1d linear admixture, $\eta= \xi = 0.5$}}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures: \anna{structureX is independent subpops, $\eta= 0.5, \xi = 0.8$}}
     \label{fig:big_small_gamma}
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{Caption goes here \anna{$\xi = 0.5$, structureX = 1d linear, structureGamma is half and half heterogeneous}}
+     \caption{Caption goes here \anna{$\xi = 0.5$, structureX = independent subpops, structureGamma is dichotomous heterogeneous, large blocks}}
     \label{fig:eta_beta_mse}
@@
     \includegraphics[scale = 0.8]{figures/time_ratio.png}
-    \caption{Caption goes here \anna{$\eta = \xi = 0.5$, structureX = 1d linear, structureGamma is half and half heterogeneous}}
+    \caption{Caption goes here \anna{$\eta = 0.5, \xi = 0.8$, structureX = independent subpops, structureGamma is dichotomous heterogeneous, large blocks}}
     \label{fig:time}

### 2020-06-29 lmm-genetic-review.tex
2.2 edits

@@
 \begin{abstract}
-Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to \anna{unobserved confounding factors, such as environmental heterogeneity.} This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
+Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to unobserved confounding factors, such as environmental heterogeneity.
+This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
 Penalized linear mixed models (LMMs) are an attractive method to correct for unobserved confounding.
 These methods correct for varying levels of relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data.
-\anna{Despite extensive literature on penalized regression, and LMMs, there is a need for a statistically rigorous review of penalized LMMs.
+Despite extensive literature on penalized regression, and LMMs, there is a need for a statistically rigorous review of penalized LMMs.
 Specifically, the ability of LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
-Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.} Considering this, we evaluate the performance of several penalized regression methods in terms of MSE in the presence of a number of confounding structures.
+Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.
+Considering this, we evaluate the performance of several penalized regression methods in terms of MSE in the presence of a number of confounding structures.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.

### 2020-06-29 main.tex
2.2 edits

@@
 Relatedness describes how genetically similar individuals are, and can refer to both ancient and recent relatedness.
-Recent relatedness is characterized by small clusters of high levels of similarity and is indicative of familial or cryptic relatedness.
+Recent relatedness is characterized by small clusters of high levels of similarity and is indicative of familial or \textbf{cryptic relatedness}.
 Ancient relatedness, or common ancestry, is generally characterized by large clusters of low levels of similarity.
-The presence of multiple distinct ancestry groups is known as population stratification.
-\pb{Should some of these terms, like cryptic relatedness, be bolded?}
+The presence of multiple distinct ancestry groups is known as \textbf{population stratification}.
 
@@
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
-Otherwise stated, we do not assume that environmental factors \anna{directly} influence population-specific allele frequencies, or vice versa, only that the two may be related.
-\anna{often end up related for complex reasons...
-the environment and genetics have become associated over time for complicated reasons that we aren't going to get into...
-air quality isn't changing people's genetics, genetics aren't causing people to move to areas of better/worse air quality}
+Otherwise stated, we do not assume that environmental factors directly influence population-specific allele frequencies, or vice versa, only that the two may be related.
+In the context of our air quality example, this means that we assume air quality is not directly altering individuals' allele frequencies, nor are allele frequencies causing individuals to move to regions of better or worse air quality.
+Genetic relatedness and environmental effects become associated over time for complex reasons, which are beyond the scope of this paper.
+As we will discuss, many methods of correcting for population structure rely on this relationship and do not formally differentiate between effects of population structure and those of environment.
+Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure in question.
 
@@
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
-While we acknowledge the challenges presented by genetic background, this work will focus on correction methods for confounding due to environmental heterogeneity, for which genetic similarity may serve as a proxy.\pb{Overall comment about this paragraph: I feel like we *are* addressing genetic background, by conducting a genomewide multivariate analysis.
-I think its worth pointing out that this is not the only way to deal with genetic background, but it's not as though we're not addressing it at all.}
+Although there are a number of methods in which to account for genetic background \anna{what are they?}, the methods upon which we will focus address this by conducting a genomewide multivariate analysis.
 

### 2020-06-29 main.tex
2.3 edits

@@
 It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs.
+This approach thus seeks to remove the effect of environmental heterogeneity on observed SNPs indirectly.
 
-The genomic control method is a univariate testing procedure which modifies the Cochran-Armitage test for trend by a correction factor \pb{genomic control is not specific to trend tests; it can be used for any type of test statistic inflation}.
-This correction attempts to quantify and remove the effect of population structure on the distribution of the chi-square statistic using markers unrelated to the phenotype of interest and independent of the marker being tested for association \cite{devlin1999genomic, bacanu2000power, wang2009testing}.
-The genomic control method retains the disadvantages of univariate SNP testing and its power to detect true associations is likely to suffer in the presence of linkage disequilibrium \pb{reference?}.
-\pb{This paragraph is a little too negative about genomic control...needs to be softened slightly}
+\anna{DAG here to illustrate this?}
+
+The genomic control (GC) method is a univariate testing procedure which modifies a test statistic by a deflation factor.
+This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \cite{devlin1999genomic, bacanu2000power, wang2009testing}.
+However, this deflation factor has been reported to be sensitive to the nature and quantity of the loci used to estimate it \cite{hellwege2017population, marchini2004effects}.
+Additionally, GC does not correct effect size estimates, so resulting coefficient estimates will be unreliable after GC is applied, even if their corresponding test statistics and p-values have been appropriately adjusted \cite{hellwege2017population}.
 
@@
 This assumption may not accurately reflect the characteristics of a particular sample, such as one consisting of admixed populations, or recent relatedness.
-Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall somewhere between these two extremes \pb{seems useful to bring in idea of a gradient, or continuously varying relatedness, here}.
-By clustering those individuals within in one subpopulation or another we lose information about more high dimensional relatedness \pb{I think ``high dimensional relatedness'' is confusing/ambiguous here.
-What do you mean by it?}.
+Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall between these two extremes and are characterized by continuously varying relatedness and a gradient of air quality.
+By clustering those individuals within in one subpopulation or another we lose information about their fine-grained, subtle relatedness.
 
@@
 
-Like PC-adjustment methods, LMMs rely on a pair-wise similarity matrix \pb{as we discussed, this is a bit odd to say here} to estimate and subsequently correct for relatedness among subjects.
+% rely on a pair-wise similarity matrix
+Like PC-adjustment methods, LMMs utilize the left singular vectors of the design matrix to estimate and subsequently correct for relatedness among subjects.
 However, the term LMM has been used to describe a number of distinct procedures for analyzing structured genetic data which differ in their objectives and methods.

### 2020-06-29 main.tex
3.1: edits in progress

@@
 
-\anna{typically Z is an incidence matrix... subpops q ... but not the only possibility}
-
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects belonging to $q \le p$ subpopulations and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$ \pb{do we need to explicitly assume q discrete subpopulations here?} \anna{In order to show the equivalence of $u \sim N(0, \sigmagg \bK) \text{ and } Z\gamma \sim N(0, \sigmagg \bK)$ I think so, but maybe I can point out that this is the assumption the equivalence is arising from, but we don't' have to have discrete suppopulations in practice...\textbf{do this}}.
-We denote the standardized version \pb{need more explicit details on standardization} \anna{\textbf{do this} get details from Hayes 2009} of this genotype matrix as $\bX$.
+To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
+We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the observed minor allele frequency (MAF) of SNP $j$.
 We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
@@
 \end{equation}
-where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that indicates population structure, $\bgamma$ is an $q \times 1$ vector of subpopulation-specific environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects, $\bgamma$ is an $q \times 1$ vector of environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+Often it is assumed that subjects belong to $q \le p$ subpopulations, and $\bZ$ is a corresponding incidence matrix of 0s and 1s.
+However, as we will discuss, this is not the only possibility.
 
 Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
-Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM: \anna{does this sentence need to change?}
+Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
+
 \begin{align}
@@
 \end{align}
-where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the singular values of $\bK$.
+\anna{See Lippert 2011 for details about the dimensions and contents of U} where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the singular values of $\bK$.
 The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.

### 2020-06-29 main.tex
3.1 - 3.4 edits

@@
 \end{equation}
-where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects, $\bgamma$ is an $q \times 1$ vector of environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bgamma$ is a $q \times 1$ vector of environmental effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
 Often it is assumed that subjects belong to $q \le p$ subpopulations, and $\bZ$ is a corresponding incidence matrix of 0s and 1s.
@@
 \end{align}
-\anna{See Lippert 2011 for details about the dimensions and contents of U} where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the singular values of $\bK$.
+where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the singular values of $\bK$.
 The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
@@
 \end{equation}
-where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \ref{eqn:model}.
+where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \eqref{eqn:model}.
 In a penalized regression framework, this model yields the objective function
@@
 \end{equation}
-where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, \anna{and $||.||_2$ the $\ell_2$ norm.} Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
+where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||.||_2$ the $\ell_2$ norm.
+Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
 Modeling unobserved confounding using fixed effects in this manner thus requires the estimation of $p + k + 1$ parameters.
@@
 LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
-$$\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p).$$  Typically, $\bZ$ is not observed directly, and a common convention \citep[e.g.,][]{Rakitsch2012} is to replace $\bZ$ with $\bX$.
+$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \cite{wang2018multiplex, lippert2011fast, yang2014advantages}.
 Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.
@@
 \anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
-Using the spectral decomposition of $\bK$ described in equation \ref{eqn:k}, we can gain further insight into the mechanism of this transformation.
-Note that $\bV^{-1}$ can be expressed using the following matrix factorization \pb{ONLY WORKS IF $\bX$ IS FULL RANK} \anna{I understand this on paper, but why does it work numerically?
-Even if $n > p$, svd(K) or eigen(K) gives a U that is $n \times n$...
-what are the vectors corresponding to the 0-eigenvalues?
-Toy example...
-PROVIDED U IS N X N}
+Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation.
+Note that $\bV^{-1}$ can be expressed using the following matrix factorization \anna{PROVIDED U IS N X N...
+see Lippert 2011 for details about the dimensions and contents of U}
 \begin{align*}
@@
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses \anna{a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$.
-These variance components and the corresponding weights are estimated once under the assumption of a null model.} A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.
+LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$.
+These variance components and the corresponding weights are estimated once under the assumption of a null model.
+A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.
 LMM-Lasso-ggmix \cite{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm.
@@
 %------------------------------%
-\subsection{Equivalence of PC-Lasso and LMM-Lasso}
+\subsection{Relationship between PC-Lasso and LMM-Lasso}
 
-\pb{I don't like calling these two methods ``equivalent''.
-It immediately begs the question: if they're equivalent, why is one of the main aims of this paper to compare them?
-I propose renaming this section ``Relationship between PC-Lasso and LMM-Lasso'' and then talking about both their similarities AND their differences.} \anna{sounds good}
+\pb{talk about both their similarities AND their differences.}
 
-Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \ref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
+Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
 \cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}.
-\pb{use \textbackslash eqref, not \textbackslash ref, for equations} \anna{ok, but why?}
 
@@
 
-In carrying out simulation studies of the LMM regression model \ref{eqn:lmm_reg}, a common practice \citep{Rakitsch2012, bhatnagar2019simultaneous} is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis.
+In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice \citep{Rakitsch2012, bhatnagar2019simultaneous} is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis.
 We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method.
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
-To see why this is the case, consider the OLS \pb{have to define OLS somewhere, if it's not already defined} estimates of $\hat{\bbeta}$ based on model \ref{eqn:lmm_reg}:
+To see why this is the case, consider the OLS \pb{have to define OLS somewhere, if it's not already defined} estimates of $\hat{\bbeta}$ based on model \eqref{eqn:lmm_reg}:
 \begin{align*}
@@
 
-Although we derive these quantities under model \ref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \ref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of $\bX$.
+Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of $\bX$.
 Assume \pb{NO ASSUMPTION IS NECESSARY HERE!} $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ where $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ \pb{need to say projection matrix, null space, column space}.
@@
 where $\boldsymbol{\tau}$ quantifies bias in the estimated $\boldsymbol{\hat{\beta}'}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
-In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{tau}$.
+In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
 
@@
 \anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership \pb{Is this true for the linear mixing case as well?}.
-500 simulations were performed for each data type and parameter combination.
+1000 simulations were performed for each data type and parameter combination.
 

### 2020-06-29 main.tex
3.3 - 3.5: edits, explicit intercept term

@@
 
-\pb{Seems like we should explicitly include an intercept in all these models?}
-
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
-We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the observed minor allele frequency (MAF) of SNP $j$.
+We denote the normalized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \cite{zhang2015principal, price2006principal}.
 We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
@@
     \label{eqn:model}
-    \by = \bX\bbeta + \bZ\bgamma + \beps
+    \by = \Int + \bX\bbeta + \bZ\bgamma + \beps
 \end{equation}
-where $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bgamma$ is a $q \times 1$ vector of environmental effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+where $\Int$ is the intercept, $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bgamma$ is a $q \times 1$ vector of environmental effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
 Often it is assumed that subjects belong to $q \le p$ subpopulations, and $\bZ$ is a corresponding incidence matrix of 0s and 1s.
@@
 
-Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
-Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
+LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
+Using the singular value decomposition (SVD) of the normalized genotype matrix, we can derive the spectral decomposition of the RRM:
 
@@
 \end{align}
-where the columns of $\bU$ are the principal components of $\bX$ (equivalently, those of $\bK$) and $\boldsymbol{D}^2$ is a diagonal matrix of the singular values of $\bK$.
-The columns of $\boldsymbol{R}$ are therefore proportional to the principal components weighted by their corresponding singular values.
+where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are therefore proportional to the principal components of $\bX$ or $\bK$.
 
@@
     \label{eqn:pca_reg}
-    \by = \bX\bbeta + \bU_{1:k}\boldsymbol{\alpha} + \beps 
+    \by = \Int + \bX\bbeta + \bU_{1:k}\boldsymbol{\alpha} + \beps 
 \end{equation}
@@
     \label{eqn:pca_obj}
-    Q(\boldsymbol{\beta, \alpha| X, y}) = \frac{1}{2n} ||\boldsymbol{y} - \boldsymbol{X \beta} - \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \boldsymbol{\beta} ||_1
+    Q(\boldsymbol{\beta, \alpha| X, y}) = \frac{1}{2n} ||\by - \Int - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1
 \end{equation}
@@
 $$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \cite{wang2018multiplex, lippert2011fast, yang2014advantages}.
-Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigmagg \bK)$.
+Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$.
 This results in the LMM regression model
@@
     \label{eqn:lmm_reg}
-    \by = \bX\bbeta + \bX\bgamma+ \beps,
+    \by = \Int + \bX\bbeta + \bX\bgamma+ \beps,
 \end{equation}
-where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
-LMM-Lasso then performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
+where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\Int + \bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
+LMM-Lasso performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
 \begin{equation}
-\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}}\bX \bbeta, \bI).
+\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \left(\Int +\bX \bbeta\right), \bI).
 \end{equation}
 \anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
+\begin{comment}
+For ease of notation, we incorporate the intercept into the design matrix and let $\bX' = \begin{bmatrix} \textbf{1} & \bX \end{bmatrix}$ and $\bbeta' = \begin{pmatrix} \mu & \bbeta \end{pmatrix}^T$ so that we can express this more compactly as
+\begin{equation}
+\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX' \bbeta', \bI).
+\end{equation}
+\end{comment}
 Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation.
-Note that $\bV^{-1}$ can be expressed using the following matrix factorization \anna{PROVIDED U IS N X N...
-see Lippert 2011 for details about the dimensions and contents of U}
+Note that $\bV^{-1}$ can be expressed using the following matrix factorization
 \begin{align*}
@@
 \end{align*}
-By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$, $\byT = \bU^T \by$, and $\bW = \text{diag}\{w_1, ...
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left( \by - \Int \right)$, where we center $\by$ to simplify notation.
+$\bW = \text{diag}\{w_1, ...
 w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1}$.
@@
 \label{eqn:lmm_obj}
-Q(\bbeta | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
+Q(\bbeta' | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
 \end{equation}
@@
 
-\pb{talk about both their similarities AND their differences.}
-
-Despite the apparent dissimilarities between PC-Lasso and LMM-Lasso, it has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
+It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
 \cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}.
@@
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
-To see why this is the case, consider the OLS \pb{have to define OLS somewhere, if it's not already defined} estimates of $\hat{\bbeta}$ based on model \eqref{eqn:lmm_reg}:
+To see why this is the case, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg}: \anna{intercept stuff gets messy here - need to include}
 \begin{align*}
@@
 
-Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of $\bX$.
-Assume \pb{NO ASSUMPTION IS NECESSARY HERE!} $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ where $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ \pb{need to say projection matrix, null space, column space}.
+Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.
+
+Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
+$\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
 Then,
@@
 
-\pb{should this paragraph move to next section?}
-\anna{yes} Thus, given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%------------------------------%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results} \label{sec:results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%------------------------------%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Overview of simulation... briefly define gamma structure, subpops size... more details in supplement...
+
+%------------------------------%
+\subsection{Simulation study}
+%------------------------------%
+
+Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
 \begin{equation}
@@
 1000 simulations were performed for each data type and parameter combination.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%------------------------------%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results} \label{sec:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%------------------------------%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Overview of simulation... briefly define gamma structure, subpops size... more details in supplement...
 

### 2020-06-29 main.tex
Replacing '\cite' with '\citep'

@@
 This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci can be identified and their effects estimated.
-In addition, it has been shown that in the worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \cite{prive2018efficient}.
+In addition, it has been shown that in the worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \citep{prive2018efficient}.
 However, LD-pruning does allow for the use of methods which assume uncorrelated loci such as \anna{BLAHHHHH: \url{https://www.nature.com/articles/srep27644/tables/1}}.
@@
 
-The Lasso \cite{tibshirani1996regression} is a widely used penalized regression method.
+The Lasso \citep{tibshirani1996regression} is a widely used penalized regression method.
 It relies on an assumption of sparsity, which means that coefficient estimation and variable selection can be accomplished simultaneously, and also provides considerable computational advantages, meaning that it scales up well to very large data sets.
@@
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \cite{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, there remain open and relevant problems \pb{this is vague} in analyzing structured genetic data \cite{lawson2019population, barton2019population}.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, there remain open and relevant problems \pb{this is vague} in analyzing structured genetic data \citep{lawson2019population, barton2019population}.
 
@@
 In other words, random genetic variation is likely to be associated with non-genetic factors that also influence the trait of interest, thereby confounding the genotype-phenotype relationship and leading to spurious associations and biased estimates of SNP effects.
-Biased estimates of this nature severely hinder prediction and understanding differences among populations \cite{barton2019population}.
-Although bias at individual loci may be small, this bias may become magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \cite{barton2019population}.
+Biased estimates of this nature severely hinder prediction and understanding differences among populations \citep{barton2019population}.
+Although bias at individual loci may be small, this bias may become magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \citep{barton2019population}.
 
@@
 In section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
-We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \cite{Sillanpaeae2011, sul2018population}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures.
+We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \citep{Sillanpaeae2011, sul2018population}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures.
 In so doing, we refine the definition of confounding due to population structure in terms of a non-genetic mechanism.
@@
 Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations.
-For example, European American \cite{campbell2005demonstrating}, Han Chinese \cite{xu2009genomic, chen2009genetic}, and most recently, cohorts within the UK Biobank data \cite{haworth2019apparent}.
+For example, European American \citep{campbell2005demonstrating}, Han Chinese \citep{xu2009genomic, chen2009genetic}, and most recently, cohorts within the UK Biobank data \citep{haworth2019apparent}.
 
@@
 However, the mechanism of this phenomenon warrants some discussion.
-In particular, it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense, rather there must exist some non-genetic mechanism by which population stratification affects both the phenotype and genotype \cite{barton2019population, vilhjalmsson2012nature}.
+In particular, it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense, rather there must exist some non-genetic mechanism by which population stratification affects both the phenotype and genotype \citep{barton2019population, vilhjalmsson2012nature}.
 \\
@@
 Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
-As \cite{vilhjalmsson2012nature} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
+As \citep{vilhjalmsson2012nature} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
@@
 The genomic control (GC) method is a univariate testing procedure which modifies a test statistic by a deflation factor.
-This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \cite{devlin1999genomic, bacanu2000power, wang2009testing}.
-However, this deflation factor has been reported to be sensitive to the nature and quantity of the loci used to estimate it \cite{hellwege2017population, marchini2004effects}.
-Additionally, GC does not correct effect size estimates, so resulting coefficient estimates will be unreliable after GC is applied, even if their corresponding test statistics and p-values have been appropriately adjusted \cite{hellwege2017population}.
+This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
+However, this deflation factor has been reported to be sensitive to the nature and quantity of the loci used to estimate it \citep{hellwege2017population, marchini2004effects}.
+Additionally, GC does not correct effect size estimates, so resulting coefficient estimates will be unreliable after GC is applied, even if their corresponding test statistics and p-values have been appropriately adjusted \citep{hellwege2017population}.
 
 Stratified analysis attempts to infer subject membership within a discrete number of nonoverlapping subpopulations.
-Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \cite{pritchard1999use, pritchard2000association}.
+Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.
 By partitioning the overall sample size in this manner, the structured association method is particularly vulnerable to a loss of power.
@@
 Another approach is to create an approximation of population structure using surrogate variables, and to adjust for these as an additional model covariates.
-Principal component-adjustment (PC-adjustment) methods use the first several principal components of the design matrix as these surrogate variables \cite{price2006principal}.
+Principal component-adjustment (PC-adjustment) methods use the first several principal components of the design matrix as these surrogate variables \citep{price2006principal}.
 This approach may be used in univariate or multivariate frameworks, and allows for analysis on the entirety of a structured sample.
@@
 
-Among LMM methods that aim to estimate heritability, the most well-known example is genome-wide complex trait analysis (GCTA) \cite{yang2011gcta}.
+Among LMM methods that aim to estimate heritability, the most well-known example is genome-wide complex trait analysis (GCTA) \citep{yang2011gcta}.
 In this framework, all observed SNPs are treated as random effects.
-These effects are integrated out in order to estimate a genetic variance component, which in turn, is used to quantify the total narrow-sense heritability of a trait \cite{yang2010common}.
+These effects are integrated out in order to estimate a genetic variance component, which in turn, is used to quantify the total narrow-sense heritability of a trait \citep{yang2010common}.
 
 LMMs can also be used to estimate SNP effects, and this can be done in either a univariate or multivariate manner.
-Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure using a genetic similarity matrix \cite{yu2006unified, kang2010variance, kang2008efficient}.
+Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure using a genetic similarity matrix \citep{yu2006unified, kang2010variance, kang2008efficient}.
 However, as with all univariate testing approaches univariate LMM implementations may still produce spurious associations and biased effect estimates in the presence of LD.
 
-Multivariate mixed model approaches reduce these problems by assessing the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while controlling for environmental confounding  \cite{Rakitsch2012, bhatnagar2019simultaneous}.
+Multivariate mixed model approaches reduce these problems by assessing the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while controlling for environmental confounding  \citep{Rakitsch2012, bhatnagar2019simultaneous}.
 This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes.
@@
 Specifically, PC-adjustment and LMM methods, which we will refer to as PC-Lasso and LMM-Lasso, respectively.
-We consider two similar implementations of the LMM-Lasso method, one developed by \cite{Rakitsch2012} and a more recent implementation from \cite{bhatnagar2019simultaneous}.
+We consider two similar implementations of the LMM-Lasso method, one developed by \citep{Rakitsch2012} and a more recent implementation from \citep{bhatnagar2019simultaneous}.
 Subsequent references to LMM-Lasso apply to both implementations, unless otherwise stated.
@@
 In section \ref{sec:results}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
-Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \cite{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\
+Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \citep{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\
 
@@
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
-We denote the normalized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \cite{zhang2015principal, price2006principal}.
+We denote the normalized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
 We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
@@
 
-LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \cite{hayes2009increased}.
+LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
 Using the singular value decomposition (SVD) of the normalized genotype matrix, we can derive the spectral decomposition of the RRM:
@@
 PC-Lasso uses the first $k$ principal components of $\bX$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates.
-It has been shown \cite{hoffman2013correcting} that this corresponds to the regression model
+It has been shown \citep{hoffman2013correcting} that this corresponds to the regression model
 \begin{equation}
@@
 LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
-$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \cite{wang2018multiplex, lippert2011fast, yang2014advantages}.
+$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
 Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$.
@@
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-LMM-Lasso-Rakitsch \cite{Rakitsch2012} uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$.
+LMM-Lasso-Rakitsch \citep{Rakitsch2012} uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$.
 These variance components and the corresponding weights are estimated once under the assumption of a null model.
-A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \cite{glmnet} or \texttt{ncvreg} \cite{ncvreg}.
-LMM-Lasso-ggmix \cite{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm.
-This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \cite{ggmix, bhatnagar2019simultaneous}.
+A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
+LMM-Lasso-ggmix \citep{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm.
+This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix, bhatnagar2019simultaneous}.
 We postpone further discussion of the differences between these methods to the results section.
@@
 It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
-\cite{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \cite{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}.
+\citep{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citep{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}.
 
 Thus, whether modeled as a fixed or random effect, the underlying effect of $\bZ \bgamma$ on phenotype remains the same, but its treatment as a fixed or random effect has important implications in terms of model fitting and estimation.
-Including principal components as fixed effects requires estimation of a greater number of parameters, which may result in a loss of power \cite{zhang2015principal}.
-Additionally, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \cite{patterson2006population, zhao2018practical}.
-It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \cite{zhang2015principal}.
+Including principal components as fixed effects requires estimation of a greater number of parameters, which may result in a loss of power \citep{zhang2015principal}.
+Additionally, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \citep{patterson2006population, zhao2018practical}.
+It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \citep{zhang2015principal}.
 Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}, so this is the way we implement PC-Lasso throughout Section~\ref{sec:results}.

### 2020-07-01 main.tex
3.3 - 3.5 edits

@@
 LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
-$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
-Then, rather than assuming anything about $\bgamma$ directly, let $\bu=\bX\bgamma$ represent the random effect and assume $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$.
-This results in the LMM regression model
+$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.This results in the LMM regression model
 \begin{equation}
@@
 \end{equation}
-where we assume $\beps \independent \bX\bgamma$ and thus $\by  \sim N(\Int + \bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
+where we assume $\beps \independent \bX\bgamma$.
+Rather than assuming anything about $\bgamma$ directly, an equivalent representation lets $\bu=\bX\bgamma$ such that $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent LMM model
+\begin{equation}
+    \label{eqn:lmm_reg_u}
+    \by = \Int + \bX\bbeta + \bu + \beps.
+\end{equation}
+Regardless of representation, the LMM assumes $\by \sim N(\Int + \bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
 LMM-Lasso performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
@@
 \end{align*}
-By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left( \by - \Int \right)$, where we center $\by$ to simplify notation.
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation.
 $\bW = \text{diag}\{w_1, ...
@@
 \label{eqn:lmm_obj}
-Q(\bbeta' | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
+Q(\bbeta | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
 \end{equation}
@@
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
-To see why this is the case, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg}: \anna{intercept stuff gets messy here - need to include}
+To see why this is the case, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\mu = 0$ for ease of notation: \anna{intercept stuff would get messy here...
+does this work?}
 \begin{align*}
@@
 \end{align*}
-Under the assumption that $\bgamma \sim N(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p)$, $\mathbb{E}(\hat{\bbeta}) = \bbeta, \mathbb{V}(\hat{\bbeta}) = \frac{\sigmagg}{p}\bI + \sigmaee (\bX^T \bX)^{-1}$ \pb{this is not a sentence}.
-To return to our earlier example involving confounding due to air quality, randomly drawing a new value of $\bX\bgamma$ for each replication would mean that subpopulation A may be exposed to good air quality in one replication, and bad air quality in the next.
+The assumption that environmental effects follow a random distribution, $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$, yields $\mathbb{E}(\hat{\bbeta}) = \bbeta$, and $\mathbb{V}(\hat{\bbeta}) = \sigmagg/p \bI + \sigmaee (\bX^T \bX)^{-1}$.
+Returning to our earlier example involving confounding due to air quality, randomly drawing a new value of $\bX\bgamma$ for each replication would mean that subpopulation A may be exposed to good air quality in one replication, and bad air quality in the next.
 Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from fixed populations A and B?
@@
 Considering $\bgamma$ as a fixed quantity is more consistent with the usual meaning of environmental confounding.
-Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\hat{\bbeta}) = \bbeta + \bgamma, \mathbb{V}(\hat{\bbeta}) = \sigmaee (\bX^T \bX)^{-1}$.
+Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\hat{\bbeta}) = \bbeta + \bgamma, \mathbb{V}(\hat{\bbeta}) = \sigmaee (\bX^T \bX)^{-1}$, which allows us to quantify bias in our estimates.
 
-Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$, by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.
+Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$.
+We see this by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.
 
 Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
-$\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
-Then,
+Then, $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ such that
 \begin{align*}

### 2020-07-01 main.tex
4.1 edits

@@
 
-Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
+Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy and bias specifically, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
 \begin{equation}
@@
 \end{equation}
-where $\by$ represents the phenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
+where $\by$ represents the pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
 \begin{equation}
@@
 Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-\pb{So, for example, $\eta=0.5$ and $\xi=0.7$ means that\ldots} Given our modeling objective of accurately estimating the effect sizes of observed SNPs, $\eta$ also represents our data-generating model's signal to noise ratio (SNR).
+$\eta$ also represents our data-generating model's signal to noise ratio (SNR).
+So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to environmental effects.
 
 We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
-\anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership \pb{Is this true for the linear mixing case as well?}.
+\anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership \pb{Is this true for the linear mixing case as well?} \anna{yes}.
 1000 simulations were performed for each data type and parameter combination.

### 2020-07-01 main.tex
4.1 - 4.2 edits

@@
 We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
-\anna{cite Vanderwas data} For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership \pb{Is this true for the linear mixing case as well?} \anna{yes}.
-1000 simulations were performed for each data type and parameter combination.
+\anna{cite Vanderwas data} $p = 1000$ for all three data types, $n = 200$ for the simulated data types, and 50 causal SNPs were used throughout.
+The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
+For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership, \anna{or founding population membership in the case of the admixed data.} 1000 simulations were performed for each data type and parameter combination.
 
@@
 PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome.
-The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$ (i.e., 40\% of the variability in the phenotype is attributable to environmental confounding).
+The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
 In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
-This improvement is more substantial in the independent subpopulations simulated data due to the fact that the subpopulations are more easily distinguishable based on their observed genetic data.
-
-\pb{Need to describe how we select $\lambda$}
-
-\anna{
-\begin{itemize}
-    \item Need details about the real data once it's run
-    \item Need to fix facet labels
-\end{itemize}
-}
+This improvement is more substantial in the independent subpopulations simulated data than the admixed simulated data.
+This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
 
@@
 
-\pb{Just FYI, it's hard to tell bias and var apart in black and white}
+\pb{Just FYI, it's hard to tell bias and var apart in black and white} \anna{will fix this and greek letters}
 
@@
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example, and (2) a large number of small, closely related subpopulations to approximate finer grained environmental heterogeneity, such as household exposures or behavioral norms.
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) coarse \anna{cut this?: relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example,} and (2) fine \anna{cut?: a large number of small subpopulations corresponding to finer grained environmental heterogeneity, such as household exposures or behavioral norms}.
 

### 2020-07-01 main.tex
Fleshing out and moving 4.1 (simulation study) to methods section 3.6

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%------------------------------%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results} \label{sec:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%------------------------------%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Overview of simulation... briefly define gamma structure, subpops size... more details in supplement...
-
 %------------------------------%
@@
 %------------------------------%
+\subsubsection{Pseudophenotype generation}
 
@@
 So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to environmental effects.
+The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
+1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
 
+\subsubsection{Design matrix generation}
 We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
 \anna{cite Vanderwas data} $p = 1000$ for all three data types, $n = 200$ for the simulated data types, and 50 causal SNPs were used throughout.
-The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
-For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership, \anna{or founding population membership in the case of the admixed data.} 1000 simulations were performed for each data type and parameter combination.
+
+\subsubsection{Subpopulation structure}
+For the fully simulated data, two subpopulation structures were considered: (1) fine and (2) coarse.
+We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example.
+Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, or founding populations, which corresponds to a finer grained continuum of environmental heterogeneity.
+Whether fine or coarse, the simulated data were generated such that subpopulations, or founding populations, are of equal size.
+We note that the fine and coarse subpopulation structures are distinct from the level of admixture in the simulated data.
+For example, the discrete nonoverlapping subpopulations in the simulated independent subpopulation data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated admixture data.
+The empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
+Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture.
+Supplementary figures \anna{?, ?, and ?} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.
+
+\subsubsection{Structure of environmental effects}
+For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership, or founding population membership in the case of the admixed data, which assigns the corresponding environmental effect found in $\bgamma$.
+Various $\bgamma$ structures were considered in order to investigate the effects of various forms of environmental confounding.
+
+ Our initial hypothesis was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that are nonlinear in nature and as such do not lend themselves to such an approximation.
+To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous.
+\anna{more details?} In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
+For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to \anna{spatial population drift, cline...
+(???)}.
+However, more behavior-based environmental exposures may be less likely to be shared by genetically similar subpopulations.
+Consider cigarette smoking rates in France, England, and China.
+While French and English populations are more genetically similar than French and Chinese populations \anna{need citaiton}, the environmental exposure of cigarette smoking is more similar between French and Chinese populations due to cultural norms and practices \anna{need citaiton}.
+
+ To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis (?) - same direction}, or (2) \anna{heterogeneous/trans (?)} in nature for a total of four distinct $\bgamma$ structures.
+We define \anna{homogenous/cis} effects as those that are similar among genetically similar subpopulations, and \anna{heterogeneous/trans} effects as those that are not.
+Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures.
+\anna{is this still unclear?
+Would a PCA picture help?}
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
+    \caption{$\bgamma$ structures}
+    \label{fig:gamma_structures}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%------------------------------%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results} \label{sec:results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%------------------------------%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\pb{I find ``small'' subpopulations unclear; how do you feel about ``fine'' and ``coarse'' structure?}
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of fine and coarse subpopulation structures.
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of two forms of simulated population structure: (1) coarse \anna{cut this?: relatedness characterized by a small number of large subpopulations to model large-scale environmental heterogeneity or population stratification, similar to that described in our motivating air quality example,} and (2) fine \anna{cut?: a large number of small subpopulations corresponding to finer grained environmental heterogeneity, such as household exposures or behavioral norms}.
+Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying environmental confounding, $\xi \in \{0, 0.2, 0.5,0.8\}$.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding simulated using a dichotomous-heterogeneous $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
 
-\anna{the following sentence needs to change now that kinship figs are gone}
-Figure \anna{KINSHIP FIGS IN SUPP NOT HERE} illustrates the simulated kinship structures in these scenarios, as well as that present in the empirical SNP data.
-We note that we show results for simulated data generated such that subpopulations are of equal size, while the empirical SNP data illustrates a scenario of unequal subpopulation sizes.
-
-\pb{should describe our empirical structure here -- is it fine or coarse?}
-
-Phenotypes for these simulations were again simulated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$.
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
+with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
 The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.
@@
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
-
-\anna{admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g.
-4 environmental exposures.
-the small subpopulations scenario more closely mimics gradations of environmental effects - what is usually simulated with a random draw from the MVN distn...
-we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}
-
 \begin{figure}[H]
@@
 
-\pb{We should discuss this section; to me, this section comes across as overly complicated and the point is not really coming across}
-\anna{homogeneous reflects groups that are genetically similar have more similar environmental exposures, heterogeneous says this is not the case.
-For example...
-negative lung exposures that are geographic make more sense as a homogeneous exposure but something more behavior-based like smoking behaviors are more plausibly be heterogenous: e.g.
-smoking in China vs.
-France - similar behavior, different genetically}
-
-In addition to considering the effect of population structure, or forms of $\bZ$, on these methods' estimation accuracy, we also evaluated the effect of varying forms of $\bgamma$, the subpopulation-specific effect sizes.
-Our initial thinking was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that do not lend themselves to linear approximations.
-To that end we investigated six distinct $\bgamma$ structures: linear homogeneous, linear heterogeneous, nonlinear homogeneous, nonlinear heterogeneous, exponential homogeneous, and exponential heterogeneous.
-In our naming convention, linear/nonlinear refers to whether the coefficients follow a linear pattern versus a nonlinear form, while homogeneous/heterogeneous indicates whether coefficients have the same, or differing signs, respectively.
-For example, (1, -2, 3, -4) is an example of a linear-heterogeneous structure, whereas (2, 2, 1, 1) is a nonlinear-homogeneous form.
-Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures.
-\\
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
-    \caption{$\bgamma$ structures}
-    \label{fig:gamma_structures}
-\end{figure}
-
-We thus evaluate PC-Lasso and LMM-Lasso in the presence of large and small subpopulations as well as distinct coefficient structures.
+We evaluate PC-Lasso and LMM-Lasso in the presence of large and small subpopulations as well as distinct coefficient structures.
 Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) for all simulations, broken down by both subpopulation and coefficient structure.

### 2020-07-01 main.tex
3.6.4 edits

@@
 \anna{more details?} In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
-For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to \anna{spatial population drift, cline...
-(???)}.
-However, more behavior-based environmental exposures may be less likely to be shared by genetically similar subpopulations.
-Consider cigarette smoking rates in France, England, and China.
-While French and English populations are more genetically similar than French and Chinese populations \anna{need citaiton}, the environmental exposure of cigarette smoking is more similar between French and Chinese populations due to cultural norms and practices \anna{need citaiton}.
+For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
+However, behavior-based environmental exposures may be less likely to be shared by subpopulations which are genetically and geographically similar.
+\anna{another citation?} As an example, \citep{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
 
- To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis (?) - same direction}, or (2) \anna{heterogeneous/trans (?)} in nature for a total of four distinct $\bgamma$ structures.
+ \anna{in organic chemistry cis means 'on the same side' and trans 'on the other side' so I thought cis/trans might be more descriptive for gamma structure than homo/heterogeneous.
+These are also used to describe gender identity, which is obviously not how I'm using it, so I don't want it to be confusing.}
+ To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis}, or (2) \anna{heterogeneous/trans} in nature for a total of four distinct $\bgamma$ structures.
 We define \anna{homogenous/cis} effects as those that are similar among genetically similar subpopulations, and \anna{heterogeneous/trans} effects as those that are not.
@@
     \includegraphics[scale = 1.1]{figures/beta_mse.png}
-    \caption{MSE for different data types in the presence and absence of environmental confounding. \anna{$\gamma$ structure is half-and-half-decreasing-heterogeneous, subpop size is large}}
+    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-\anna{heterogeneous/trans} environmental effects.}
     \label{fig:mse}
@@
      \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy:  \anna{structureX is independent subpops, $\bgamma$ structure is dichotomous and heterogeneous, $\eta= 0.5, \xi$ is varying}}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-\anna{heterogeneous/trans} environmental effects.}
     \label{fig:big_vs_small}
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures: \anna{structureX is independent subpops, $\eta= 0.5, \xi = 0.8$}}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with independent subpopulation data, $\eta = 0.5$, and $\xi = 0.8$}
     \label{fig:big_small_gamma}
@@
 
-\anna{Fix factor label greek letters - R update broke some things}
 \begin{figure}[H]
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{Caption goes here \anna{$\xi = 0.5$, structureX = independent subpops, structureGamma is dichotomous heterogeneous, large blocks}}
+     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with independent subpopulation data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-\anna{heterogeneous/trans} environmental effect structure.}
     \label{fig:eta_beta_mse}
@@
     \includegraphics[scale = 0.8]{figures/time_ratio.png}
-    \caption{Caption goes here \anna{$\eta = 0.5, \xi = 0.8$, structureX = independent subpops, structureGamma is dichotomous heterogeneous, large blocks}}
+    \caption{Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-\anna{heterogeneous/trans} environmental effect structure.}
     \label{fig:time}

### 2020-07-01 main.tex
Introduction edits, notational tweaks

@@
 This occurs when the variant being tested is correlated with other variants, so that the effects of these correlated variants are erroneously attributed to the variant in question.
-\anna{One method for analyzing data in the presence of correlated loci, or linkage disequilibrium (LD), is LD-pruning.
-In LD-pruning methods, SNPs are sequentially thinned, so that only SNPs correlated with each other below some desired threshold are used in modeling and testing procedures.} LD-pruning limits analysis to SNPs which are representative of a particular region.
-This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci can be identified and their effects estimated.
-In addition, it has been shown that in the worst-case scenario, pruning algorithms may not select any representative SNPs in a particular region of the genome \citep{prive2018efficient}.
-However, LD-pruning does allow for the use of methods which assume uncorrelated loci such as \anna{BLAHHHHH: \url{https://www.nature.com/articles/srep27644/tables/1}}.
-In addition, LD-pruning may result in improved predictive accuracy \anna{because why}, which can be of particular benefit to polygenic risk score methods.
+One method for analyzing data in the presence of correlated loci, or linkage disequilibrium (LD), is LD pruning.
+LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
+The overarching approach of these methods is to sequentially thin SNPs, so that only SNPs correlated with each other below some desired threshold are used in modeling and testing procedures.
+This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci, as opposed to the causal loci themselves, can be identified and their effects estimated.
+
+% LD pruning thus limits analysis to SNPs which are representative of a particular region.
+
+% In some scenarios pruning algorithms may not select any representative SNPs in a particular region of the genome \citep{prive2018efficient}.
+However, LD-pruning has been shown to improve performance of principal component analysis (PCA) methods \citep{abdellaoui2013population} and \anna{what} \anna{because why}.
 
@@
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, there remain open and relevant problems \pb{this is vague} in analyzing structured genetic data \citep{lawson2019population, barton2019population}.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, until recently the majority of this work has focused on plant and animal breeding data, much of which experimentally controlled.
+While many standard correction techniques may in principal be applied to human genetic data, their performance must be studied within the context of human ancestry as well as non-experimental conditions \citep{lawson2019population, barton2019population}.
 
@@
     \label{eqn:pca_obj}
-    Q(\boldsymbol{\beta, \alpha| X, y}) = \frac{1}{2n} ||\by - \Int - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1
+    Q(\bbeta, \boldsymbol{\alpha}|\bX, \by) = \frac{1}{2n} ||\by - \Int - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1
 \end{equation}
@@
 where we assume $\beps \independent \bX\bgamma$.
-Rather than assuming anything about $\bgamma$ directly, an equivalent representation lets $\bu=\bX\bgamma$ such that $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent LMM model
+Rather than assuming anything about $\bgamma$ directly, an equivalent representation lets $\bu=\bX\bgamma$ such that $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent model
 \begin{equation}
@@
 \end{align*}
-By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation.
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation \anna{does this make sense?}.
 $\bW = \text{diag}\{w_1, ...
@@
 \end{align*}
-The assumption that environmental effects follow a random distribution, $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$, yields $\mathbb{E}(\hat{\bbeta}) = \bbeta$, and $\mathbb{V}(\hat{\bbeta}) = \sigmagg/p \bI + \sigmaee (\bX^T \bX)^{-1}$.
+The assumption that environmental effects follow a random distribution, $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$, yields $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = \sigmagg/p \bI + \sigmaee (\bX^T \bX)^{-1}$.
 Returning to our earlier example involving confounding due to air quality, randomly drawing a new value of $\bX\bgamma$ for each replication would mean that subpopulation A may be exposed to good air quality in one replication, and bad air quality in the next.
@@
 Considering $\bgamma$ as a fixed quantity is more consistent with the usual meaning of environmental confounding.
-Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\hat{\bbeta}) = \bbeta + \bgamma, \mathbb{V}(\hat{\bbeta}) = \sigmaee (\bX^T \bX)^{-1}$, which allows us to quantify bias in our estimates.
+Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^T \bX)^{-1}$, which allows us to quantify bias in our estimates.
 
@@
 \end{align*}
-where $\boldsymbol{\tau}$ quantifies bias in the estimated $\boldsymbol{\hat{\beta}'}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
+where $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat'$ and $\boldsymbol{\psi}$ is absorbed by the error term.
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
@@
 
-\pb{Just FYI, it's hard to tell bias and var apart in black and white} \anna{will fix this and greek letters}
+\pb{Just FYI, it's hard to tell bias and var apart in black and white} \anna{will fix this, greek letters, labels}
 

### 2020-07-03 main.tex
A few notes from 2020-07-02 meeting

@@
 
- \anna{in organic chemistry cis means 'on the same side' and trans 'on the other side' so I thought cis/trans might be more descriptive for gamma structure than homo/heterogeneous.
-These are also used to describe gender identity, which is obviously not how I'm using it, so I don't want it to be confusing.}
+ \anna{CONCORDANT/DISCORDANT... synchronous/asynchronous}
  To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis}, or (2) \anna{heterogeneous/trans} in nature for a total of four distinct $\bgamma$ structures.

### 2020-07-03 main.tex
Replacing 'homo/heterogeneous' with 'concordant/discordant' throughout.

@@
 \end{align*}
-By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation \anna{does this make sense?}.
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation.
 $\bW = \text{diag}\{w_1, ...
@@
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
-To see why this is the case, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\mu = 0$ for ease of notation: \anna{intercept stuff would get messy here...
-does this work?}
+To see why this is the case, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\mu = 0$ for ease of notation:
 \begin{align*}
@@
 
- \anna{CONCORDANT/DISCORDANT... synchronous/asynchronous}
- To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis}, or (2) \anna{heterogeneous/trans} in nature for a total of four distinct $\bgamma$ structures.
-We define \anna{homogenous/cis} effects as those that are similar among genetically similar subpopulations, and \anna{heterogeneous/trans} effects as those that are not.
+ To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) concordant, or (2) discordant in nature for a total of four distinct $\bgamma$ structures.
+We define concordant effects as those that are similar among genetically similar subpopulations, and discordant effects as those that are not.
 Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures.
@@
     \includegraphics[scale = 1.1]{figures/beta_mse.png}
-    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-\anna{heterogeneous/trans} environmental effects.}
+    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
     \label{fig:mse}
@@
      \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-\anna{heterogeneous/trans} environmental effects.}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
     \label{fig:big_vs_small}
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with independent subpopulation data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-\anna{heterogeneous/trans} environmental effect structure.}
+     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with independent subpopulation data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
     \label{fig:eta_beta_mse}
@@
     \includegraphics[scale = 0.8]{figures/time_ratio.png}
-    \caption{Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-\anna{heterogeneous/trans} environmental effect structure.}
+    \caption{Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
     \label{fig:time}

### 2020-07-07 main.tex
4 and 5 edits

@@
 $\bW = \text{diag}\{w_1, ...
-w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1}$.
+w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$.
 Together, this yields the objective function
@@
 \end{equation}
-An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of substantially fewer parameters: $p + 2$ for LMM-Lasso-Rakitsch, and $p + 3$ for LMM-Lasso-ggmix.
+An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of only $p + 3$ parameters.
 
@@
 This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix, bhatnagar2019simultaneous}.
-We postpone further discussion of the differences between these methods to the results section.
+We postpone further discussion of the differences between these methods to section \ref{sec:results}.
 
@@
 %------------------------------%
-\subsection{LMM-Lasso outperforms PC-Lasso when simulated subpopulations are small}
+\subsection{LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure}
 %------------------------------%
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of fine and coarse subpopulation structures.
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures.
+Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0, 0.2, 0.5,0.8\}$.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
+The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
+The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.
 
-Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying environmental confounding, $\xi \in \{0, 0.2, 0.5,0.8\}$.
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding simulated using a dichotomous-heterogeneous $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
-
-with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
-The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.
-In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line.
-However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations.
-In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
+In the presence of coarse population structure, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding boxes straddling the 0 line.
+However, the majority of these boxes fall above 0, particularly for lower $\eta$ values, indicating the superior performance of LMM-Lasso.
+In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
 This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
-As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.
+As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
 \begin{figure}[H]
@@
 %------------------------------%
-\subsection{LMM-Lasso outperforms PC-Lasso when $\gamma$ is heterogeneous}
+\subsection{LMM-Lasso outperforms PC-Lasso when environmental effects are discordant}
 %------------------------------%
 
-We evaluate PC-Lasso and LMM-Lasso in the presence of large and small subpopulations as well as distinct coefficient structures.
-Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) for all simulations, broken down by both subpopulation and coefficient structure.
-The right-hand panels of \ref{fig:big_small_gamma} are consistent with figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of small subpopulations, regardless of coefficient linearity or homogeneity.
-The panel of figure \ref{fig:big_small_gamma} corresponding to large subpopulations with heterogeneous coefficients also appears to be a setting where LMM-Lasso consistently outperforms PC-Lasso, regardless of coefficient linearity.
-Only in the setting of large supopulations with homogeneous coefficients does PC-Lasso outperform LMM-Lasso.
-In line with our initial intuition, PC-Lasso performs best within this setting when coefficients are linear in nature.
+In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider distinct patterns of environmental effects.
+Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure.
+Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the independent subpopulation simulated $\bX$.
+The right-hand panels of figure \ref{fig:big_small_gamma} are consistent with figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
+The upper left hand panel of figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso.
+The lower left hand panel of figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients.
+In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while PC-Lasso outperforms LMM-Lasso in the majority of dichotomous-concordant environmental effects.
+These findings are in line with our initial hypothesis that PC-Lasso will perform well in the presence of environmental effects that can be well-approximated by a linear combination of the observed data.
+When that is not the case, either due to the presence of fine subpopulation structure, nonlinear environmental effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.
 
@@
 
-Section~\ref{sec:rak-ggmix} briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+In section \ref{sec:rak-ggmix} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
 As we will see, the two generally perform quite similarly.
-Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of either LMM-Lasso method.
-
+Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods.
 The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.
-However, given the generally strong performance of LMM-Lasso, we also wish to briefly discuss the key differences between these two ways of implementing it, and provide guidance for potential users.
+However, given the generally strong performance of LMM-Lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.
 
@@
 LMM-Lasso-ggmix seeks to improve upon this by re-estimating the variance components and $\bbeta$ iteratively.
-Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$ or $\eta$.
+Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$.
+Figure \ref{fig:eta_beta_mse} shows the MSE for estimation of both $\bbeta$ and $\eta$ in the presence of varying $\eta$ values and sample sizes, $n$, for both LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+The bottom panel of this figure shows a modest improvement in the estimation of $\eta$ for the LMM-Lasso-ggmix method, particularly in the presence of lower $\eta$ values.
+However, these apparent improvements in estimating $\eta$ do not improve the $\bbeta$ estimation accuracy of LMM-Lasso-ggmix.
+This is evidenced by the top panel of figure \ref{fig:eta_beta_mse}, which shows nearly identical performances of both methods in terms of $\bbeta$ MSE.
 
@@
 The horizontal dashed line at 1 indicates both methods require the same computing time.
-Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes approximately 71 to 9 times as long as LMM-Lasso-Rakitsch.
+Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes requires as much as 75 times the computing time of LMM-Lasso-Rakitsch.
 Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this work, we have reviewed the concept of population stratification and defined explicitly the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship.
+In this work, we have reviewed the concept of population stratification and explicitly defined the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
@@
 We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively.
-We have also detailed the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity.
+We have also described the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity.
 In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
@@
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
-Additionally, we find that LMM-Lasso outperforms PC-Lasso when (1) small genetic subpopulations are present which correspond to environmental heterogeneity and (2) when confounding effects are nonlinear and heterogeneous in nature.
-In the presence of large subpopulations which correspond to homogeneous, linear environmental effects, PC-Lasso and LMM-Lasso perform comparably in estimation accuracy.
-While this scenario may be plausible, given that LMM-Lasso is more robust to subpopulation sizes and nonlinear effects, it may present a 'safer choice' for practitioners seeking to obtain unbiased estimates when the exact nature of population structure and corresponding environmental heterogeneity are unknown.
+Additionally, we find that LMM-Lasso outperforms PC-Lasso when (1) fine subpopulation structure is present and (2) when confounding effects are nonlinear and discordant in nature.
+In the presence of coarse subpopulation structure which corresponds to concordant, dichotomous environmental effects, PC-Lasso may outperform LMM-Lasso in estimation accuracy.
+However, LMM-Lasso is more robust to subpopulation structure as well as nonlinear and discordant environmental effects.
+Therefore, LMM-Lasso is a more robust method for practitioners seeking to obtain unbiased estimates when the exact nature of the population structure and corresponding environmental heterogeneity are unknown.
 
-We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
-Although we found that both methods perform similarly in terms of estimation accuracy in many scenarios, LMM-Lasso-ggmix has several distinct disadvantages.
+Finally, we considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
+Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks.
 Firstly, it suffers from instability in low dimensions.
-Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix takes approximately \anna{x} times longer due to its iterative estimation procedure.
-Thirdly, while the additional time required of LMM-Lasso-ggmix could be worthwhile if it resulted in a more accurate estimation of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are superior to those of LMM-Lasso-Rakitsch.
-We have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM (final name tbd)}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.
+Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires as much as \anna{75 - check} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure.
+Thirdly, while the additional time required of LMM-Lasso-ggmix may be worthwhile if it yielded more accurate estimates of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are greatly superior to those of LMM-Lasso-Rakitsch.
+We have implemented the LMM-Lasso-Rakitsch method using the equivalent, but more interpretable, $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM (final name tbd)}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.
 
@@
 
-Even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
+It is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
 Because of this, bias-reduction techniques for penalized LMMs may be an attractive avenue for future research.

### 2020-07-10 main.tex
A few notes from 2020-07-09 meeting about including unpenalized covariates in the model.

@@
 
+\anna{We show this in the absence of fixed covariates for the sake of a clear illustration.. $\bX_0 \bbeta_0$ unpenalized things}
+
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
@@
     \label{eqn:model}
-    \by = \Int + \bX\bbeta + \bZ\bgamma + \beps
+    \by = \text{\anna{unpenalized stuff}} + \bX\bbeta + \bZ\bgamma + \beps
 \end{equation}
@@
 \subsection{LMM-Lasso}
+
+\anna{We show this in the absence of fixed covariates for the sake of a clear illustration.. $\bX_0 \bbeta_0$ unpenalized things}
+
 LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
@@
 
+\anna{In the presence of unpenalized covariates to adjust for like age and sex... equations 7 and 8 and X-tilde become this... penalty term only penalizes beta1 - betas corresponding to SNPs}
+
 %------------------------------%
@@
 Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}, so this is the way we implement PC-Lasso throughout Section~\ref{sec:results}.
+
+\anna{put derivation comparing lmmlasso and pclasso back}
 

### 2020-07-20 main.tex
Figures: updating figures to use 1) new simulation data that incorporated a manual intercept, 2) black and white color scheme

@@
 
-\pb{Just FYI, it's hard to tell bias and var apart in black and white} \anna{will fix this, greek letters, labels}
-
 %------------------------------%
@@
 \end{figure}
+\anna{beta-hat and eta-hat seems like a bug with ggplot2 update (?) - checking on this}
 
@@
 \end{figure}
+\anna{Note figure \ref{fig:time} is on raw scale, not log10. This is because the negative values on the error bars create problems and become NaN on the log10 scale}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 0.8]{figures/time_ratio_box.png}
+    \caption{\anna{Alternate time fig} Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
+    \label{fig:time_box}
+\end{figure}
 
@@
 The horizontal dashed line at 1 indicates both methods require the same computing time.
-Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes requires as much as 75 times the computing time of LMM-Lasso-Rakitsch.
+Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes requires an average of \anna{73.4} times the computing time of LMM-Lasso-Rakitsch.
 Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
@@
 Firstly, it suffers from instability in low dimensions.
-Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires as much as \anna{75 - check} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure.
+Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires an average of 73.4 and as much as 167.57 \anna{73.4 is the mean ratio + sd, but the max absolute ratio is 167.57 - which should I report?} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure.
 Thirdly, while the additional time required of LMM-Lasso-ggmix may be worthwhile if it yielded more accurate estimates of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are greatly superior to those of LMM-Lasso-Rakitsch.

### 2020-07-20 main.tex
3.1 - 3.5: adding unpenalized covariates to models and derivations where appropriate.

@@
 
-\anna{We show this in the absence of fixed covariates for the sake of a clear illustration.. $\bX_0 \bbeta_0$ unpenalized things}
-
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
 We denote the normalized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
-We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
+Additionally, we allow $\bX_0$ to represent an $n \times p_0$ matrix of unpenalized covariates.
+At a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
+However, $\bX_0$ may additionally include covariates, such as age and sex, which are of interest to include in the model without penalty.
+We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of unpenalized covariates, additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
 \begin{equation}
     \label{eqn:model}
-    \by = \text{\anna{unpenalized stuff}} + \bX\bbeta + \bZ\bgamma + \beps
+    \by = \bX_0 \bbeta_0 + \bX\bbeta + \bZ\bgamma + \beps
 \end{equation}
-where $\Int$ is the intercept, $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bgamma$ is a $q \times 1$ vector of environmental effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+where $\bbeta_0$ is a $p_0 \times 1$ vector which includes the intercept term and the effects of any unpenalized covariates, $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bgamma$ is a $q \times 1$ vector of environmental effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects among subjects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
 Often it is assumed that subjects belong to $q \le p$ subpopulations, and $\bZ$ is a corresponding incidence matrix of 0s and 1s.
@@
     \label{eqn:k}
-    \bK &= \frac{1}{p} \bX \bX^T \nonumber \\
-                   &= \frac{1}{p} \bU \bD^2 \bU^T \nonumber \\
-                   &= \frac{1}{p} \bU \bD (\bU \bD)^T \nonumber \\
-                   &\propto \bR \bR^T,
+    \bK &= \frac{1}{p} \bX \bX^\T \nonumber \\
+                   &= \frac{1}{p} \bU \bD^2 \bU^\T \nonumber \\
+                   &= \frac{1}{p} \bU \bD (\bU \bD)^\T \nonumber \\
+                   &\propto \bR \bR^\T,
 \end{align}
@@
     \label{eqn:pca_reg}
-    \by = \Int + \bX\bbeta + \bU_{1:k}\boldsymbol{\alpha} + \beps 
+    \by = \bX_0\bbeta_0 + \bX\bbeta + \bU_{1:k}\boldsymbol{\alpha} + \beps 
 \end{equation}
@@
     \label{eqn:pca_obj}
-    Q(\bbeta, \boldsymbol{\alpha}|\bX, \by) = \frac{1}{2n} ||\by - \Int - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1
+    Q(\bbeta, \boldsymbol{\alpha}|\bX, \by) = \frac{1}{2n} ||\by - \bX_0\bbeta_0 - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1
 \end{equation}
@@
 Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
-Modeling unobserved confounding using fixed effects in this manner thus requires the estimation of $p + k + 1$ parameters.
 
@@
 \subsection{LMM-Lasso}
-
-\anna{We show this in the absence of fixed covariates for the sake of a clear illustration.. $\bX_0 \bbeta_0$ unpenalized things}
 
@@
     \label{eqn:lmm_reg}
-    \by = \Int + \bX\bbeta + \bX\bgamma+ \beps,
+    \by = \bX_0\bbeta_0 + \bX\bbeta + \bX\bgamma+ \beps,
 \end{equation}
@@
     \label{eqn:lmm_reg_u}
-    \by = \Int + \bX\bbeta + \bu + \beps.
+    \by = \bX_0\bbeta_0 + \bX\bbeta + \bu + \beps.
 \end{equation}
-Regardless of representation, the LMM assumes $\by \sim N(\Int + \bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
+We will show LMM-Lasso derivations in the absence of unpenalized covariates for the sake of a clear illustration.
+
+In the absence of unpenalized covariates, the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
 LMM-Lasso performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
 \begin{equation}
-\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \left(\Int +\bX \bbeta\right), \bI).
+\label{eqn:rtY_distn}
+\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX \bbeta, \bI).
 \end{equation}
 \anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
-\begin{comment}
-For ease of notation, we incorporate the intercept into the design matrix and let $\bX' = \begin{bmatrix} \textbf{1} & \bX \end{bmatrix}$ and $\bbeta' = \begin{pmatrix} \mu & \bbeta \end{pmatrix}^T$ so that we can express this more compactly as
-\begin{equation}
-\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX' \bbeta', \bI).
-\end{equation}
-\end{comment}
 Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation.
@@
     \bV^{-1} &= (\sigmagg \bK + \sigmaee \bI)^{-1}\\
-    &=(\sigmagg \bU \bD^2 \bU^T + \sigmaee \bU \bU^T)^{-1}\\
-    &= (\bU (\sigmagg \bD^2 + \sigmaee \bI) \bU^T)^{-1}\\
-    &=\bU (\sigmagg \bD^2 + \sigmaee \bI)^{-1} \bU^T.
+    &=(\sigmagg \bU \bD^2 \bU^\T + \sigmaee \bU \bU^\T)^{-1}\\
+    &= (\bU (\sigmagg \bD^2 + \sigmaee \bI) \bU^\T)^{-1}\\
+    &=\bU (\sigmagg \bD^2 + \sigmaee \bI)^{-1} \bU^\T\\
+    &=\bU \bW^2 \bU^\T.
 \end{align*}
-By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation.
-$\bW = \text{diag}\{w_1, ...
-w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$.
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^\T \bX$ and $\byT = \bU^\T \by$, where $\bW = \text{diag}\{w_1, ...
+w_n\}$ and $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$.
 Together, this yields the objective function
@@
 \end{equation}
-An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of only $p + 3$ parameters.
-
-\anna{In the presence of unpenalized covariates to adjust for like age and sex... equations 7 and 8 and X-tilde become this... penalty term only penalizes beta1 - betas corresponding to SNPs}
+In the presence of an intercept and unpenalized covariates, \eqref{eqn:rtY_distn} becomes 
+\begin{equation}
+\bV^{-\frac{1}{2}}\by \sim N\left(\bV^{-\frac{1}{2}} \bX{^\prime} \bbeta{^\prime}, \bI \right)
+\end{equation}
+where $\bX{^\prime} = \begin{bmatrix} \bX_0 & \bX \end{bmatrix}$ and $\bbeta{^\prime} = \begin{bmatrix} \bbeta_0^\T & \bbeta^\T \end{bmatrix}^\T$.
+This in turn leads to a new version of \eqref{eqn:lmm_obj}, our objective function,
+\begin{equation}
+Q(\bbeta | \bX{^\prime}, \byT, \bW) = || \bW (\byT - \bXT{^\prime} \bbeta{^\prime})||_2^2 + \lambda || \bbeta ||_1
+\end{equation}
+where $\bXT{^\prime} = \bU^\T \bX{^\prime}$, and only the elements of $\bbeta{^\prime}$ corresponing to SNPs are penalized.
 
@@
 Thus, whether modeled as a fixed or random effect, the underlying effect of $\bZ \bgamma$ on phenotype remains the same, but its treatment as a fixed or random effect has important implications in terms of model fitting and estimation.
-Including principal components as fixed effects requires estimation of a greater number of parameters, which may result in a loss of power \citep{zhang2015principal}.
+Modeling unobserved confounding using fixed effects requires estimation of $p_0 + p + k$ parameters, compared to the $p_0 + p + 2$ required of an LMM.
+This may result in a loss of power \citep{zhang2015principal}.
 Additionally, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \citep{patterson2006population, zhao2018practical}.
 It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \citep{zhang2015principal}.
-Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}, so this is the way we implement PC-Lasso throughout Section~\ref{sec:results}.
+Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}.
+This is the implementation of PC-Lasso throughout Section \ref{sec:results}.
 
@@
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
-To see why this is the case, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\mu = 0$ for ease of notation:
+To see why this is the case, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
 \begin{align*}
-    \mathbb{E}(\hat{\bbeta}) &= (\bX^T \bX)^{-1} \bX^T \mathbb{E}(\by)\\
-    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{E}(\bX\bbeta + \bX\bgamma + \beps)\\
+    \mathbb{E}(\hat{\bbeta}) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{E}(\by)\\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{E}(\bX\bbeta + \bX\bgamma + \beps)\\
     &\quad\\
-    \mathbb{V}(\hat{\bbeta}) &= (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\by)\\
-    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps)\\
-    &=  (\bX^T \bX)^{-1} \bX^T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^T \bX)^{-1}.
+    \mathbb{V}(\hat{\bbeta}) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\by)\\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps)\\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^\T \bX)^{-1}.
 \end{align*}
-The assumption that environmental effects follow a random distribution, $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$, yields $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = \sigmagg/p \bI + \sigmaee (\bX^T \bX)^{-1}$.
+The assumption that environmental effects follow a random distribution, $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$, yields $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = \sigmagg/p \bI + \sigmaee (\bX^\T \bX)^{-1}$.
 Returning to our earlier example involving confounding due to air quality, randomly drawing a new value of $\bX\bgamma$ for each replication would mean that subpopulation A may be exposed to good air quality in one replication, and bad air quality in the next.
@@
 Considering $\bgamma$ as a fixed quantity is more consistent with the usual meaning of environmental confounding.
-Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^T \bX)^{-1}$, which allows us to quantify bias in our estimates.
+Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, which allows us to quantify bias in our estimates.
 
@@
 
-Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^T \bX)^{-1} \bX^T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
+Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
 Then, $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ such that
@@
     &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps\\
-    &=  \bX \bbeta' + \beps'
+    &=  \bX \bbeta{^*} + \beps{^*}
 \end{align*}
-where $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat'$ and $\boldsymbol{\psi}$ is absorbed by the error term.
+where $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat{^\prime}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.

### 2020-08-12 lmm-genetic-review.tex
3.3 - 3.5: edits, explicit intercept term
3.1 - 3.5: adding unpenalized covariates to models and derivations where appropriate.
Abstract and Section 1: revising language

@@
 \providecommand{\anna}[1]{\textcolor{blue}{#1}}
-\newcommand{\bbeta}{\boldsymbol{\beta}}
+\newcommand{\bbeta}[1][]{\boldsymbol{\beta}{#1}}
 \newcommand{\bbetaT}{\widetilde{\boldsymbol{\beta}}}
-\newcommand{\bbetaHat}{\widehat{\boldsymbol{\beta}}}
+\newcommand{\bbetaHat}[1][]{\widehat{\boldsymbol{\beta}{#1}}}
 \newcommand{\bgamma}{\boldsymbol{\gamma}}
-\newcommand{\beps}{\boldsymbol{\varepsilon}}
-\newcommand{\bepsT}{\widetilde{\boldsymbol{\varepsilon}}}
-\newcommand{\bX}{\textbf{X}}
-\newcommand{\bXT}{\widetilde{\textbf{X}}}
+\newcommand{\beps}[1][]{\boldsymbol{\varepsilon}{#1}}
+\newcommand{\bepsT}[1][]{\widetilde{\boldsymbol{\varepsilon}{#1}}}
+\newcommand{\bX}[1][]{\textbf{X}#1}
+\newcommand{\bXT}[1][]{\widetilde{\textbf{X}{#1}}}
 \newcommand{\bZ}{\textbf{Z}}
@@
 \newcommand{\sigmagg}{\sigma^2_g}
+\newcommand{\T}{\text{T}}
 
@@
 Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.
-Considering this, we evaluate the performance of several penalized regression methods in terms of MSE in the presence of a number of confounding structures.
+Considering this, we evaluate the performance of several penalized regression methods in terms of MSE \anna{Do we want to revisit any of the precision-recall curves?} in the presence of a number of confounding structures.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.

### 2020-08-12 main.tex
Abstract and Section 1: revising language

@@
 
-An objective of many genetic studies is to accurately identify variants associated with a particular trait or disease, and to estimate the effects of those variants.
-Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a disease, and identifying potential drug targets.
-However, this effort may be hindered by several factors.
+An objective of many genetic studies is to accurately identify variants associated with complex traits or disease, and to estimate the effects of those variants.
+Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.
+However, this effort may be hindered by several factors such as non-independent errors, high dimensions, and linkage disequilibrium (LD).
 
-Genetic variants are often correlated with each other, meaning that univariate testing of genetic variants may result in false positives and biased effect estimates.
-This occurs when the variant being tested is correlated with other variants, so that the effects of these correlated variants are erroneously attributed to the variant in question.
-One method for analyzing data in the presence of correlated loci, or linkage disequilibrium (LD), is LD pruning.
+Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
+When a genetic variant of interest is highly correlated with others, the effects of these will be attributed to the variant in question.
+\anna{I feel like I'm saying 'variant' way too much...
+SNP?
+gene?} One method for analyzing data in the presence of correlated loci is LD pruning.
 LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
@@
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
-
 The Lasso \citep{tibshirani1996regression} is a widely used penalized regression method.
-It relies on an assumption of sparsity, which means that coefficient estimation and variable selection can be accomplished simultaneously, and also provides considerable computational advantages, meaning that it scales up well to very large data sets.
+It relies on an assumption of sparsity, which allows coefficient estimation and variable selection to be accomplished simultaneously.
+It also provides considerable computational advantages that scale up well to very large data sets.
 These features make the Lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
 
-Lasso penalized regression models typically assume independent observations and uncorrelated errors.
-In the presence of population structure, some samples are more correlated than others, violating the independent errors assumption.
+As with multiple linear regression models, Lasso penalized regression assumes independent observations and uncorrelated errors.
+In the presence of population structure or cryptic relatedness, some samples are more correlated than others, violating this assumption.
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, until recently the majority of this work has focused on plant and animal breeding data, much of which experimentally controlled.
-While many standard correction techniques may in principal be applied to human genetic data, their performance must be studied within the context of human ancestry as well as non-experimental conditions \citep{lawson2019population, barton2019population}.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled.
+While many standard correction techniques may in principal be applied to human genetic data, their performance has not been well-studied within the context of human ancestry as well as non-experimental conditions \citep{lawson2019population, barton2019population}.
 
@@
 Biased estimates of this nature severely hinder prediction and understanding differences among populations \citep{barton2019population}.
-Although bias at individual loci may be small, this bias may become magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \citep{barton2019population}.
+Although bias at individual loci may be small, this bias becomes magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \citep{barton2019population}.
 
-In order to clarify and emphasize the importance of this subtle distinction, we present a detailed review of relevant concepts and methods.
+In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
 The remainder of this paper is organized as follows.

### 2020-08-12 main.tex
Section 2: revisions

@@
 Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations.
-For example, European American \citep{campbell2005demonstrating}, Han Chinese \citep{xu2009genomic, chen2009genetic}, and most recently, cohorts within the UK Biobank data \citep{haworth2019apparent}.
+For example, European American \citep{campbell2005demonstrating}, Han Chinese \citep{xu2009genomic, chen2009genetic}, and most recently, cohorts within the UK Biobank data \citep{haworth2019apparent} have been shown to exhibit patterns of population and geographic structure despiet their seemingly similar subjects.
 
-Relatedness describes how genetically similar individuals are, and can refer to both ancient and recent relatedness.
-Recent relatedness is characterized by small clusters of high levels of similarity and is indicative of familial or \textbf{cryptic relatedness}.
+Population structure is broadly categorized bsed on whether it describes recent or ancient relatedness.
+Recent relatedness describes familial or \textbf{cryptic relatedness}.
 Ancient relatedness, or common ancestry, is generally characterized by large clusters of low levels of similarity.
-The presence of multiple distinct ancestry groups is known as \textbf{population stratification}.
+\anna{I need some sort of subjective cutoff for recent vs.
+ancient relatedness rather than describing them in terms of clusters...
+more details} The presence of multiple distinct ancestry groups in a sample is known as \textbf{population stratification}.
 
@@
 Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations.
-This phenomenon is commonly described as \textbf{confounding due to population stratification}.
-However, the mechanism of this phenomenon warrants some discussion.
-In particular, it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense, rather there must exist some non-genetic mechanism by which population stratification affects both the phenotype and genotype \citep{barton2019population, vilhjalmsson2012nature}.
+This phenomenon is commonly described as \textbf{confounding due to population stratification or ancestry}.
+The mechanism of this phenomenon warrants some discussion as it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense.
+While population structure may directly affect observed allele frequencies, there must also exist some mechanism by which population stratification affects the phenotype \citep{barton2019population, vilhjalmsson2012nature}, namely, the environment.
 \\
@@
 
-As an example, consider a genetic study to assess genetic variants associated with lung cancer in a sample comprised of subjects from two distinct subpopulations A and B.
-Assume the major allele of SNP $X$ is present with higher frequency in subpopulation A compared to subpopulation B, but it has no direct effect on lung cancer.
-Now suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality.
-Consequently, a GWAS analysis of data from these subpopulations would be predisposed to find SNP $X$ significantly associated with lung cancer.
+As an example, consider a genome wide association study (GWAS) to assess genetic variants associated with lung cancer in a sample comprised of subjects from two distinct subpopulations, A and B.
+Assume the minor allele of SNP $X$ is present with higher frequency in subpopulation A compared to subpopulation B, but has no direct effect on lung cancer.
+Also suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
+A GWAS of data from these subpopulations would find SNP $X$ to be significantly associated with lung cancer.
 This spurious association is not identified because of the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
@@
 
-The mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env}.
+This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env}.
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
@@
 Genetic relatedness and environmental effects become associated over time for complex reasons, which are beyond the scope of this paper.
-As we will discuss, many methods of correcting for population structure rely on this relationship and do not formally differentiate between effects of population structure and those of environment.
-Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure in question.
+However, as we will discuss, many methods of correcting for population structure rely on this relationship and do not formally differentiate between effects of population structure and those of environment.
+Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure and population in question.
 
@@
 \end{tikzpicture}
-\caption{Directed acyclic graph (DAG) of confounding due to population structure}
+\caption{Directed acyclic graph (DAG) of confounding due to population structure.}
 \label{fig:ps_env}
@@
 An additional consideration in genetic studies that we will briefly mention is that of genetic background.
-Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
+\anna{make the following sentence better} Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
 As \citep{vilhjalmsson2012nature} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
@@
 Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts.
-It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs.
-This approach thus seeks to remove the effect of environmental heterogeneity on observed SNPs indirectly.
+In order to reduce confounding these methods attempt to block the causal pathway between population structure and observed SNP data.
+This is illustrated by the blocked arrow in figure \ref{fig:ps_env_block}.
+These methods, in general, do not specifically address the environmental component of confounding due to population structure.
+Nevertheless, they may be effective if they are able to indirectly block the causal pathway between the environment and observed SNPs.
+Whether or not they are effective depends upon whether population structure can be accurately inferred, and if this is a good proxy for environmental effects.
 
-\anna{DAG here to illustrate this?}
+\begin{figure}[H]
+\centering
+\begin{tikzpicture}[%
+>=latex',
+circ/.style={draw, shape=circle, node distance=1cm, line width=1pt}]%Define the arrow type and style for circled nodes
+\draw[dashed] (0,1) node[left] (P) {Population Stratification} -- (2,1) node[right] (E) {Environment};
+\draw[->] (0,0) node[left] (X) {Observed SNPs} -- (2,0) node[right] (Y) {Phenotype}; %Line between X and Y
+\draw[->, red] (P) --node {|} (X) ; 
+\draw[->] (E)   -- (Y) ; 
+% \filldraw[color=red!60, fill=none, very thick] (0,1) ellipse (4.5 and 0.35);
+\end{tikzpicture}
+\caption{The causal pathway between the environment and observed SNPs may be indirectly blocked by many methods that attempt to reduce confounding due to population structure.}
+\label{fig:ps_env_block}
+\end{figure}
 
-The genomic control (GC) method is a univariate testing procedure which modifies a test statistic by a deflation factor.
+One of the first methods to address thie problem of confounding due to population structure is the genomic control (GC) method.
+GC is a univariate testing procedure which modifies a test statistic by a deflation factor.
 This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
@@
 Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall between these two extremes and are characterized by continuously varying relatedness and a gradient of air quality.
-By clustering those individuals within in one subpopulation or another we lose information about their fine-grained, subtle relatedness.
+By clustering those individuals within in one subpopulation or another we lose information about their more fine-grained, subtle relatedness.
 
 Another approach is to create an approximation of population structure using surrogate variables, and to adjust for these as an additional model covariates.
-Principal component-adjustment (PC-adjustment) methods use the first several principal components of the design matrix as these surrogate variables \citep{price2006principal}.
+Principal component-adjustment (PC-adjustment) methods use the first several left singular vectors of a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
 This approach may be used in univariate or multivariate frameworks, and allows for analysis on the entirety of a structured sample.
@@
 % rely on a pair-wise similarity matrix
-Like PC-adjustment methods, LMMs utilize the left singular vectors of the design matrix to estimate and subsequently correct for relatedness among subjects.
-However, the term LMM has been used to describe a number of distinct procedures for analyzing structured genetic data which differ in their objectives and methods.
+Like PC-adjustment methods, LMMs utilize the left singular vectors of a similarity matrix to estimate and subsequently correct for relatedness among subjects.
+In an LMM, however, these left singular vectors are treated as random, rather than fixed effects.
+Additionally, the term LMM has been used to describe a number of distinct but related procedures for analyzing structured genetic data which differ in their objectives and methods.
 There are two main objectives of LMMs used in a structured genetic context: estimating narrow-sense heritability, and estimating individual SNP effects.
@@
 LMMs can also be used to estimate SNP effects, and this can be done in either a univariate or multivariate manner.
-Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure using a genetic similarity matrix \citep{yu2006unified, kang2010variance, kang2008efficient}.
-However, as with all univariate testing approaches univariate LMM implementations may still produce spurious associations and biased effect estimates in the presence of LD.
+Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure by modeling it as a random effect using a genetic similarity matrix \citep{yu2006unified, kang2010variance, kang2008efficient}.
+However, as with all univariate testing approaches, univariate LMM implementations are prone to spurious associations and biased effect estimates in the presence of LD.
 
-Multivariate mixed model approaches reduce these problems by assessing the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while controlling for environmental confounding  \citep{Rakitsch2012, bhatnagar2019simultaneous}.
+Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while controlling for dependent error structures  \citep{Rakitsch2012, bhatnagar2019simultaneous}.
 This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes.
-These SNPs will be included in the model as both fixed and random effects, while SNPs with smaller effect sizes are modeled only as part of the random effect in order to control for environmental confounding.
+These SNPs will be included in the model as both fixed and random effects, while SNPs with smaller effect sizes are modeled only as part of the random effect in order to control for dependent errors due to relatedness and sample structure.
 
-The remainder of this review will focus on the performance of multivariate Lasso-penalized methods for the analysis of continuous, normally distributed outcomes.
+The remainder of this review will focus on the performance of such multivariate Lasso-penalized methods for the analysis of continuous, normally distributed outcomes.
 Specifically, PC-adjustment and LMM methods, which we will refer to as PC-Lasso and LMM-Lasso, respectively.
@@
 Subsequent references to LMM-Lasso apply to both implementations, unless otherwise stated.
-Where necessary, we will differentiate between these implementations as LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, respectively.
+Where necessary, we will differentiate between these implementations by referring to them as LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, respectively.
 
-In section \ref{sec:results}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
-Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \citep{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and which formalizes the non-genetic confounding mechanism of population stratification.\\
+%In section \ref{sec:results}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
+Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \citep{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and the environmental mechanism of confounding due to population structure.\\
 

### 2020-08-12 main.tex
Section 3: revisions

@@
 We denote the normalized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
-Additionally, we allow $\bX_0$ to represent an $n \times p_0$ matrix of unpenalized covariates.
-At a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
-However, $\bX_0$ may additionally include covariates, such as age and sex, which are of interest to include in the model without penalty.
-We assume that an $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of unpenalized covariates, additive SNP effects, environmental confounding, and observational noise and consider the linear mixed model,
+Additionally, $\bX_0$ represents an $n \times p_0$ matrix of unpenalized covariates and the intercept.
+Thus, at a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
+However, $\bX_0$ may additionally include covariates such as age and sex, which are of interest to include in the model without penalty.
+We assume that the $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of these covariates, additive SNP effects, environmental effects, and observational noise, and consider the linear mixed model,
 \begin{equation}
@@
 
-LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
-Using the singular value decomposition (SVD) of the normalized genotype matrix, we can derive the spectral decomposition of the RRM:
+LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
+Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
 
@@
 \end{align}
-where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are therefore proportional to the principal components of $\bX$ or $\bK$.
+where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are therefore proportional to the principal components of $\bK$.
 
@@
 \subsection{PC-Lasso}
-PC-Lasso uses the first $k$ principal components of $\bX$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates.
+PC-Lasso uses the first $k$ principal components of $\bK$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates.
 It has been shown \citep{hoffman2013correcting} that this corresponds to the regression model
@@
 
-LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
-$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.This results in the LMM regression model
+Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N\left(\boldsymbol{0}, p^{-1} \sigmagg \bI_p \right).$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
+This results in the LMM regression model
 \begin{equation}
@@
 where we assume $\beps \independent \bX\bgamma$.
-Rather than assuming anything about $\bgamma$ directly, an equivalent representation lets $\bu=\bX\bgamma$ such that $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent model
+Rather than assuming anything about $\bgamma$ directly, an equivalent representation paramaterizes the model in terms of $\bu=\bX\bgamma$ where $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent model
 \begin{equation}
@@
 \end{equation}
-We will show LMM-Lasso derivations in the absence of unpenalized covariates for the sake of a clear illustration.
+We will first derive the LMM-Lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
 
-In the absence of unpenalized covariates, the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
-LMM-Lasso performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
+Following equations \eqref{eqn:lmm_reg} and \eqref{eqn:lmm_reg_u}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
+LMM-Lasso uses \anna{an estimate of} the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
+Specifically, left-multiplying the original data by $\bV^{-1/2}$ yields
+
 \begin{equation}
 \label{eqn:rtY_distn}
-\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX \bbeta, \bI).
+\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX \bbeta, \bI)
 \end{equation}
+and a regular \anna{vanilla, naive, regular?} Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
 \anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
+
 Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation.
@@
 \end{equation}
-where $\bXT{^\prime} = \bU^\T \bX{^\prime}$, and only the elements of $\bbeta{^\prime}$ corresponing to SNPs are penalized.
+where $\bXT{^\prime} = \bU^\T \bX{^\prime}$, so that all of the data are transformed, but only the elements of $\bbeta{^\prime}$ corresponing to SNPs are penalized.
 
@@
 It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
-\citep{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citep{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ shown in equation \eqref{eqn:k}.
+\citep{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citep{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ \anna{which we briefly review here...}
 
 Thus, whether modeled as a fixed or random effect, the underlying effect of $\bZ \bgamma$ on phenotype remains the same, but its treatment as a fixed or random effect has important implications in terms of model fitting and estimation.
-Modeling unobserved confounding using fixed effects requires estimation of $p_0 + p + k$ parameters, compared to the $p_0 + p + 2$ required of an LMM.
-This may result in a loss of power \citep{zhang2015principal}.
-Additionally, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \citep{patterson2006population, zhao2018practical}.
+Modeling unobserved confounding using fixed effects requires estimation of $p_0 + p + k$ parameters, compared to the $p_0 + p + 2$ required of an LMM, which may result in a loss of power \citep{zhang2015principal}.
+Additionally, when modeling $\bZ \bgamma$ using fixed effects, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \citep{patterson2006population, zhao2018practical}.
 It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \citep{zhang2015principal}.
@@
 This is the implementation of PC-Lasso throughout Section \ref{sec:results}.
-
-\anna{put derivation comparing lmmlasso and pclasso back}
 
@@
 We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method.
-Averaging over situations in which the environmental confounding is constantly changing direction and magnitude ``washes out'' bias in $\hat{\bbeta}$ while inflating the variance.
-To see why this is the case, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
-\begin{align*}
-    \mathbb{E}(\hat{\bbeta}) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{E}(\by)\\
-    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{E}(\bX\bbeta + \bX\bgamma + \beps)\\
-    &\quad\\
-    \mathbb{V}(\hat{\bbeta}) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\by)\\
-    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps)\\
-    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^\T \bX)^{-1}.
-\end{align*}
-The assumption that environmental effects follow a random distribution, $\bgamma \sim N(\boldsymbol{0}, \sigmagg / p \bI_p)$, yields $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = \sigmagg/p \bI + \sigmaee (\bX^\T \bX)^{-1}$.
-Returning to our earlier example involving confounding due to air quality, randomly drawing a new value of $\bX\bgamma$ for each replication would mean that subpopulation A may be exposed to good air quality in one replication, and bad air quality in the next.
-Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from fixed populations A and B?
-Those repeated samples would show a systematic bias due to the difference in air quality, but this is masked if we generate new values of $\bX\bgamma$ with each replication.
-Considering $\bgamma$ as a fixed quantity is more consistent with the usual meaning of environmental confounding.
-Specifically, in the low-dimensional OLS setting, under the assumption that the quantity $\bgamma$ is fixed, $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, which allows us to quantify bias in our estimates.
+Averaging over situations in which the environmental confounding is constantly changing direction and magnitude is inconsistent with the confounding environmental effects.
+This has the effect of ``washing out'' bias in $\hat{\bbeta}$, which is of interest, while inflating the variance.
+To illustrate this concept, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
+\begin{align}
+    \mathbb{E}(\bbetaHat) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{E}(\by) \notag \\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{E}(\bX\bbeta + \bX\bgamma + \beps) \label{eqn:expectation}\\
+    &\quad \notag\\
+    \mathbb{V}(\bbetaHat) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\by) \notag\\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps) \notag\\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^\T \bX)^{-1} \label{eqn:variance}.
+\end{align}
+Under the assumption that environmental effects arise from a random process, $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg I_p)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
+Consider our earlier example involving confounding due to air quality.
+Simulating $\bgamma$ as a random quantity for each replication corresponds to subpopulation A being exposed to good air quality in one replication, and bad air quality in the next.
+Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from populations A and B, when these populations are subject to different environmental exposures?
+Those repeated samples would show a systematic bias due to the difference in air quality, but this is masked if $\bgamma$ is assumed to be random.
+Treating $\bgamma$ as a fixed quantity so that environmental effects are consistent across replications more closely approximates environmental confounding.
+Under the assumption that the quantity $\bgamma$ is fixed, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, allowing us to quantify estimation bias due to environmental heterogeneity.
 
@@
 We see this by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.
-
 Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
@@
 
-Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy and bias specifically, we simulate according to a fixed $\bgamma$ framework and using the following data-generating model
+Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy and bias specifically, we simulate according to a fixed $\bgamma$ framework using the following data-generating model
 \begin{equation}
@@
 \end{equation}
-where $\by$ represents the pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1, and is partitioned into signal, environmental confounding, and random noise components,
+where $\by$ represents the pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
 \begin{equation}
@@
 \end{equation}
-Here, $\eta$ is the proportion of the variance of $\by$ attributable to genetic factors, commonly referred to as the narrow-sense heritability, and $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-$\eta$ also represents our data-generating model's signal to noise ratio (SNR).
+$\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$.
+$\eta$ also describes the data-generating model's signal to noise ratio (SNR).
+$\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
 So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to environmental effects.
-The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
+\anna{more details about how/why this lambda selection method - cite CV as bad for LMMs, BIC not great bc different likelihoods} The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
+
+%This method for selecting $\lambda$ was chosen to fairly compare the
+
+%Cross-validation has generally been found to be inadvisable for penalized LMM methods \anna{cite thiiiiis}
+%Using an information criterion such as BIC to select $\lambda$ is not comparable across methods since Lasso, PC-Lasso, LMM-Lasso-Rakitsch, and LMM-Lasso-ggmix methods each use a different likelihood.
 
@@
 We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
-\anna{cite Vanderwas data} $p = 1000$ for all three data types, $n = 200$ for the simulated data types, and 50 causal SNPs were used throughout.
+\anna{cite Vanderwas data} $p = 1000$, $n = \{200, 400, 600, 800 \}$ for the simulated data types, and 50 causal SNPs were used throughout.
 
@@
 \subsubsection{Structure of environmental effects}
-For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation membership, or founding population membership in the case of the admixed data, which assigns the corresponding environmental effect found in $\bgamma$.
+For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$.
 Various $\bgamma$ structures were considered in order to investigate the effects of various forms of environmental confounding.
@@
 To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous.
-\anna{more details?} In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
+In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
 For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
 However, behavior-based environmental exposures may be less likely to be shared by subpopulations which are genetically and geographically similar.
-\anna{another citation?} As an example, \citep{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
+As an example, \citep{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
 
- To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) concordant, or (2) discordant in nature for a total of four distinct $\bgamma$ structures.
+ To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are concordant or discordant in nature, for a total of four distinct $\bgamma$ structures.
 We define concordant effects as those that are similar among genetically similar subpopulations, and discordant effects as those that are not.
-Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures.
-\anna{is this still unclear?
-Would a PCA picture help?}
+Figure \ref{fig:gamma_structures} displays a visual summary of these four coefficient structures.
 \begin{figure}[H]

### 2020-08-13 main.tex
Sections 4 and 5: small edits, notation fixes

@@
 \begin{equation}
-Q(\bbeta | \bX{^\prime}, \byT, \bW) = || \bW (\byT - \bXT{^\prime} \bbeta{^\prime})||_2^2 + \lambda || \bbeta ||_1
+Q(\bbeta | \bXT{^\prime}, \byT, \bW) = || \bW (\byT - \bXT{^\prime} \bbeta{^\prime})||_2^2 + \lambda || \bbeta ||_1
 \end{equation}
@@
 
-We have so far used the term $\bW$ to describe the general LMM-Lasso method.
+We have so far used the term $\bW$ to describe the LMM-Lasso method.
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
@@
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude is inconsistent with the confounding environmental effects.
-This has the effect of ``washing out'' bias in $\hat{\bbeta}$, which is of interest, while inflating the variance.
+This has the effect of ``washing out'' bias in $\bbetaHat$, which is of interest, while inflating the variance.
 To illustrate this concept, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
@@
 \end{figure}
-\anna{beta-hat and eta-hat seems like a bug with ggplot2 update (?) - checking on this}
 
@@
 However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications.
-Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.
+\anna{time interpretation depends on reason ggmix is slow...} Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.
 The horizontal dashed line at 1 indicates both methods require the same computing time.
-Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes requires an average of \anna{73.4} times the computing time of LMM-Lasso-Rakitsch.
+Despite its comparable performance in accurately estimating $\bbetaHat$, LMM-Lasso-ggmix takes requires an average of \anna{73.4} times the computing time of LMM-Lasso-Rakitsch.
 Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this work, we have reviewed the concept of population stratification and explicitly defined the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship.
+In this work, we have reviewed the concept of population stratification and explicitly defined the environmental mechanism by which it may confound the genotype-phenotype relationship.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
@@
 In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
-We then compare the estimation accuracy of a naive lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
+We then compare the estimation accuracy of a \anna{naive} Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
 
-Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the naive lasso whenever environmental confounding effects are present.
-However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the naive lasso.
+Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the \anna{naive} Lasso whenever environmental confounding effects are present.
+However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the \anna{naive} Lasso.
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
@@
 Firstly, it suffers from instability in low dimensions.
-Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires an average of 73.4 and as much as 167.57 \anna{73.4 is the mean ratio + sd, but the max absolute ratio is 167.57 - which should I report?} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure.
+\anna{depends on ggmix time results} Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires an average of 73.4 and as much as 167.57 \anna{73.4 is the mean ratio + sd, but the max absolute ratio is 167.57 - which should I report?} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure.
 Thirdly, while the additional time required of LMM-Lasso-ggmix may be worthwhile if it yielded more accurate estimates of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are greatly superior to those of LMM-Lasso-Rakitsch.
@@
 It is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
-Because of this, bias-reduction techniques for penalized LMMs may be an attractive avenue for future research.
+Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.

### 2020-08-13 lmm-genetic-review.tex
Edits to 1 through 2.2

@@
 \title{Review Paper Draft}
-\author{Anna Reisetter}
+\author{Anna Reisetter \qquad Patrick Breheny\\
+  Department of Biostatistics\\University of Iowa}
 \date{\today}
@@
 Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to unobserved confounding factors, such as environmental heterogeneity.
-This poses a challenge to the detection of multivariate associations of interest and is known to induce spurious associations when left unaccounted for.
+This poses a challenge to detecting associations of interest and is known to induce spurious associations when left unaccounted for.
 Penalized linear mixed models (LMMs) are an attractive method to correct for unobserved confounding.
 These methods correct for varying levels of relatedness and population structure by modeling it as a random effect with a covariance structure estimated from observed genetic data.
-Despite extensive literature on penalized regression, and LMMs, there is a need for a statistically rigorous review of penalized LMMs.
-Specifically, the ability of LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
+Despite an extensive literature on penalized regression and LMMs separately, the two are rarely discussed together.
+The aim of this review is to do so while offering a rigorous study of the statistical properties of penalized LMMs.
+Specifically, the ability of penalized LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
 Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.
-Considering this, we evaluate the performance of several penalized regression methods in terms of MSE \anna{Do we want to revisit any of the precision-recall curves?} in the presence of a number of confounding structures.
+Considering this, we evaluate the performance of several penalized regression methods in terms of MSE \anna{Do we want to revisit any of the precision-recall curves?} \pb {yes, let's talk about this today -- it is highly likely a reviewer will bring it up} in the presence of a number of confounding structures, comparing penalized LMMs to other approaches.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.

### 2020-08-13 main.tex
Edits to 1 through 2.2

@@
 
-An objective of many genetic studies is to accurately identify variants associated with complex traits or disease, and to estimate the effects of those variants.
+The objective of many genetic studies is to accurately identify variants associated with complex traits or disease, and to estimate the effects of those variants.
 Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.
-However, this effort may be hindered by several factors such as non-independent errors, high dimensions, and linkage disequilibrium (LD).
+However, this effort may be hindered by several factors such as non-independent errors \pb{is ``error'' the best term here?}, high dimensions \pb{is ``dimension'' the best term here?}, and linkage disequilibrium (LD).
 
-Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
+\pb{This paragraph seems like too much of a digression \ldots let's get to the point sooner} Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
 When a genetic variant of interest is highly correlated with others, the effects of these will be attributed to the variant in question.
@@
 An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
-However, such methods rely on data where the number of observations is less than the number of features.
-This is often infeasible in genetic studies where the number of features typically exceeds the number of observations.
+However, classical approaches rely on collecting data with more observations than unknown parameters.
+This is typically infeasible in genome-wide studies where the number of features greatly exceeds the number of observations.
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
-The Lasso \citep{tibshirani1996regression} is a widely used penalized regression method.
+The Lasso \citep{tibshirani1996regression} is a widely used penalized regression method \pb{refs at end?}.
 It relies on an assumption of sparsity, which allows coefficient estimation and variable selection to be accomplished simultaneously.
@@
 As with multiple linear regression models, Lasso penalized regression assumes independent observations and uncorrelated errors.
-In the presence of population structure or cryptic relatedness, some samples are more correlated than others, violating this assumption.
+In the presence of population structure or cryptic relatedness, however, samples are correlated, violating this assumption.
 If unaccounted for, this dependency can result in the identification of spurious associations.
 Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled.
-While many standard correction techniques may in principal be applied to human genetic data, their performance has not been well-studied within the context of human ancestry as well as non-experimental conditions \citep{lawson2019population, barton2019population}.
+While many standard correction techniques may in principal be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and non-experimental conditions \citep{lawson2019population, barton2019population}.
 
 One reason that population stratification warrants such concern in genetic studies is that due to ethnic and geographic segregation, population stratification tends to be associated with differing environmental exposures and cultural practices.
-In other words, random genetic variation is likely to be associated with non-genetic factors that also influence the trait of interest, thereby confounding the genotype-phenotype relationship and leading to spurious associations and biased estimates of SNP effects.
+In other words, random genetic variation is likely to be correlated with non-genetic factors that also influence the trait of interest, thereby confounding the genotype-phenotype relationship and leading to spurious associations and biased estimates of SNP effects.
 Biased estimates of this nature severely hinder prediction and understanding differences among populations \citep{barton2019population}.
@@
 
-In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
+In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods \pb{I really like this sentence \ldots can we work it into the abstract?}.
 The remainder of this paper is organized as follows.
@@
 In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-Lasso) and lasso penalized multivariate LMMs (LMM-Lasso).
-We review the statistical details of these methods as well as the consequences of the ways in which they attempt to correct for population structure and environmental heterogeneity.
+We review the statistical details of these methods as well as the consequences \pb{do we review \emph{consequences} in section 3?} of the ways in which they attempt to correct for population structure and environmental heterogeneity.
 In section \ref{sec:results}, we illustrate the concepts described in section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
-These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods.
+These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods \pb{are we still illustrating this?}.
 In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
@@
 
-Population structure is broadly categorized bsed on whether it describes recent or ancient relatedness.
+Population structure is broadly categorized based on whether it describes recent or ancient relatedness.
 Recent relatedness describes familial or \textbf{cryptic relatedness}.
@@
 ancient relatedness rather than describing them in terms of clusters...
-more details} The presence of multiple distinct ancestry groups in a sample is known as \textbf{population stratification}.
+more details} \pb{I would suggest something along the lines of: cryptic relatedness means that only a few generations, maybe 3 or so, separate common ancestors} The presence of multiple distinct ancestry groups in a sample is known as \textbf{population stratification}.
 
@@
 While population structure may directly affect observed allele frequencies, there must also exist some mechanism by which population stratification affects the phenotype \citep{barton2019population, vilhjalmsson2012nature}, namely, the environment.
-\\
 
@@
 This spurious association is not identified because of the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
-Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be found to be associated with the phenotype.\\
+Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be found to be associated with the phenotype.
 
@@
 
-This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env}.
+This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env} \pb{typically, this is capitalized so it matches how Figure 1 appears in the caption}.
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
@@
 In the context of our air quality example, this means that we assume air quality is not directly altering individuals' allele frequencies, nor are allele frequencies causing individuals to move to regions of better or worse air quality.
-Genetic relatedness and environmental effects become associated over time for complex reasons, which are beyond the scope of this paper.
+Genetic relatedness and environmental effects may become associated over time for any number of complex reasons beyond the scope of this paper.
 However, as we will discuss, many methods of correcting for population structure rely on this relationship and do not formally differentiate between effects of population structure and those of environment.
 Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure and population in question.
+\pb{is this an important point?}
 
@@
 \anna{make the following sentence better} Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
-As \citep{vilhjalmsson2012nature} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
+As \citep{vilhjalmsson2012nature} \pb{Use citet, not citep here} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.

### 2020-08-13 lmm-genetic-review.tex
2020-08-13 meeting notes/edits

@@
 Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.
-Considering this, we evaluate the performance of several penalized regression methods in terms of MSE \anna{Do we want to revisit any of the precision-recall curves?} \pb {yes, let's talk about this today -- it is highly likely a reviewer will bring it up} in the presence of a number of confounding structures, comparing penalized LMMs to other approaches.
+Considering this, we evaluate the performance of several penalized regression methods in terms of MSE \anna{Do we want to revisit any of the precision-recall curves?} \pb {yes, let's talk about this today -- it is highly likely a reviewer will bring it up} \anna{yes - work in the proportion of the first 50 vars that are correct, not necessarily a prc curve.
+Consider partial AUC...
+area under curve up to something like 50\% false discovery rate - focus on more relevant part in the path - google 'partial auc'} in the presence of a number of confounding structures, comparing penalized LMMs to other approaches.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.

### 2020-08-13 main.tex
2020-08-13 meeting notes/edits

@@
 
+\anna{get to the point sooner}
+
 The objective of many genetic studies is to accurately identify variants associated with complex traits or disease, and to estimate the effects of those variants.
 Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.
-However, this effort may be hindered by several factors such as non-independent errors \pb{is ``error'' the best term here?}, high dimensions \pb{is ``dimension'' the best term here?}, and linkage disequilibrium (LD).
+However, this effort may be hindered by several factors such as non-independent samples, high dimensions \pb{is ``dimension'' the best term here?} \anna{large numbers of variants >> n}, and linkage disequilibrium (LD).
 
-\pb{This paragraph seems like too much of a digression \ldots let's get to the point sooner} Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
+\pb{This paragraph seems like too much of a digression \ldots let's get to the point sooner} \anna{this could be more like a sentence} Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
 When a genetic variant of interest is highly correlated with others, the effects of these will be attributed to the variant in question.
@@
 In section \ref{sec:results}, we illustrate the concepts described in section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
-These simulation results are used to illustrate the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods \pb{are we still illustrating this?}.
 In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
@@
 Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure and population in question.
-\pb{is this an important point?}
+\pb{is this an important point?} \anna{hmm...
+maybe?
+robust to this assumption: if it's not a good proxy it's not going to cause confounding}
 
@@
 An additional consideration in genetic studies that we will briefly mention is that of genetic background.
-\anna{make the following sentence better} Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
-As \citep{vilhjalmsson2012nature} \pb{Use citet, not citep here} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
+\anna{make the following sentence better: correlated snps are the issue - talk about ld pruning} Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
+As \citet{vilhjalmsson2012nature} \pb{Use citet, not citep here} \anna{double check all citations for whether they should be (1) citet, citep or (2) better at the end of the sentence} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
-Although there are a number of methods in which to account for genetic background \anna{what are they?}, the methods upon which we will focus address this by conducting a genomewide multivariate analysis.
+Although there are a number of methods in which to account for genetic background \anna{what are they?
+\textbf{LD pruning} or block-wise analysis, gene-level test, \textbf{haplotype block}...group correlated variants}, the methods upon which we will focus address this by conducting a genomewide multivariate analysis.
 
@@
 
+\anna{switch first and second paragraphs... the current first paragraph doesn't describe GC}
+
 Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts.
@@
 This is illustrated by the blocked arrow in figure \ref{fig:ps_env_block}.
-These methods, in general, do not specifically address the environmental component of confounding due to population structure.
+These methods do not directly address the environmental component of confounding due to population structure.
 Nevertheless, they may be effective if they are able to indirectly block the causal pathway between the environment and observed SNPs.
 Whether or not they are effective depends upon whether population structure can be accurately inferred, and if this is a good proxy for environmental effects.
+\anna{maybe say something about how 'if it's a good proxy - great!
+they'll work', but 'if it's not a good proxy, also fine!
+because then it's not a confounder!
+...SE's potentially a problem like in GLS...
+but this is different with lasso.
+(maybe not worth getting into here)}
+
+\anna{make a better slashed arrow}
 
@@
 
+\anna{in admixed data less of the environmental effect is in the column space of X so it's less of a confounder}
+
 \begin{figure}[H]
@@
 \end{figure}
+\anna{add time to this figure? third row...}
 
@@
 \end{figure}
+
+\anna{make new time box plot - two boxes instead of ratio - also run for n = 1000, 1200, maybe 1400 and show MSE along with time... compare convergence results to ncvfit}
 
@@
 Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.
+
+\anna{the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods - do we want to get into the importance of not conflating environmental effects with heritability?}

### 2020-08-15 lmm-genetic-review.tex
Abstract edits

--- 2020-08-13 lmm-genetic-review.tex
@@
 Specifically, the ability of penalized LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
-Additionally, the non-genetic mechanism by which population stratification may confound the phenotype-genotype relationship is rarely discussed or treated explicitly in the statistical genetics literature.
-Considering this, we evaluate the performance of several penalized regression methods in terms of MSE \anna{Do we want to revisit any of the precision-recall curves?} \pb {yes, let's talk about this today -- it is highly likely a reviewer will bring it up} in the presence of a number of confounding structures, comparing penalized LMMs to other approaches.
+In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
+In addition, we evaluate the performance of penalized LMMs and competing methods in terms of MSE \anna{and variable selction (specifically how we quantify this tbd)} in the presence of a number of confounding structures.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.

### 2020-08-15 main.tex
Abstract edits

@@
 Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.
-However, this effort may be hindered by several factors such as non-independent samples, high dimensions \pb{is ``dimension'' the best term here?} \anna{large numbers of variants >> n}, and linkage disequilibrium (LD).
+However, this effort may be hindered by several factors such as non-independent samples, large numbers of variants, and linkage disequilibrium (LD).
 
@@
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
-The Lasso \citep{tibshirani1996regression} is a widely used penalized regression method \pb{refs at end?}.
+The Lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
 It relies on an assumption of sparsity, which allows coefficient estimation and variable selection to be accomplished simultaneously.

### 2020-08-15 main.tex
Section 1 edits

@@
 
-\anna{get to the point sooner}
-
 The objective of many genetic studies is to accurately identify variants associated with complex traits or disease, and to estimate the effects of those variants.
@@
 
-\pb{This paragraph seems like too much of a digression \ldots let's get to the point sooner} \anna{this could be more like a sentence} Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
+% \anna{keep this for later use when talking about genetic background?} Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
 When a genetic variant of interest is highly correlated with others, the effects of these will be attributed to the variant in question.
-\anna{I feel like I'm saying 'variant' way too much...
-SNP?
-gene?} One method for analyzing data in the presence of correlated loci is LD pruning.
+This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci, as opposed to the causal loci themselves, can be identified and their effects estimated.
+
+One method for analyzing data in the presence of correlated loci is LD pruning.
 LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
 The overarching approach of these methods is to sequentially thin SNPs, so that only SNPs correlated with each other below some desired threshold are used in modeling and testing procedures.
-This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci, as opposed to the causal loci themselves, can be identified and their effects estimated.
-
-% LD pruning thus limits analysis to SNPs which are representative of a particular region.
-
-% In some scenarios pruning algorithms may not select any representative SNPs in a particular region of the genome \citep{prive2018efficient}.
-However, LD-pruning has been shown to improve performance of principal component analysis (PCA) methods \citep{abdellaoui2013population} and \anna{what} \anna{because why}.
-
 An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
@@
 
-In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods \pb{I really like this sentence \ldots can we work it into the abstract?}.
+In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
 The remainder of this paper is organized as follows.
@@
 In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-Lasso) and lasso penalized multivariate LMMs (LMM-Lasso).
-We review the statistical details of these methods as well as the consequences \pb{do we review \emph{consequences} in section 3?} of the ways in which they attempt to correct for population structure and environmental heterogeneity.
+We review the statistical details of these methods as well as the consequences \pb{do we review \emph{consequences} in section 3?} \anna{tbd} of the ways in which they attempt to correct for population structure and environmental heterogeneity.
 In section \ref{sec:results}, we illustrate the concepts described in section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.

### 2020-08-15 main.tex
Section 1: more edits

@@
 
-One method for analyzing data in the presence of correlated loci is LD pruning.
+One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs, so that only those correlated with each other below some desired threshold are used in modeling and testing procedures.
 LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
-The overarching approach of these methods is to sequentially thin SNPs, so that only SNPs correlated with each other below some desired threshold are used in modeling and testing procedures.
 An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.

### 2020-08-15 main.tex
All sections: updating citation and figure refs.

@@
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}.
 While many standard correction techniques may in principal be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and non-experimental conditions \citep{lawson2019population, barton2019population}.
 
-One reason that population stratification warrants such concern in genetic studies is that due to ethnic and geographic segregation, population stratification tends to be associated with differing environmental exposures and cultural practices.
+One reason that population stratification warrants such concern in genetic studies is that due to ethnic and geographic segregation, population stratification tends to be associated with differing environmental exposures and cultural practices \citep{thornton2015statistical}.
 In other words, random genetic variation is likely to be correlated with non-genetic factors that also influence the trait of interest, thereby confounding the genotype-phenotype relationship and leading to spurious associations and biased estimates of SNP effects.
 Biased estimates of this nature severely hinder prediction and understanding differences among populations \citep{barton2019population}.
-Although bias at individual loci may be small, this bias becomes magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \citep{barton2019population}.
+Although bias at individual loci may be small, this bias becomes magnified when aggregated across thousands of SNPs, as is done when calculating polygenic risk scores \citep{barton2019population, peterson2019genome}.
 
@@
 In section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
-We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations \citep{Sillanpaeae2011, sul2018population}, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures.
+We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures \citep{Sillanpaeae2011, sul2018population, vilhjalmsson2012nature, barton2019population}.
 In so doing, we refine the definition of confounding due to population structure in terms of a non-genetic mechanism.
@@
 Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations.
-For example, European American \citep{campbell2005demonstrating}, Han Chinese \citep{xu2009genomic, chen2009genetic}, and most recently, cohorts within the UK Biobank data \citep{haworth2019apparent} have been shown to exhibit patterns of population and geographic structure despiet their seemingly similar subjects.
+For example, European American, Han Chinese, and most recently, cohorts within the UK Biobank data have been shown to exhibit patterns of population and geographic structure despiet their seemingly similar subjects \citep{campbell2005demonstrating, xu2009genomic, chen2009genetic, haworth2019apparent}.
 
@@
 The mechanism of this phenomenon warrants some discussion as it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense.
-While population structure may directly affect observed allele frequencies, there must also exist some mechanism by which population stratification affects the phenotype \citep{barton2019population, vilhjalmsson2012nature}, namely, the environment.
+While population structure may directly affect observed allele frequencies, there must also exist some mechanism by which population stratification affects the phenotype, namely, the environment \citep{barton2019population, vilhjalmsson2012nature}.
 
@@
 
-This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env} \pb{typically, this is capitalized so it matches how Figure 1 appears in the caption}.
+This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in Figure \ref{fig:ps_env}.
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
@@
 In order to reduce confounding these methods attempt to block the causal pathway between population structure and observed SNP data.
-This is illustrated by the blocked arrow in figure \ref{fig:ps_env_block}.
+This is illustrated by the blocked arrow in Figure \ref{fig:ps_env_block}.
 These methods do not directly address the environmental component of confounding due to population structure.
@@
 
-Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while controlling for dependent error structures  \citep{Rakitsch2012, bhatnagar2019simultaneous}.
+Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependent error structures  \citep{Rakitsch2012, bhatnagar2019simultaneous}.
 This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes.
@@
 Specifically, PC-adjustment and LMM methods, which we will refer to as PC-Lasso and LMM-Lasso, respectively.
-We consider two similar implementations of the LMM-Lasso method, one developed by \citep{Rakitsch2012} and a more recent implementation from \citep{bhatnagar2019simultaneous}.
+We consider two similar implementations of the LMM-Lasso method, one developed by \citet{Rakitsch2012} and a more recent implementation from \citet{bhatnagar2019simultaneous}.
 Subsequent references to LMM-Lasso apply to both implementations, unless otherwise stated.
@@
 
-%In section \ref{sec:results}, we compare the performance of PC-Lasso and LMM-Lasso in terms of their abilities to accurately estimate SNP effects and in terms of their true sign rates in the presence of varying levels of relatedness and environmental confounding.
-Although several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \citep{wang2013analytical, kang2010variance, zhao2007arabidopsis}, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and the environmental mechanism of confounding due to population structure.\\
+Several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \citep{wang2013analytical, kang2010variance, zhao2007arabidopsis}.
+However, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and the environmental mechanism of confounding due to population structure.\\
 
@@
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
-We denote the normalized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
+We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{2 p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
 Additionally, $\bX_0$ represents an $n \times p_0$ matrix of unpenalized covariates and the intercept.
@@
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-LMM-Lasso-Rakitsch \citep{Rakitsch2012} uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$.
+LMM-Lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
 These variance components and the corresponding weights are estimated once under the assumption of a null model.
 A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
-LMM-Lasso-ggmix \citep{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm.
-This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix, bhatnagar2019simultaneous}.
+LMM-Lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2019simultaneous}.
+This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix}.
 We postpone further discussion of the differences between these methods to section \ref{sec:results}.
@@
 It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
-\citep{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citep{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ \anna{which we briefly review here...}
+\citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ \anna{which we briefly review here...}
 
@@
 
-In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice \citep{Rakitsch2012, bhatnagar2019simultaneous} is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis.
+In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis \citep{Rakitsch2012, bhatnagar2019simultaneous}.
 We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method.
@@
 
+\anna{bring this back to why it may not matter if genetic similarity isn't a good proxy for population structure - in that case it's not a confounder anyway}
+
 %------------------------------%
@@
 So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to environmental effects.
-\anna{more details about how/why this lambda selection method - cite CV as bad for LMMs, BIC not great bc different likelihoods} The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
+\anna{more details about how/why this lambda selection method - cite CV as bad for LMMs, BIC not great bc different likelihoods...
+tie this into evaluating performance of each method in terms of variable selection} The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
@@
 \subsubsection{Design matrix generation}
-We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations.
-\anna{cite Vanderwas data} $p = 1000$, $n = \{200, 400, 600, 800 \}$ for the simulated data types, and 50 causal SNPs were used throughout.
+We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data}.
+$p = 1000$, $n = \{200, 400, 600, 800 \}$ for the simulated data types, and 50 causal SNPs were used throughout.
 
@@
 Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture.
-Supplementary figures \anna{?, ?, and ?} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.
+Supplementary Figures \anna{?, ?, and ?} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.
 
@@
 However, behavior-based environmental exposures may be less likely to be shared by subpopulations which are genetically and geographically similar.
-As an example, \citep{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
+As an example, \citet{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
 
@@
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
-The top row of figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably.
+The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably.
 PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome.
-The bottom row of figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
+The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
 In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
@@
 In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
-This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
+This is consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
@@
 Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the independent subpopulation simulated $\bX$.
-The right-hand panels of figure \ref{fig:big_small_gamma} are consistent with figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
-The upper left hand panel of figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso.
-The lower left hand panel of figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients.
+The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
+The upper left hand panel of Figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso.
+The lower left hand panel of Figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients.
 In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while PC-Lasso outperforms LMM-Lasso in the majority of dichotomous-concordant environmental effects.
@@
 However, these apparent improvements in estimating $\eta$ do not improve the $\bbeta$ estimation accuracy of LMM-Lasso-ggmix.
-This is evidenced by the top panel of figure \ref{fig:eta_beta_mse}, which shows nearly identical performances of both methods in terms of $\bbeta$ MSE.
+This is evidenced by the top panel of Figure \ref{fig:eta_beta_mse}, which shows nearly identical performances of both methods in terms of $\bbeta$ MSE.
 
@@
 \end{figure}
-\anna{Note figure \ref{fig:time} is on raw scale, not log10. This is because the negative values on the error bars create problems and become NaN on the log10 scale}
+\anna{Note Figure \ref{fig:time} is on raw scale, not log10. This is because the negative values on the error bars create problems and become NaN on the log10 scale}
 
@@
 Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners.
-However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications.
+However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in Figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications.
 \anna{time interpretation depends on reason ggmix is slow...} Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.

### 2020-08-15 main.tex
Sections 2.2 and 2.3 edits

@@
 
-An additional consideration in genetic studies that we will briefly mention is that of genetic background.
-\anna{make the following sentence better: correlated snps are the issue - talk about ld pruning} Genetic background describes when SNPs that have a causal relationship with the phenotype of interest are not directly observed, but are correlated with SNPs that are observed.
-As \citet{vilhjalmsson2012nature} \pb{Use citet, not citep here} \anna{double check all citations for whether they should be (1) citet, citep or (2) better at the end of the sentence} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
+An additional source of confounding in genetic studies that we will briefly mention is that of confounding due to genetic background.
+This occurs in univariate association methods when a non-causal SNP is tested for association with a trait of interest, and that SNP is correlated with a causal SNP.
+As \citet{vilhjalmsson2012nature} note, any trait that has some genetic basis will be subject to genetic background effects, just as traits that are influenced by the environment may be subject to environmental confounding.
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
-Although there are a number of methods in which to account for genetic background \anna{what are they?
-\textbf{LD pruning} or block-wise analysis, gene-level test, \textbf{haplotype block}...group correlated variants}, the methods upon which we will focus address this by conducting a genomewide multivariate analysis.
+LD pruning is commonly employed to reduce confounding due to genetic background in univariate association testing.
+However, we will focus on methods which address this via genomewide multivariate analysis.
 
@@
 
-\anna{switch first and second paragraphs... the current first paragraph doesn't describe GC}
+One of the first methods to address the problem of confounding due to population structure is the genomic control (GC) method.
+GC is a univariate testing procedure which modifies a test statistic by a deflation factor.
+This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
+However, this deflation factor has been reported to be sensitive to the nature and quantity of the loci used to estimate it \citep{hellwege2017population, marchini2004effects}.
+Additionally, GC does not correct effect size estimates, so resulting coefficient estimates will be unreliable after GC is applied, even if their corresponding test statistics and p-values have been appropriately adjusted \citep{hellwege2017population}.
 
-Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts.
-In order to reduce confounding these methods attempt to block the causal pathway between population structure and observed SNP data.
+New statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts.
+Many of these methods attempt to block the causal pathway between population structure and observed SNP data.
 This is illustrated by the blocked arrow in Figure \ref{fig:ps_env_block}.
@@
 Nevertheless, they may be effective if they are able to indirectly block the causal pathway between the environment and observed SNPs.
-Whether or not they are effective depends upon whether population structure can be accurately inferred, and if this is a good proxy for environmental effects.
-\anna{maybe say something about how 'if it's a good proxy - great!
-they'll work', but 'if it's not a good proxy, also fine!
-because then it's not a confounder!
-...SE's potentially a problem like in GLS...
-but this is different with lasso.
-(maybe not worth getting into here)}
 
-\anna{make a better slashed arrow}
+\anna{make better slashed arrows / DAGs}
 
@@
 
-One of the first methods to address thie problem of confounding due to population structure is the genomic control (GC) method.
-GC is a univariate testing procedure which modifies a test statistic by a deflation factor.
-This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
-However, this deflation factor has been reported to be sensitive to the nature and quantity of the loci used to estimate it \citep{hellwege2017population, marchini2004effects}.
-Additionally, GC does not correct effect size estimates, so resulting coefficient estimates will be unreliable after GC is applied, even if their corresponding test statistics and p-values have been appropriately adjusted \citep{hellwege2017population}.
+Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred, and if this is a good proxy for environmental effects.
+In the presence of unobserved heterogeneous environmental effects which are not well-approximated by inferred population structure, these effects will not confound SNP-phenotype associations.
+This is due to the absence of a causal pathway between population stratification and environment, as depicted in Figure \ref{fig:ps_env_block2}.
+In such a scenario, failing to account for unobserved environmental effects will result in greater phenotypic variability \anna{and as a result, greater SE in beta-hats...
+if we're talking about gls}, however, it will not bias SNP estimates \anna{cite something from causal inference here}.
+We will return to this idea and its mathematical justification in section \ref{sec:methods}.
 
-Stratified analysis attempts to infer subject membership within a discrete number of nonoverlapping subpopulations.
+\begin{figure}[H]
+\centering
+\begin{tikzpicture}[%
+>=latex',
+circ/.style={draw, shape=circle, node distance=1cm, line width=1pt}]%Define the arrow type and style for circled nodes
+\draw[dashed, red] (0,1) node[left] (P) {Population Stratification} --node {X} (2,1) node[right] (E) {Environment};
+\draw[->] (0,0) node[left] (X) {Observed SNPs} -- (2,0) node[right] (Y) {Phenotype}; %Line between X and Y
+\draw[->] (P) -- (X) ; 
+\draw[->] (E)   -- (Y) ; 
+% \filldraw[color=red!60, fill=none, very thick] (0,1) ellipse (4.5 and 0.35);
+\end{tikzpicture}
+\caption{The causal pathway between the environment and observed SNPs may be blocked if genetic information does not serve as an accurate proxy for it.}
+\label{fig:ps_env_block2}
+\end{figure}
+
+\anna{need transition to bring it back to the first DAG for now...} Stratified analysis attempts to infer subject membership within a discrete number of nonoverlapping subpopulations.
 Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.

### 2020-08-17 lmm-genetic-review.tex
Section 2.3: better DAGs

@@
 \usepackage{float}
+\usepackage{color}
 \usepackage{tikz}
+\usepackage{tkz-euclide}
+
+% Tikz settings optimized for causal graphs.
+% Just copy-paste this part
+\usetikzlibrary{shapes,decorations,decorations.markings,arrows,calc,arrows.meta,fit,positioning}
+\tikzset{
+    -Latex,auto,node distance =1 cm and 3 cm,semithick,
+    state/.style ={ellipse, draw, minimum width = 0.7 cm},
+    point/.style = {circle, draw, inner sep=0.04cm,fill,node contents={}},
+    bidirected/.style={Latex-Latex,dashed},
+    el/.style = {inner sep=2pt, align=left, sloped},
+    strike through/.append style={
+    decoration={markings, mark=at position 0.5 with {
+    \draw[-] ++ (-4pt,-4pt) -- (4pt,4pt);}
+  },postaction={decorate}}
+}
+
 \usepackage{subcaption}

### 2020-08-17 main.tex
Section 2.3: better DAGs

@@
 \centering
-\begin{tikzpicture}[%
->=latex',
-circ/.style={draw, shape=circle, node distance=1cm, line width=1pt}]%Define the arrow type and style for circled nodes
-\draw[dashed] (0,1) node[left] (P) {Population Stratification} -- (2,1) node[right] (E) {Environment}; 
-\draw[->] (0,0) node[left] (X) {Observed SNPs} -- (2,0) node[right] (Y) {Phenotype}; %Line between X and Y
-\draw[->] (P)   -- (X) ; 
-\draw[->] (E)   -- (Y) ; 
-% \filldraw[color=red!60, fill=none, very thick] (0,1) ellipse (4.5 and 0.35);
+\begin{tikzpicture}
+    \node (1) at (0,0) {Population Stratification};
+    \node (2) [right = of 1] {Environment};
+    \node (3) [below = of 1] {Observbed SNPs};
+    \node (4) [below = of 2] {Phenotype};
+    \path[bidirected, -] (1) edge (2);
+    \path (1) edge (3);
+    \path (2) edge (4);
+    \path (3) edge (4);
 \end{tikzpicture}
@@
 
-\anna{make better slashed arrows / DAGs}
-
 \begin{figure}[H]
 \centering
-\begin{tikzpicture}[%
->=latex',
-circ/.style={draw, shape=circle, node distance=1cm, line width=1pt}]%Define the arrow type and style for circled nodes
-\draw[dashed] (0,1) node[left] (P) {Population Stratification} -- (2,1) node[right] (E) {Environment};
-\draw[->] (0,0) node[left] (X) {Observed SNPs} -- (2,0) node[right] (Y) {Phenotype}; %Line between X and Y
-\draw[->, red] (P) --node {|} (X) ; 
-\draw[->] (E)   -- (Y) ; 
-% \filldraw[color=red!60, fill=none, very thick] (0,1) ellipse (4.5 and 0.35);
+\begin{tikzpicture}
+    \node (1) at (0,0) {Population Stratification};
+    \node (2) [right = of 1] {Environment};
+    \node (3) [below = of 1] {Observbed SNPs};
+    \node (4) [below = of 2] {Phenotype};
+    \path[bidirected, -] (1) edge (2);
+    \path[red, line width = 0.5] (1) edge (3);
+    \path (2) edge (4);
+    \path (3) edge (4);
+    \tkzMarkSegment[color=red,pos=.5,mark=x](1,3)
 \end{tikzpicture}
@@
 \centering
-\begin{tikzpicture}[%
->=latex',
-circ/.style={draw, shape=circle, node distance=1cm, line width=1pt}]%Define the arrow type and style for circled nodes
-\draw[dashed, red] (0,1) node[left] (P) {Population Stratification} --node {X} (2,1) node[right] (E) {Environment};
-\draw[->] (0,0) node[left] (X) {Observed SNPs} -- (2,0) node[right] (Y) {Phenotype}; %Line between X and Y
-\draw[->] (P) -- (X) ; 
-\draw[->] (E)   -- (Y) ; 
-% \filldraw[color=red!60, fill=none, very thick] (0,1) ellipse (4.5 and 0.35);
+\begin{tikzpicture}
+    \node (1) at (0,0) {Population Stratification};
+    \node (2) [right = of 1] {Environment};
+    \node (3) [below = of 1] {Observbed SNPs};
+    \node (4) [below = of 2] {Phenotype};
+    \path[bidirected, -, red] (1) edge (2);
+    \path (1) edge (3);
+    \path (2) edge (4);
+    \path (3) edge (4);
+    \tkzMarkSegment[color=red,pos=.5,mark=x](1,2)
 \end{tikzpicture}

### 2020-08-18 main.tex
Title; making DAG accents pink because girl power
Section 2 edits

@@
 
-One reason that population stratification warrants such concern in genetic studies is that due to ethnic and geographic segregation, population stratification tends to be associated with differing environmental exposures and cultural practices \citep{thornton2015statistical}.
+One reason that population stratification warrants such concern in genetic studies is that due to ethnic and geographic segregation, population stratification tends to be associated with differing environmental exposures and cultural practices \citep{thornton2015statistical, browning2011population}.
 In other words, random genetic variation is likely to be correlated with non-genetic factors that also influence the trait of interest, thereby confounding the genotype-phenotype relationship and leading to spurious associations and biased estimates of SNP effects.
@@
 
+Population structure is defined by the existence of allele frequency differences that define sub-populations and is driven by the combined effects of evolutionary processes such as genetic drift, migration, demographic history, and natural selection \citep{gibson2015primer, tibayrenc2017genetics}.
 Population structure is broadly categorized based on whether it describes recent or ancient relatedness.
-Recent relatedness describes familial or \textbf{cryptic relatedness}.
-Ancient relatedness, or common ancestry, is generally characterized by large clusters of low levels of similarity.
-\anna{I need some sort of subjective cutoff for recent vs.
-ancient relatedness rather than describing them in terms of clusters...
-more details} \pb{I would suggest something along the lines of: cryptic relatedness means that only a few generations, maybe 3 or so, separate common ancestors} The presence of multiple distinct ancestry groups in a sample is known as \textbf{population stratification}.
+Ancient relatedness describes the presence of a common ancestor many generations previously.
+The presence of distinct ancestry groups with different allele frequencies in a sample is known as \textbf{population stratification}.
+Recent relatedness describes the sharing of a common ancestor only several generations previously.
+Pedigree-based methods may be used to explicitly model recent relatedness if familial relationships are known.
+In this absence of known familial relationships, recent relatedness is referred to as \textbf{cryptic relatedness.} In this review we focus on methods to account for unobserved relatedness such as population stratification and cryptic relatedness.
 
-Pedigree-based methods may be used to explicitly model recent relatedness if familial relationships are known.
-However, in this review we focus on methods to account for unobserved relatedness.
 Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are present.
-Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations.
-This phenomenon is commonly described as \textbf{confounding due to population stratification or ancestry}.
-The mechanism of this phenomenon warrants some discussion as it is often overlooked that population structure itself does not confound the genotype-phenotype relationship in the classical sense.
-While population structure may directly affect observed allele frequencies, there must also exist some mechanism by which population stratification affects the phenotype, namely, the environment \citep{barton2019population, vilhjalmsson2012nature}.
-
-% \anna{DAG here with only population stratification, observed snps, and phenotype ... maybe an arrow from population stratification to phenotype with a question mark (?)}
+Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations when population structure is associated with differences in both allele frequency and the trait or disease \citep{gibson2015primer}.
+This phenomenon is commonly described as \textbf{confounding due to population stratification}.
+The mechanism of this phenomenon warrants some discussion as it is often overlooked that population structure in and of itself does not confound the genotype-phenotype relationship.
+While population structure may directly affect observed allele frequencies, there must also exist some non-genetic mechanism by which population stratification affects the phenotype, namely, the environment \citep{barton2019population, vilhjalmsson2012nature}.
 
@@
 This is illustrated by the blocked arrow in Figure \ref{fig:ps_env_block}.
-These methods do not directly address the environmental component of confounding due to population structure.
-Nevertheless, they may be effective if they are able to indirectly block the causal pathway between the environment and observed SNPs.
+As the figure shows, these methods do not directly address the environmental component of confounding due to population structure.
+Nevertheless, they may be effective in controlling confounding by indirectly blocking the causal pathway between the environment and observed SNPs.
 
@@
     \path[bidirected, -] (1) edge (2);
-    \path[red, line width = 0.5] (1) edge (3);
+    \path[magenta, line width = 0.5] (1) edge (3);
     \path (2) edge (4);
     \path (3) edge (4);
-    \tkzMarkSegment[color=red,pos=.5,mark=x](1,3)
+    \tkzMarkSegment[color=magenta,pos=.5,mark=x](1,3)
 \end{tikzpicture}
@@
 
-Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred, and if this is a good proxy for environmental effects.
-In the presence of unobserved heterogeneous environmental effects which are not well-approximated by inferred population structure, these effects will not confound SNP-phenotype associations.
+Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred, and if population structure is a good proxy for environmental effects.
+In the presence of environmental heterogeneity which affects the phenotype of interest, and which is not well-approximated by inferred population structure, SNP-phenotype associations will not be confounded.
 This is due to the absence of a causal pathway between population stratification and environment, as depicted in Figure \ref{fig:ps_env_block2}.
-In such a scenario, failing to account for unobserved environmental effects will result in greater phenotypic variability \anna{and as a result, greater SE in beta-hats...
-if we're talking about gls}, however, it will not bias SNP estimates \anna{cite something from causal inference here}.
+In such a scenario, failing to account for unobserved environmental effects is likely to result in greater residual variability, however, SNP effect estimates will be unbiased \citep{greenland1999causal}.
 We will return to this idea and its mathematical justification in section \ref{sec:methods}.
@@
     \node (4) [below = of 2] {Phenotype};
-    \path[bidirected, -, red] (1) edge (2);
+    \path[bidirected, -, magenta] (1) edge (2);
     \path (1) edge (3);
@@
     \path (3) edge (4);
-    \tkzMarkSegment[color=red,pos=.5,mark=x](1,2)
+    \tkzMarkSegment[color=magenta,pos=.5,mark=x](1,2)
 \end{tikzpicture}
@@
 
-\anna{need transition to bring it back to the first DAG for now...} Stratified analysis attempts to infer subject membership within a discrete number of nonoverlapping subpopulations.
+Assuming for now that population structure is a good proxy for environmental effects, we return to our consideration of methods that attempt to control for confounding via the mechanism depicted in Figure \ref{fig:ps_env_block}.
+Such methods include LMMs, principal-component-based approaches, and stratified analyses.
+Stratified analyses refer to a general class of methods which attempt to infer subject membership within a discrete number of nonoverlapping subpopulations.
 Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.

### 2020-08-19 main.tex
Sections 3.1 - 3.4

@@
                    &= \frac{1}{p} \bU \bD (\bU \bD)^\T \nonumber \\
-                   &\propto \bR \bR^\T,
+                   &= \frac{1}{p} \bR \bR^\T,
 \end{align}
-where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are therefore proportional to the principal components of $\bK$.
+where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors or principal components of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are then the principal components of $\bK$ weighted by their corresponding singular values.
 
@@
 \subsection{PC-Lasso}
-PC-Lasso uses the first $k$ principal components of $\bK$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates.
-It has been shown \citep{hoffman2013correcting} that this corresponds to the regression model
+PC-Lasso uses the first $k$ principal components of $\bK$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates. This corresponds to the regression model
 \begin{equation}
@@
 \end{equation}
+$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI), $$
+
 where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \eqref{eqn:model}.
@@
 
-Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect assuming $\bgamma \sim N\left(\boldsymbol{0}, p^{-1} \sigmagg \bI_p \right).$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
+Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect.
+Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
 This results in the LMM regression model
@@
 \end{equation}
-where we assume $\beps \independent \bX\bgamma$.
-Rather than assuming anything about $\bgamma$ directly, an equivalent representation paramaterizes the model in terms of $\bu=\bX\bgamma$ where $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent model
+$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI) $$
+$$ \bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bX \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK), $$
+where $\sigmagg$ is the additive genetic variance, and we assume $\beps \independent \bgamma$.
+Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields the equivalent model
 \begin{equation}
     \label{eqn:lmm_reg_u}
-    \by = \bX_0\bbeta_0 + \bX\bbeta + \bu + \beps.
+    \by = \bX_0\bbeta_0 + \bX\bbeta + \bu + \beps
 \end{equation}
+$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI) $$
+$$ \bu \sim N(\boldsymbol{0}, \sigmagg \bK).$$
 We will first derive the LMM-Lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
@@
 Following equations \eqref{eqn:lmm_reg} and \eqref{eqn:lmm_reg_u}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
-LMM-Lasso uses \anna{an estimate of} the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
+LMM-Lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
 Specifically, left-multiplying the original data by $\bV^{-1/2}$ yields
@@
 It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
-\citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$ \anna{which we briefly review here...}
+\citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$, which we briefly review.
 
-Thus, whether modeled as a fixed or random effect, the underlying effect of $\bZ \bgamma$ on phenotype remains the same, but its treatment as a fixed or random effect has important implications in terms of model fitting and estimation.
+Using the result of the eigendecomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg_u} can be written equivalently as
+
+\begin{equation}
+    \label{eqn:pc_lmm_equiv}
+    \by = \bX_0 \bbeta_0 + \bX \bbeta + \bR \bgamma + \beps
+\end{equation}
+
+$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI) $$
+$$ \bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bR \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK).$$
+
+Based on the relationship between equations \eqref{eqn:pca_reg} and \eqref{eqn:pc_lmm_equiv}, we can see that modeling the principal components of $\bK$ as fixed or random effects relies on the same underlying regression model.
+However, while the LMM includes all the principal components, only $k < n$ are included in the fixed effects model.
+
+Additionally, while the fixed effects model considers the principal components, $\bU$, of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
+The eigenvalues have been shown to serve as a measure of biological relevance of each principal component in relation to underlying population structure \citep{hoffman2013correcting, price2006principal}.
+
+From a Bayesian perspective, using the first $k$ principal components as regression covariates can be viewed as assigning uniform priors to the first $k$ components of $\bgamma$, while the remaining $n - k$ components receive a prior with unit mass at zero.
+These assignments imply certainty that the effect of population structure on the phenotype is fully captured by the first $k$ PCs.
+In contrast, the LMM approach can be viewed as assigning a Gaussian prior on each component of $\bgamma$, with variances proportional to the corresponding eigenvalues \citep{astle2009population}.
+
+Whether modeled as a fixed or random effect, the underlying effect of $\bZ \bgamma$ on phenotype remains the same, but its treatment as a fixed or random effect has important implications in terms of model fitting and estimation.
 Modeling unobserved confounding using fixed effects requires estimation of $p_0 + p + k$ parameters, compared to the $p_0 + p + 2$ required of an LMM, which may result in a loss of power \citep{zhang2015principal}.
-Additionally, when modeling $\bZ \bgamma$ using fixed effects, determining the appropriate number of principal components to adjust for is non-trivial, and has been the subject of numerous studies \citep{patterson2006population, zhao2018practical}.
+Additionally, when modeling $\bZ \bgamma$ using fixed effects, determining the appropriate number of principal components, $k$, to adjust for is non-trivial, and has been the subject of numerous studies \citep{patterson2006population, zhao2018practical}.
 It has been shown that including too many or too few principal components can result in power loss or increased Type I errors, respectively \citep{zhang2015principal}.

### 2020-08-19 main.tex
3.5 edits

@@
 \end{align}
-Under the assumption that environmental effects arise from a random process, $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg I_p)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
+Under the assumption that environmental effects arise from a random process, $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
 Consider our earlier example involving confounding due to air quality.
@@
 Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
-Then, $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ such that
+Then, $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ such that
 \begin{align*}
@@
 \end{align*}
-where $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat{^\prime}$ and $\boldsymbol{\psi}$ is absorbed by the error term.
+where $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term.
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
 
-\anna{bring this back to why it may not matter if genetic similarity isn't a good proxy for population structure - in that case it's not a confounder anyway}
+This result also allows us to mathematically justify the statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
+The proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ is equal to $||\bX \boldsymbol{\tau}||_F / ||\bZ \bgamma||_F \in [0, 1]$.
+Since $\bX$ is fixed, this quantity is reflected in $||\boldsymbol{\tau}||_F$, which also quantifies the estimation bias due environmental heterogeneity.
+In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
+Thus the SNP-phenotype associations will be unbiased, but the residual variance will be increased.
+Although we focus our attention on accurately estimating fixed SNP effects, we note that this result could lead to substantial bias in narrow-sense heritability estimates.
+
+In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
+However, the degree to which it reflects population structure that can be inferred from $\bX$ will impact the extent to which environmental heterogeneity confounds SNP-phenotype associations.
 
@@
 So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to environmental effects.
-\anna{more details about how/why this lambda selection method - cite CV as bad for LMMs, BIC not great bc different likelihoods...
-tie this into evaluating performance of each method in terms of variable selection} The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
-1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
-
-%This method for selecting $\lambda$ was chosen to fairly compare the
-
-%Cross-validation has generally been found to be inadvisable for penalized LMM methods \anna{cite thiiiiis}
-%Using an information criterion such as BIC to select $\lambda$ is not comparable across methods since Lasso, PC-Lasso, LMM-Lasso-Rakitsch, and LMM-Lasso-ggmix methods each use a different likelihood.
 
@@
 
+\subsubsection{$\lambda$ selection}
+\anna{more details about how/why this lambda selection method - cite CV as bad for LMMs, BIC not great bc different likelihoods...
+tie this into evaluating performance of each method in terms of variable selection} The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
+1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
+
+%This method for selecting $\lambda$ was chosen to fairly compare the
+
+%Cross-validation has generally been found to be inadvisable for penalized LMM methods \anna{cite thiiiiis}
+%Using an information criterion such as BIC to select $\lambda$ is not comparable across methods since Lasso, PC-Lasso, LMM-Lasso-Rakitsch, and LMM-Lasso-ggmix methods each use a different likelihood.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
+In other words, a greater amount of the environmental effect is in the column space of the projection matrix of $\bX$ for the independent subpopulations data than in the admixed data.
+As such, there is a greater amount of confounding in the independent subpopulations which can be corrected by the PC-Lasso and LMM-Lasso methods.
 
-\anna{in admixed data less of the environmental effect is in the column space of X so it's less of a confounder}
+\anna{add variable selection tables for these settings}
 

### 2020-08-20 main.tex
3.6.5 - lambda selection method

@@
 We define concordant effects as those that are similar among genetically similar subpopulations, and discordant effects as those that are not.
-Figure \ref{fig:gamma_structures} displays a visual summary of these four coefficient structures.
+Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient structures.
+1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
 \begin{figure}[H]
@@
 \subsubsection{$\lambda$ selection}
-\anna{more details about how/why this lambda selection method - cite CV as bad for LMMs, BIC not great bc different likelihoods...
-tie this into evaluating performance of each method in terms of variable selection} The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method.
-1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
-
-%This method for selecting $\lambda$ was chosen to fairly compare the
-
-%Cross-validation has generally been found to be inadvisable for penalized LMM methods \anna{cite thiiiiis}
-%Using an information criterion such as BIC to select $\lambda$ is not comparable across methods since Lasso, PC-Lasso, LMM-Lasso-Rakitsch, and LMM-Lasso-ggmix methods each use a different likelihood.
+The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method, where 50 was the true number of causal SNPs.
+This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.
+Alternative $\lambda$ selection methods were also considered.
+Cross-validation has been found to be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here.
+While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
+In addition, because Lasso, PC-Lasso, and LMM-Lasso use different likelihoods an information-criterion-based approach would not lend itself to comparable results across methods.
 
@@
 
-\anna{add variable selection tables for these settings}
-
 \begin{figure}[H]
@@
 \end{figure}
+
+\anna{add variable selection tables for these settings}
 

### 2020-08-20 main.tex
4.1 - ugly placeholder true/false positive tables

@@
 
-\anna{add variable selection tables for these settings}
+\anna{Obviously these need to be formatted...}\\
+
+\anna{True positive rates}
+\begin{verbatim}
+structureX    xi    method      `round(mean(value), 2)` `round(sd(value), 2)`
+ 1 1d_linear     0     lasso                         0.25                   0.05
+ 2 1d_linear     0     lasso_pca10                   0.23                   0.05
+ 3 1d_linear     0     lmm_ggmix                     0.25                   0.05
+ 4 1d_linear     0     lmm_lasso                     0.25                   0.05
+ 5 1d_linear     0     plmm_lasso                    0.25                   0.05
+ 6 1d_linear     0.8   lasso                         0.23                   0.05
+ 7 1d_linear     0.8   lasso_pca10                   0.28                   0.05
+ 8 1d_linear     0.8   lmm_ggmix                     0.28                   0.05
+ 9 1d_linear     0.8   lmm_lasso                     0.28                   0.05
+10 1d_linear     0.8   plmm_lasso                    0.290                  0.05
+11 indep_subpops 0     lasso                         0.23                   0.05
+12 indep_subpops 0     lasso_pca10                   0.21                   0.05
+13 indep_subpops 0     lmm_ggmix                     0.22                   0.05
+14 indep_subpops 0     lmm_lasso                     0.23                   0.05
+15 indep_subpops 0     plmm_lasso                    0.23                   0.05
+16 indep_subpops 0.8   lasso                         0.25                   0.05
+17 indep_subpops 0.8   lasso_pca10                   0.36                   0.06
+18 indep_subpops 0.8   lmm_ggmix                     0.38                   0.06
+19 indep_subpops 0.8   lmm_lasso                     0.38                   0.06
+20 indep_subpops 0.8   plmm_lasso                    0.39                   0.06
+21 other         0     lasso                         0.37                   0.06
+22 other         0     lasso_pca10                   0.36                   0.06
+23 other         0     lmm_ggmix                     0.4                    0.06
+24 other         0     lmm_lasso                     0.4                    0.06
+25 other         0     plmm_lasso                    0.41                   0.06
+26 other         0.8   lasso                         0.290                  0.06
+27 other         0.8   lasso_pca10                   0.47                   0.06
+28 other         0.8   lmm_ggmix                     0.61                   0.06
+29 other         0.8   lmm_lasso                     0.580                  0.06
+30 other         0.8   plmm_lasso                    0.580                  0.06
+\end{verbatim}
+
+\anna{False positive rates}
+\begin{verbatim}
+   structureX    xi    method      `round(mean(value), 3)` `sd(value)`
+ 1 1d_linear     0     lasso                         0.038     0.00309
+ 2 1d_linear     0     lasso_pca10                   0.038     0.00317
+ 3 1d_linear     0     lmm_ggmix                     0.037     0.00343
+ 4 1d_linear     0     lmm_lasso                     0.037     0.00320
+ 5 1d_linear     0     plmm_lasso                    0.038     0.00311
+ 6 1d_linear     0.8   lasso                         0.039     0.00295
+ 7 1d_linear     0.8   lasso_pca10                   0.036     0.00307
+ 8 1d_linear     0.8   lmm_ggmix                     0.035     0.00328
+ 9 1d_linear     0.8   lmm_lasso                     0.035     0.00327
+10 1d_linear     0.8   plmm_lasso                    0.036     0.00298
+11 indep_subpops 0     lasso                         0.039     0.00301
+12 indep_subpops 0     lasso_pca10                   0.039     0.00310
+13 indep_subpops 0     lmm_ggmix                     0.038     0.00367
+14 indep_subpops 0     lmm_lasso                     0.038     0.00315
+15 indep_subpops 0     plmm_lasso                    0.039     0.00301
+16 indep_subpops 0.8   lasso                         0.038     0.00307
+17 indep_subpops 0.8   lasso_pca10                   0.032     0.00321
+18 indep_subpops 0.8   lmm_ggmix                     0.03      0.00338
+19 indep_subpops 0.8   lmm_lasso                     0.03      0.00338
+20 indep_subpops 0.8   plmm_lasso                    0.031     0.00318
+21 other         0     lasso                         0.031     0.00326
+22 other         0     lasso_pca10                   0.031     0.00335
+23 other         0     lmm_ggmix                     0.025     0.00451
+24 other         0     lmm_lasso                     0.028     0.00358
+25 other         0     plmm_lasso                    0.029     0.00335
+26 other         0.8   lasso                         0.036     0.00300
+27 other         0.8   lasso_pca10                   0.026     0.00338
+28 other         0.8   lmm_ggmix                     0.016     0.00348
+29 other         0.8   lmm_lasso                     0.02      0.00298
+30 other         0.8   plmm_lasso                    0.021     0.00291
+\end{verbatim}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\anna{this to change somewhat based on time/convergence results...}
+
 In this work, we have reviewed the concept of population stratification and explicitly defined the environmental mechanism by which it may confound the genotype-phenotype relationship.

### 2020-08-20 main.tex
Small punctuation tweaks

@@
 \end{align}
-where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors or principal components of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are then the principal components of $\bK$ weighted by their corresponding singular values.
+where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors or principal components of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are thus the principal components of $\bK$ weighted by their corresponding singular values.
 
@@
 $$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI), $$
-
 where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \eqref{eqn:model}.
@@
 We will first derive the LMM-Lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
-
 Following equations \eqref{eqn:lmm_reg} and \eqref{eqn:lmm_reg_u}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
@@
 \label{eqn:rtY_distn}
-\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX \bbeta, \bI)
+\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX \bbeta, \bI),
 \end{equation}
-and a regular \anna{vanilla, naive, regular?} Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
+and thus a regular \anna{vanilla, naive, regular?} Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
 \anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
@@
 \begin{equation}
-\bV^{-\frac{1}{2}}\by \sim N\left(\bV^{-\frac{1}{2}} \bX{^\prime} \bbeta{^\prime}, \bI \right)
+\bV^{-\frac{1}{2}}\by \sim N\left(\bV^{-\frac{1}{2}} \bX{^\prime} \bbeta{^\prime}, \bI \right),
 \end{equation}
@@
 $$ \bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bR \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK).$$
-
 Based on the relationship between equations \eqref{eqn:pca_reg} and \eqref{eqn:pc_lmm_equiv}, we can see that modeling the principal components of $\bK$ as fixed or random effects relies on the same underlying regression model.
@@
     &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps\\
-    &=  \bX \bbeta{^*} + \beps{^*}
+    &=  \bX \bbeta{^*} + \beps{^*},
 \end{align*}

### 2020-08-20 main.tex
2020-08-20 meeting notes/edits

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\anna{REPLACE FIGURES WITH RESULTS FROM PLMM-LASSO ONCE FINISHED}
 
@@
 This result also allows us to mathematically justify the statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
-The proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ is equal to $||\bX \boldsymbol{\tau}||_F / ||\bZ \bgamma||_F \in [0, 1]$.
-Since $\bX$ is fixed, this quantity is reflected in $||\boldsymbol{\tau}||_F$, which also quantifies the estimation bias due environmental heterogeneity.
-In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
+We can quantify the proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ using the ratio $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2 \in [0, 1]$.
+In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
 Thus the SNP-phenotype associations will be unbiased, but the residual variance will be increased.
 Although we focus our attention on accurately estimating fixed SNP effects, we note that this result could lead to substantial bias in narrow-sense heritability estimates.
-
 In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
-However, the degree to which it reflects population structure that can be inferred from $\bX$ will impact the extent to which environmental heterogeneity confounds SNP-phenotype associations.
+However, the degree to which it reflects population structure that can be inferred from $\bX$, as quantified by $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$, will impact the extent to which environmental heterogeneity confounds SNP-phenotype associations.
 
@@
 This improvement is more substantial in the independent subpopulations simulated data than the admixed simulated data.
-This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
+\anna{quantify this for all 3 data settings using $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$} This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
 In other words, a greater amount of the environmental effect is in the column space of the projection matrix of $\bX$ for the independent subpopulations data than in the admixed data.
@@
 
-\anna{Obviously these need to be formatted...}\\
+\anna{Obviously these need to be formatted... just use TP rates - FP not adding much}\\
 
@@
 30 other         0.8   plmm_lasso                    0.580                  0.06
-\end{verbatim}
-
-\anna{False positive rates}
-\begin{verbatim}
-   structureX    xi    method      `round(mean(value), 3)` `sd(value)`
- 1 1d_linear     0     lasso                         0.038     0.00309
- 2 1d_linear     0     lasso_pca10                   0.038     0.00317
- 3 1d_linear     0     lmm_ggmix                     0.037     0.00343
- 4 1d_linear     0     lmm_lasso                     0.037     0.00320
- 5 1d_linear     0     plmm_lasso                    0.038     0.00311
- 6 1d_linear     0.8   lasso                         0.039     0.00295
- 7 1d_linear     0.8   lasso_pca10                   0.036     0.00307
- 8 1d_linear     0.8   lmm_ggmix                     0.035     0.00328
- 9 1d_linear     0.8   lmm_lasso                     0.035     0.00327
-10 1d_linear     0.8   plmm_lasso                    0.036     0.00298
-11 indep_subpops 0     lasso                         0.039     0.00301
-12 indep_subpops 0     lasso_pca10                   0.039     0.00310
-13 indep_subpops 0     lmm_ggmix                     0.038     0.00367
-14 indep_subpops 0     lmm_lasso                     0.038     0.00315
-15 indep_subpops 0     plmm_lasso                    0.039     0.00301
-16 indep_subpops 0.8   lasso                         0.038     0.00307
-17 indep_subpops 0.8   lasso_pca10                   0.032     0.00321
-18 indep_subpops 0.8   lmm_ggmix                     0.03      0.00338
-19 indep_subpops 0.8   lmm_lasso                     0.03      0.00338
-20 indep_subpops 0.8   plmm_lasso                    0.031     0.00318
-21 other         0     lasso                         0.031     0.00326
-22 other         0     lasso_pca10                   0.031     0.00335
-23 other         0     lmm_ggmix                     0.025     0.00451
-24 other         0     lmm_lasso                     0.028     0.00358
-25 other         0     plmm_lasso                    0.029     0.00335
-26 other         0.8   lasso                         0.036     0.00300
-27 other         0.8   lasso_pca10                   0.026     0.00338
-28 other         0.8   lmm_ggmix                     0.016     0.00348
-29 other         0.8   lmm_lasso                     0.02      0.00298
-30 other         0.8   plmm_lasso                    0.021     0.00291
 \end{verbatim}
@@
 
+\anna{add some info/figure on convergence and number of iterations for lmm-lasso vs ggmix, delete the following time figures and references to them}
+
 \begin{figure}[H]
@@
 \end{figure}
-\anna{add time to this figure? third row...}
+\anna{replace this figure with one that includes time in the third row}
 

### 2020-08-21 main.tex
4.1 - formatted variable selection table

@@
 
-\anna{Obviously these need to be formatted... just use TP rates - FP not adding much}\\
-
-\anna{True positive rates}
-\begin{verbatim}
-structureX    xi    method      `round(mean(value), 2)` `round(sd(value), 2)`
- 1 1d_linear     0     lasso                         0.25                   0.05
- 2 1d_linear     0     lasso_pca10                   0.23                   0.05
- 3 1d_linear     0     lmm_ggmix                     0.25                   0.05
- 4 1d_linear     0     lmm_lasso                     0.25                   0.05
- 5 1d_linear     0     plmm_lasso                    0.25                   0.05
- 6 1d_linear     0.8   lasso                         0.23                   0.05
- 7 1d_linear     0.8   lasso_pca10                   0.28                   0.05
- 8 1d_linear     0.8   lmm_ggmix                     0.28                   0.05
- 9 1d_linear     0.8   lmm_lasso                     0.28                   0.05
-10 1d_linear     0.8   plmm_lasso                    0.290                  0.05
-11 indep_subpops 0     lasso                         0.23                   0.05
-12 indep_subpops 0     lasso_pca10                   0.21                   0.05
-13 indep_subpops 0     lmm_ggmix                     0.22                   0.05
-14 indep_subpops 0     lmm_lasso                     0.23                   0.05
-15 indep_subpops 0     plmm_lasso                    0.23                   0.05
-16 indep_subpops 0.8   lasso                         0.25                   0.05
-17 indep_subpops 0.8   lasso_pca10                   0.36                   0.06
-18 indep_subpops 0.8   lmm_ggmix                     0.38                   0.06
-19 indep_subpops 0.8   lmm_lasso                     0.38                   0.06
-20 indep_subpops 0.8   plmm_lasso                    0.39                   0.06
-21 other         0     lasso                         0.37                   0.06
-22 other         0     lasso_pca10                   0.36                   0.06
-23 other         0     lmm_ggmix                     0.4                    0.06
-24 other         0     lmm_lasso                     0.4                    0.06
-25 other         0     plmm_lasso                    0.41                   0.06
-26 other         0.8   lasso                         0.290                  0.06
-27 other         0.8   lasso_pca10                   0.47                   0.06
-28 other         0.8   lmm_ggmix                     0.61                   0.06
-29 other         0.8   lmm_lasso                     0.580                  0.06
-30 other         0.8   plmm_lasso                    0.580                  0.06
-\end{verbatim}
+\anna{var selection results}
+\begin{table}[H]
+\centering
+\begin{tabularx}{400pt}{cclYYY}
+\toprule
+$\eta$ & $\xi$ & \multicolumn{1}{c}{Method} & 1D Linear Admixture & Independent Subpopulations & Semi-Empirical \\ 
+\midrule
+\multirow{6}{*}{0.5} & \multirow{3}{*}{0} & Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.37 (0.06) \\ 
+& & PC-Lasso & 0.23 (0.05) & 0.21 (0.05) & 0.36 (0.06) \\ 
+& & LMM-Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.41 (0.06) \\
+\cmidrule{2-6}
+& \multirow{3}{*}{0.8} & Lasso & 0.23 (0.05) & 0.25 (0.05) & 0.29 (0.06) \\ 
+& & PC-Lasso & 0.28 (0.05) & 0.36 (0.06) & 0.47 (0.06) \\ 
+& & LMM-Lasso & 0.29 (0.05) & 0.39 (0.06) & 0.58 (0.06) \\ 
+\bottomrule
+\end{tabularx}
+\label{tab:var_sel}
+\caption{True positive rates, mean (SD), for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
+\end{table}
 

### 2020-08-22 main.tex
4.1 - variable selection table results description

@@
 We begin by comparing the ordinary lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
-Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) in various settings and broken down into bias and variance components.
+Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
@@
 In other words, a greater amount of the environmental effect is in the column space of the projection matrix of $\bX$ for the independent subpopulations data than in the admixed data.
-As such, there is a greater amount of confounding in the independent subpopulations which can be corrected by the PC-Lasso and LMM-Lasso methods.
+As such, there is a greater amount of confounding and potential estimation bias in the independent subpopulations which can be corrected by the PC-Lasso and LMM-Lasso methods.
 
@@
 
-\anna{var selection results}
+In addition to MSE, these methods' performance in terms of variable selection accuracy were quantified using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
+Table \ref{tab:var_sel} displays the mean true positive rate and corresponding standard deviation in the same settings as those illustrated in Figure \ref{fig:mse}.
+The TPRs echo the MSE results in that Lasso and LMM-Lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
+In the presence of environmental confounding, both LMM-Lasso and PC-Lasso achieve substantial improvements in TPR over Lasso across data types.
+As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Semi-Empirical data sets than in the 1D Linear Admixture data.
+Again, this is due to the reduced proportion of environmental confounding that is in the column space of $\boldsymbol{P}_{\bX}$, which results in a lesser degree of confounding of the SNP-phenotype associations.
+
 \begin{table}[H]
@@
 \end{tabularx}
+\caption{True positive rates, mean (SD), for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
 \label{tab:var_sel}
-\caption{True positive rates, mean (SD), for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
 \end{table}

### 2020-08-22 lmm-genetic-review.tex
Title; making DAG accents pink because girl power
4.1 - formatted variable selection table
Abstract edits

@@
 \usepackage{tkz-euclide}
+
+% table packages and commands
+\usepackage{booktabs}
+\usepackage{multirow}
+\usepackage{tabularx}
+\renewcommand\tabularxcolumn[1]{m{#1}} % for vertical centering in tables
+\newcolumntype{Y}{>{\centering\arraybackslash}X} % centering for X columns
 
@@
 
-\title{Review Paper Draft}
+\title{A comparative analysis of penalized linear mixed models in structured genetic data}
 \author{Anna Reisetter \qquad Patrick Breheny\\
@@
 In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
-In addition, we evaluate the performance of penalized LMMs and competing methods in terms of MSE \anna{and variable selction (specifically how we quantify this tbd)} in the presence of a number of confounding structures.
+In addition, we evaluate the performance of penalized LMMs and competing methods in terms of MSE and variable selection accuracy in the presence of a number of confounding structures.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.

### 2020-08-24 main.tex
Updating figures with plmm_lasso and updated time results

@@
 Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures.
-Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0, 0.2, 0.5,0.8\}$.
+Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
 Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
@@
 \end{figure}
-\anna{replace this figure with one that includes time in the third row}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.8]{figures/time_ratio.png}
-    \caption{Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
-    \label{fig:time}
-\end{figure}
-\anna{Note Figure \ref{fig:time} is on raw scale, not log10. This is because the negative values on the error bars create problems and become NaN on the log10 scale}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.8]{figures/time_ratio_box.png}
-    \caption{\anna{Alternate time fig} Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
-    \label{fig:time_box}
-\end{figure}
-
-\anna{make new time box plot - two boxes instead of ratio - also run for n = 1000, 1200, maybe 1400 and show MSE along with time... compare convergence results to ncvfit}
 
@@
 However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in Figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications.
-\anna{time interpretation depends on reason ggmix is slow...} Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.
+\anna{time interpretation depends on reason ggmix is slow...} Figure \ref{fig:eta_beta_mse} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.
 The horizontal dashed line at 1 indicates both methods require the same computing time.

### 2020-08-24 main.tex
4.4 edits - in progress

@@
 In section \ref{sec:rak-ggmix} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-As we will see, the two generally perform quite similarly.
-Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods.
+Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection.
+Becasue of this, the preceding results, though generated using the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods in the $n = 200, p = 1000$ setting.
 The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.
@@
 
-LMM-Lasso-Rakitsch estimates the variance components once, under the assumption of a null model.
-LMM-Lasso-ggmix seeks to improve upon this by re-estimating the variance components and $\bbeta$ iteratively.
+LMM-Lasso-Rakitsch uses a two-step approach.
+First, the variance components are estimated under the assumption of a null model.
+Then, the estimated variance components are used to transform the data, and the $\bbeta$ are estimated using coordinate descent.
+LMM-Lasso-ggmix seeks to improve upon this procedure by re-estimating the variance components and $\bbeta$ iteratively.
 Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$.
-Figure \ref{fig:eta_beta_mse} shows the MSE for estimation of both $\bbeta$ and $\eta$ in the presence of varying $\eta$ values and sample sizes, $n$, for both LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-The bottom panel of this figure shows a modest improvement in the estimation of $\eta$ for the LMM-Lasso-ggmix method, particularly in the presence of lower $\eta$ values.
-However, these apparent improvements in estimating $\eta$ do not improve the $\bbeta$ estimation accuracy of LMM-Lasso-ggmix.
-This is evidenced by the top panel of Figure \ref{fig:eta_beta_mse}, which shows nearly identical performances of both methods in terms of $\bbeta$ MSE.
+In order to systematically compare both methods, we carried out simulations where $p = 1000$, $n = \{200, 400, 600, 800, 1000, 1200, 1400\}$, and $\eta = \{0.2, 0.5, 0.8\}$.
+Figure \ref{fig:eta_beta_mse} shows the MSE for estimation of the parameters $\bbeta$ and $\eta$, in addition to the corresponding computing time required.
 
-\anna{add some info/figure on convergence and number of iterations for lmm-lasso vs ggmix, delete the following time figures and references to them}
+LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n << p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
+When $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-Lasso-Rakitsch.
+This is most dramatic in the presence of substantial environmental environmental confounding, $\xi = 0.8$, and $n = \{1200, 1400\}$.
+LMM-Lasso-ggmix achieves a moderate improvement in the estimation of $\eta$, particularly in the presence of lower $\eta$ values and where $n << p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
+Interestingly, these improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix.
+When $n \ge 1000$, LMM-Lasso-ggmix's estimation of $\eta$ in suffers.
+When $n \ge 1000$ and in the presence of substantial environmental confounding, $\xi = 0.8$,  LMM-Lasso-ggmix's is hindered to the point that LMM-Lasso-Rakitsch estimates $\eta$ more accurately.
+
+Overall these results show that LMM-Lasso Rakitsch outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ where $n \ge p$, likely due to the ill-conditioning of the RRM in these settings \anna{cite French paper (?)}.
+% Given the intended use of these methods for the analysis of genetic data, where the number of genetic variants often outnumber the number of subjects, these results could be of little consequence to practitioners.
+When $n < p$, LMM-Lasso-ggmix estimates $\bbeta$ as well as LMM-Lasso-Rakitsch, and $\eta$ better than LMM-Lasso-Rakitsch.
+However, in this setting the computational time LMM-Lasso-ggmix reqiures may be prohibitive for practical applications.
+The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required for each method in the various simulation settings.
+
+\anna{ below in progress}
+
+Unfortunately, the settings where ggmix estimates things well coincide with excessive times and when it works well in terms of time, it estimates poorly
+
+Despite its comparable performance in accurately estimating $\bbetaHat$, LMM-Lasso-ggmix takes requires an average of \anna{73.4} times the computing time of LMM-Lasso-Rakitsch.
+Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
 
@@
 
-In addition, we have found that LMM-Lasso-ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the RRM.
-Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners.
-However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in Figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications.
-\anna{time interpretation depends on reason ggmix is slow...} Figure \ref{fig:eta_beta_mse} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes.
-The horizontal dashed line at 1 indicates both methods require the same computing time.
-Despite its comparable performance in accurately estimating $\bbetaHat$, LMM-Lasso-ggmix takes requires an average of \anna{73.4} times the computing time of LMM-Lasso-Rakitsch.
-Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
+\anna{add some info/figure on convergence and number of iterations for lmm-lasso vs ggmix}
 

### 2020-08-25 main.tex
4.4 and 5 edits

@@
 LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n << p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
-When $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-Lasso-Rakitsch.
+However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-Lasso-Rakitsch.
 This is most dramatic in the presence of substantial environmental environmental confounding, $\xi = 0.8$, and $n = \{1200, 1400\}$.
-LMM-Lasso-ggmix achieves a moderate improvement in the estimation of $\eta$, particularly in the presence of lower $\eta$ values and where $n << p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
-Interestingly, these improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix.
-When $n \ge 1000$, LMM-Lasso-ggmix's estimation of $\eta$ in suffers.
-When $n \ge 1000$ and in the presence of substantial environmental confounding, $\xi = 0.8$,  LMM-Lasso-ggmix's is hindered to the point that LMM-Lasso-Rakitsch estimates $\eta$ more accurately.
 
-Overall these results show that LMM-Lasso Rakitsch outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ where $n \ge p$, likely due to the ill-conditioning of the RRM in these settings \anna{cite French paper (?)}.
+LMM-Lasso-ggmix achieves a moderate improvement in the estimation of $\eta$ compared to LMM-Lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n << p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
+These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix.
+In addition, when $n \ge 1000$, LMM-Lasso-ggmix's estimation of $\eta$ suffers.
+When $n \ge 1000$ and in the presence of substantial environmental confounding, $\xi = 0.8$,  LMM-Lasso-ggmix's performance is hindered to the point that LMM-Lasso-Rakitsch estimates $\eta$ more accurately.
+
+Overall these results show that LMM-Lasso-ggmix performs as well as LMM-Lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-Lasso-Rakitsch in estimating $\eta$ when $n << p$.
+However, these are also the settings in which LMM-Lasso-ggmix is most computationally inefficient.
+The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n << p$, LMM-Lasso-ggmix is up to 60 times slower than LMM-Lasso-Rakitsch.
+
+Conversely, in settings where $n \ge p$, LMM-Lasso-ggmix is up to 10 times faster than LMM-Lasso-Rakitsch.
+However, these are also settings in which LMM-Lasso Rakitsch outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently, likely due to the ill-conditioning of the RRM \anna{cite French paper (?)}.
+
 % Given the intended use of these methods for the analysis of genetic data, where the number of genetic variants often outnumber the number of subjects, these results could be of little consequence to practitioners.
-When $n < p$, LMM-Lasso-ggmix estimates $\bbeta$ as well as LMM-Lasso-Rakitsch, and $\eta$ better than LMM-Lasso-Rakitsch.
-However, in this setting the computational time LMM-Lasso-ggmix reqiures may be prohibitive for practical applications.
-The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required for each method in the various simulation settings.
-
-\anna{ below in progress}
-
-Unfortunately, the settings where ggmix estimates things well coincide with excessive times and when it works well in terms of time, it estimates poorly
-
-Despite its comparable performance in accurately estimating $\bbetaHat$, LMM-Lasso-ggmix takes requires an average of \anna{73.4} times the computing time of LMM-Lasso-Rakitsch.
-Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\anna{this to change somewhat based on time/convergence results...}
 
@@
 Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks.
-Firstly, it suffers from instability in low dimensions.
-\anna{depends on ggmix time results} Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires an average of 73.4 and as much as 167.57 \anna{73.4 is the mean ratio + sd, but the max absolute ratio is 167.57 - which should I report?} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure.
-Thirdly, while the additional time required of LMM-Lasso-ggmix may be worthwhile if it yielded more accurate estimates of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are greatly superior to those of LMM-Lasso-Rakitsch.
-We have implemented the LMM-Lasso-Rakitsch method using the equivalent, but more interpretable, $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM (final name tbd)}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.
+Firstly, it suffers from estimation instability in $n \ge p$ settings.
+Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably when $n << p$, LMM-Lasso-ggmix requires up to 60 times the computing time of LMM-Lasso-Rakitsch in these settings due to its iterative estimation procedure.
+The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n << p$ settings.
+However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus considered by the authors to be more broadly applicable.
 
-\pb{Introduce penalizedLMM here, or in next paper?}
+\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-Lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
 

### 2020-08-26 main.tex
Final figure updates

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\anna{REPLACE FIGURES WITH RESULTS FROM PLMM-LASSO ONCE FINISHED}
 
@@
 Conversely, in settings where $n \ge p$, LMM-Lasso-ggmix is up to 10 times faster than LMM-Lasso-Rakitsch.
-However, these are also settings in which LMM-Lasso Rakitsch outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently, likely due to the ill-conditioning of the RRM \anna{cite French paper (?)}.
+However, these are also settings in which LMM-Lasso Rakitsch outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently, likely due to the ill-conditioned RRM \citep{ledoit2004well}.
 
@@
 \end{figure}
-
-\anna{add some info/figure on convergence and number of iterations for lmm-lasso vs ggmix}
 

### 2020-08-28 main.tex
Edits to 2.3, 3,1

@@
 One of the first methods to address the problem of confounding due to population structure is the genomic control (GC) method.
-GC is a univariate testing procedure which modifies a test statistic by a deflation factor.
-This deflation factor attempts to quantify and remove the effect of population structure on the distribution of the appropriate test statistic, and is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
-However, this deflation factor has been reported to be sensitive to the nature and quantity of the loci used to estimate it \citep{hellwege2017population, marchini2004effects}.
+Because the effect of this confounding is to inflate the distribution of test statistics, GC introduces a factor to counteract this effect and restore the appropriate null distribution.
+This deflation factor is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
+However, quantifying the extent of inflation is not trivial, and is sensitive to the nature and quantity of the loci used to estimate it \citep{hellwege2017population, marchini2004effects}.
 Additionally, GC does not correct effect size estimates, so resulting coefficient estimates will be unreliable after GC is applied, even if their corresponding test statistics and p-values have been appropriately adjusted \citep{hellwege2017population}.
@@
 This is illustrated by the blocked arrow in Figure \ref{fig:ps_env_block}.
-As the figure shows, these methods do not directly address the environmental component of confounding due to population structure.
-Nevertheless, they may be effective in controlling confounding by indirectly blocking the causal pathway between the environment and observed SNPs.
+As the figure shows, these methods do not directly address the environmental component of confounding due to population structure, instead controlling confounding by indirectly blocking the causal pathway between the environment and observed SNPs.
 
@@
     \node (2) [right = of 1] {Environment};
-    \node (3) [below = of 1] {Observbed SNPs};
+    \node (3) [below = of 1] {Observed SNPs};
     \node (4) [below = of 2] {Phenotype};
@@
 
-Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred, and if population structure is a good proxy for environmental effects.
+\pb{The flow of ideas here does not make sense to me; let's discuss} Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred, and if population structure is a good proxy for environmental effects.
 In the presence of environmental heterogeneity which affects the phenotype of interest, and which is not well-approximated by inferred population structure, SNP-phenotype associations will not be confounded.
@@
 Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.
-By partitioning the overall sample size in this manner, the structured association method is particularly vulnerable to a loss of power.
+By partitioning the overall sample size in this manner, the structured association method is vulnerable to a loss of power.
 Additionally, it relies on the assumption that these subpopulations are distinct and nonoveralapping.
@@
 Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall between these two extremes and are characterized by continuously varying relatedness and a gradient of air quality.
-By clustering those individuals within in one subpopulation or another we lose information about their more fine-grained, subtle relatedness.
+By clustering those individuals within in one subpopulation or another, we lose information about relatedness by oversimplifying the population structure.
 
-Another approach is to create an approximation of population structure using surrogate variables, and to adjust for these as an additional model covariates.
+A different approach is to create an approximation of population structure using surrogate variables, and to adjust for these as an additional model covariates.
 Principal component-adjustment (PC-adjustment) methods use the first several left singular vectors of a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
@@
 In an LMM, however, these left singular vectors are treated as random, rather than fixed effects.
-Additionally, the term LMM has been used to describe a number of distinct but related procedures for analyzing structured genetic data which differ in their objectives and methods.
+The term LMM has been used to describe a number of distinct but related procedures for analyzing structured genetic data which differ in their objectives and methods.
 There are two main objectives of LMMs used in a structured genetic context: estimating narrow-sense heritability, and estimating individual SNP effects.
@@
 
-Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependent error structures  \citep{Rakitsch2012, bhatnagar2019simultaneous}.
+Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependencies among subjects \citep{Rakitsch2012, bhatnagar2019simultaneous}.
 This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes.
@@
 %------------------------------%
+\pb{I think we need a more descriptive section title than ``Model''}
 \subsection{Model}
+
+\pb{(1) should we use something besides p for MAF?  Seems like we're overloading p. (2) Mention motivation for standardizing?}
 
@@
 Thus, at a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
-However, $\bX_0$ may additionally include covariates such as age and sex, which are of interest to include in the model without penalty.
+However, $\bX_0$ may additionally include covariates such as age, sex, or smoking history, which are included in the model without penalty.
 We assume that the $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of these covariates, additive SNP effects, environmental effects, and observational noise, and consider the linear mixed model,
@@
 
+\pb{Need to discuss fact that Z is not observed.  Perhaps move discussion of decomposition to here?  Or at least mention it?}
+
 LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
@@
 
+\pb{make one line?}
 \begin{align}
@@
 \end{equation}
-where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||.||_2$ the $\ell_2$ norm.
+where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm.
 Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
@@
 Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect.
-Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
-This results in the LMM regression model
+\pb{I think this needs to be discuss much earlier, in Section 3.1}
+Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}. This results in the LMM regression model
 \begin{equation}
@@
 where $\sigmagg$ is the additive genetic variance, and we assume $\beps \independent \bgamma$.
-Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields the equivalent model
+Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields the equivalent model \pb{Writing these equations out twice seems redundant}
 \begin{equation}
@@
 \end{equation}
-and thus a regular \anna{vanilla, naive, regular?} Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
-\anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
+and thus a regular \anna{vanilla, naive, regular?} \pb{how about ``ordinary''} Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
+\anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?} \pb{I don't think we need to}
 
@@
 \end{align*}
-By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^\T \bX$ and $\byT = \bU^\T \by$, where $\bW = \text{diag}\{w_1, ...
-w_n\}$ and $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$.
+By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^\T \bX$, $\byT = \bU^\T \by$, $\bW = \text{diag}\{w_1, ...
+w_n\}$, and $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$.
 Together, this yields the objective function
@@
 \end{equation}
-where $\bX{^\prime} = \begin{bmatrix} \bX_0 & \bX \end{bmatrix}$ and $\bbeta{^\prime} = \begin{bmatrix} \bbeta_0^\T & \bbeta^\T \end{bmatrix}^\T$.
-This in turn leads to a new version of \eqref{eqn:lmm_obj}, our objective function,
+where $\bX{^\prime} = \begin{bmatrix} \bX_0 & \bX \end{bmatrix}$ and $\bbeta{^\prime} = \begin{bmatrix} \bbeta_0^\T & \bbeta^\T \end{bmatrix}^\T$. This in turn leads the objective function
 \begin{equation}
-Q(\bbeta | \bXT{^\prime}, \byT, \bW) = || \bW (\byT - \bXT{^\prime} \bbeta{^\prime})||_2^2 + \lambda || \bbeta ||_1
+Q(\bbeta | \bXT{^\prime}, \byT, \bW) = || \bW (\byT - \bXT{^\prime} \bbeta{^\prime})||_2^2 + \lambda || \bbeta ||_1,
 \end{equation}

### 2020-08-28 main.tex
2020-08-28 meeting notes/annotations

@@
 
-\pb{The flow of ideas here does not make sense to me; let's discuss} Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred, and if population structure is a good proxy for environmental effects.
+\pb{The flow of ideas here does not make sense to me; let's discuss} \anna{earlier or later paragraph} Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
+\anna{an important point to note, and one that we will return to later in 3.5, is that this only leads to confounding if pop structure is correlated with the environment...}
 In the presence of environmental heterogeneity which affects the phenotype of interest, and which is not well-approximated by inferred population structure, SNP-phenotype associations will not be confounded.
@@
 Thus, at a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
-However, $\bX_0$ may additionally include covariates such as age, sex, or smoking history, which are included in the model without penalty.
+However, $\bX_0$ may additionally include \anna{observed (?)} covariates such as age, sex, or smoking history, which are included in the model without penalty.
 We assume that the $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of these covariates, additive SNP effects, environmental effects, and observational noise, and consider the linear mixed model,
@@
 
-\pb{Need to discuss fact that Z is not observed.  Perhaps move discussion of decomposition to here?  Or at least mention it?}
+\pb{Need to discuss fact that Z is not observed.
+Perhaps move discussion of decomposition to here?
+Or at least mention it?} \anna{how is Z different from X0?
+introduce this earlier...
+explicityly state that Z is not directly observed}
 
@@
 where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors or principal components of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are thus the principal components of $\bK$ weighted by their corresponding singular values.
+\anna{cut some of this?}
 
@@
 \label{sec:rak-ggmix}
+\anna{remove this as a section}
 
-We have so far used the term $\bW$ to describe the LMM-Lasso method.
+\anna{make this sentence less weird...
+how to estimate W leads to different ways leads to 2 different variants...} We have so far used the term $\bW$ to describe the LMM-Lasso method.
 However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.

### 2020-09-01 main.tex
Changing references to Lasso to 'ordinary Lasso'

@@
 \end{equation}
-and thus a regular \anna{vanilla, naive, regular?} \pb{how about ``ordinary''} Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
-\anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?} \pb{I don't think we need to}
+and thus an ordinary Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
 
@@
 %------------------------------%
-\subsection{PC-Lasso and LMM-Lasso outperform Lasso in the presence of environmental confounding}
+\subsection{PC-Lasso and LMM-Lasso outperform the ordinary Lasso in the presence of environmental confounding}
 %------------------------------%
 
-We begin by comparing the ordinary lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
+We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
 Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components.
@@
 \end{figure}
-
+\anna{talk about karl rohe stuff here - if anywhere}
 In addition to MSE, these methods' performance in terms of variable selection accuracy were quantified using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
@@
 In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
-We then compare the estimation accuracy of a \anna{naive} Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
+We then compare the estimation accuracy of an ordinary Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
 
-Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the \anna{naive} Lasso whenever environmental confounding effects are present.
-However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the \anna{naive} Lasso.
+Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the ordinary Lasso whenever environmental confounding effects are present.
+However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the ordinary Lasso.
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.

### 2020-09-01 main.tex
2.3 edits

@@
 
-\pb{The flow of ideas here does not make sense to me; let's discuss} \anna{earlier or later paragraph} Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
-\anna{an important point to note, and one that we will return to later in 3.5, is that this only leads to confounding if pop structure is correlated with the environment...}
+Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
+An important concept to note, and one that we will return to in section \ref{sec:sim_env_conf}, is that population structure only leads to confounding if it is correlated with the environment.
 In the presence of environmental heterogeneity which affects the phenotype of interest, and which is not well-approximated by inferred population structure, SNP-phenotype associations will not be confounded.
@@
 \subsection{Simulating environmental confounding}
+\label{sec:sim_env_conf}
 

### 2020-09-01 main.tex
3.1 - 3.3 edits

@@
 %------------------------------%
-\pb{I think we need a more descriptive section title than ``Model''}
-\subsection{Model}
+\subsection{LMM and PC models}
 
-\pb{(1) should we use something besides p for MAF?  Seems like we're overloading p. (2) Mention motivation for standardizing?}
+\pb{(2) Mention motivation for standardizing?}
 
 To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
-We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 p_j) / \sqrt{2 p_j (1 - p_j)}$ and $p_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
+\anna{in practice we standardize this matrix because reasons...} We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 \pi_j) / \sqrt{\pi_j (1 - \pi_j)}$ and $\pi_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
 Additionally, $\bX_0$ represents an $n \times p_0$ matrix of unpenalized covariates and the intercept.
 Thus, at a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
-However, $\bX_0$ may additionally include \anna{observed (?)} covariates such as age, sex, or smoking history, which are included in the model without penalty.
-We assume that the $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of these covariates, additive SNP effects, environmental effects, and observational noise, and consider the linear mixed model,
+However, $\bX_0$ may additionally include observed covariates such as age, sex, or smoking history, which are included in the model without penalty.
+We assume that the $n \times 1$ vector of continuous phenotype values, $\by$, can be expressed as the sum of these covariates, additive SNP effects, environmental effects, and observational noise, and consider the linear model,
 \begin{equation}
@@
 
-\pb{Need to discuss fact that Z is not observed.
-Perhaps move discussion of decomposition to here?
-Or at least mention it?} \anna{how is Z different from X0?
-introduce this earlier...
-explicityly state that Z is not directly observed}
+A key feature which differentiates \eqref{eqn:model} from traditional linear mixed models is that in our setting, $\bZ$ is not observed directly, but must be corrected for in order to remove the effects of confounding influences.
+A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
+This, in turn, motivates the ways in which LMM-Lasso and PC-Lasso attempt to model unobserved confounding effects.
 
-LMM-Lasso relies on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
+Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
 Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
 
-\pb{make one line?}
-\begin{align}
+\begin{equation}
     \label{eqn:k}
-    \bK &= \frac{1}{p} \bX \bX^\T \nonumber \\
-                   &= \frac{1}{p} \bU \bD^2 \bU^\T \nonumber \\
-                   &= \frac{1}{p} \bU \bD (\bU \bD)^\T \nonumber \\
-                   &= \frac{1}{p} \bR \bR^\T,
-\end{align}
-where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ (equivalently, the $k$ nonzero squared singular values of $\bX$) on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal, and $\bU$ is the corresponding $n \times n$ orthonormal matrix of the left eigenvectors or principal components of $\bK$ (equivalently, the left singular vectors of $\bX$.) The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$ (squared singular values of $\bX$.) The columns of $\bR$ are thus the principal components of $\bK$ weighted by their corresponding singular values.
-\anna{cut some of this?}
+    \bK = \frac{1}{p} \bX \bX^\T = \frac{1}{p} \bU \bD^2 \bU^\T = \frac{1}{p} \bR \bR^\T,
+\end{equation}
+where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal.
+$\bU$ is the corresponding $n \times n$ orthonormal matrix of left eigenvectors, or principal components, of $\bK$.
+The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$, and the columns of $\bR$ are thus the principal components of $\bK$ weighted by their corresponding singular values.
 
@@
 \subsection{LMM-Lasso}
+\label{sec:lmmlasso}
 
-Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect.
-\pb{I think this needs to be discuss much earlier, in Section 3.1}
-Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}. This results in the LMM regression model
+Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect, which corresponds to the LMM model
 \begin{equation}
@@
 where $\sigmagg$ is the additive genetic variance, and we assume $\beps \independent \bgamma$.
-Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields the equivalent model \pb{Writing these equations out twice seems redundant}
-\begin{equation}
-    \label{eqn:lmm_reg_u}
-    \by = \bX_0\bbeta_0 + \bX\bbeta + \bu + \beps
-\end{equation}
-$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI) $$
-$$ \bu \sim N(\boldsymbol{0}, \sigmagg \bK).$$
+Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields an equivalent model.
+
 We will first derive the LMM-Lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
-Following equations \eqref{eqn:lmm_reg} and \eqref{eqn:lmm_reg_u}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
+Following equation \eqref{eqn:lmm_reg}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
 LMM-Lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
@@
 %------------------------------%
-\subsubsection{LMM-Lasso variants}
-\label{sec:rak-ggmix}
-\anna{remove this as a section}
 
-\anna{make this sentence less weird...
-how to estimate W leads to different ways leads to 2 different variants...} We have so far used the term $\bW$ to describe the LMM-Lasso method.
-However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+Different variance component estimation methods lead to two different forms of $\widehat{\bW}$ and LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
 LMM-Lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
@@
 
-Using the result of the eigendecomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg_u} can be written equivalently as
+Using the result of the eigendecomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
 
@@
 
-In section \ref{sec:rak-ggmix} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+In section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
 Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection.

### 2020-09-02 main.tex
Notes to self

@@
 
-Different variance component estimation methods lead to two different forms of $\widehat{\bW}$ and LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+Different variance component estimation methods lead to different forms of $\widehat{\bW}$ and two distinct LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
 LMM-Lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
@@
 However, the degree to which it reflects population structure that can be inferred from $\bX$, as quantified by $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$, will impact the extent to which environmental heterogeneity confounds SNP-phenotype associations.
+\anna{this calculation is still having issues for the admixture and independent supbopulation data - the ratio is slightly larger for the independent data - as we would expect - but only on the order of 1e-15...
+correlation of Xtau with Zgamma?
+euclidian distance between vectors?}
 
@@
 $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to environmental effects.
+So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to unobserved environmental effects.
 
@@
 While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
-In addition, because Lasso, PC-Lasso, and LMM-Lasso use different likelihoods an information-criterion-based approach would not lend itself to comparable results across methods.
+In addition, because Lasso, PC-Lasso, and LMM-Lasso use different likelihoods, an information-criterion-based approach would not lend itself to comparable results across methods.
 
@@
 However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus considered by the authors to be more broadly applicable.
+\anna{talk about how ggmix seems to get stuck in a local minimum because it's esimating more parameters the optimization problem is not necessarily convex, this its prone to getting stuck.
+While this may make it fast, it makes it terrible for estimation accuracy.
+Reference iterations plots?}
 

### 2020-09-02 main.tex
3.1 edits

@@
 
-\pb{(2) Mention motivation for standardizing?}
-
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\boldsymbol{G}$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
-\anna{in practice we standardize this matrix because reasons...} We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 \pi_j) / \sqrt{\pi_j (1 - \pi_j)}$ and $\pi_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
+To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\bG$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
+In order to equitably compare SNP effects, regardless of allele frequency, in practice $\bG$ is standardized.
+We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 \pi_j) / \sqrt{\pi_j (1 - \pi_j)}$ and $\pi_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
 Additionally, $\bX_0$ represents an $n \times p_0$ matrix of unpenalized covariates and the intercept.
@@
 A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
-This, in turn, motivates the ways in which LMM-Lasso and PC-Lasso attempt to model unobserved confounding effects.
+This, in turn, motivates the methods by which LMM-Lasso and PC-Lasso attempt to model unobserved confounding effects using observed genetic data.
 

### 2020-09-02 main.tex
4.4 - 5 edits

@@
 
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\bG$ for $n$ subjects and $p$ SNPs, where each $G_{i,j} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
+To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\bG$ for $n$ subjects and $p$ SNPs, where each $G_{ij} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
 In order to equitably compare SNP effects, regardless of allele frequency, in practice $\bG$ is standardized.
@@
 \end{figure}
-\anna{talk about karl rohe stuff here - if anywhere}
+% \anna{talk about karl rohe stuff here - if anywhere}
+
 In addition to MSE, these methods' performance in terms of variable selection accuracy were quantified using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
@@
 
-LMM-Lasso-ggmix achieves a moderate improvement in the estimation of $\eta$ compared to LMM-Lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n << p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
-These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix.
+LMM-Lasso-ggmix achieves an improvement in the estimation of $\eta$ compared to LMM-Lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n << p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
+These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix and thus do not seem to improve its estimation of the fixed genetic effects.
 In addition, when $n \ge 1000$, LMM-Lasso-ggmix's estimation of $\eta$ suffers.
-When $n \ge 1000$ and in the presence of substantial environmental confounding, $\xi = 0.8$,  LMM-Lasso-ggmix's performance is hindered to the point that LMM-Lasso-Rakitsch estimates $\eta$ more accurately.
+When $n \ge 1000$ and substantial environmental confounding is present, $\xi = 0.8$,  LMM-Lasso-ggmix's performance is hindered to the point that LMM-Lasso-Rakitsch estimates $\eta$ more accurately.
 
@@
 The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n << p$, LMM-Lasso-ggmix is up to 60 times slower than LMM-Lasso-Rakitsch.
+Conversely, in settings where $n \ge p$, LMM-Lasso-ggmix is up to 10 times faster than LMM-Lasso-Rakitsch.
+However, these are also settings in which LMM-Lasso Rakitsch substantially outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently.
 
-Conversely, in settings where $n \ge p$, LMM-Lasso-ggmix is up to 10 times faster than LMM-Lasso-Rakitsch.
-However, these are also settings in which LMM-Lasso Rakitsch outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently, likely due to the ill-conditioned RRM \citep{ledoit2004well}.
+LMM-Lasso-ggmix's poor performance in these scenarios are likely attributable to a number of a factors such as an ill-conditioned RRM  in $n \ge p$ settings \citep{ledoit2004well}.
+In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
+An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure ?}.
+Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.
 
-% Given the intended use of these methods for the analysis of genetic data, where the number of genetic variants often outnumber the number of subjects, these results could be of little consequence to practitioners.
+\anna{re-create niter\_parLambda fig focusing on the eta = xi = 0.8 seting and displaying all data points - not just averages}
 
@@
 
-In this work, we have reviewed the concept of population stratification and explicitly defined the environmental mechanism by which it may confound the genotype-phenotype relationship.
+In this work, we have reviewed the concept of population stratification and explicitly defined an environmental mechanism by which it may confound the genotype-phenotype relationship.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
@@
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
-Additionally, we find that LMM-Lasso outperforms PC-Lasso when (1) fine subpopulation structure is present and (2) when confounding effects are nonlinear and discordant in nature.
+Additionally, we find that LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure and when confounding effects are nonlinear and discordant in nature.
 In the presence of coarse subpopulation structure which corresponds to concordant, dichotomous environmental effects, PC-Lasso may outperform LMM-Lasso in estimation accuracy.
@@
 
-Finally, we considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
+We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
 Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks.
@@
 However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus considered by the authors to be more broadly applicable.
-\anna{talk about how ggmix seems to get stuck in a local minimum because it's esimating more parameters the optimization problem is not necessarily convex, this its prone to getting stuck.
-While this may make it fast, it makes it terrible for estimation accuracy.
-Reference iterations plots?}
 
@@
 
-It is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
+As an additional point of interest, recall that we provide qualitative and mathematical explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure.
+While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, it also highlights the manner in which such unobserved environmental effects have the potential to inflate residual model variance and bias narrow-sense heritability estimates.
+A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper.
+However, we hope this emphasizes the broad implications and importance of the subtle distinction between the effect of population structure and unobserved environmental heterogeneity.
+
+Finally, it is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
 Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.
-
-\anna{the importance of distinguishing between population structure and heterogeneous environmental exposures in penalized regression methods - do we want to get into the importance of not conflating environmental effects with heritability?}

### 2020-09-03 main.tex
Edits to Section 3

@@
 An important concept to note, and one that we will return to in section \ref{sec:sim_env_conf}, is that population structure only leads to confounding if it is correlated with the environment.
-In the presence of environmental heterogeneity which affects the phenotype of interest, and which is not well-approximated by inferred population structure, SNP-phenotype associations will not be confounded.
+In the presence of environmental heterogeneity which affects the phenotype of interest in a manner that is orthogonal to population structure, SNP-phenotype associations will not be confounded.
 This is due to the absence of a causal pathway between population stratification and environment, as depicted in Figure \ref{fig:ps_env_block2}.
-In such a scenario, failing to account for unobserved environmental effects is likely to result in greater residual variability, however, SNP effect estimates will be unbiased \citep{greenland1999causal}.
+In such a scenario, failing to account for unobserved environmental effects is likely to result in greater variability, but SNP effect estimates will remain unbiased \citep{greenland1999causal}.
 We will return to this idea and its mathematical justification in section \ref{sec:methods}.
@@
 
-Assuming for now that population structure is a good proxy for environmental effects, we return to our consideration of methods that attempt to control for confounding via the mechanism depicted in Figure \ref{fig:ps_env_block}.
+Assuming for now that population structure is correlated with environmental exposures and can be accurately inferred from available data, we return to our consideration of methods that attempt to control for confounding via the mechanism depicted in Figure \ref{fig:ps_env_block}.
 Such methods include LMMs, principal-component-based approaches, and stratified analyses.
@@
 
-A key feature which differentiates \eqref{eqn:model} from traditional linear mixed models is that in our setting, $\bZ$ is not observed directly, but must be corrected for in order to remove the effects of confounding influences.
-A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.
+A key feature which differentiates \eqref{eqn:model} from traditional linear mixed models is that in our setting, $\bZ$ is not observed directly, although we still wish to correct for it in order to remove the effects of confounding influences.
+A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}\pb{; we will discuss the rationale for this approach in Section~\ref{sec:sim_env_conf}}.
 This, in turn, motivates the methods by which LMM-Lasso and PC-Lasso attempt to model unobserved confounding effects using observed genetic data.
@@
 LMM-Lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2019simultaneous}.
-This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix}.
-We postpone further discussion of the differences between these methods to section \ref{sec:results}.
+This iterative update scheme prevents the use of standard Lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
+We further discuss differences between these methods in section \ref{sec:results}.
 
@@
 
-It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects.
-\citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, and the decomposition of $\bK$, which we briefly review.
+It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects \pb{in the absence of penalization?}.
+\citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, which we briefly review.
 
-Using the result of the eigendecomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
+Using the result of the eigendecomposition \pb{in what sense is this an eigendecomposition?
+I would describe it as a Cholesky decomposition} from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
 
@@
 
-Additionally, while the fixed effects model considers the principal components, $\bU$, of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
+Additionally, while the fixed effects model considers the principal components $\bU$ of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
 The eigenvalues have been shown to serve as a measure of biological relevance of each principal component in relation to underlying population structure \citep{hoffman2013correcting, price2006principal}.
@@
 Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}.
-This is the implementation of PC-Lasso throughout Section \ref{sec:results}.
+This is the approach we take when implementing PC-Lasso in Section \ref{sec:results}.
 
@@
 
+\pb{We need a new section title here; the next section is the one that focuses on simulation.  This one talks about a number of important concepts that go well beyond simulation.}
+
 In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis \citep{Rakitsch2012, bhatnagar2019simultaneous}.
 We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method.
-Averaging over situations in which the environmental confounding is constantly changing direction and magnitude is inconsistent with the confounding environmental effects.
-This has the effect of ``washing out'' bias in $\bbetaHat$, which is of interest, while inflating the variance.
+Averaging over situations in which the environmental confounding is constantly changing direction and magnitude has the effect of ``washing out'' bias in $\bbetaHat$.
 To illustrate this concept, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
@@
 Under the assumption that environmental effects arise from a random process, $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
-Consider our earlier example involving confounding due to air quality.
+Consider our earlier example involving confounding due to air quality \pb{I feel that this should be moved up to before the mathematical derivations}.
 Simulating $\bgamma$ as a random quantity for each replication corresponds to subpopulation A being exposed to good air quality in one replication, and bad air quality in the next.
@@
 We see this by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.
-Define the projection matrix of $\bX$ by $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$, and let $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denote the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
-Then, $\exists \boldsymbol{\tau, \psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$ , $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$, and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$ such that
+Let $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denoting the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
+Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
+This in turn means that
 \begin{align*}
@@
     &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps\\
-    &=  \bX \bbeta{^*} + \beps{^*},
+    &=  \bX \bbeta{^*} + \beps{^*};
 \end{align*}
-where $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term.
+note that $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term.
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
@@
 
-Given the nature of environmental confounding in our setting of interest, and in order to assess these methodological performance in terms of estimation accuracy and bias specifically, we simulate according to a fixed $\bgamma$ framework using the following data-generating model
+Given the nature of environmental confounding in our setting of interest, and in order to assess methodological performance in terms of estimation accuracy and bias, we simulate from the data-generating model
 \begin{equation}
@@
 \end{equation}
-where $\by$ represents the pseudophenotype, and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
+where $\by$ represents the pseudophenotype, $\bgamma$ is a fixed quantity (see section~\ref{sec:sim_env_conf}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
 \begin{equation}

### 2020-09-03 main.tex
2020-09-03 meeting edits

@@
 Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure and population in question.
-\pb{is this an important point?} \anna{hmm...
-maybe?
-robust to this assumption: if it's not a good proxy it's not going to cause confounding}
 
@@
 
-Using the result of the eigendecomposition \pb{in what sense is this an eigendecomposition?
-I would describe it as a Cholesky decomposition} from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
+Using the result of the decomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
 
@@
 
+\anna{(1) fixed vs. random, (2) decomposition}
+
 \pb{We need a new section title here; the next section is the one that focuses on simulation.  This one talks about a number of important concepts that go well beyond simulation.}
+
+\anna{qualitative explanation, then quantitative}
 
@@
 Under the assumption that the quantity $\bgamma$ is fixed, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, allowing us to quantify estimation bias due to environmental heterogeneity.
-
+\anna{talk about a low-rank approximation of xtx-inv for a unique solution}
 Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$.

### 2020-09-08 main.tex
3.5 edits

--- 2020-09-03 main.tex
@@
 A key feature which differentiates \eqref{eqn:model} from traditional linear mixed models is that in our setting, $\bZ$ is not observed directly, although we still wish to correct for it in order to remove the effects of confounding influences.
-A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}\pb{; we will discuss the rationale for this approach in Section~\ref{sec:sim_env_conf}}.
+A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}; we will discuss the rationale for this approach in Section~\ref{sec:sim_env_conf}.
 This, in turn, motivates the methods by which LMM-Lasso and PC-Lasso attempt to model unobserved confounding effects using observed genetic data.
@@
 
-It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects \pb{in the absence of penalization?}.
+It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects in the absence of penalization?.
 \citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, which we briefly review.
 
-Using the result of the eigendecomposition \pb{in what sense is this an eigendecomposition?
-I would describe it as a Cholesky decomposition} from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
+Using the result of the decomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
 
@@
 %------------------------------%
-\subsection{Simulating environmental confounding}
+\subsection{\anna{Decomposing (Quantifying?) the effect of environmental confounding}}
 \label{sec:sim_env_conf}
 
-\pb{We need a new section title here; the next section is the one that focuses on simulation.  This one talks about a number of important concepts that go well beyond simulation.}
+In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis \citep{Rakitsch2012, bhatnagar2019simultaneous}.
+We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method in the presence of environmental heterogeneity.
+Averaging over situations in which the environmental confounding is constantly changing direction and magnitude has the effect of ``washing out'' bias in $\bbetaHat$.
 
-In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis \citep{Rakitsch2012, bhatnagar2019simultaneous}.
-We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method.
-Averaging over situations in which the environmental confounding is constantly changing direction and magnitude has the effect of ``washing out'' bias in $\bbetaHat$.
-To illustrate this concept, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
+Consider our earlier example involving confounding due to air quality.
+Simulating $\bgamma$ as a random quantity for each replication corresponds to subpopulation A being exposed to good air quality in one replication, and bad air quality in the next.
+Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from populations A and B, when these populations are subject to different environmental exposures?
+Those repeated samples would show a systematic bias due to the difference in air quality, but this is masked if $\bgamma$ is assumed to be random.
+Treating $\bgamma$ as a fixed quantity so that environmental effects are consistent across replications more closely approximates environmental confounding.
+
+To illustrate this concept mathematically, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
 \begin{align}
@@
 Under the assumption that environmental effects arise from a random process, $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
-Consider our earlier example involving confounding due to air quality \pb{I feel that this should be moved up to before the mathematical derivations}.
-Simulating $\bgamma$ as a random quantity for each replication corresponds to subpopulation A being exposed to good air quality in one replication, and bad air quality in the next.
-Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from populations A and B, when these populations are subject to different environmental exposures?
-Those repeated samples would show a systematic bias due to the difference in air quality, but this is masked if $\bgamma$ is assumed to be random.
-Treating $\bgamma$ as a fixed quantity so that environmental effects are consistent across replications more closely approximates environmental confounding.
-Under the assumption that the quantity $\bgamma$ is fixed, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, allowing us to quantify estimation bias due to environmental heterogeneity.
+However, under the assumption that the quantity $\bgamma$ is fixed, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, allowing us to quantify estimation bias due to environmental heterogeneity.
 
@@
 This result also allows us to mathematically justify the statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
-We can quantify the proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ using the ratio $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2 \in [0, 1]$.
+We can quantify the proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ \anna{and thus biases the $\bbeta$ estimates} using the ratio \anna{$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2 \in [0, 1]$}.
 In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
@@
 Although we focus our attention on accurately estimating fixed SNP effects, we note that this result could lead to substantial bias in narrow-sense heritability estimates.
+
 In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
-However, the degree to which it reflects population structure that can be inferred from $\bX$, as quantified by $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$, will impact the extent to which environmental heterogeneity confounds SNP-phenotype associations.
-\anna{this calculation is still having issues for the admixture and independent supbopulation data - the ratio is slightly larger for the independent data - as we would expect - but only on the order of 1e-15...
-correlation of Xtau with Zgamma?
-euclidian distance between vectors?}
+However, the degree to which it reflects population structure that can be inferred from $\bX$, as quantified by \anna{$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2$}, will impact the extent to which environmental heterogeneity \anna{has the potential to bias} SNP-phenotype associations.
+\anna{We will use this metric in Section \ref{sec:results}.
+Given the high-dimensional nature of our data, we will approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
+We use this approach as opposed to that of a generalized inverse in order to ensure a unique solution.}
 
@@
 Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture.
-Supplementary Figures \anna{?, ?, and ?} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.
+Supplementary Figures \anna{?, ?, and ?
+MAKE THESE AND MAKE THEM PRETTY} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.
 
@@
 This improvement is more substantial in the independent subpopulations simulated data than the admixed simulated data.
-\anna{quantify this for all 3 data settings using $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$} This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
+\anna{quantify this for all 3 data settings using $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$: Admixture ratio = 0.46412, Independent ratio = 0.98615, Vanderwas ratio = 0.64325} This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
 In other words, a greater amount of the environmental effect is in the column space of the projection matrix of $\bX$ for the independent subpopulations data than in the admixed data.
@@
 In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
-An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure ?}.
+An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure ?
+MAKE THIS AND MAKE IT PRETTY}.
 Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.

### 2020-09-08 main.tex
3.5 edits

@@
 This result also allows us to mathematically justify the statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
-We can quantify the proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ \anna{and thus biases the $\bbeta$ estimates} using the ratio \anna{$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2 \in [0, 1]$}.
+We can quantify the proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ using the ratio
+\begin{equation}
+    \label{eqn:ratio}
+    \anna{||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2 \in [0, 1]}.
+\end{equation} 
 In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
@@
 In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
-However, the degree to which it reflects population structure that can be inferred from $\bX$, as quantified by \anna{$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2$}, will impact the extent to which environmental heterogeneity \anna{has the potential to bias} SNP-phenotype associations.
-\anna{We will use this metric in Section \ref{sec:results}.
+However, the degree to which it reflects population structure \anna{and has the potential to bias SNP-phenotype associations can be inferred from $\bX$ and quantified by \anna{\eqref{eqn:ratio}}.
+We will use this metric in our analysis of our simulation results in Section \ref{sec:results}.
 Given the high-dimensional nature of our data, we will approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
@@
 \subsubsection{Design matrix generation}
-We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data}.
+We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
 $p = 1000$, $n = \{200, 400, 600, 800 \}$ for the simulated data types, and 50 causal SNPs were used throughout.
@@
 In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
-This improvement is more substantial in the independent subpopulations simulated data than the admixed simulated data.
-\anna{quantify this for all 3 data settings using $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$: Admixture ratio = 0.46412, Independent ratio = 0.98615, Vanderwas ratio = 0.64325} This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
-In other words, a greater amount of the environmental effect is in the column space of the projection matrix of $\bX$ for the independent subpopulations data than in the admixed data.
+
+\anna{Using the metric $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$}
+
+\anna{quantify this for all 3 data settings using $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$: Admixture ratio = 0.46412, Independent ratio = 0.98615, Vanderwas ratio = 0.64325}
+
+This improvement is more substantial in the semi-empirical and independent subpopulations data than the admixed data.
+This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
+In other words, a greater proportion of the environmental effect is in the column space of the projection matrix of $\bX$ for the semi-empirical and independent subpopulations data than in the admixed data.
+
 As such, there is a greater amount of confounding and potential estimation bias in the independent subpopulations which can be corrected by the PC-Lasso and LMM-Lasso methods.

### 2020-09-08 main.tex
4.1 edits

@@
 In addition, we briefly review historical methods of correcting for population stratification and relatedness.
-In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-Lasso) and lasso penalized multivariate LMMs (LMM-Lasso).
+In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted Lasso penalized regression (PC-Lasso) and Lasso penalized multivariate LMMs (LMM-Lasso).
 We review the statistical details of these methods as well as the consequences \pb{do we review \emph{consequences} in section 3?} \anna{tbd} of the ways in which they attempt to correct for population structure and environmental heterogeneity.
@@
 \subsubsection{Design matrix generation}
-We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from independent subpopulations, (2) simulated SNP data from admixed subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
-$p = 1000$, $n = \{200, 400, 600, 800 \}$ for the simulated data types, and 50 causal SNPs were used throughout.
+We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
+We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Semi-Empirical data, respectively.
+$p = 1000$ for all data types, $n = \{200, 400, 600, 800 \}$ for the 1D Linear Admixture and Independent Subpopulations data, and 50 causal SNPs were used for all data types.
 
@@
 We note that the fine and coarse subpopulation structures are distinct from the level of admixture in the simulated data.
-For example, the discrete nonoverlapping subpopulations in the simulated independent subpopulation data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated admixture data.
-The empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
+For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data.
+The Semi-Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
 Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture.
@@
 We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
-Estimation accuracy for lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components.
+Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
-The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-Lasso perform comparably.
-PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors have been added to the model, but none of them actually explain any variability in the outcome.
+The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), Lasso and LMM-Lasso perform comparably.
+PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
 The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
-In this case, both PC-Lasso and LMM-Lasso clearly outperform the naive lasso.
-
-\anna{Using the metric $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$}
-
-\anna{quantify this for all 3 data settings using $||\bX \boldsymbol{\tau}||_2 / ||\bZ \bgamma||_2$: Admixture ratio = 0.46412, Independent ratio = 0.98615, Vanderwas ratio = 0.64325}
-
-This improvement is more substantial in the semi-empirical and independent subpopulations data than the admixed data.
-This is due to the fact that the discrete, nonoverlapping genetic subpopulations more directly correspond to the distinct fixed environmental effects than do the more continuous and blended subpopulations in the admixture data.
-In other words, a greater proportion of the environmental effect is in the column space of the projection matrix of $\bX$ for the semi-empirical and independent subpopulations data than in the admixed data.
-
-As such, there is a greater amount of confounding and potential estimation bias in the independent subpopulations which can be corrected by the PC-Lasso and LMM-Lasso methods.
+In this case, both PC-Lasso and LMM-Lasso clearly outperform the ordinary Lasso.
+This improvement is more substantial in the Semi-Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
+\anna{Qualitatively, this is due to the fact that the Independent Subpopulations and Semi-Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the distinct fixed environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
+In other words, a greater proportion of the environmental effect is in the column space of the projection matrix of $\bX$ for the Semi-Empirical and Independent Subpopulations data than for the 1D Linear Admixture data.
+We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the ratio \eqref{eqn:ratio} introduced in Section \ref{sec:sim_env_conf}.
+$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2$, is equal to 0.98615, 0.64325, and 0.46412 for the Independent Subpopulations, Semi-Empirical, and 1D Linear Admixture data, respectively.
+As such there is a greater amount of confounding and estimation bias which the PC-Lasso and LMM-Lasso methods are able to correct for in the Independent Subpopulations, while there is less confounding which can be corrected for in the 1D Linear Admixture data.}
 
@@
 In addition to MSE, these methods' performance in terms of variable selection accuracy were quantified using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
-Table \ref{tab:var_sel} displays the mean true positive rate and corresponding standard deviation in the same settings as those illustrated in Figure \ref{fig:mse}.
-The TPRs echo the MSE results in that Lasso and LMM-Lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
+Table \ref{tab:var_sel} displays the mean TPR and corresponding standard deviation in the same settings as those illustrated in Figure \ref{fig:mse}.
+The TPRs echo the MSE results in that Lasso and LMM-Lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
 In the presence of environmental confounding, both LMM-Lasso and PC-Lasso achieve substantial improvements in TPR over Lasso across data types.
 As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Semi-Empirical data sets than in the 1D Linear Admixture data.
-Again, this is due to the reduced proportion of environmental confounding that is in the column space of $\boldsymbol{P}_{\bX}$, which results in a lesser degree of confounding of the SNP-phenotype associations.
+\anna{Again, this is due to the smaller proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ for the 1D Linear Admixture data, which results in less confounded SNP-phenotype associations to begin with.}
 
@@
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures.
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures.
 Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
 The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
@@
      \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
     \label{fig:big_vs_small}
@@
 Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure.
-Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the independent subpopulation simulated $\bX$.
+Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the Independent Subpopulations simulated $\bX$.
 The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with independent subpopulation data, $\eta = 0.5$, and $\xi = 0.8$}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$}
     \label{fig:big_small_gamma}
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with independent subpopulation data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
+     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
     \label{fig:eta_beta_mse}

### 2020-09-10 main.tex
1 tweak

@@
 In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted Lasso penalized regression (PC-Lasso) and Lasso penalized multivariate LMMs (LMM-Lasso).
-We review the statistical details of these methods as well as the consequences \pb{do we review \emph{consequences} in section 3?} \anna{tbd} of the ways in which they attempt to correct for population structure and environmental heterogeneity.
+We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
 In section \ref{sec:results}, we illustrate the concepts described in section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.

### 2020-09-10 main.tex
3.5 wording

@@
 We will use this metric in our analysis of our simulation results in Section \ref{sec:results}.
-Given the high-dimensional nature of our data, we will approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
-We use this approach as opposed to that of a generalized inverse in order to ensure a unique solution.}
+Given the high-dimensional nature of our data, we use a low-rank approximation of $(\bX^\T \bX)^{-1}$.
+We use this approach as opposed to a generalized inverse in order to ensure a unique solution.}
 

### 2020-09-10 main.tex
4.4 note
Edits to 3.6

@@
 
+\pb{How about ``Partitioning the effects of environmental confounding''}
+
 In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis \citep{Rakitsch2012, bhatnagar2019simultaneous}.
@@
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
+This result provides the explicit mathematical justification for our statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
 
-This result also allows us to mathematically justify the statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
-We can quantify the proportion of the environmental effect which is in the column space of $\boldsymbol{P}_{\bX}$ using the ratio
+Furthermore, the proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ (i.e., the portion that will cause confounding) is given by 
 \begin{equation}
     \label{eqn:ratio}
-    \anna{||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2 \in [0, 1]}.
+    \anna{||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2},
 \end{equation}
+\pb{should we have a symbol for this?}
+which can range from 0 (no confounding) to 1 (entirely confounding).
 In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
-Thus the SNP-phenotype associations will be unbiased, but the residual variance will be increased.
+In this scenario, the SNP-phenotype associations will be unbiased, but the residual variance will be increased.
 Although we focus our attention on accurately estimating fixed SNP effects, we note that this result could lead to substantial bias in narrow-sense heritability estimates.
+\pb{Hmm...could this sentence be moved somewhere else?
+I feel that narrow-sense heritability deserves a paragraph somewhere}
 
-In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
-However, the degree to which it reflects population structure \anna{and has the potential to bias SNP-phenotype associations can be inferred from $\bX$ and quantified by \anna{\eqref{eqn:ratio}}.
+\anna{In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
+However, the degree to which it reflects population structure and has the potential to bias SNP-phenotype associations can be inferred from $\bX$ and quantified by \eqref{eqn:ratio}.
 We will use this metric in our analysis of our simulation results in Section \ref{sec:results}.
@@
 
+  \pb{Need to mention that this ratio is difficult/impossible to estimate with real data, but can be calculated in simulations and is informative, etc.}
+In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
+However, the degree to which it reflects population structure \anna{and has the potential to bias SNP-phenotype associations can be inferred from $\bX$ and quantified by \anna{\eqref{eqn:ratio}}.
+We will use this metric in our analysis of our simulation results in Section \ref{sec:results}.
+Given the high-dimensional nature of our data, we will approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
+We use this approach as opposed to that of a generalized inverse in order to ensure a unique solution.}\pb{no need to mention generalized inverses}
+
 %------------------------------%
@@
 %------------------------------%
-\subsubsection{Pseudophenotype generation}
+
+\subsubsection{Generation of synthetic phenotypes}
+
+\pb{I prefer ``synthetic'' phenotype to pseudophenotype}
 
@@
 \subsubsection{Design matrix generation}
+
+\pb{Let's not call it a ``design matrix''}
+
+\pb{Also, we're generating $G$ matrices, correct?  And then obtaining $X$ from standardization?  Wouldn't hurt to remind people of that here}
+
+\pb{Vanderwas citation: https://doi.org/10.1186/s12890-015-0040-0}
+
 We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
 We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Semi-Empirical data, respectively.
-$p = 1000$ for all data types, $n = \{200, 400, 600, 800 \}$ for the 1D Linear Admixture and Independent Subpopulations data, and 50 causal SNPs were used for all data types.
+\pb{Don't start sentences with mathematical symbols} $p = 1000$ for all data types, $n = \{200, 400, 600, 800 \}$ for the 1D Linear Admixture and Independent Subpopulations data, and 50 causal SNPs were used for all data types.
+
+\pb{Can't we just call the last one ``Empirical''?}
 
 \subsubsection{Subpopulation structure}
-For the fully simulated data, two subpopulation structures were considered: (1) fine and (2) coarse.
+For the fully simulated data, two subpopulation structures were considered, which we refer to as ``fine'' and ``coarse''.
 We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example.
-Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, or founding populations, which corresponds to a finer grained continuum of environmental heterogeneity.
-Whether fine or coarse, the simulated data were generated such that subpopulations, or founding populations, are of equal size.
-We note that the fine and coarse subpopulation structures are distinct from the level of admixture in the simulated data.
-For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data.
+Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, representing a more subtle continuum of environmental heterogeneity \pb{do you know, are the subpopulations more similar in the fine case?}.
+Whether fine or coarse, the simulated data were generated such that subpopulations (also known as founding populations) are of equal size.
+We note that the distinction between fine and coarse structure is an entirely different concept than the level of admixture in the simulated data.
+For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data \pb{?????}.
 The Semi-Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
@@
 \subsubsection{Structure of environmental effects}
-For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$.
+
+\pb{Perhaps ``Patterns of environmental effects'' would be better?}
+
+For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership \pb{you don't need to keep saying ``or founding population'' over and over again}, which assigns the corresponding environmental effect found in $\bgamma$.
 Various $\bgamma$ structures were considered in order to investigate the effects of various forms of environmental confounding.
 
- Our initial hypothesis was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that are nonlinear in nature and as such do not lend themselves to such an approximation.
-To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous.
-In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
+Our initial hypothesis was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that are nonlinear in nature and as such do not lend themselves to such an approximation \pb{is the ``linearity'' the key issue here?
+I'd say it's more a matter of complexity: PC-Lasso would work for simple patterns, but can't represent more complex patterns}.
+To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous \pb{neither of which are linear, so this doesn't transition well}.
+In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects \pb{Shouldn't we lead with this?
+I mean, isn't this the most important concept in this section?}.
 For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
@@
  To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are concordant or discordant in nature, for a total of four distinct $\bgamma$ structures.
-We define concordant effects as those that are similar among genetically similar subpopulations, and discordant effects as those that are not.
+We define concordant environmental effects as those that are similar among genetically similar subpopulations, and discordant environmental effects as those that are not.
 Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient structures.
 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
+\pb{need to describe somehow that subpopulations are ordered with respect to how similar they are?}
 \begin{figure}[H]
@@
     \includegraphics[scale = 0.9]{figures/gamma_structure.png}
-    \caption{$\bgamma$ structures}
+    \caption{$\bgamma$ structures \pb{probably want a better caption}}
     \label{fig:gamma_structures}
@@
 
-\subsubsection{$\lambda$ selection}
-The $\lambda$ that allowed for the first 50 variables to enter the model was selected for each method, where 50 was the true number of causal SNPs.
+\subsubsection{Tuning parameter selection}
+Selection of the tuning parameter $\lambda$ plays an important role in all penalized regression methods, with many possible approaches.
+Here, for each model, we selected the $\lam$ value such that 50 SNPs were selected (recall that 50 is also the true number of causal SNPs).
 This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.
 Alternative $\lambda$ selection methods were also considered.
-Cross-validation has been found to be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here.
+\pb{This seems way too strong}  Cross-validation has been found to be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here.
 While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
@@
 
-\anna{re-create niter\_parLambda fig focusing on the eta = xi = 0.8 seting and displaying all data points - not just averages}
+\anna{See figure options: niter\_parLambda fig focusing on the eta = xi = 0.8 seting and displaying all data points - not just averages}
 

### 2020-09-10 main.tex
2020-09-10 meeting notes

@@
 \pb{Hmm...could this sentence be moved somewhere else?
-I feel that narrow-sense heritability deserves a paragraph somewhere}
+I feel that narrow-sense heritability deserves a paragraph somewhere} \anna{move to discussion do a paragraph on narrow-sense heritability...
+reference this eq - and give it a label}
 
@@
 We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example.
-Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, representing a more subtle continuum of environmental heterogeneity \pb{do you know, are the subpopulations more similar in the fine case?}.
+Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, representing a continuum of environmental heterogeneity.
 Whether fine or coarse, the simulated data were generated such that subpopulations (also known as founding populations) are of equal size.
 We note that the distinction between fine and coarse structure is an entirely different concept than the level of admixture in the simulated data.
-For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data \pb{?????}.
+\anna{size of the subpopulations that determines fine/course - cut into 2 sentences.
+similarly for the 1d amixture...} For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data \pb{?????}.
 The Semi-Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
@@
 
+\anna{complex/noncomplex ... can be complex in different ways... discordant vs concordant ... dichotomous vs exponential}
+
  To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are concordant or discordant in nature, for a total of four distinct $\bgamma$ structures.
@@
 This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.
-Alternative $\lambda$ selection methods were also considered.
-\pb{This seems way too strong}  Cross-validation has been found to be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here.
+\anna{make this sentence more getting at the point that we had concerns about using these common lambda selection methods...
+Although CV and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons across methods.} Alternative $\lambda$ selection methods were also considered.
+\anna{I know i have a citation for this somewhere FIND IT} Cross-validation may be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here.
 While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
@@
 As such there is a greater amount of confounding and estimation bias which the PC-Lasso and LMM-Lasso methods are able to correct for in the Independent Subpopulations, while there is less confounding which can be corrected for in the 1D Linear Admixture data.}
+
+\anna{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes? at least to pin down the issue}
 

### 2020-09-21 main.tex
3.5 edits

@@
 %------------------------------%
-\subsection{\anna{Decomposing (Quantifying?) the effect of environmental confounding}}
+\subsection{Partitioning the effects of environmental confounding}
 \label{sec:sim_env_conf}
-
-\pb{How about ``Partitioning the effects of environmental confounding''}
 
@@
 This in turn means that
-\begin{align*}
-    \by &= \bX\bbeta + \bZ\bgamma + \beps\\
-    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps\\
+\begin{align}
+    \label{eqn:partition_e}    \by &= \bX\bbeta + \bZ\bgamma + \beps \notag \\
+    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps \notag\\
     &=  \bX \bbeta{^*} + \beps{^*};
-\end{align*}
+\end{align}
 note that $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term.
@@
     \label{eqn:ratio}
-    \anna{||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2},
+    ||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2,
 \end{equation}
-\pb{should we have a symbol for this?}
-which can range from 0 (no confounding) to 1 (entirely confounding).
-In the case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
+which we refer to as the Confounding Proportion (CP).
+CP can range from 0 (no confounding) to 1 (entirely confounded).
+In the extreme case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
 In this scenario, the SNP-phenotype associations will be unbiased, but the residual variance will be increased.
-Although we focus our attention on accurately estimating fixed SNP effects, we note that this result could lead to substantial bias in narrow-sense heritability estimates.
-\pb{Hmm...could this sentence be moved somewhere else?
-I feel that narrow-sense heritability deserves a paragraph somewhere} \anna{move to discussion do a paragraph on narrow-sense heritability...
-reference this eq - and give it a label}
 
-\anna{In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
-However, the degree to which it reflects population structure and has the potential to bias SNP-phenotype associations can be inferred from $\bX$ and quantified by \eqref{eqn:ratio}.
-We will use this metric in our analysis of our simulation results in Section \ref{sec:results}.
-Given the high-dimensional nature of our data, we use a low-rank approximation of $(\bX^\T \bX)^{-1}$.
-We use this approach as opposed to a generalized inverse in order to ensure a unique solution.}
-
-  \pb{Need to mention that this ratio is difficult/impossible to estimate with real data, but can be calculated in simulations and is informative, etc.}
-In practical applications $\bZ \bgamma$ is unlikely to be entirely orthogonal to population structure.
-However, the degree to which it reflects population structure \anna{and has the potential to bias SNP-phenotype associations can be inferred from $\bX$ and quantified by \anna{\eqref{eqn:ratio}}.
-We will use this metric in our analysis of our simulation results in Section \ref{sec:results}.
-Given the high-dimensional nature of our data, we will approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
-We use this approach as opposed to that of a generalized inverse in order to ensure a unique solution.}\pb{no need to mention generalized inverses}
+In real data settings CP cannot be calculated since $\bZ \bgamma$ is not directly observed.
+In simulation settings, however, this ratio is a useful metric for quantifying the degree to which $\bZ \bgamma$ reflects population structure and has the potential to bias SNP-phenotype associations.
+We will use CP in our interpretation of the simulation results in Section \ref{sec:results}.
+Given the high-dimensional nature of our data, we approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
+We use this approach in order to ensure a unique solution.
 
@@
 Finally, it is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
-Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.
+Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.\\
+
+\anna{HERITABILITY PARAGRAPH}\\
+Although we focus our attention on accurately estimating fixed SNP effects, we note that the result shown in \ref{eqn:partition_e} could lead to substantial bias in narrow-sense heritability estimates.
+\pb{Hmm...could this sentence be moved somewhere else?
+I feel that narrow-sense heritability deserves a paragraph somewhere}

### 2020-09-21 main.tex
3.6 edits

@@
 
-\pb{I prefer ``synthetic'' phenotype to pseudophenotype}
-
 Given the nature of environmental confounding in our setting of interest, and in order to assess methodological performance in terms of estimation accuracy and bias, we simulate from the data-generating model
@@
 
-\subsubsection{Design matrix generation}
-
-\pb{Let's not call it a ``design matrix''}
-
-\pb{Also, we're generating $G$ matrices, correct?  And then obtaining $X$ from standardization?  Wouldn't hurt to remind people of that here}
+\subsubsection{SNP matrix generation}
 
@@
 
-We generated these pseudophenotypes using three types of high-dimensional ($p > n$) $\bX$ matrices: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
-We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Semi-Empirical data, respectively.
-\pb{Don't start sentences with mathematical symbols} $p = 1000$ for all data types, $n = \{200, 400, 600, 800 \}$ for the 1D Linear Admixture and Independent Subpopulations data, and 50 causal SNPs were used for all data types.
-
-\pb{Can't we just call the last one ``Empirical''?}
+We generated these synthetic phenotypes using three types of high-dimensional ($p > n$) $\bG$ matrices, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
+We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
+For all data types $p = 1000$ with 50 causal SNPs.
+For the simulated 1D Linear Admixture and Independent Subpopulations data $n = \{200, 400, 600, 800 \}$, and $n = 588$ for the Empirical data.
 
@@
 similarly for the 1d amixture...} For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data \pb{?????}.
-The Semi-Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
+The Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
 Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture.
@@
 In this case, both PC-Lasso and LMM-Lasso clearly outperform the ordinary Lasso.
-This improvement is more substantial in the Semi-Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
-\anna{Qualitatively, this is due to the fact that the Independent Subpopulations and Semi-Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the distinct fixed environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
-In other words, a greater proportion of the environmental effect is in the column space of the projection matrix of $\bX$ for the Semi-Empirical and Independent Subpopulations data than for the 1D Linear Admixture data.
+This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
+\anna{Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the distinct fixed environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
+In other words, a greater proportion of the environmental effect is in the column space of the projection matrix of $\bX$ for the Empirical and Independent Subpopulations data than for the 1D Linear Admixture data.
 We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the ratio \eqref{eqn:ratio} introduced in Section \ref{sec:sim_env_conf}.
-$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2$, is equal to 0.98615, 0.64325, and 0.46412 for the Independent Subpopulations, Semi-Empirical, and 1D Linear Admixture data, respectively.
+$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2$, is equal to 0.98615, 0.64325, and 0.46412 for the Independent Subpopulations, Empirical, and 1D Linear Admixture data, respectively.
 As such there is a greater amount of confounding and estimation bias which the PC-Lasso and LMM-Lasso methods are able to correct for in the Independent Subpopulations, while there is less confounding which can be corrected for in the 1D Linear Admixture data.}
@@
 In the presence of environmental confounding, both LMM-Lasso and PC-Lasso achieve substantial improvements in TPR over Lasso across data types.
-As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Semi-Empirical data sets than in the 1D Linear Admixture data.
+As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Empirical data sets than in the 1D Linear Admixture data.
 \anna{Again, this is due to the smaller proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ for the 1D Linear Admixture data, which results in less confounded SNP-phenotype associations to begin with.}
@@
 \toprule
-$\eta$ & $\xi$ & \multicolumn{1}{c}{Method} & 1D Linear Admixture & Independent Subpopulations & Semi-Empirical \\ 
+$\eta$ & $\xi$ & \multicolumn{1}{c}{Method} & 1D Linear Admixture & Independent Subpopulations & Empirical \\ 
 \midrule

### 2020-09-21 main.tex
3.6.2 edits

@@
 
-\pb{Vanderwas citation: https://doi.org/10.1186/s12890-015-0040-0}
-
-We generated these synthetic phenotypes using three types of high-dimensional ($p > n$) $\bG$ matrices, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \anna{cite Vanderwas data PATRICK TO DO}.
+We generated these synthetic phenotypes using three types of high-dimensional ($p > n$) $\bG$ matrices, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
 We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.

### 2020-09-22 main.tex
3.6.3 edits

@@
 which we refer to as the Confounding Proportion (CP).
-CP can range from 0 (no confounding) to 1 (entirely confounded).
+The CP can range from 0 (no confounding) to 1 (entirely confounded).
 In the extreme case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
-In this scenario, the SNP-phenotype associations will be unbiased, but the residual variance will be increased.
+In this scenario, the SNP-phenotype associations will be unbiased, but the residual variance will be inflated.
 
@@
 We note that the distinction between fine and coarse structure is an entirely different concept than the level of admixture in the simulated data.
-\anna{size of the subpopulations that determines fine/course - cut into 2 sentences.
-similarly for the 1d amixture...} For example, the discrete nonoverlapping subpopulations in the simulated Independent Subpopulations data result in an RRM which more directly corresponds to the discrete and fixed environmental effects, fine or coarse, than do the continuous and overlapping subpopulations in the simulated 1D Linear Admixture data \pb{?????}.
+It is the size of the subpopulations, not the level of admixture, that determines what we refer to as fine or coarse structure.
+For example, both the Independent Subpopulation and 1D Linear Admixture simulated data may be comprised of fine or coarse structure.
+However, the subpopulations will be more distinct in the unadmixed Independent Subpopulation data than in the 1D Linear Admixed data.
 The Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
@@
 
-\anna{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes? at least to pin down the issue}
+\anna{\textbf{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes?
+at least to pin down the issue - more comparable with same sampesize and subpop sizes hmm}}
 

### 2020-09-23 main.tex
3.6.4 edits

@@
 
-\subsubsection{Structure of environmental effects}
+\subsubsection{Patterns of environmental effects}
 
-\pb{Perhaps ``Patterns of environmental effects'' would be better?}
+For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$.
+Various $\bgamma$ patterns were considered in order to investigate the effects of distinct forms of environmental confounding.
+Our initial hypothesis was that simple patterns of environmental effects could be well-approximated by a linear combination of the observed genetic data and well-corrected for by both PC-Lasso and LMM-Lasso, while PC-Lasso would be unable to represent and correct for more complex patterns of environmental effects.
 
-For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership \pb{you don't need to keep saying ``or founding population'' over and over again}, which assigns the corresponding environmental effect found in $\bgamma$.
-Various $\bgamma$ structures were considered in order to investigate the effects of various forms of environmental confounding.
-
-Our initial hypothesis was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that are nonlinear in nature and as such do not lend themselves to such an approximation \pb{is the ``linearity'' the key issue here?
-I'd say it's more a matter of complexity: PC-Lasso would work for simple patterns, but can't represent more complex patterns}.
-To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous \pb{neither of which are linear, so this doesn't transition well}.
-In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects \pb{Shouldn't we lead with this?
-I mean, isn't this the most important concept in this section?}.
+Environmental effects may be complex in different ways.
+Firstly, we considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
 For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
@@
 As an example, \citet{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
+We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.
+Secondly, we considered two patterns of environmental effect sizes, which we refer to as exponential and dichotomous.
+We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations.
+While exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes.
 
-\anna{complex/noncomplex ... can be complex in different ways... discordant vs concordant ... dichotomous vs exponential}
-
- To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are concordant or discordant in nature, for a total of four distinct $\bgamma$ structures.
-We define concordant environmental effects as those that are similar among genetically similar subpopulations, and discordant environmental effects as those that are not.
-Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient structures.
-1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
-\pb{need to describe somehow that subpopulations are ordered with respect to how similar they are?}
+ To evaluate the performance of penalized LMM methods in the presence of these different forms of confounding complexity, we consider environmental effects which are concordant or discordant in nature in the presence of both exponential and dichotomous effect sizes, for a total of four distinct $\bgamma$ patterns.
+Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient patterns, where the subpopulations on the x-axis are ordered by genetic similarity so that subpopulation 1 is most genetically similar to subpopulation 2, and least similar to subpopulation 10.
+This may also be thought of as subpopulations ordered spatially along a 1-dimensional geographic region.
 \begin{figure}[H]
@@
     \includegraphics[scale = 0.9]{figures/gamma_structure.png}
-    \caption{$\bgamma$ structures \pb{probably want a better caption}}
+    \caption{An illustration of four distinct patterns of environmental effects and corresponding $\bgamma$ values in the presence of 10 subpopulations ordered by genetic similarity.}
     \label{fig:gamma_structures}
 \end{figure}
+\noindent In the case of concordant and dichotomous effects, Figure \ref{fig:gamma_structures} shows that the first five ordered subpopulations are subject to an effect size of 1, while the second five are subject to an effect size of $-$1.
+In the case of discordant dichotomous effects, the first five subpopulations experience an effect size with an absolute value of \anna{1}, and the second five an absolute effect size of \anna{0.5}.
+However, the directions of these absolute effects sizes alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects.
+\anna{not quite 1 and 0.5 because of centering and standardization...how to word this so it's not confusing?}
 
@@
 In addition, because Lasso, PC-Lasso, and LMM-Lasso use different likelihoods, an information-criterion-based approach would not lend itself to comparable results across methods.
+
+1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
 

### 2020-09-23 main.tex
4.1

@@
 PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
+
 The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
@@
 This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
-\anna{Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the distinct fixed environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
-In other words, a greater proportion of the environmental effect is in the column space of the projection matrix of $\bX$ for the Empirical and Independent Subpopulations data than for the 1D Linear Admixture data.
-We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the ratio \eqref{eqn:ratio} introduced in Section \ref{sec:sim_env_conf}.
-$||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2$, is equal to 0.98615, 0.64325, and 0.46412 for the Independent Subpopulations, Empirical, and 1D Linear Admixture data, respectively.
-As such there is a greater amount of confounding and estimation bias which the PC-Lasso and LMM-Lasso methods are able to correct for in the Independent Subpopulations, while there is less confounding which can be corrected for in the 1D Linear Admixture data.}
+
+Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the nonoverlapping environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
+In other words, a greater proportion of the environmental effect is in the column space of $\boldsymbol{P}_{\bX}$ for the Empirical and Independent Subpopulations data than for the 1D Linear Admixture data.
+We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}.
+CP $= 0.46, 0.99, 0.64$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
+The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
+Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-Lasso and LMM-Lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.
 
 \anna{\textbf{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes?
-at least to pin down the issue - more comparable with same sampesize and subpop sizes hmm}}
+at least to pin down the issue - more comparable with same sampesize and subpop sizes but still different...}}
+
+\anna{with the same sample size and subpopulation structure as the vanderwas data, I can replicate the CP for the indepenent subpopulations data when simulating using a lower Fst value (as defined in the ochoa and storey papers: the mean inbreeding coefficient of the individuals’ local populations relative to their last common ancestral population.
+This FST naturally yields a useful pairwise FST between individuals.) So, this suggests that the greater improvement in the semi-empirical data is because there is less cryptic relatedness/inbreeding (?????) still thinking about this...}
 
@@
 As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Empirical data sets than in the 1D Linear Admixture data.
-\anna{Again, this is due to the smaller proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ for the 1D Linear Admixture data, which results in less confounded SNP-phenotype associations to begin with.}
+Again, this is due to the smaller proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ for the 1D Linear Admixture data, which results in less confounded SNP-phenotype associations to begin with.
 

### 2020-09-23 main.tex
4.4 - note to self

@@
 In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
-An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure ?
-MAKE THIS AND MAKE IT PRETTY}.
+An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure - in progress, stuck on dumb facet labeling issue}.
 Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.
-
-\anna{See figure options: niter\_parLambda fig focusing on the eta = xi = 0.8 seting and displaying all data points - not just averages}
 

### 2020-09-24 main.tex
3.5 small tweaks

@@
 The CP can range from 0 (no confounding) to 1 (entirely confounded).
-In the extreme case where $\bZ \bgamma$ is entirely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
-In this scenario, the SNP-phenotype associations will be unbiased, but the residual variance will be inflated.
+In the extreme case where $\bZ \bgamma$ is completely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
+In this scenario, the SNP-phenotype associations will be unbiased, but the residual error will be inflated.
 
@@
 $\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$.
-$\eta$ also describes the data-generating model's signal to noise ratio (SNR).
-$\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
+The parameter $\eta$ also describes the data-generating model's signal to noise ratio (SNR), while $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
 So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to unobserved environmental effects.

### 2020-09-24 main.tex
3.5 more wording tweaks

@@
 We will use CP in our interpretation of the simulation results in Section \ref{sec:results}.
-Given the high-dimensional nature of our data, we approximate $(\bX^\T \bX)^{-1}$ using a low-rank approximation.
-We use this approach in order to ensure a unique solution.
+Given the high-dimensional nature of our data, we use a low-rank approximation of $(\bX^\T \bX)^{-1}$ in order to ensure a unique solution.
 

### 2020-09-24 main.tex
3.6.5 edits - in progress

@@
 This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.
-\anna{make this sentence more getting at the point that we had concerns about using these common lambda selection methods...
-Although CV and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons across methods.} Alternative $\lambda$ selection methods were also considered.
-\anna{I know i have a citation for this somewhere FIND IT} Cross-validation may be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here.
+
+Although cross-validation and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons across methods.
+Cross-validation may be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here \anna{I know i have a citation for this somewhere FIND IT}.
 While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
@@
 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
+\anna{somewhere else to put this that makes more sense?}
 

### 2020-09-24 main.tex
Small wording/punctuation tweaks throughout

@@
 The remainder of this paper is organized as follows.
-In section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
+In Section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
 We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures \citep{Sillanpaeae2011, sul2018population, vilhjalmsson2012nature, barton2019population}.
@@
 In addition, we briefly review historical methods of correcting for population stratification and relatedness.
-In section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted Lasso penalized regression (PC-Lasso) and Lasso penalized multivariate LMMs (LMM-Lasso).
+In Section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted Lasso penalized regression (PC-Lasso) and Lasso penalized multivariate LMMs (LMM-Lasso).
 We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
-In section \ref{sec:results}, we illustrate the concepts described in section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
+In Section \ref{sec:results}, we illustrate the concepts described in Section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
 In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
@@
 Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
-An important concept to note, and one that we will return to in section \ref{sec:sim_env_conf}, is that population structure only leads to confounding if it is correlated with the environment.
+An important concept to note, and one that we will return to in Section \ref{sec:sim_env_conf}, is that population structure only leads to confounding if it is correlated with the environment.
 In the presence of environmental heterogeneity which affects the phenotype of interest in a manner that is orthogonal to population structure, SNP-phenotype associations will not be confounded.
@@
 In such a scenario, failing to account for unobserved environmental effects is likely to result in greater variability, but SNP effect estimates will remain unbiased \citep{greenland1999causal}.
-We will return to this idea and its mathematical justification in section \ref{sec:methods}.
+We will return to this idea and its mathematical justification in Section \ref{sec:methods}.
 
@@
 This iterative update scheme prevents the use of standard Lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
-We further discuss differences between these methods in section \ref{sec:results}.
+We further discuss differences between these methods in Section \ref{sec:results}.
 
@@
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
-This result provides the explicit mathematical justification for our statement from section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
+This result provides the explicit mathematical justification for our statement from Section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
 
@@
 \end{equation}
-where $\by$ represents the pseudophenotype, $\bgamma$ is a fixed quantity (see section~\ref{sec:sim_env_conf}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
+where $\by$ represents the synthetic phenotype, $\bgamma$ is a fixed quantity (see Section~\ref{sec:sim_env_conf}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
 \begin{equation}
@@
 
+Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is more complex...
+
 \anna{\textbf{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes?
@@
 \anna{with the same sample size and subpopulation structure as the vanderwas data, I can replicate the CP for the indepenent subpopulations data when simulating using a lower Fst value (as defined in the ochoa and storey papers: the mean inbreeding coefficient of the individuals’ local populations relative to their last common ancestral population.
-This FST naturally yields a useful pairwise FST between individuals.) So, this suggests that the greater improvement in the semi-empirical data is because there is less cryptic relatedness/inbreeding (?????) still thinking about this...}
+This FST naturally yields a useful pairwise FST between individuals.) So, this suggests that the greater improvement in the semi-empirical data is because there is less inbreeding (?????) still thinking about this...}
 
@@
 Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\bgamma$ structures.
 The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
@@
 
-In section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+In Section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
 Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection.
@@
 In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
-An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure - in progress, stuck on dumb facet labeling issue}.
+An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure - in progress, stuck on dumb facet labeling with a factor issue}.
 Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.

### 2020-09-24 main.tex
20200924 meeting notes

@@
     \label{eqn:ratio}
-    ||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2,
+    \text{CP} = ||\bX \boldsymbol{\tau}||_2^2 / ||\bZ \bgamma||_2^2,
 \end{equation}
@@
 However, the directions of these absolute effects sizes alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects.
-\anna{not quite 1 and 0.5 because of centering and standardization...how to word this so it's not confusing?}
+\anna{not quite 1 and 0.5 because of centering and standardization...how to word this so it's not confusing?
+- trying better version of dichotomous-discordant...}
 
@@
 Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is more complex...
+this increases with equal subpopulation sizes
 
@@
 at least to pin down the issue - more comparable with same sampesize and subpop sizes but still different...}}
+
+ \anna{correlation between columns of X..., mspe vs mse- for various reasons CP is a reasonable measure but not perfect...}
 

### 2020-09-28 main.tex
2.1 edits - reorganizing a few paragraphs, correcting typos

@@
 
-Population structure is a common phenomenon in genetic data.
-Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations.
-For example, European American, Han Chinese, and most recently, cohorts within the UK Biobank data have been shown to exhibit patterns of population and geographic structure despiet their seemingly similar subjects \citep{campbell2005demonstrating, xu2009genomic, chen2009genetic, haworth2019apparent}.
-
 Population structure is defined by the existence of allele frequency differences that define sub-populations and is driven by the combined effects of evolutionary processes such as genetic drift, migration, demographic history, and natural selection \citep{gibson2015primer, tibayrenc2017genetics}.
@@
 Pedigree-based methods may be used to explicitly model recent relatedness if familial relationships are known.
-In this absence of known familial relationships, recent relatedness is referred to as \textbf{cryptic relatedness.} In this review we focus on methods to account for unobserved relatedness such as population stratification and cryptic relatedness.
+In the absence of known familial relationships, recent relatedness is referred to as \textbf{cryptic relatedness.}
 
+Population structure is a common phenomenon in genetic data.
+Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations.
+For example, European American, Han Chinese, and most recently, cohorts within the UK Biobank data have been shown to exhibit patterns of population and geographic structure despite their seemingly similar subjects \citep{campbell2005demonstrating, xu2009genomic, chen2009genetic, haworth2019apparent}.
+
+In this review we focus on methods to account for unobserved relatedness such as population stratification and cryptic relatedness.
 Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are present.

### 2020-09-29 main.tex
3.6.5 - citing CV  maybe inadvisable for LMMs

@@
 Although cross-validation and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons across methods.
-Cross-validation may be inadvisable for LMMs, and in the presence of non-inedpendent samples as is the case here \anna{I know i have a citation for this somewhere FIND IT}.
+Cross-validation may be inadvisable for LMMs, and in the presence of samples with underlying structure, as is the case here \citep{roberts2017cross}.
 While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
@@
 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
-\anna{somewhere else to put this that makes more sense?}
 

### 2020-09-29 main.tex
Updating paper figs (dichotomous-discordant change) and descriptions

@@
 \noindent In the case of concordant and dichotomous effects, Figure \ref{fig:gamma_structures} shows that the first five ordered subpopulations are subject to an effect size of 1, while the second five are subject to an effect size of $-$1.
-In the case of discordant dichotomous effects, the first five subpopulations experience an effect size with an absolute value of \anna{1}, and the second five an absolute effect size of \anna{0.5}.
-However, the directions of these absolute effects sizes alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects.
-\anna{not quite 1 and 0.5 because of centering and standardization...how to word this so it's not confusing?
-- trying better version of dichotomous-discordant...}
+In the case of discordant dichotomous effects, all subpopulations again experience an effect size with an absolute value of 1, however, the directions of these effects alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects.
 
@@
 The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), Lasso and LMM-Lasso perform comparably.
-PC-Lasso results in a slight inflation in MSE compared to the other methods due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
+PC-Lasso results in a slight inflation in MSE compared to the other methods in the simulated data sets due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
 
@@
 This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
-
 Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the nonoverlapping environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
@@
 We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}.
-CP $= 0.46, 0.99, 0.64$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
+CP $= 0.47, 0.99, 0.61$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
 The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
@@
 
-Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is more complex...
-this increases with equal subpopulation sizes
+Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is not straightforward.
+We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as in the Empirical data.
 
-\anna{\textbf{check this value for sampsizes of 588 for admixture and independent...maybe even diff subpop sizes?
-at least to pin down the issue - more comparable with same sampesize and subpop sizes but still different...}}
-
- \anna{correlation between columns of X..., mspe vs mse- for various reasons CP is a reasonable measure but not perfect...}
-
-\anna{with the same sample size and subpopulation structure as the vanderwas data, I can replicate the CP for the indepenent subpopulations data when simulating using a lower Fst value (as defined in the ochoa and storey papers: the mean inbreeding coefficient of the individuals’ local populations relative to their last common ancestral population.
-This FST naturally yields a useful pairwise FST between individuals.) So, this suggests that the greater improvement in the semi-empirical data is because there is less inbreeding (?????) still thinking about this...}
+% \anna{MSPE here (?)}
 
@@
 \cmidrule{2-6}
-& \multirow{3}{*}{0.8} & Lasso & 0.23 (0.05) & 0.25 (0.05) & 0.29 (0.06) \\ 
-& & PC-Lasso & 0.28 (0.05) & 0.36 (0.06) & 0.47 (0.06) \\ 
-& & LMM-Lasso & 0.29 (0.05) & 0.39 (0.06) & 0.58 (0.06) \\ 
+& \multirow{3}{*}{0.8} & Lasso & 0.24 (0.05) & 0.26 (0.06) & 0.29 (0.06) \\
+& & PC-Lasso & 0.28 (0.05) & 0.36 (0.06) & 0.46 (0.06) \\ 
+& & LMM-Lasso & 0.28 (0.05) & 0.39 (0.06) & 0.57 (0.05) \\
 \bottomrule
@@
 Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\bgamma$ structures.
+
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bG$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\bgamma$ structures.
 The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
@@
 In the presence of coarse population structure, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding boxes straddling the 0 line.
-However, the majority of these boxes fall above 0, particularly for lower $\eta$ values, indicating the superior performance of LMM-Lasso.
-In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy.
+However, the majority of these boxes fall above 0, indicating the superior performance of LMM-Lasso.
+In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy across all levels of environmental confounding.
 This is consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
@@
 Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure.
-Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the Independent Subpopulations simulated $\bX$.
+Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the Independent Subpopulations simulated $\bG$.
+
 The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
@@
 The lower left hand panel of Figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients.
-In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while PC-Lasso outperforms LMM-Lasso in the majority of dichotomous-concordant environmental effects.
-These findings are in line with our initial hypothesis that PC-Lasso will perform well in the presence of environmental effects that can be well-approximated by a linear combination of the observed data.
-When that is not the case, either due to the presence of fine subpopulation structure, nonlinear environmental effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.
+In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while LMM-Lasso outperforms PC-Lasso in the majority of dichotomous-concordant environmental effects.
+These findings align with our initial hypothesis that PC-Lasso will perform well in the presence of simple environmental effect patterns that can be well-approximated by a linear combination of the observed data.
+When that is not the case, either due to the presence of fine subpopulation structure, dichotomous effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.
 
@@
 In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
-We then compare the estimation accuracy of an ordinary Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
+We then compare the estimation and variable selection accuracy of the ordinary Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
 
@@
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
-Additionally, we find that LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure and when confounding effects are nonlinear and discordant in nature.
-In the presence of coarse subpopulation structure which corresponds to concordant, dichotomous environmental effects, PC-Lasso may outperform LMM-Lasso in estimation accuracy.
-However, LMM-Lasso is more robust to subpopulation structure as well as nonlinear and discordant environmental effects.
-Therefore, LMM-Lasso is a more robust method for practitioners seeking to obtain unbiased estimates when the exact nature of the population structure and corresponding environmental heterogeneity are unknown.
+Additionally, we find that LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure and when confounding effects are discordant in nature.
+In the presence of coarse subpopulation structure and exponential environmental effects, PC-Lasso performs comparably to LMM-Lasso in estimation accuracy.
+However, LMM-Lasso is more robust to subpopulation structure and complex patterns of environmental effects.
+Therefore, LMM-Lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure and environmental heterogeneity are unknown.
 
@@
 The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n << p$ settings.
-However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus considered by the authors to be more broadly applicable.
+However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and is considered by the authors to be more broadly applicable.
 
-\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-Lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
+\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
 
-As an additional point of interest, recall that we provide qualitative and mathematical explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure.
-While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, it also highlights the manner in which such unobserved environmental effects have the potential to inflate residual model variance and bias narrow-sense heritability estimates.
+As an additional point of interest, recall that we provide qualitative and quantitative explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure.
+While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, the result shown in \ref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimatess.
 A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper.
@@
 Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.\\
-
-\anna{HERITABILITY PARAGRAPH}\\
-Although we focus our attention on accurately estimating fixed SNP effects, we note that the result shown in \ref{eqn:partition_e} could lead to substantial bias in narrow-sense heritability estimates.
-\pb{Hmm...could this sentence be moved somewhere else?
-I feel that narrow-sense heritability deserves a paragraph somewhere}

### 2020-09-29 main.tex
Supplement and S4 (ggmix vs lmmlasso)

@@
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$.}
     \label{fig:big_small_gamma}
@@
 In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
-An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings \anna{Supplementary figure - in progress, stuck on dumb facet labeling with a factor issue}.
+An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:s4}).
 Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.

### 2020-09-29 supplement.tex
Supplement and S4 (ggmix vs lmmlasso)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Supplementary material}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.45]{figures/niter_perLambda_point.png}
    \caption{The number of iterations per $\lambda$ value for LMM-Lasso-Rakitsch and LMM-Lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure. Note that 100 is the maximum number of iterations for both implementations of LMM-Lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
    \label{fig:s4}
\end{figure}

### 2020-09-29 main.tex
Supplementary figures added and incorporated into fig refs

@@
 Thus it represents a hybrid of our simulated coarse and fine structures, with little apparent admixture.
-Supplementary Figures \anna{?, ?, and ?
-MAKE THESE AND MAKE THEM PRETTY} illustrate the simulated kinship structures in the fine and coarse subpopulation settings, as well as that present in the empirical SNP data.
+Supplementary Figures \ref{fig:admixed} - \ref{fig:empirical} illustrate the simulated kinship structures, as well as that for the Empirical SNP data.
 
@@
 In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
-An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:s4}).
+An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:niter}).
 Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.

### 2020-09-29 supplement.tex
Supplementary figures added and incorporated into fig refs

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Supplementary material}
+\section{Supplementary materials}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% 1D LINEAR ADMIXTURE KINSHIP
 \begin{figure}[H]
     \centering
-    \includegraphics[scale = 0.45]{figures/niter_perLambda_point.png}
+    \includegraphics[scale = 1]{figures/admixed_kinship.png}
+    \caption{1D Linear Admixture with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
+    \label{fig:admixed}
+\end{figure}
+
+% INDEPENDENT SUBPOPULATIONS, COARSE, KINSHIP
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/indep_coarse_kinship.png}
+    \caption{Independent Subpopulations with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
+    \label{fig:indep_coarse}
+\end{figure}
+
+% INDEPENDENT SUBPOPULATIONS, FINE, KINSHIP
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/indep_fine_kinship.png}
+    \caption{Independent Subpopulations with subpopulation structure (2 subjects per subpopulation), estimated kinship matrix.}
+    \label{fig:indep_fine}
+\end{figure}
+
+% EMPIRICAL KINSHIP
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/empirical_kinship.png}
+    \caption{Empirical estimated kinship matrix.}
+    \label{fig:empirical}
+\end{figure}
+
+\begin{landscape}
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 0.8]{figures/niter_perLambda_point.png}
     \caption{The number of iterations per $\lambda$ value for LMM-Lasso-Rakitsch and LMM-Lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
 Note that 100 is the maximum number of iterations for both implementations of LMM-Lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
-    \label{fig:s4}
+    \label{fig:niter}
 \end{figure}
+\end{landscape}

### 2020-09-29 main.tex
4.4 edits

@@
 An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:niter}).
-Getting stuck in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, as we see in Figure \ref{fig:eta_beta_mse}.
+Converging to a local minimum in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, which is what we observe in Figure \ref{fig:eta_beta_mse}.
 

### 2020-10-08 lmm-genetic-review.tex
3.1 edits
Edits to 3.6
Updating paper figs (dichotomous-discordant change) and descriptions
Supplement and S4 (ggmix vs lmmlasso)
\newpage for supplement
Edits to abstract

@@
 \usepackage{tkz-euclide}
+\usepackage{lscape}
 
@@
 \newcommand{\bD}{\textbf{D}}
+\newcommand{\bG}{\textbf{G}}
 \newcommand{\bK}{\textbf{K}}
@@
 \newcommand{\T}{\text{T}}
+\newcommand{\lam}{\lambda}
 
@@
 Despite an extensive literature on penalized regression and LMMs separately, the two are rarely discussed together.
-The aim of this review is to do so while offering a rigorous study of the statistical properties of penalized LMMs.
+The aim of this review is to do so while rigorously examining the statistical properties of penalized LMMs.
 Specifically, the ability of penalized LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
-In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
-In addition, we evaluate the performance of penalized LMMs and competing methods in terms of MSE and variable selection accuracy in the presence of a number of confounding structures.
+In order to clarify the importance yet subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
+In addition, we evaluate the performance of penalized LMMs and competing methods in terms of estimation and selection accuracy in the presence of a number of confounding structures.
 % in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.
@@
 \newpage
-
 \bibliographystyle{ims}
@@
 
+\newpage
+\input{supplement}
+
 \end{document}

### 2020-10-08 main.tex
Edits to abstract

@@
 
-The objective of many genetic studies is to accurately identify variants associated with complex traits or disease, and to estimate the effects of those variants.
+The objective of many genetic studies is to accurately identify variants associated with complex traits or diseases, and to estimate the effects of those variants.
 Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.

### 2020-10-09 main.tex
Edits to 4 and 5

@@
 Such methods assume that genetic relatedness serves as an accurate proxy for environmental exposures, which may or may not be the case depending on the exposure and population in question.
+
+\pb{Just an FYI, you may want not to call this diagram a ``Figure''.
+Sometimes you're restricted in terms of the number of figures you can include, and this doesn't really need to be a figure, just a diagram (like an equation).
+Same thing for the other DAG}
 
@@
 We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
-Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components.
+Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components \pb{let's just report MSE (and maybe MSPE too); breakdown blurs ``bias'' concepts a bit}.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1.1]{figures/beta_mse.png}
+    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
+    \label{fig:mse}
+\end{figure}
+
 The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
@@
 This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
-Qualitatively, this is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by more discrete, nonoverlapping genetic subpopulations, which more directly correspond to the nonoverlapping environmental effects than do the continuous and blended subpopulations of the 1D Linear Admixture data.
-In other words, a greater proportion of the environmental effect is in the column space of $\boldsymbol{P}_{\bX}$ for the Empirical and Independent Subpopulations data than for the 1D Linear Admixture data.
+This is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by nonoverlapping genetic subpopulations, meaning that a greater proportion of the environmental effect is in the column space of $\boldsymbol{P}_{\bX}$.
 We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}.
-CP $= 0.47, 0.99, 0.61$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
-The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
+Here, we have CP $= 0.47, 0.99, 0.61$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
+\pb{let's move the 0.61 to the next paragraph} The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
 Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-Lasso and LMM-Lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.
@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 1.1]{figures/beta_mse.png}
-    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
-    \label{fig:mse}
-\end{figure}
 % \anna{talk about karl rohe stuff here - if anywhere}
 
-In addition to MSE, these methods' performance in terms of variable selection accuracy were quantified using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
-Table \ref{tab:var_sel} displays the mean TPR and corresponding standard deviation in the same settings as those illustrated in Figure \ref{fig:mse}.
+In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
+Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
 The TPRs echo the MSE results in that Lasso and LMM-Lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
-In the presence of environmental confounding, both LMM-Lasso and PC-Lasso achieve substantial improvements in TPR over Lasso across data types.
+In the presence of environmental confounding, both LMM-Lasso and PC-Lasso achieve substantial improvements in TPR over the ordinary Lasso.
 As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Empirical data sets than in the 1D Linear Admixture data.
-Again, this is due to the smaller proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ for the 1D Linear Admixture data, which results in less confounded SNP-phenotype associations to begin with.
 
@@
 %------------------------------%
-\subsection{LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure}
+\subsection{LMM-Lasso outperforms PC-Lasso for finer population structures}
 %------------------------------%
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures.
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures \pb{as defined in section....}.
 Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
 
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations simulated $\bG$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\bgamma$ structures.
-The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
-The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.
-
-In the presence of coarse population structure, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding boxes straddling the 0 line.
-However, the majority of these boxes fall above 0, indicating the superior performance of LMM-Lasso.
-In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy across all levels of environmental confounding.
-This is consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
-As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
 \begin{figure}[H]
@@
     % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
-     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
+     \includegraphics[width=0.8\textwidth]{figures/mse_diff_subpops.png}
+    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
+\pb{This figure could maybe be shorter -- takes up a lot of vertical room that isn't really necessary}}
     \label{fig:big_vs_small}
 \end{figure}
+
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations and a dichotomous-discordant environmental confounding pattern, though qualitatively similar results were observed for other settings.
+The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
+The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences above 0 correspond to LMM-Lasso outperforming PC-Lasso.
+
+Overall, in all of the settings, LMM-Lasso outperformed PC-Lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
+These findings are consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
+As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
 
@@
 
-In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider distinct patterns of environmental effects.
-Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure.
-Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the Independent Subpopulations simulated $\bG$.
+In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider various patterns of environmental effects.
+Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and \pb{``coefficient structure'' is vague; can we describe this better?} coefficient structure.
+Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and Independent Subpopulations.
+
+\pb{What do you think about dropping the exponential/dichotomous thing and just going with dichotomous throughout?
+Doesn't really seem super important.
+Well, we can just keep it in mind and if we need to cut down on space / figures, we can think about it.}
 
@@
 
+\pb{Not sure I love this figure, as it emphasizes exp vs dichotomous, which isn't really very interesting.  Can we re-panel it so that it emphasizes discordant vs concordant?}
+
 %------------------------------%
@@
 LMM-Lasso-ggmix seeks to improve upon this procedure by re-estimating the variance components and $\bbeta$ iteratively.
-Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$.
+Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix reliably outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$.
 In order to systematically compare both methods, we carried out simulations where $p = 1000$, $n = \{200, 400, 600, 800, 1000, 1200, 1400\}$, and $\eta = \{0.2, 0.5, 0.8\}$.
@@
 
-LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n << p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
+LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n << p$ \pb{you want $n \ll p$, not $n << p$; this occurs a bunch}, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
 However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-Lasso-Rakitsch.
@@
 Overall these results show that LMM-Lasso-ggmix performs as well as LMM-Lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-Lasso-Rakitsch in estimating $\eta$ when $n << p$.
-However, these are also the settings in which LMM-Lasso-ggmix is most computationally inefficient.
+However, these are also the settings in which LMM-Lasso-ggmix is least computationally efficient.
 The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n << p$, LMM-Lasso-ggmix is up to 60 times slower than LMM-Lasso-Rakitsch.
@@
 
-In this work, we have reviewed the concept of population stratification and explicitly defined an environmental mechanism by which it may confound the genotype-phenotype relationship.
+In this work, we have reviewed the concept of population stratification and explicitly defined an environmental mechanism by which it may confound the genotype-phenotype relationship \pb{not sure what this sentence means; what are you referring to?}.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
 In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.
+\pb{emphasize the \emph{penalized} nature of our work here, which is I think what makes it novel}
 
@@
 The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n << p$ settings.
-However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and is considered by the authors to be more broadly applicable.
+However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus more broadly applicable.
 
@@
 As an additional point of interest, recall that we provide qualitative and quantitative explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure.
-While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, the result shown in \ref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimatess.
+While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimatess.
 A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper.
@@
 Finally, it is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
-Because of this, bias-reduction techniques for penalized LMMs are an attractive avenue for future research.\\
+Because of this, bias-reduction techniques for penalized LMMs are a potentially attractive avenue for future research.

### 2020-10-12 main.tex
sec 4 updates - in progress

@@
 \subsubsection{Subpopulation structure}
+\label{sec:subpopulation_structure}
 For the fully simulated data, two subpopulation structures were considered, which we refer to as ``fine'' and ``coarse''.
@@
 We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}.
-Here, we have CP $= 0.47, 0.99, 0.61$ for the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
-\pb{let's move the 0.61 to the next paragraph} The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
+Here, we have CP $= 0.47, 0.99$ for the 1D Linear Admixture and Independent Subpopulations, respectively.
+The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
 Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-Lasso and LMM-Lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.
 
-Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP with those from the simulated data sets is not straightforward.
+Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP $= 0.61$ with those from the simulated data sets is not straightforward.
 We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as in the Empirical data.
@@
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures \pb{as defined in section....}.
-Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures as defined in section \ref{sec:subpopulation_structure}.
+Synthetic phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
 
@@
     \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
-\pb{This figure could maybe be shorter -- takes up a lot of vertical room that isn't really necessary}}
+}
     \label{fig:big_vs_small}
@@
 In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider various patterns of environmental effects.
-Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and \pb{``coefficient structure'' is vague; can we describe this better?} coefficient structure.
+Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and \anna{environmental effect pattern}.
 Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and Independent Subpopulations.
@@
 Doesn't really seem super important.
-Well, we can just keep it in mind and if we need to cut down on space / figures, we can think about it.}
+Well, we can just keep it in mind and if we need to cut down on space / figures, we can think about it.} \anna{so leave for now?}
 
-The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
+\anna{RE-WRITE THIS TO MAKE SURE IT MAKES SENSE WITH NEW FIGURES} The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
 The upper left hand panel of Figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso.
@@
 \end{figure}
-
-\pb{Not sure I love this figure, as it emphasizes exp vs dichotomous, which isn't really very interesting.  Can we re-panel it so that it emphasizes discordant vs concordant?}
 
@@
 
-LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n << p$ \pb{you want $n \ll p$, not $n << p$; this occurs a bunch}, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
+LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n \ll p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
 However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-Lasso-Rakitsch.
@@
 
-LMM-Lasso-ggmix achieves an improvement in the estimation of $\eta$ compared to LMM-Lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n << p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
+LMM-Lasso-ggmix achieves an improvement in the estimation of $\eta$ compared to LMM-Lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
 These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix and thus do not seem to improve its estimation of the fixed genetic effects.
@@
 
-Overall these results show that LMM-Lasso-ggmix performs as well as LMM-Lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-Lasso-Rakitsch in estimating $\eta$ when $n << p$.
+Overall these results show that LMM-Lasso-ggmix performs as well as LMM-Lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-Lasso-Rakitsch in estimating $\eta$ when $n \ll p$.
 However, these are also the settings in which LMM-Lasso-ggmix is least computationally efficient.
-The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n << p$, LMM-Lasso-ggmix is up to 60 times slower than LMM-Lasso-Rakitsch.
+The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n \ll p$, LMM-Lasso-ggmix is up to 60 times slower than LMM-Lasso-Rakitsch.
 Conversely, in settings where $n \ge p$, LMM-Lasso-ggmix is up to 10 times faster than LMM-Lasso-Rakitsch.
@@
 Firstly, it suffers from estimation instability in $n \ge p$ settings.
-Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably when $n << p$, LMM-Lasso-ggmix requires up to 60 times the computing time of LMM-Lasso-Rakitsch in these settings due to its iterative estimation procedure.
-The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n << p$ settings.
+Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably when $n \ll p$, LMM-Lasso-ggmix requires up to 60 times the computing time of LMM-Lasso-Rakitsch in these settings due to its iterative estimation procedure.
+The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n \ll p$ settings.
 However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus more broadly applicable.

### 2020-10-12 main.tex
4.1, 4.3 edits

@@
 We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
-Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) and broken down into bias and variance components \pb{let's just report MSE (and maybe MSPE too); breakdown blurs ``bias'' concepts a bit}.
+Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) \anna{MSPE sims running - maybe add here}.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
@@
     \centering
-    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
-     \includegraphics[width=0.8\textwidth]{figures/mse_diff_subpops.png}
+    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
     \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
@@
 
-\pb{What do you think about dropping the exponential/dichotomous thing and just going with dichotomous throughout?
+% \pb{What do you think about dropping the exponential/dichotomous thing and just going with dichotomous throughout?
 Doesn't really seem super important.
@@
 
-\anna{RE-WRITE THIS TO MAKE SURE IT MAKES SENSE WITH NEW FIGURES} The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
-The upper left hand panel of Figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso.
-The lower left hand panel of Figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients.
-In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while LMM-Lasso outperforms PC-Lasso in the majority of dichotomous-concordant environmental effects.
+The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
+
+LMM-Lasso appears to perform as well, or better than, PC-Lasso in the presence of coarse subpopulation structure, as shown by the left-hand panels of Figure \ref{fig:big_small_gamma}.
+The upper left-hand facet in particular shows that PC-Lasso and LMM-Lasso perform more comparably when environmental effect magnitudes are exponential and concordant in nature.
 These findings align with our initial hypothesis that PC-Lasso will perform well in the presence of simple environmental effect patterns that can be well-approximated by a linear combination of the observed data.
@@
     \centering
-    % \includegraphics[scale = 0.9]{figures/fig2b.png}
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
@@
 As an additional point of interest, recall that we provide qualitative and quantitative explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure.
-While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimatess.
+While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimates.
 A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper.

### 2020-10-12 main.tex
5 edits

@@
 
-In this work, we have reviewed the concept of population stratification and explicitly defined an environmental mechanism by which it may confound the genotype-phenotype relationship \pb{not sure what this sentence means; what are you referring to?}.
+In this work, we have reviewed the concept of population structure in a framework which \anna{formally differentiates between effects of population structure and those of the environment}.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
-In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.
-\pb{emphasize the \emph{penalized} nature of our work here, which is I think what makes it novel}
+In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data, \anna{now applied within a novel penalized framework}: PC-Lasso and LMM-Lasso.
 

### 2020-10-12 main.tex
Section 2: putting DAGs in eq environment

@@
 
-This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in Figure \ref{fig:ps_env}.
+This mechanism of environmental confounding is summarized in the directed acyclic graph displayed in diagram \eqref{diag:ps_env}.
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
@@
 
-\pb{Just an FYI, you may want not to call this diagram a ``Figure''.
-Sometimes you're restricted in terms of the number of figures you can include, and this doesn't really need to be a figure, just a diagram (like an equation).
-Same thing for the other DAG}
-
-\begin{figure}[H]
+\begin{equation}
 \centering
-\begin{tikzpicture}
+\begin{tikzpicture}[baseline=(current  bounding  box.center)]
     \node (1) at (0,0) {Population Stratification};
@@
 \end{tikzpicture}
-\caption{Directed acyclic graph (DAG) of confounding due to population structure.}
-\label{fig:ps_env}
-\end{figure}
+%\caption{Directed acyclic graph (DAG) of confounding due to population structure.}
+\label{diag:ps_env}
+\end{equation}
 
@@
 Many of these methods attempt to block the causal pathway between population structure and observed SNP data.
-This is illustrated by the blocked arrow in Figure \ref{fig:ps_env_block}.
+This is illustrated by the blocked arrow in diagram \eqref{diag:ps_env_block}.
 As the figure shows, these methods do not directly address the environmental component of confounding due to population structure, instead controlling confounding by indirectly blocking the causal pathway between the environment and observed SNPs.
 
-\begin{figure}[H]
+\begin{equation}
 \centering
-\begin{tikzpicture}
+\begin{tikzpicture}[baseline=(current  bounding  box.center)]
     \node (1) at (0,0) {Population Stratification};
@@
     \path[bidirected, -] (1) edge (2);
-    \path[magenta, line width = 0.5] (1) edge (3);
+    \path[red, line width = 0.5] (1) edge (3);
     \path (2) edge (4);
     \path (3) edge (4);
-    \tkzMarkSegment[color=magenta,pos=.5,mark=x](1,3)
+    \tkzMarkSegment[color=red,pos=.5,mark=x](1,3)
 \end{tikzpicture}
-\caption{The causal pathway between the environment and observed SNPs may be indirectly blocked by many methods that attempt to reduce confounding due to population structure.}
-\label{fig:ps_env_block}
-\end{figure}
+%\caption{The causal pathway between the environment and observed SNPs may be indirectly blocked by many methods that attempt to reduce confounding due to population structure.}
+\label{diag:ps_env_block}
+\end{equation}
 
@@
 In the presence of environmental heterogeneity which affects the phenotype of interest in a manner that is orthogonal to population structure, SNP-phenotype associations will not be confounded.
-This is due to the absence of a causal pathway between population stratification and environment, as depicted in Figure \ref{fig:ps_env_block2}.
+This is due to the absence of a causal pathway between population stratification and environment, as depicted in diagram \eqref{diag:ps_env_block2}.
 In such a scenario, failing to account for unobserved environmental effects is likely to result in greater variability, but SNP effect estimates will remain unbiased \citep{greenland1999causal}.
@@
 
-\begin{figure}[H]
+\begin{equation}
 \centering
-\begin{tikzpicture}
+\begin{tikzpicture}[baseline=(current  bounding  box.center)]
     \node (1) at (0,0) {Population Stratification};
@@
     \node (4) [below = of 2] {Phenotype};
-    \path[bidirected, -, magenta] (1) edge (2);
+    \path[bidirected, -, red] (1) edge (2);
     \path (1) edge (3);
@@
     \path (3) edge (4);
-    \tkzMarkSegment[color=magenta,pos=.5,mark=x](1,2)
+    \tkzMarkSegment[color=red,pos=.5,mark=x](1,2)
 \end{tikzpicture}
-\caption{The causal pathway between the environment and observed SNPs may be blocked if genetic information does not serve as an accurate proxy for it.}
-\label{fig:ps_env_block2}
-\end{figure}
+%\caption{The causal pathway between the environment and observed SNPs may be blocked if genetic information does not serve as an accurate proxy for it.}
+\label{diag:ps_env_block2}
+\end{equation}
 
-Assuming for now that population structure is correlated with environmental exposures and can be accurately inferred from available data, we return to our consideration of methods that attempt to control for confounding via the mechanism depicted in Figure \ref{fig:ps_env_block}.
+Assuming for now that population structure is correlated with environmental exposures and can be accurately inferred from available data, we return to our consideration of methods that attempt to control for confounding via the mechanism depicted in diagram \eqref{diag:ps_env_block}.
 Such methods include LMMs, principal-component-based approaches, and stratified analyses.

### 2020-10-15 main.tex
Small typos

@@
 Let $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denoting the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
-Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
+Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
 This in turn means that
@@
 Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection.
-Becasue of this, the preceding results, though generated using the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods in the $n = 200, p = 1000$ setting.
+Because of this, the preceding results, though generated using the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods in the $n = 200, p = 1000$ setting.
 The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.

### 2020-10-15 supplement.tex
Small typos

@@
     \includegraphics[scale = 1]{figures/admixed_kinship.png}
-    \caption{1D Linear Admixture with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
+    \caption{1D Linear Admixture with coarse subpopulation structure (50 subjects per subpopulation), RRM.}
     \label{fig:admixed}
@@
     \includegraphics[scale = 1]{figures/indep_coarse_kinship.png}
-    \caption{Independent Subpopulations with coarse subpopulation structure (50 subjects per subpopulation), estimated kinship matrix.}
+    \caption{Independent Subpopulations with coarse subpopulation structure (50 subjects per subpopulation), RRM.}
     \label{fig:indep_coarse}
@@
     \includegraphics[scale = 1]{figures/indep_fine_kinship.png}
-    \caption{Independent Subpopulations with subpopulation structure (2 subjects per subpopulation), estimated kinship matrix.}
+    \caption{Independent Subpopulations with fine subpopulation structure (2 subjects per subpopulation), RRM.}
     \label{fig:indep_fine}
@@
     \includegraphics[scale = 1]{figures/empirical_kinship.png}
-    \caption{Empirical estimated kinship matrix.}
+    \caption{Empirical RRM.}
     \label{fig:empirical}

### 2020-10-15 main.tex
Wording tweaks

@@
 
-Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation.
+Using the decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation.
 Note that $\bV^{-1}$ can be expressed using the following matrix factorization
@@
 
-It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects in the absence of penalization?.
+It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects, in the absence of penalization.
 \citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, which we briefly review.
@@
 
-Additionally, while the fixed effects model considers the principal components $\bU$ of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
+Additionally, while the fixed effects model considers the principal components of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
 The eigenvalues have been shown to serve as a measure of biological relevance of each principal component in relation to underlying population structure \citep{hoffman2013correcting, price2006principal}.
@@
 Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) \anna{MSPE sims running - maybe add here}.
-Figure \ref{fig:mse} presents these results for all three data types in a 1:1 signal to noise (SNR) setting, where $\eta = 0.5$.
+Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
 The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), Lasso and LMM-Lasso perform comparably.
@@
 
-Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical CP $= 0.61$ with those from the simulated data sets is not straightforward.
-We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as in the Empirical data.
+Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.61$, with those from the simulated data sets is not straightforward.
+We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as is the case in the Empirical data.
 
@@
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures as defined in section \ref{sec:subpopulation_structure}.
+Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures as defined in Section \ref{sec:subpopulation_structure}.
 Synthetic phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
@@
 Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations and a dichotomous-discordant environmental confounding pattern, though qualitatively similar results were observed for other settings.
-The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures.
-The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences above 0 correspond to LMM-Lasso outperforming PC-Lasso.
+The y-axis displays the difference in MSE (PC-Lasso $-$ LMM-Lasso) for each simulation and for the various subpopulation structures.
+The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences above 0 correspond to LMM-Lasso outperforming PC-Lasso, and vice versa.
 
-Overall, in all of the settings, LMM-Lasso outperformed PC-Lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
-These findings are consistent with \citet{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models.
+In all of these settings, LMM-Lasso outperformed PC-Lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
+These findings are consistent with \citet{hoffman2013correcting}, who showed that PC adjustment methods represent a low dimensional approximation to LMMs.
 As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
@@
 In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider various patterns of environmental effects.
-Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and \anna{environmental effect pattern}.
+Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso $-$ LMM-Lasso) broken down by both subpopulation and \anna{environmental effect pattern}.
 Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and Independent Subpopulations.
@@
 The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
-
 LMM-Lasso appears to perform as well, or better than, PC-Lasso in the presence of coarse subpopulation structure, as shown by the left-hand panels of Figure \ref{fig:big_small_gamma}.

### 2020-10-16 main.tex
Lasso -> lasso

@@
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
-The Lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
+The lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
 It relies on an assumption of sparsity, which allows coefficient estimation and variable selection to be accomplished simultaneously.
 It also provides considerable computational advantages that scale up well to very large data sets.
-These features make the Lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
+These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
 
-As with multiple linear regression models, Lasso penalized regression assumes independent observations and uncorrelated errors.
+As with multiple linear regression models, lasso penalized regression assumes independent observations and uncorrelated errors.
 In the presence of population structure or cryptic relatedness, however, samples are correlated, violating this assumption.
@@
 In addition, we briefly review historical methods of correcting for population stratification and relatedness.
-In Section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted Lasso penalized regression (PC-Lasso) and Lasso penalized multivariate LMMs (LMM-Lasso).
+In Section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-lasso) and lasso penalized multivariate LMMs (LMM-lasso).
 We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
@@
 
-The remainder of this review will focus on the performance of such multivariate Lasso-penalized methods for the analysis of continuous, normally distributed outcomes.
-Specifically, PC-adjustment and LMM methods, which we will refer to as PC-Lasso and LMM-Lasso, respectively.
-We consider two similar implementations of the LMM-Lasso method, one developed by \citet{Rakitsch2012} and a more recent implementation from \citet{bhatnagar2019simultaneous}.
-Subsequent references to LMM-Lasso apply to both implementations, unless otherwise stated.
-Where necessary, we will differentiate between these implementations by referring to them as LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, respectively.
+The remainder of this review will focus on the performance of such multivariate lasso-penalized methods for the analysis of continuous, normally distributed outcomes.
+Specifically, PC-adjustment and LMM methods, which we will refer to as PC-lasso and LMM-lasso, respectively.
+We consider two similar implementations of the LMM-lasso method, one developed by \citet{Rakitsch2012} and a more recent implementation from \citet{bhatnagar2019simultaneous}.
+Subsequent references to LMM-lasso apply to both implementations, unless otherwise stated.
+Where necessary, we will differentiate between these implementations by referring to them as LMM-lasso-Rakitsch and LMM-lasso-ggmix, respectively.
 
@@
 
-To describe and compare PC-Lasso and LMM-Lasso, we consider an $n \times p$ matrix of genotype data $\bG$ for $n$ subjects and $p$ SNPs, where each $G_{ij} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
+To describe and compare PC-lasso and LMM-lasso, we consider an $n \times p$ matrix of genotype data $\bG$ for $n$ subjects and $p$ SNPs, where each $G_{ij} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
 In order to equitably compare SNP effects, regardless of allele frequency, in practice $\bG$ is standardized.
@@
 A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}; we will discuss the rationale for this approach in Section~\ref{sec:sim_env_conf}.
-This, in turn, motivates the methods by which LMM-Lasso and PC-Lasso attempt to model unobserved confounding effects using observed genetic data.
+This, in turn, motivates the methods by which LMM-lasso and PC-lasso attempt to model unobserved confounding effects using observed genetic data.
 
-Both PC-Lasso and LMM-Lasso rely on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
+Both PC-lasso and LMM-lasso rely on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
 Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
@@
 %------------------------------%
-\subsection{PC-Lasso}
-PC-Lasso uses the first $k$ principal components of $\bK$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates. This corresponds to the regression model
+\subsection{PC-lasso}
+PC-lasso uses the first $k$ principal components of $\bK$ to approximate $\bZ\bgamma$ and includes them in the model as unpenalized covariates. This corresponds to the regression model
 \begin{equation}
@@
 %------------------------------%
-\subsection{LMM-Lasso}
+\subsection{LMM-lasso}
 \label{sec:lmmlasso}
 
-Conversely, LMM-Lasso models $\bZ\bgamma$ as a random effect, which corresponds to the LMM model
+Conversely, LMM-lasso models $\bZ\bgamma$ as a random effect, which corresponds to the LMM model
 \begin{equation}
@@
 
-We will first derive the LMM-Lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
+We will first derive the LMM-lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
 Following equation \eqref{eqn:lmm_reg}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
-LMM-Lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
+LMM-lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
 Specifically, left-multiplying the original data by $\bV^{-1/2}$ yields
@@
 \end{equation}
-and thus an ordinary Lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
+and thus an ordinary lasso model can be fit on the transformed data, ($\bV^{-\frac{1}{2}}\bX, \bV^{-\frac{1}{2}}\by$).
 
@@
 
-Different variance component estimation methods lead to different forms of $\widehat{\bW}$ and two distinct LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
-LMM-Lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
+Different variance component estimation methods lead to different forms of $\widehat{\bW}$ and two distinct LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
+LMM-lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
 These variance components and the corresponding weights are estimated once under the assumption of a null model.
-A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
-LMM-Lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2019simultaneous}.
-This iterative update scheme prevents the use of standard Lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
+A lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
+LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2019simultaneous}.
+This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
 We further discuss differences between these methods in Section \ref{sec:results}.
@@
 %------------------------------%
-\subsection{Relationship between PC-Lasso and LMM-Lasso}
+\subsection{Relationship between PC-lasso and LMM-lasso}
 
-It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-Lasso can be equivalently expressed in terms of fixed effects, in the absence of penalization.
+It has been shown that the random effect formulation of model \eqref{eqn:model} used by LMM-lasso can be equivalently expressed in terms of fixed effects, in the absence of penalization.
 \citet{zhang2015principal} derived this equivalence using a probabilistic PCA formulation, while \citet{hoffman2013correcting} used the singular value decomposition of $\bX$, which we briefly review.
@@
 Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}.
-This is the approach we take when implementing PC-Lasso in Section \ref{sec:results}.
+This is the approach we take when implementing PC-lasso in Section \ref{sec:results}.
 
@@
 Various $\bgamma$ patterns were considered in order to investigate the effects of distinct forms of environmental confounding.
-Our initial hypothesis was that simple patterns of environmental effects could be well-approximated by a linear combination of the observed genetic data and well-corrected for by both PC-Lasso and LMM-Lasso, while PC-Lasso would be unable to represent and correct for more complex patterns of environmental effects.
+Our initial hypothesis was that simple patterns of environmental effects could be well-approximated by a linear combination of the observed genetic data and well-corrected for by both PC-lasso and LMM-lasso, while PC-lasso would be unable to represent and correct for more complex patterns of environmental effects.
 
@@
 While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
-In addition, because Lasso, PC-Lasso, and LMM-Lasso use different likelihoods, an information-criterion-based approach would not lend itself to comparable results across methods.
+In addition, because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach would not lend itself to comparable results across methods.
 
@@
 %------------------------------%
-\subsection{PC-Lasso and LMM-Lasso outperform the ordinary Lasso in the presence of environmental confounding}
+\subsection{PC-lasso and LMM-lasso outperform the ordinary lasso in the presence of environmental confounding}
 %------------------------------%
 
-We begin by comparing the ordinary Lasso, PC-Lasso, and LMM-Lasso as the level of environmental confounding is varied.
-Estimation accuracy for Lasso, PC-Lasso, and LMM-Lasso was evaluated based on mean squared error (MSE) \anna{MSPE sims running - maybe add here}.
+We begin by comparing the ordinary lasso, PC-lasso, and LMM-lasso as the level of environmental confounding is varied.
+Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE) \anna{MSPE sims running - maybe add here}.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
-The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), Lasso and LMM-Lasso perform comparably.
-PC-Lasso results in a slight inflation in MSE compared to the other methods in the simulated data sets due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
+The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-lasso perform comparably.
+PC-lasso results in a slight inflation in MSE compared to the other methods in the simulated data sets due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
 
@@
 The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
-In this case, both PC-Lasso and LMM-Lasso clearly outperform the ordinary Lasso.
+In this case, both PC-lasso and LMM-lasso clearly outperform the ordinary lasso.
 This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
@@
 Here, we have CP $= 0.47, 0.99$ for the 1D Linear Admixture and Independent Subpopulations, respectively.
-The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-Lasso and LMM-Lasso methods are able to correct.
-Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-Lasso and LMM-Lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.
+The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-lasso and LMM-lasso methods are able to correct.
+Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-lasso and LMM-lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.
 
@@
 Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
-The TPRs echo the MSE results in that Lasso and LMM-Lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
-In the presence of environmental confounding, both LMM-Lasso and PC-Lasso achieve substantial improvements in TPR over the ordinary Lasso.
+The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
+In the presence of environmental confounding, both LMM-lasso and PC-lasso achieve substantial improvements in TPR over the ordinary lasso.
 As with the MSE data, these improvements are more substantial in the Independent Subpopulations and Empirical data sets than in the 1D Linear Admixture data.
@@
 \midrule
-\multirow{6}{*}{0.5} & \multirow{3}{*}{0} & Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.37 (0.06) \\ 
-& & PC-Lasso & 0.23 (0.05) & 0.21 (0.05) & 0.36 (0.06) \\ 
-& & LMM-Lasso & 0.25 (0.05) & 0.23 (0.05) & 0.41 (0.06) \\ 
+\multirow{6}{*}{0.5} & \multirow{3}{*}{0} & lasso & 0.25 (0.05) & 0.23 (0.05) & 0.37 (0.06) \\ 
+& & PC-lasso & 0.23 (0.05) & 0.21 (0.05) & 0.36 (0.06) \\ 
+& & LMM-lasso & 0.25 (0.05) & 0.23 (0.05) & 0.41 (0.06) \\ 
 \cmidrule{2-6}
-& \multirow{3}{*}{0.8} & Lasso & 0.24 (0.05) & 0.26 (0.06) & 0.29 (0.06) \\
-& & PC-Lasso & 0.28 (0.05) & 0.36 (0.06) & 0.46 (0.06) \\ 
-& & LMM-Lasso & 0.28 (0.05) & 0.39 (0.06) & 0.57 (0.05) \\
+& \multirow{3}{*}{0.8} & lasso & 0.24 (0.05) & 0.26 (0.06) & 0.29 (0.06) \\
+& & PC-lasso & 0.28 (0.05) & 0.36 (0.06) & 0.46 (0.06) \\ 
+& & LMM-lasso & 0.28 (0.05) & 0.39 (0.06) & 0.57 (0.05) \\
 \bottomrule
@@
 %------------------------------%
-\subsection{LMM-Lasso outperforms PC-Lasso for finer population structures}
+\subsection{LMM-lasso outperforms PC-lasso for finer population structures}
 %------------------------------%
 
-Given the superior performance of PC-Lasso and LMM-Lasso compared to the ordinary Lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures as defined in Section \ref{sec:subpopulation_structure}.
+Given the superior performance of PC-lasso and LMM-lasso compared to the ordinary lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures as defined in Section \ref{sec:subpopulation_structure}.
 Synthetic phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
+    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
 }
@@
 Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations and a dichotomous-discordant environmental confounding pattern, though qualitatively similar results were observed for other settings.
-The y-axis displays the difference in MSE (PC-Lasso $-$ LMM-Lasso) for each simulation and for the various subpopulation structures.
-The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences above 0 correspond to LMM-Lasso outperforming PC-Lasso, and vice versa.
+The y-axis displays the difference in MSE (PC-lasso $-$ LMM-lasso) for each simulation and for the various subpopulation structures.
+The dashed horizontal line at 0 indicates no difference in the performance of PC-lasso and LMM-lasso, while differences above 0 correspond to LMM-lasso outperforming PC-lasso, and vice versa.
 
-In all of these settings, LMM-Lasso outperformed PC-Lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
+In all of these settings, LMM-lasso outperformed PC-lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
 These findings are consistent with \citet{hoffman2013correcting}, who showed that PC adjustment methods represent a low dimensional approximation to LMMs.
-As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
+As such, it is reasonable that PC-lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
 
 %------------------------------%
-\subsection{LMM-Lasso outperforms PC-Lasso when environmental effects are discordant}
+\subsection{LMM-lasso outperforms PC-lasso when environmental effects are discordant}
 %------------------------------%
 
-In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider various patterns of environmental effects.
-Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso $-$ LMM-Lasso) broken down by both subpopulation and \anna{environmental effect pattern}.
+In addition to evaluating PC-lasso and LMM-lasso in the presence of coarse and fine subpopulation structure, we also consider various patterns of environmental effects.
+Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-lasso $-$ LMM-lasso) broken down by both subpopulation and \anna{environmental effect pattern}.
 Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and Independent Subpopulations.
@@
 
-The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
-LMM-Lasso appears to perform as well, or better than, PC-Lasso in the presence of coarse subpopulation structure, as shown by the left-hand panels of Figure \ref{fig:big_small_gamma}.
-The upper left-hand facet in particular shows that PC-Lasso and LMM-Lasso perform more comparably when environmental effect magnitudes are exponential and concordant in nature.
-These findings align with our initial hypothesis that PC-Lasso will perform well in the presence of simple environmental effect patterns that can be well-approximated by a linear combination of the observed data.
-When that is not the case, either due to the presence of fine subpopulation structure, dichotomous effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.
+The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
+LMM-lasso appears to perform as well, or better than, PC-lasso in the presence of coarse subpopulation structure, as shown by the left-hand panels of Figure \ref{fig:big_small_gamma}.
+The upper left-hand facet in particular shows that PC-lasso and LMM-lasso perform more comparably when environmental effect magnitudes are exponential and concordant in nature.
+These findings align with our initial hypothesis that PC-lasso will perform well in the presence of simple environmental effect patterns that can be well-approximated by a linear combination of the observed data.
+When that is not the case, either due to the presence of fine subpopulation structure, dichotomous effects, or genetic similarity that does not reflect similar environmental effects, LMM-lasso exhibits estimation accuracy superior to that of PC-lasso.
 
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$.}
+    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$.}
     \label{fig:big_small_gamma}
@@
 %------------------------------%
-\subsection{Different implementations of LMM-Lasso}
+\subsection{Different implementations of LMM-lasso}
 %------------------------------%
 
-In Section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix.
+In Section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
 Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection.
-Because of this, the preceding results, though generated using the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods in the $n = 200, p = 1000$ setting.
-The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.
-However, given the generally strong performance of LMM-Lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.
+Because of this, the preceding results, though generated using the LMM-lasso-Rakitsch implementation, are representative of both LMM-lasso methods in the $n = 200, p = 1000$ setting.
+The primary purpose of this review is to compare LMM-lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.
+However, given the generally strong performance of LMM-lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.
 
-LMM-Lasso-Rakitsch uses a two-step approach.
+LMM-lasso-Rakitsch uses a two-step approach.
 First, the variance components are estimated under the assumption of a null model.
 Then, the estimated variance components are used to transform the data, and the $\bbeta$ are estimated using coordinate descent.
-LMM-Lasso-ggmix seeks to improve upon this procedure by re-estimating the variance components and $\bbeta$ iteratively.
-Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix reliably outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$.
+LMM-lasso-ggmix seeks to improve upon this procedure by re-estimating the variance components and $\bbeta$ iteratively.
+Although this seems a reasonable approach, we have been unable to find a scenario where LMM-lasso-ggmix reliably outperforms the LMM-lasso-Rakitsch in estimation accuracy of $\bbeta$.
 In order to systematically compare both methods, we carried out simulations where $p = 1000$, $n = \{200, 400, 600, 800, 1000, 1200, 1400\}$, and $\eta = \{0.2, 0.5, 0.8\}$.
@@
 
-LMM-Lasso-Rakitsh and LMM-Lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n \ll p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
-However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-Lasso-Rakitsch.
+LMM-lasso-Rakitsh and LMM-lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n \ll p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
+However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-lasso-Rakitsch.
 This is most dramatic in the presence of substantial environmental environmental confounding, $\xi = 0.8$, and $n = \{1200, 1400\}$.
 
-LMM-Lasso-ggmix achieves an improvement in the estimation of $\eta$ compared to LMM-Lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
-These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-Lasso-ggmix and thus do not seem to improve its estimation of the fixed genetic effects.
-In addition, when $n \ge 1000$, LMM-Lasso-ggmix's estimation of $\eta$ suffers.
-When $n \ge 1000$ and substantial environmental confounding is present, $\xi = 0.8$,  LMM-Lasso-ggmix's performance is hindered to the point that LMM-Lasso-Rakitsch estimates $\eta$ more accurately.
+LMM-lasso-ggmix achieves an improvement in the estimation of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
+These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-lasso-ggmix and thus do not seem to improve its estimation of the fixed genetic effects.
+In addition, when $n \ge 1000$, LMM-lasso-ggmix's estimation of $\eta$ suffers.
+When $n \ge 1000$ and substantial environmental confounding is present, $\xi = 0.8$,  LMM-lasso-ggmix's performance is hindered to the point that LMM-lasso-Rakitsch estimates $\eta$ more accurately.
 
-Overall these results show that LMM-Lasso-ggmix performs as well as LMM-Lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-Lasso-Rakitsch in estimating $\eta$ when $n \ll p$.
-However, these are also the settings in which LMM-Lasso-ggmix is least computationally efficient.
-The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n \ll p$, LMM-Lasso-ggmix is up to 60 times slower than LMM-Lasso-Rakitsch.
-Conversely, in settings where $n \ge p$, LMM-Lasso-ggmix is up to 10 times faster than LMM-Lasso-Rakitsch.
-However, these are also settings in which LMM-Lasso Rakitsch substantially outperforms LMM-Lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently.
+Overall these results show that LMM-lasso-ggmix performs as well as LMM-lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-lasso-Rakitsch in estimating $\eta$ when $n \ll p$.
+However, these are also the settings in which LMM-lasso-ggmix is least computationally efficient.
+The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n \ll p$, LMM-lasso-ggmix is up to 60 times slower than LMM-lasso-Rakitsch.
+Conversely, in settings where $n \ge p$, LMM-lasso-ggmix is up to 10 times faster than LMM-lasso-Rakitsch.
+However, these are also settings in which LMM-lasso Rakitsch substantially outperforms LMM-lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently.
 
-LMM-Lasso-ggmix's poor performance in these scenarios are likely attributable to a number of a factors such as an ill-conditioned RRM  in $n \ge p$ settings \citep{ledoit2004well}.
-In addition, because LMM-Lasso-ggmix attempts to estimate more parameters than LMM-Lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
+LMM-lasso-ggmix's poor performance in these scenarios are likely attributable to a number of a factors such as an ill-conditioned RRM  in $n \ge p$ settings \citep{ledoit2004well}.
+In addition, because LMM-lasso-ggmix attempts to estimate more parameters than LMM-lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
 An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:niter}).
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
+     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
     \label{fig:eta_beta_mse}
@@
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
-In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data, \anna{now applied within a novel penalized framework}: PC-Lasso and LMM-Lasso.
+In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data, \anna{now applied within a novel penalized framework}: PC-lasso and LMM-lasso.
 
-We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively.
+We have detailed the assumptions made by PC-lasso and LMM-lasso, which model confounding influences as fixed and random effects, respectively.
 We have also described the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity.
 In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
-We then compare the estimation and variable selection accuracy of the ordinary Lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.
+We then compare the estimation and variable selection accuracy of the ordinary lasso, PC-lasso, and LMM-lasso in various confounding scenarios.
 
-Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the ordinary Lasso whenever environmental confounding effects are present.
-However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the ordinary Lasso.
+Unsurprisingly, we find that PC-lasso and LMM-lasso outperform the ordinary lasso whenever environmental confounding effects are present.
+However, in the absence of such effects, PC-lasso may hinder estimation accuracy and perform more poorly than even the ordinary lasso.
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
-Additionally, we find that LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure and when confounding effects are discordant in nature.
-In the presence of coarse subpopulation structure and exponential environmental effects, PC-Lasso performs comparably to LMM-Lasso in estimation accuracy.
-However, LMM-Lasso is more robust to subpopulation structure and complex patterns of environmental effects.
-Therefore, LMM-Lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure and environmental heterogeneity are unknown.
+Additionally, we find that LMM-lasso outperforms PC-lasso in the presence of fine subpopulation structure and when confounding effects are discordant in nature.
+In the presence of coarse subpopulation structure and exponential environmental effects, PC-lasso performs comparably to LMM-lasso in estimation accuracy.
+However, LMM-lasso is more robust to subpopulation structure and complex patterns of environmental effects.
+Therefore, LMM-lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure and environmental heterogeneity are unknown.
 
-We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
-Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks.
+We considered two implementations of the LMM-lasso method: LMM-lasso-Rakitsch and LMM-lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
+Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-lasso-ggmix has several drawbacks.
 Firstly, it suffers from estimation instability in $n \ge p$ settings.
-Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably when $n \ll p$, LMM-Lasso-ggmix requires up to 60 times the computing time of LMM-Lasso-Rakitsch in these settings due to its iterative estimation procedure.
-The additional time required of LMM-Lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n \ll p$ settings.
-However, the LMM-Lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus more broadly applicable.
+Secondly, although LMM-lasso-Rakitsch and LMM-lasso-ggmix perform comparably when $n \ll p$, LMM-lasso-ggmix requires up to 60 times the computing time of LMM-lasso-Rakitsch in these settings due to its iterative estimation procedure.
+The additional time required of LMM-lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n \ll p$ settings.
+However, the LMM-lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus more broadly applicable.
 
-\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
+\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-lasso-Rakitsch method using the equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
 
@@
 
-Finally, it is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
+Finally, it is worth nothing that even in scenarios where LMM-lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
 Because of this, bias-reduction techniques for penalized LMMs are a potentially attractive avenue for future research.

### 2020-11-09 main.tex
Small tweaks - avoid merge conflicts
Formatting for genetic epi submission; using endfloat

@@
 In order to equitably compare SNP effects, regardless of allele frequency, in practice $\bG$ is standardized.
-We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 \pi_j) / \sqrt{\pi_j (1 - \pi_j)}$ and $\pi_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
+We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 \pi_j) / \sqrt{2\pi_j (1 - \pi_j)}$ and $\pi_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
 Additionally, $\bX_0$ represents an $n \times p_0$ matrix of unpenalized covariates and the intercept.
@@
 Let $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denoting the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
-Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
+Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
 This in turn means that
@@
 We begin by comparing the ordinary lasso, PC-lasso, and LMM-lasso as the level of environmental confounding is varied.
-Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE) \anna{MSPE sims running - maybe add here}.
+Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE).
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
@@
 
-In this work, we have reviewed the concept of population structure in a framework which \anna{formally differentiates between effects of population structure and those of the environment}.
+In this work, we have reviewed the concept of population structure in a framework which formally differentiates between effects of population structure and those of the environment.
 We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences.
-In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data, \anna{now applied within a novel penalized framework}: PC-lasso and LMM-lasso.
+In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data, now applied within a novel penalized framework: PC-lasso and LMM-lasso.
 

### 2020-11-18 main.tex
3.6.5, 4.3, 5 - removing concordant vs discordant and exponential vs dichotomous discussion and results

@@
 For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$.
-Various $\bgamma$ patterns were considered in order to investigate the effects of distinct forms of environmental confounding.
-Our initial hypothesis was that simple patterns of environmental effects could be well-approximated by a linear combination of the observed genetic data and well-corrected for by both PC-lasso and LMM-lasso, while PC-lasso would be unable to represent and correct for more complex patterns of environmental effects.
+Various $\bgamma$ patterns were considered in order to investigate the effects of distinct and complex patterns of environmental confounding.
+Firstly, we considered two different patterns of effect magnitudes, which we refer to as exponential and dichotomous.
+We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations, while exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes.
+Secondly, we considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
+We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.
 
-Environmental effects may be complex in different ways.
-Firstly, we considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
-For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
+% For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
 However, behavior-based environmental exposures may be less likely to be shared by subpopulations which are genetically and geographically similar.
 As an example, \citet{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
-We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.
-Secondly, we considered two patterns of environmental effect sizes, which we refer to as exponential and dichotomous.
-We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations.
-While exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes.
-
- To evaluate the performance of penalized LMM methods in the presence of these different forms of confounding complexity, we consider environmental effects which are concordant or discordant in nature in the presence of both exponential and dichotomous effect sizes, for a total of four distinct $\bgamma$ patterns.
-Figure \ref{fig:gamma_structures} provides a visual summary of the four coefficient patterns, where the subpopulations on the x-axis are ordered by genetic similarity so that subpopulation 1 is most genetically similar to subpopulation 2, and least similar to subpopulation 10.
-This may also be thought of as subpopulations ordered spatially along a 1-dimensional geographic region.
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
-    \caption{An illustration of four distinct patterns of environmental effects and corresponding $\bgamma$ values in the presence of 10 subpopulations ordered by genetic similarity.}
-    \label{fig:gamma_structures}
-\end{figure}
-\noindent In the case of concordant and dichotomous effects, Figure \ref{fig:gamma_structures} shows that the first five ordered subpopulations are subject to an effect size of 1, while the second five are subject to an effect size of $-$1.
-In the case of discordant dichotomous effects, all subpopulations again experience an effect size with an absolute value of 1, however, the directions of these effects alternate among the ordered subpopulations so that genetically similar subpopulations do not experience similar environmental effects.
 
@@
 
-Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.61$, with those from the simulated data sets is not straightforward.
-We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as is the case in the Empirical data.
+\anna{Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.61$, with those from the simulated data sets is not straightforward.
+We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as is the case in the Empirical data.}
 
@@
 
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations and a dichotomous-discordant environmental confounding pattern, though qualitatively similar results were observed for other settings.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations and a dichotomous-discordant environmental confounding pattern.
 The y-axis displays the difference in MSE (PC-lasso $-$ LMM-lasso) for each simulation and for the various subpopulation structures.
@@
 
+\anna{Similar results were observed in the presence of other patterns of environmental effects.
+No meaningful differences in methodological performance were observed for dichotomous versus exponential effect magnitudes or concordant versus discordant effects.
+}
+
 %------------------------------%
-\subsection{LMM-lasso outperforms PC-lasso when environmental effects are discordant}
+\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-Lasso}}
 %------------------------------%
 
-In addition to evaluating PC-lasso and LMM-lasso in the presence of coarse and fine subpopulation structure, we also consider various patterns of environmental effects.
-Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-lasso $-$ LMM-lasso) broken down by both subpopulation and \anna{environmental effect pattern}.
-Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and Independent Subpopulations.
-
-% \pb{What do you think about dropping the exponential/dichotomous thing and just going with dichotomous throughout?
-Doesn't really seem super important.
-Well, we can just keep it in mind and if we need to cut down on space / figures, we can think about it.} \anna{so leave for now?}
-
-The right-hand panels of Figure \ref{fig:big_small_gamma} are consistent with Figure \ref{fig:big_vs_small}, showing superior LMM-lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance.
-LMM-lasso appears to perform as well, or better than, PC-lasso in the presence of coarse subpopulation structure, as shown by the left-hand panels of Figure \ref{fig:big_small_gamma}.
-The upper left-hand facet in particular shows that PC-lasso and LMM-lasso perform more comparably when environmental effect magnitudes are exponential and concordant in nature.
-These findings align with our initial hypothesis that PC-lasso will perform well in the presence of simple environmental effect patterns that can be well-approximated by a linear combination of the observed data.
-When that is not the case, either due to the presence of fine subpopulation structure, dichotomous effects, or genetic similarity that does not reflect similar environmental effects, LMM-lasso exhibits estimation accuracy superior to that of PC-lasso.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$.}
-    \label{fig:big_small_gamma}
-\end{figure}
+% \begin{figure}[H]
+%    \centering
+%    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
+%    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$.}
+%    \label{fig:big_small_gamma}
+%\end{figure}
 
@@
 This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
-Additionally, we find that LMM-lasso outperforms PC-lasso in the presence of fine subpopulation structure and when confounding effects are discordant in nature.
-In the presence of coarse subpopulation structure and exponential environmental effects, PC-lasso performs comparably to LMM-lasso in estimation accuracy.
-However, LMM-lasso is more robust to subpopulation structure and complex patterns of environmental effects.
-Therefore, LMM-lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure and environmental heterogeneity are unknown.
+Additionally, we find that LMM-lasso consistently outperforms PC-lasso in the presence of fine subpopulation structure, while PC-lasso appears to its best advantage in the presence of coarse subpopulation structure.
+LMM-lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure is unknown.
 

### 2020-11-19 main.tex
4.2 figure

@@
 
-\anna{Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.61$, with those from the simulated data sets is not straightforward.
-We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as is the case in the Empirical data.}
+Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.87$, with those from the simulated data sets is not straightforward.
+We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as is the case in the Empirical data.
 
@@
 & & PC-lasso & 0.28 (0.05) & 0.36 (0.06) & 0.46 (0.06) \\ 
-& & LMM-lasso & 0.28 (0.05) & 0.39 (0.06) & 0.57 (0.05) \\
+& &  LMM-lasso & 0.28 (0.05) & 0.39 (0.06) & 0.61 (0.06) \\ 
 \bottomrule
@@
 \end{table}
+
+%------------------------------%
+\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-Lasso}}
+%------------------------------%
+
+ \begin{figure}[H]
+    \centering
+    \includegraphics[scale = 0.9]{figures/popInfSNPs_n=200.png}
+    \caption{Causal and non-causal SNP selection in the presence of varying levels of population informative SNPs with a coarse 2 Independent Subpopulation structure, dichotomous-discordant effects, $\eta = 0.5$, and $\xi = 0.8$.}
+    \label{fig:big_small_gamma}
+\end{figure}
 
@@
 }
-
-%------------------------------%
-\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-Lasso}}
-%------------------------------%
-
-% \begin{figure}[H]
-%    \centering
-%    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
-%    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying environmental effect structures with Independent Subpopulations data, $\eta = 0.5$, and $\xi = 0.8$.}
-%    \label{fig:big_small_gamma}
-%\end{figure}
 

### 2020-11-21 main.tex
4.2 figure and interpretation (in progress)

@@
 %------------------------------%
-\subsection{Motivating example}
+\subsection{Motivating example} \label{sec:example}
 
@@
 %------------------------------%
-\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-Lasso}}
+\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-lasso}}
+\anna{LMM-lasso is more robust to spurious associations due to allele frequency differences than lasso and PC-lasso }\\
 %------------------------------%
+
+As described in Section \ref{sec:example}, differences in allele frequencies have the potential to lead to spurious associations.
+In the previous section we saw how in the presence of confounding, LMM-lasso and PC-lasso outperform the ordinary lasso in terms of selection accuracy.
+It is of interest to characterize these methods' selection accuracy in terms of how robust they are to allele frequency differences.
+To that end, we considered a simplified scenario, with a coarse 2 Independent Subpopulations structure and dichotomous-discordant environmental effects.
+Figure \ref{fig:pop_inf} shows the relationship between how often a given SNP is selected by the various methods and its absolute difference in  standardized allele frequencies between the two supopulations.
 
@@
     \caption{Causal and non-causal SNP selection in the presence of varying levels of population informative SNPs with a coarse 2 Independent Subpopulation structure, dichotomous-discordant effects, $\eta = 0.5$, and $\xi = 0.8$.}
-    \label{fig:big_small_gamma}
+    \label{fig:pop_inf}
 \end{figure}
+
+A method which is robust to differences in allele frequencies should show no systematic association between selection proportion and differences in allele frequencies.
+We see that for both causal and non-causal SNPs, the LMM-lasso and PC-lasso methods exhibit no such association, while the ordinary lasso demonstrates a \anna{substantial} positive trend.
+\anna{pearson corr with pvalues?}
+
+The presence of this trend for the lasso, and the lack thereof for LMM-lasso and PC-lasso, illustrates the underlying mechanism leading to the superior TPRs of LMM-lasso and PC-lasso compared to that of the lasso in Table \ref{tab:var_sel}.
 

### 2020-11-23 main.tex
4.2 figure and description

--- 2020-11-19 main.tex
@@
 %------------------------------%
-\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-Lasso}}
+\subsection{\anna{Lasso selects more population-informative SNPs than PC-lasso and LMM-lasso}}
 %------------------------------%
+
+In the previous section we saw how in the presence of confounding, LMM-lasso and PC-lasso outperform the ordinary lasso in terms of selection accuracy.
+Additionally, as we described in Section \ref{sec:example}, differences in allele frequencies are known to have the potential to lead to spurious associations.
+As such, it is also of interest to characterize these methods' selection accuracy in terms of how robust they are to allele frequency differences.
+A method which is robust to differences in allele frequencies and thus show no association between selection proportion and differences in allele frequencies.
+
+In order to assess the behavior of these methods in this context, we considered the same parameter settings as those presented in Figure \ref{fig:mse} and Table \ref{tab:var_sel}, albeit with a slightly simpler  population structure.
+We examined the case of two distinct Independent Subpopulations in order to quantify the absolute difference in subpopulation-specific allele frequencies in an interpretable manner.
+
+Figure \ref{fig:pop_inf} shows the relationship between how often a given SNP is selected by the various methods and its absolute difference in  standardized allele frequencies between the two supopulations in the presence and absence of environmental confounding.
 
@@
     \includegraphics[scale = 0.9]{figures/popInfSNPs_n=200.png}
-    \caption{Causal and non-causal SNP selection in the presence of varying levels of population informative SNPs with a coarse 2 Independent Subpopulation structure, dichotomous-discordant effects, $\eta = 0.5$, and $\xi = 0.8$.}
-    \label{fig:big_small_gamma}
+    \caption{Causal and non-causal SNP selection in the presence of varying levels of population informative SNPs with a coarse 2 Independent Subpopulation structure, dichotomous-discordant effects, $\eta = 0.5$.}
+    \label{fig:pop_inf}
 \end{figure}
+
+\anna{maybe no table ? If table, make more like table 1}
+\begin{table}[H]
+\centering
+\begin{tabularx}{225pt}{llrrr}
+\toprule
+\multicolumn{1}{c}{Method} & \multicolumn{1}{c}{SNPs} & \multicolumn{1}{c}{$\xi$}& \multicolumn{1}{c}{$r$} & \multicolumn{1}{c}{p} \\ 
+\midrule
+lasso & Causal & 0.0 & 0.13 & 0.3862 \\ 
+  lasso & Causal & 0.8 & 0.46 & 0.0009 \\ 
+  lasso & Non-causal & 0.0 & 0.04 & 0.1928 \\ 
+  lasso & Non-causal & 0.8 & 0.12 & 0.0003 \\ 
+  PC-lasso & Causal & 0.0 & 0.10 & 0.4963 \\ 
+  PC-lasso & Causal & 0.8 & -0.01 & 0.9422 \\ 
+  PC-lasso & Non-causal & 0.0 & -0.01 & 0.8467 \\ 
+  PC-lasso & Non-causal & 0.8 & -0.04 & 0.2371 \\ 
+  LMM-lasso & Causal & 0.0 & 0.06 & 0.6950 \\ 
+  LMM-lasso & Causal & 0.8 & -0.11 & 0.4469 \\ 
+  LMM-lasso & Non-causal & 0.0 & 0.03 & 0.3233 \\ 
+  LMM-lasso & Non-causal & 0.8 & 0.01 & 0.7716 \\ 
+\bottomrule
+\end{tabularx}
+\caption{Pearson correlation between average SNP selection and absolute difference in standardized allele frequencies across methods.
+Coarse 2 Independent Subpopulation structure, dichotomous-discordant effects, $\eta = 0.5$.}
+\label{tab:pop_inf}
+\end{table}
+
+Underscoring the results from Figure \ref{fig:mse} and Table \ref{tab:var_sel}, we see that in the absence of environmental confounding, none of the three methods demonstrates a significant association between selection proportion and differences in allele frequencies.
+Table \ref{tab:pop_inf} displays the Pearson correlation coefficients and associated p-values in order to quantify these relationships.
+
+In the presence of confounding, however, the LMM-lasso and PC-lasso methods are able to remain robust to allele frequency differences, \anna{while the ordinary lasso demonstrates a statistically significant positive correlation for both causal ($r = 0.46, p = 0.0009$) and non-causal ($r = 0.12, p = 0.0003$) SNPs}.
+The presence of this trend for the lasso, and the lack thereof for LMM-lasso and PC-lasso, illustrates the underlying mechanism leading to the superior TPRs of LMM-lasso and PC-lasso compared to that of the lasso in Table \ref{tab:var_sel}.
 

### 2020-12-02 main.tex
4.2 edits - still in progress

--- 2020-11-19 main.tex
@@
 
-Both PC-lasso and LMM-lasso rely on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{hayes2009increased}.
+Both PC-lasso and LMM-lasso rely on matrix decompositions of $\bX$ and the realized relationship matrix (RRM), $\bK = p^{-1} \bX \bX^\T$, which describes the pair-wise genetic similarity among subjects \citep{patterson2006population, hayes2009increased}.
 Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
@@
 
-In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
+In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) \anna{technically this is true sign rate - only estimates with the correct sign are counted.
+Explain this?
+Calculate overall selection instead?} for the first 50 variables each method allowed into the model.
 Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
@@
 %------------------------------%
-\subsection{\anna{Lasso and PC-lasso select more population-informative SNPs than LMM-Lasso}}
+\subsection{\anna{PC-lasso and LMM-lasso reduce bias towards selecting population-informative SNPs}}
 %------------------------------%
+
+In the previous section we saw how in the presence of confounding, LMM-lasso and PC-lasso outperform the ordinary lasso in terms of selection accuracy.
+Additionally, as we described in Section \ref{sec:example}, subpopulation differences in allele frequencies are known to have the potential to lead to spurious associations.
+As such it is of interest to characterize these methods' selection accuracy in terms of how robust they are to allele frequency differences.
+
+In order to assess the behavior of these methods in this context, we conducted a simulation using the parameter settings presented in Figure \ref{fig:mse} and Table \ref{tab:var_sel} and a modified version of the Empirical data.
+In order to quantify the absolute difference in subpopulation-specific allele frequencies in an straightforward, interpretable manner, we limited the Empirical data to that of the white and African American subpopulations.
+Figure \ref{fig:pop_inf} shows the relationship between a given SNP's false positive rate (FPR) and its absolute difference in standardized allele frequency between the two supopulations.
 
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/popInfSNPs_n=200.png}
-    \caption{Causal and non-causal SNP selection in the presence of varying levels of population informative SNPs with a coarse 2 Independent Subpopulation structure, dichotomous-discordant effects, $\eta = 0.5$, and $\xi = 0.8$.}
-    \label{fig:big_small_gamma}
+    \includegraphics[scale = 0.9]{figures/pop_inf_snps_loess.png}
+    \caption{\anna{Non-causal SNP selection / SNP FPR (?)} in the presence of varying levels of population informative SNPs and environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
+    \label{fig:pop_inf}
 \end{figure}
+
+\anna{maybe no table ? If table, make more like table 1}
+\begin{table}[H]
+\centering
+\begin{tabular}{rrlr}
+  \hline
+eta & xi & Method & $r_s$ \\ 
+  \hline
+0.5 & 0.0 & lasso & 0.34 \\ 
+  0.5 & 0.0 & PC-lasso & -0.92 \\ 
+  0.5 & 0.0 & LMM-lasso & -0.89 \\ 
+  0.5 & 0.8 & lasso & 0.78 \\ 
+  0.5 & 0.8 & PC-lasso & -0.82 \\ 
+  0.5 & 0.8 & LMM-lasso & -0.67 \\ 
+   \hline
+\end{tabular}
+\caption{Spearman's rank correlation coefficient between \anna{non-causal SNP selection / SNP FPR (?)} and absolute difference in standardized allele frequencies across methods.}
+\label{tab:pop_inf}
+\end{table}
+
+\anna{Another way to quantify this...(?)}
+\begin{table}[H]
+\centering
+\begin{tabular}{rlrrr}
+  \hline
+xi & chunks & lasso & lasso\_pca10 & plmm\_lasso \\ 
+  \hline
+0.0 & 1 & 0.108 & 0.187 & 0.135 \\ 
+  0.0 & 2 & 0.394 & 0.132 & 0.221 \\ 
+  0.0 & 3 & 0.214 & 0.049 & 0.071 \\ 
+  0.8 & 1 & 0.005 & 0.437 & 0.161 \\ 
+  0.8 & 2 & 0.504 & 0.188 & 0.348 \\ 
+  0.8 & 3 & 0.966 & 0.084 & 0.157 \\ 
+   \hline
+\end{tabular}
+\caption{Max FPR for each method, $\xi$, and level of difference (low, med, high)}
+\end{table}
+
+In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.34$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
+
+Overall, however, the FPR does not exceed 0.39, 0.19, and 0.22 for the lasso, PC-lasso, and LMM-lasso methods, respectively.  
+%\anna{We attribute these associations to the presence of population structure (in other words, correlated observations) and thus artificially shrunken errors in the case of the lasso.
+For PC-lasso and LMM-lasso, the absence of a non-genetic mechanism by which population structure directly affects the outcome, results in a slight over-correction for both, resulting in their moderately negative correlations.
+}
+
+Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.
+
+In the presence of confounding, however, lasso's correlation with allele frequency increases, and particularly struggles to correctly select the most population-informative SNPs.
+PC-lasso and LMM-lasso are able to remain more robust to allele frequency differences, yet remain somewhat negatively correlated with allele frequency differences.
+This is is advantageous in preventing dramatically high FPRs in the presence of highly population-informative SNPs.
+
+In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{apparent cryptic relatedness (?)} \anna{max absolute kinship?
+histograms of off-diag kinship?} in the Empirical data.
+This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
 

### 2020-12-03 main.tex
4.2 tweaks

@@
 
-In the previous section we saw how in the presence of confounding, LMM-lasso and PC-lasso outperform the ordinary lasso in terms of selection accuracy.
+In the previous section we saw how in the presence of confounding, LMM-lasso and PC-lasso outperform the ordinary lasso in selection accuracy.
 Additionally, as we described in Section \ref{sec:example}, subpopulation differences in allele frequencies are known to have the potential to lead to spurious associations.
@@
 
+\begin{comment}
 \anna{Another way to quantify this...(?)}
@@
 \end{table}
+\end{comment}
 
-In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.34$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
-
+\anna{Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.} In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.34$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
 Overall, however, the FPR does not exceed 0.39, 0.19, and 0.22 for the lasso, PC-lasso, and LMM-lasso methods, respectively.
@@
 
-Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.
+In the presence of confounding, however, lasso's correlation with allele frequency increases ($r_s = 0.78$), and particularly struggles to correctly select the most population-informative SNPs.
+PC-lasso and LMM-lasso are able to remain more robust to allele frequency differences, yet remain somewhat negatively correlated with allele frequency differences.
+This is advantageous in preventing dramatically high FPRs in the presence of highly population-informative SNPs.
 
-In the presence of confounding, however, lasso's correlation with allele frequency increases, and particularly struggles to correctly select the most population-informative SNPs.
-PC-lasso and LMM-lasso are able to remain more robust to allele frequency differences, yet remain somewhat negatively correlated with allele frequency differences.
-This is is advantageous in preventing dramatically high FPRs in the presence of highly population-informative SNPs.
-
-In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{apparent cryptic relatedness (?)} \anna{max absolute kinship?
-histograms of off-diag kinship?} in the Empirical data.
-This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
+In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{cryptic relatedness in addition to population structure in the Empirical data.} \anna{max absolute kinship?
+histograms of off-diag kinship?} This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
 

### 2020-12-03 main.tex
Section 4 meeting notes

@@
 We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.
+\anna{We examined these but found they didn't make a huge difference in performance so we don't focus on them.}
 
@@
 
-In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{cryptic relatedness in addition to population structure in the Empirical data.} \anna{max absolute kinship?
-histograms of off-diag kinship?} This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
+\anna{LMM-lasso seems to exhibit even stronger control of this than PC-lasso.}
+
+%In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{finer grained relatedness in addition to population stratification in the Empirical data.} This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
 

### 2020-12-04 main.tex
Figure 2 boxplot and related edits.

--- 2020-12-03 main.tex
@@
 We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.
-
-% For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography.
-However, behavior-based environmental exposures may be less likely to be shared by subpopulations which are genetically and geographically similar.
-As an example, \citet{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.
+\anna{We examined all combinations of these patterns of environmental effects, but found they did not result in appreciable differences in performance.
+We thus focus our attention elsewhere.}
 
@@
 Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.87$, with those from the simulated data sets is not straightforward.
-We note that the CP decreases in the presence of subpopulations of unequal sizes and with more outbred subpopulations as is the case in the Empirical data.
+We note that the CP decreases in the presence of subpopulations of unequal sizes and with more complex population structure as is the case in the Empirical data.
 
-% \anna{MSPE here (?)}
-
-% \anna{talk about karl rohe stuff here - if anywhere}
-
-In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) \anna{technically this is true sign rate - only estimates with the correct sign are counted.
-Explain this?
-Calculate overall selection instead?} for the first 50 variables each method allowed into the model.
-Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
+In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
+\anna{Throughout this work, we refer to a selection as true positive only if it is selected and the sign of the coefficient is estimated correctly.} Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
 The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
@@
 In order to quantify the absolute difference in subpopulation-specific allele frequencies in an straightforward, interpretable manner, we limited the Empirical data to that of the white and African American subpopulations.
-Figure \ref{fig:pop_inf} shows the relationship between a given SNP's false positive rate (FPR) and its absolute difference in standardized allele frequency between the two supopulations.
+Figure \ref{fig:pop_inf} summarizes the relationship between a given SNP's false positive rate (FPR) and its absolute difference in MAF between the two supopulations.
+The top panel of Figure \ref{fig:pop_inf} summarizes the marginal relationship between FPR and MAF differences, while the lower panel shows the association between SNPs falsely selected by each method and MAF differences.
 
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/pop_inf_snps_loess.png}
-    \caption{\anna{Non-causal SNP selection / SNP FPR (?)} in the presence of varying levels of population informative SNPs and environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
+    \includegraphics[width = \textwidth]{figures/pop_inf_snps.png}
+    \caption{\anna{Association of population informative nature of SNPs and false selection across methods in the presence and absence of environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}}
     \label{fig:pop_inf}
 \end{figure}
+
+%\begin{figure}[H]
+%\begin{subfigure}{\textwidth}
+%  \centering
+%  % include first image
+%  \includegraphics[scale = 0.9]{figures/pop_inf_snps_loess.png}  
+%  \caption{FPR in the presence of varying levels of population informative SNPs.}
+%  \label{fig:pop_inf_scatter}
+%\end{subfigure}
+%\begin{subfigure}{\textwidth}
+%  \centering
+%  % include second image
+%  \includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}  
+%  \caption{Absolute allele frequency differences of falsely selected, non-causal SNPs across methods.}
+%  \label{fig:pop_inf_box}
+%\end{subfigure}
+%\caption{Association of population informative nature of SNPs and false selection across methods in the presence and absence of environmental confounding for for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
+%\label{fig:pop_inf}
+%\end{figure}
 
@@
   \hline
-0.5 & 0.0 & lasso & 0.34 \\ 
-  0.5 & 0.0 & PC-lasso & -0.92 \\ 
-  0.5 & 0.0 & LMM-lasso & -0.89 \\ 
-  0.5 & 0.8 & lasso & 0.78 \\ 
-  0.5 & 0.8 & PC-lasso & -0.82 \\ 
-  0.5 & 0.8 & LMM-lasso & -0.67 \\ 
+0.5 & 0.0 & lasso & 0.42 \\ 
+  0.5 & 0.0 & PC-lasso & -0.77 \\ 
+  0.5 & 0.0 & LMM-lasso & -0.71 \\ 
+  0.5 & 0.8 & lasso & 0.67 \\ 
+  0.5 & 0.8 & PC-lasso & -0.61 \\ 
+  0.5 & 0.8 & LMM-lasso & -0.34 \\ 
    \hline
 \end{tabular}
-\caption{Spearman's rank correlation coefficient between \anna{non-causal SNP selection / SNP FPR (?)} and absolute difference in standardized allele frequencies across methods.}
+\caption{\anna{Spearman's rank correlation coefficient between SNP FPR and absolute difference in standardized allele frequencies across methods.}}
 \label{tab:pop_inf}
@@
 
-\begin{comment}
-\anna{Another way to quantify this...(?)}
-\begin{table}[H]
-\centering
-\begin{tabular}{rlrrr}
-  \hline
-xi & chunks & lasso & lasso\_pca10 & plmm\_lasso \\ 
-  \hline
-0.0 & 1 & 0.108 & 0.187 & 0.135 \\ 
-  0.0 & 2 & 0.394 & 0.132 & 0.221 \\ 
-  0.0 & 3 & 0.214 & 0.049 & 0.071 \\ 
-  0.8 & 1 & 0.005 & 0.437 & 0.161 \\ 
-  0.8 & 2 & 0.504 & 0.188 & 0.348 \\ 
-  0.8 & 3 & 0.966 & 0.084 & 0.157 \\ 
-   \hline
-\end{tabular}
-\caption{Max FPR for each method, $\xi$, and level of difference (low, med, high)}
-\end{table}
-\end{comment}
+\anna{Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.} In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.42$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
+Overall, however, the FPR does not exceed 0.39, 0.19, and 0.22 for the lasso, PC-lasso, and LMM-lasso methods, respectively.
 
-\anna{Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.} In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.34$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
-Overall, however, the FPR does not exceed 0.39, 0.19, and 0.22 for the lasso, PC-lasso, and LMM-lasso methods, respectively.
-%\anna{We attribute these associations to the presence of population structure (in other words, correlated observations) and thus artificially shrunken errors in the case of the lasso.
-For PC-lasso and LMM-lasso, the absence of a non-genetic mechanism by which population structure directly affects the outcome, results in a slight over-correction for both, resulting in their moderately negative correlations.
-}
-
-In the presence of confounding, however, lasso's correlation with allele frequency increases ($r_s = 0.78$), and particularly struggles to correctly select the most population-informative SNPs.
+In the presence of confounding, however, lasso's correlation with allele frequency increases ($r_s = 0.67$), and particularly struggles to correctly select the most population-informative SNPs.
 PC-lasso and LMM-lasso are able to remain more robust to allele frequency differences, yet remain somewhat negatively correlated with allele frequency differences.
 This is advantageous in preventing dramatically high FPRs in the presence of highly population-informative SNPs.
-
-In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{cryptic relatedness in addition to population structure in the Empirical data.} \anna{max absolute kinship?
-histograms of off-diag kinship?} This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
+From a marginal perspective, LMM-lasso seems to exhibit even stronger control of this than PC-lasso.
+This echoes the superior TPRs of LMM-lasso compared to those of the PC-lasso for the Empirical data in Table \ref{tab:var_sel}.
+Conditional on selection, however, LMM-lasso may be more likely to confuse MAF differences with true signal.
 

### 2020-12-10 main.tex
First attempt at including eps version of file - required for genetic epi submission
eps figs not working... trying to avoid merge conflicts
Edits to 3.5, 4, and 5

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\pb{work this citation in; maybe others like it: https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1009141}
 
@@
 \subsection{Correcting for population structure}
+\label{Sec:correcting-structure}
 
@@
 Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
-An important concept to note, and one that we will return to in Section \ref{sec:sim_env_conf}, is that population structure only leads to confounding if it is correlated with the environment.
+An important concept to note, and one that we will return to in Section \ref{sec:partition}, is that population structure only leads to confounding if it is correlated with the environment.
 In the presence of environmental heterogeneity which affects the phenotype of interest in a manner that is orthogonal to population structure, SNP-phenotype associations will not be confounded.
@@
 \subsection{LMM and PC models}
+\label{Sec:lmm-and-pc}
 
@@
 A key feature which differentiates \eqref{eqn:model} from traditional linear mixed models is that in our setting, $\bZ$ is not observed directly, although we still wish to correct for it in order to remove the effects of confounding influences.
-A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}; we will discuss the rationale for this approach in Section~\ref{sec:sim_env_conf}.
+A common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}; we will discuss the rationale for this approach in Section~\ref{sec:partition}.
 This, in turn, motivates the methods by which LMM-lasso and PC-lasso attempt to model unobserved confounding effects using observed genetic data.
@@
 \subsection{Partitioning the effects of environmental confounding}
-\label{sec:sim_env_conf}
+\label{sec:partition}
 
-In carrying out simulation studies of the LMM regression model \eqref{eqn:lmm_reg}, a common practice is to draw a new value of the quantity $\bX\bgamma$ from a $N(\boldsymbol{0}, \sigmagg \bK)$ for each replication of the analysis \citep{Rakitsch2012, bhatnagar2019simultaneous}.
-We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method in the presence of environmental heterogeneity.
+In Section~\ref{Sec:correcting-structure}, we remarked that population structure only leads to confounding if it is correlated with the
+environment, and in Section~\ref{Sec:lmm-and-pc}, we remarked that in LMM modeling, a typical albeit somewhat mysterious assumption is that $\bZ=\bX$.
+Here, we provide explicit mathematical justifications for those statements.
+In doing so, we derive a partitioning relationship that illustrates how LMMs achieve bias reduction due to environmental confounding.
+
+We begin by noting that in the GWAS context, the $\bgamma$ term in \eqref{eqn:model} is a fixed effect, although treated as ``random'' in the linear mixed model.
+This is done because the confounding environmental effects are unobserved; by considering them to be random, we can integrate over their unknown effects and obtain a marginal likelihood.
+However, in terms of a repeated sampling perspective, they are fixed.
+If repeated samples were drawn from the same populations, the confounding environmental effects would remain the same, not randomly change from sample to sample.
+
+Consider our earlier example, where $\bgamma$ represented the effect of air quality.
+Clearly, a GWAS that failed to account for population stratification would tend to identify as risk factors SNPs that were simply more common in the population with poor air quality.
+Treating $\bgamma$ as random, however, would mean that in one sample, a given population is exposed to good air quality, while in the next, it may be exposed to bad air quality.
+Any theoretical or numerical investigation of statistical properties would therefore fail to reveal any bias in this situation, as the two situations would cancel out.
+
+%We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method in the presence of environmental heterogeneity.
 Averaging over situations in which the environmental confounding is constantly changing direction and magnitude has the effect of ``washing out'' bias in $\bbetaHat$.
 
-Consider our earlier example involving confounding due to air quality.
-Simulating $\bgamma$ as a random quantity for each replication corresponds to subpopulation A being exposed to good air quality in one replication, and bad air quality in the next.
-Reporting the average across these situations is inconsistent with the question we are truly interested in, namely: what would happen if we took repeated samples from populations A and B, when these populations are subject to different environmental exposures?
-Those repeated samples would show a systematic bias due to the difference in air quality, but this is masked if $\bgamma$ is assumed to be random.
-Treating $\bgamma$ as a fixed quantity so that environmental effects are consistent across replications more closely approximates environmental confounding.
-
-To illustrate this concept mathematically, consider the low-dimensional ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for ease of notation:
+To illustrate this concept mathematically, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for the sake of simplicity:
 \begin{align}
@@
 \end{align}
-Under the assumption that environmental effects arise from a random process, $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
-However, under the assumption that the quantity $\bgamma$ is fixed, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma, \mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$, allowing us to quantify estimation bias due to environmental heterogeneity.
+Treating the environmental effects as random with $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
+Treating them as fixed, on the other hand, we have $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma$ and $\mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$.
+Treating $\bgamma$ as random gives the misleading impression that the unadjusted association between SNP and phenotype is unbiased.
 
-Although we derive these quantities under model \eqref{eqn:lmm_reg}, where $\bZ = \bX$, these results also hold for the more general model \eqref{eqn:model}, where $\bZ \ne \bX$.
-We see this by decomposing $\bZ \bgamma$ into components in and orthogonal to the column space of the projection matrix of $\bX$.
-Let $\boldsymbol{P}_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\boldsymbol{P}_{\bX})$ and $\mathcal{N}(\boldsymbol{P}_{\bX})$ denoting the column and null spaces of $\boldsymbol{P}_{\bX}$, respectively.
-Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}: \bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\bX \boldsymbol{\tau} \in \mathcal{C}(\boldsymbol{P}_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\boldsymbol{P}_{\bX})$.
+Let us then decompose the environmental effects $\bZ \bgamma$ into two parts: a component that can be projected onto the column space of $\bX$ and a component that cannot.
+Let $\P_{\bX} = \bX (\bX^\T \bX)^{-1} \bX^\T$ denote the projection matrix of $\bX$, with $\mathcal{C}(\P_{\bX})$ and $\mathcal{N}(\P_{\bX})$ denoting the column and null spaces of $\P_{\bX}$, respectively.
+Then there exist $\boldsymbol{\tau}, \boldsymbol{\psi}$ such that $\bZ \bgamma = \bX \boldsymbol{\tau + \psi}$, where $\bX \boldsymbol{\tau} \in \mathcal{C}(\P_{\bX})$ and $\boldsymbol{\psi} \in \mathcal{N}(\P_{\bX})$.
 This in turn means that
-\begin{align}
-    \label{eqn:partition_e}    \by &= \bX\bbeta + \bZ\bgamma + \beps \notag \\
-    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps \notag\\
-    &=  \bX \bbeta{^*} + \beps{^*};
-\end{align}
-note that $\boldsymbol{\tau}$ quantifies bias in the estimated $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term.
+\begin{equation}
+  \begin{aligned}
+    \label{eqn:partition_e}
+    \by &= \bX\bbeta + \bZ\bgamma + \beps \\
+    &= \bX\bbeta + \bX \boldsymbol{\tau} + \boldsymbol{\psi} + \beps \\
+    &=  \bX \bbeta{^*} + \beps{^*},
+  \end{aligned}
+\end{equation}
+where $\bbeta^* = \bbeta + \boldsymbol{\tau}$ and $\beps{^*} = \boldsymbol{\psi} + \beps$.
+Note that $\boldsymbol{\tau}$ quantifies bias in the na\"ive estimate $\bbetaHat^*$ and $\boldsymbol{\psi}$ is absorbed by the residual error term.
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
-This result provides the explicit mathematical justification for our statement from Section \ref{sec:background}, that SNP-phenotype associations will be unbiased in the presence of environmental heterogeneity which does not correspond to population structure.
+This result provides the explicit mathematical justification for the statement in Section~\ref{sec:background} that population structure only biases the SNP-phenotype association if it is correlated with the environmental effects (in the sense that $\bZ\bgamma$ can be projected onto the column space of $\bX$).
+It furthermore illustrates why the assumption $\bZ=\bX$ in the linear mixed model is reasonable, as well as how it works to remove bias in estimation.
 
-Furthermore, the proportion of the environmental effect in the column space of $\boldsymbol{P}_{\bX}$ (i.e., the portion that will cause confounding) is given by
+Finally, it is worth notiong that the proportion of environmental effects in the column space of $\P_{\bX}$ (i.e., the portion that will cause confounding) is given by
 \begin{equation}
@@
 which we refer to as the Confounding Proportion (CP).
-The CP can range from 0 (no confounding) to 1 (entirely confounded).
-In the extreme case where $\bZ \bgamma$ is completely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\boldsymbol{P}_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
-In this scenario, the SNP-phenotype associations will be unbiased, but the residual error will be inflated.
+The CP can range from 0 (no confounding) to 1.
+In the extreme case where $\bZ \bgamma$ is completely independent of population structure, $\bZ \bgamma \in \mathcal{N}(\P_{\bX})$ and $\bX \boldsymbol{\tau} = \boldsymbol{0}$ so that $\mathbb{E}(\bbetaHat^*) = \bbeta$.
+In this scenario, the SNP-phenotype associations will be unbiased, although the residual error will be inflated.
 
 In real data settings CP cannot be calculated since $\bZ \bgamma$ is not directly observed.
-In simulation settings, however, this ratio is a useful metric for quantifying the degree to which $\bZ \bgamma$ reflects population structure and has the potential to bias SNP-phenotype associations.
-We will use CP in our interpretation of the simulation results in Section \ref{sec:results}.
-Given the high-dimensional nature of our data, we use a low-rank approximation of $(\bX^\T \bX)^{-1}$ in order to ensure a unique solution.
+In simulation settings, however, this ratio is a useful metric for quantifying the degree to which $\bZ \bgamma$ reflects population structure and has the potential to bias observed associations.
+We use CP to help interpret the simulation results in Section \ref{sec:results}.
+%In high dimensions, we use a low-rank approximation of $(\bX^\T \bX)^{-1}$ to improve stability.
 
@@
 
-\subsubsection{Generation of synthetic phenotypes}
+\subsubsection{Synthetic phenotypes}
 
@@
 \end{equation}
-where $\by$ represents the synthetic phenotype, $\bgamma$ is a fixed quantity (see Section~\ref{sec:sim_env_conf}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
+where $\by$ represents the synthetic phenotype, $\bgamma$ is a fixed quantity (see Section~\ref{sec:partition}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
 \begin{equation}
@@
 
-\subsubsection{SNP matrix generation}
+\subsubsection{Genotypes}
+\label{sec:subpopulation_structure}
 
@@
 
-\subsubsection{Subpopulation structure}
-\label{sec:subpopulation_structure}
 For the fully simulated data, two subpopulation structures were considered, which we refer to as ``fine'' and ``coarse''.
@@
 
-For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$.
-Various $\bgamma$ patterns were considered in order to investigate the effects of distinct and complex patterns of environmental confounding.
-Firstly, we considered two different patterns of effect magnitudes, which we refer to as exponential and dichotomous.
+In our simulations, $\bZ$ is a matrix of indicator functions identifying subpopulation (founding population) membership, which assigns the corresponding environmental effect $\bgamma$.
+Various $\bgamma$ patterns were considered in order to investigate the effects of distinct patterns of environmental confounding.
+Simulations were carried out with dichotomous (high/low) environmental effects, as well as more complex heterogeneous effects of all different sizes.
+In addition, we examined the ``concordance'' of the environmental effect: whether individuals in genetically similar subpopulations experience similar environmental effects or not.
+In our experiments, however, neither the heterogeneity nor the concordance of the environmental effects had any meaningful impact on the performance of the methods under consideration.
+Thus, throughout Section~\ref{sec:results}, we present results only for disconcordant, dichotomous environmental effects, as the other results are very similar.
+
+%Firstly, we considered two different patterns of effect magnitudes, which we refer to as exponential and dichotomous.
 We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations, while exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes.
@@
 \subsubsection{Tuning parameter selection}
+
 Selection of the tuning parameter $\lambda$ plays an important role in all penalized regression methods, with many possible approaches.
@@
 %------------------------------%
-\subsection{PC-lasso and LMM-lasso outperform the ordinary lasso in the presence of environmental confounding}
+\subsection{PC-lasso and LMM-lasso increase accuracy}
+\label{sec:sim-mse}
 %------------------------------%
@@
 We begin by comparing the ordinary lasso, PC-lasso, and LMM-lasso as the level of environmental confounding is varied.
-Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE).
+Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE) \pb{include a formula}.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
-The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental confounding), lasso and LMM-lasso perform comparably.
+The top row of Figure \ref{fig:mse} shows that when $\xi = 0$ (i.e., no environmental effects), lasso and LMM-lasso perform comparably.
 PC-lasso results in a slight inflation in MSE compared to the other methods in the simulated data sets due to the fact that 10 additional predictors (PCs) have been added to the model, but none of them actually explain any variability in the outcome.
@@
     \centering
-    \includegraphics[scale = 1.1]{figures/beta_mse.png}
+    \includegraphics[width = \textwidth]{figures/beta_mse.png}
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
@@
 
-The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$.
+The bottom row of Figure \ref{fig:mse} shows the MSE when $\xi = 0.8$ (substantial environmental effects).
 In this case, both PC-lasso and LMM-lasso clearly outperform the ordinary lasso.
 This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
-This is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by nonoverlapping genetic subpopulations, meaning that a greater proportion of the environmental effect is in the column space of $\boldsymbol{P}_{\bX}$.
-We can quantify the proportion of the environmental component in the column space of $\boldsymbol{P}_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:sim_env_conf}.
-Here, we have CP $= 0.47, 0.99$ for the 1D Linear Admixture and Independent Subpopulations, respectively.
-The large CP of the Independent Subpopulations model reflects the extreme confounding and estimation bias for this data, for which the PC-lasso and LMM-lasso methods are able to correct.
-Meanwhile, the less dramatic improvement in MSE for the 1D Linear Admixture data by the PC-lasso and LMM-lasso methods confirms what we know from this data's substantially smaller CP value; that there is less confounding and bias which can be corrected for.
-
-Due to the differing sample sizes and subpopulation structures, a direct comparison of the Empirical data, which has a CP $= 0.87$, with those from the simulated data sets is not straightforward.
-We note that the CP decreases in the presence of subpopulations of unequal sizes and with more complex population structure as is the case in the Empirical data.
+This is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by nonoverlapping genetic subpopulations, meaning that a greater proportion of the environmental effect is in the column space of $\P_{\bX}$.
+We can quantify the proportion of the environmental component in the column space of $\P_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:partition}.
+Here, we have CP = 0.47, 0.87, and 0.99 for the 1D Linear Admixture, Empirical, and Independent Subpopulations, respectively.
+The large CP of the Empirical and Independent Subpopulations shows that the environmental effect is more highly correlated with genotype, and therefore has more potential for introducing bias into the estimation of SNP effects; the PC-lasso and LMM-lasso methods correct for this effect, while ordinary lasso does not.
+This phenomenon occurs in 1D Linear Admixture setting as well, although the phenomenon is less pronounced because, while the total environmental effect is the same in all three settings, in the admixture setting 53\% of the environmental effect is uncorrelated with genotype and therefore does not introduce bias.
 
 In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
-\anna{Throughout this work, we refer to a selection as true positive only if it is selected and the sign of the coefficient is estimated correctly.} Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
+Throughout, we use ``true positive'' to mean both that the SNP was selected and that its direction was estimated correctly.
+Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
 The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
@@
 \end{tabularx}
-\caption{True positive rates, mean (SD), for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
+\caption{True positive rates, mean (SD), for different data types in the presence ($\xi = 0$) and absence ($\xi = 0.8$) of environmental effects.
+For the data sets with simulated genotypes, \pb{???} subpopulations are present.}
 \label{tab:var_sel}
@@
 %------------------------------%
-\subsection{\anna{PC-lasso and LMM-lasso reduce bias towards selecting population-informative SNPs}}
+\subsection{PC-lasso and LMM-lasso reduce selection bias}
 %------------------------------%
 
-In the previous section we saw how in the presence of confounding, LMM-lasso and PC-lasso outperform the ordinary lasso in selection accuracy.
-Additionally, as we described in Section \ref{sec:example}, subpopulation differences in allele frequencies are known to have the potential to lead to spurious associations.
-As such it is of interest to characterize these methods' selection accuracy in terms of how robust they are to allele frequency differences.
+As described earlier in the manuscript, correlation between environmental effects and genotype introduces the potential to observe spurious associations in which SNPs that differ in frequency between subpopulations are identified as contributing to the phenotype.
+In this section, we characterize the three lasso methods (ordinary, PC, and LMM) in terms of their susceptibility to this form of selection bias.
 
-In order to assess the behavior of these methods in this context, we conducted a simulation using the parameter settings presented in Figure \ref{fig:mse} and Table \ref{tab:var_sel} and a modified version of the Empirical data.
-In order to quantify the absolute difference in subpopulation-specific allele frequencies in an straightforward, interpretable manner, we limited the Empirical data to that of the white and African American subpopulations.
-Figure \ref{fig:pop_inf} summarizes the relationship between a given SNP's false positive rate (FPR) and its absolute difference in MAF between the two supopulations.
-The top panel of Figure \ref{fig:pop_inf} summarizes the marginal relationship between FPR and MAF differences, while the lower panel shows the association between SNPs falsely selected by each method and MAF differences.
+In order to assess the behavior of these methods in this context, we conducted a simulation based on the Empirical genotype type, but restricted to just two subpopulations, Caucasian and African-American.
+\pb{DESCRIBE HOW SNPS CHOSEN HERE}
+As in the earlier simulations, the SNR was set to 1 ($\eta = 0.5$).
 
@@
 
-\anna{maybe no table ? If table, make more like table 1}
-\begin{table}[H]
-\centering
-\begin{tabular}{rrlr}
-  \hline
-eta & xi & Method & $r_s$ \\ 
-  \hline
-0.5 & 0.0 & lasso & 0.42 \\ 
-  0.5 & 0.0 & PC-lasso & -0.77 \\ 
-  0.5 & 0.0 & LMM-lasso & -0.71 \\ 
-  0.5 & 0.8 & lasso & 0.67 \\ 
-  0.5 & 0.8 & PC-lasso & -0.61 \\ 
-  0.5 & 0.8 & LMM-lasso & -0.34 \\ 
-   \hline
-\end{tabular}
-\caption{\anna{Spearman's rank correlation coefficient between SNP FPR and absolute difference in standardized allele frequencies across methods.}}
-\label{tab:pop_inf}
-\end{table}
+%% \anna{maybe no table ? If table, make more like table 1}
+%% \begin{table}[H]
+%% \centering
+%% \begin{tabular}{rrlr}
+%%   \hline
+%% eta & xi & Method & $r_s$ \\ 
+%%   \hline
+%% 0.5 & 0.0 & lasso & 0.42 \\ 
+%%   0.5 & 0.0 & PC-lasso & -0.77 \\ 
+%%   0.5 & 0.0 & LMM-lasso & -0.71 \\ 
+%%   0.5 & 0.8 & lasso & 0.67 \\ 
+%%   0.5 & 0.8 & PC-lasso & -0.61 \\ 
+%%   0.5 & 0.8 & LMM-lasso & -0.34 \\ 
+%%    \hline
+%% \end{tabular}
+%% \caption{\anna{Spearman's rank correlation coefficient between SNP FPR and absolute difference in standardized allele frequencies across methods.}}
+%% \label{tab:pop_inf}
+%% \end{table}
 
-\anna{Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.} In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.42$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
+%% \anna{Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.} In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.42$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
 Overall, however, the FPR does not exceed 0.39, 0.19, and 0.22 for the lasso, PC-lasso, and LMM-lasso methods, respectively.
 
-In the presence of confounding, however, lasso's correlation with allele frequency increases ($r_s = 0.67$), and particularly struggles to correctly select the most population-informative SNPs.
+Figure \ref{fig:pop_inf} summarizes each method's tendency to select spurious associations driven by differences in MAF between populations.
+Even in the absence of environmental effects, the ordinary lasso shows a preference for selecting population-informative SNPs, while PC-lasso and LMM-lasso are reluctant to select such SNPs (recall that with this design, the median MAF difference between populations is 0.25).
+
+In the presence of confounding, however, the lasso shows an overwhelming tendency to select population-informative SNPs.
+Its false positive selections are almost always the ones with the largest differences between populations.
+PC-lasso and LMM-lasso, on the other, show no such such selection bias.
+
+%Nearly all of its false positive selections are lasso's correlation with allele frequency increases ($r_s = 0.67$), and particularly struggles to correctly select the most population-informative SNPs.
 PC-lasso and LMM-lasso are able to remain more robust to allele frequency differences, yet remain somewhat negatively correlated with allele frequency differences.
@@
 \subsection{LMM-lasso outperforms PC-lasso for finer population structures}
+\label{sec:sim-fine-coarse}
 %------------------------------%
 
-Given the superior performance of PC-lasso and LMM-lasso compared to the ordinary lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures as defined in Section \ref{sec:subpopulation_structure}.
-Synthetic phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
+\pb{Need to explicitly say how many subpopulations in each case, both in text and legend.}
+
+Although PC-lasso and LMM-lasso outperform the ordinary lasso in the presence of environmental confounding, they are not equally effective across all population structures.
+Here, we compare the performance of these two methods in two settings, as described in Section~\ref{sec:subpopulation_structure}: ``coarse'' (\pb{XXX subpopulations}) and ``fine'' (\pb{XXX subpopulations}).
+Synthetic phenotypes were simulated using a 1:1 SNR ($\eta = 0.5$) and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
 
@@
 
-Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the Independent Subpopulations and a dichotomous-discordant environmental confounding pattern.
-The y-axis displays the difference in MSE (PC-lasso $-$ LMM-lasso) for each simulation and for the various subpopulation structures.
-The dashed horizontal line at 0 indicates no difference in the performance of PC-lasso and LMM-lasso, while differences above 0 correspond to LMM-lasso outperforming PC-lasso, and vice versa.
+Figure \ref{fig:big_vs_small} summarizes the results of these simulations.
+The vertical axis displays the difference in MSE: PC-lasso $-$ LMM-lasso, so that positive numbers correspond to LMM-lasso outperforming PC-lasso and vice versa.
+The dashed horizontal line at 0 indicates no difference in the performance the two methods.
 
-In all of these settings, LMM-lasso outperformed PC-lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
+In all settings, LMM-lasso outperformed PC-lasso, although the difference was considerably more pronounced when the subpopulation structure was fine.
 These findings are consistent with \citet{hoffman2013correcting}, who showed that PC adjustment methods represent a low dimensional approximation to LMMs.
-As such, it is reasonable that PC-lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure.
-
-\anna{Similar results were observed in the presence of other patterns of environmental effects.
-No meaningful differences in methodological performance were observed for dichotomous versus exponential effect magnitudes or concordant versus discordant effects.
-}
+As such, it is reasonable that PC-lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in settings with coarse population structures, but fall short when the structure is more complex.
 
 %------------------------------%
-\subsection{Different implementations of LMM-lasso}
+\subsection{Implementations of LMM-lasso}
 %------------------------------%
@@
 In Section \ref{sec:lmmlasso} we briefly mentioned the existence of two LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
-Both methods perform quite similarly in terms of $\bbeta$ MSE and variable selection.
+Both methods perform quite similarly in terms of estimation and variable selection for the settings presented in Sections~\ref{sec:sim-mse}-\ref{sec:sim-fine-coarse}.
 Because of this, the preceding results, though generated using the LMM-lasso-Rakitsch implementation, are representative of both LMM-lasso methods in the $n = 200, p = 1000$ setting.
 The primary purpose of this review is to compare LMM-lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding.
-However, given the generally strong performance of LMM-lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.
+However, given the generally strong performance of LMM-lasso, in this section we describe the key differences between these two implementations and provide guidance for potential users.
 
@@
 First, the variance components are estimated under the assumption of a null model.
-Then, the estimated variance components are used to transform the data, and the $\bbeta$ are estimated using coordinate descent.
+Then, the estimated variance components are used to transform the data, and $\bbeta$ is estimated using coordinate descent.
 LMM-lasso-ggmix seeks to improve upon this procedure by re-estimating the variance components and $\bbeta$ iteratively.
-Although this seems a reasonable approach, we have been unable to find a scenario where LMM-lasso-ggmix reliably outperforms the LMM-lasso-Rakitsch in estimation accuracy of $\bbeta$.
+Although this seems a reasonable idea, our experiments reveal that it makes little to no difference in terms of estimating $\bbeta$.
 In order to systematically compare both methods, we carried out simulations where $p = 1000$, $n = \{200, 400, 600, 800, 1000, 1200, 1400\}$, and $\eta = \{0.2, 0.5, 0.8\}$.
-Figure \ref{fig:eta_beta_mse} shows the MSE for estimation of the parameters $\bbeta$ and $\eta$, in addition to the corresponding computing time required.
+Figure \ref{fig:eta_beta_mse} shows the MSE for estimation of $\bbeta$ and $\eta$, in addition to the corresponding computing time required.
 
-LMM-lasso-Rakitsh and LMM-lasso-ggmix achieve nearly identical MSE for $\bbeta$ in many scenarios, particularly where $n \ll p$, as shown in the top panel of Figure \ref{fig:eta_beta_mse}.
-However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, and in some cases worse than that of LMM-lasso-Rakitsch.
+LMM-lasso-Rakitsh and LMM-lasso-ggmix achieve nearly identical MSE for $\bbeta$ in almost all scenarios, particularly when $n \ll p$ (Figure \ref{fig:eta_beta_mse}, top panel).
+However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, occasionally far worse than that of LMM-lasso-Rakitsch.
 This is most dramatic in the presence of substantial environmental environmental confounding, $\xi = 0.8$, and $n = \{1200, 1400\}$.
 
-LMM-lasso-ggmix achieves an improvement in the estimation of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$, as shown in the middle panel of Figure \ref{fig:eta_beta_mse}.
-These improved $\eta$ estimates do not correspond to smaller $\bbeta$ MSE for LMM-lasso-ggmix and thus do not seem to improve its estimation of the fixed genetic effects.
-In addition, when $n \ge 1000$, LMM-lasso-ggmix's estimation of $\eta$ suffers.
-When $n \ge 1000$ and substantial environmental confounding is present, $\xi = 0.8$,  LMM-lasso-ggmix's performance is hindered to the point that LMM-lasso-Rakitsch estimates $\eta$ more accurately.
+LMM-lasso-ggmix does tend to provide improved estimates of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$ (\ref{fig:eta_beta_mse}, middle panel), although these improvements did not help it to estimate $\bbeta$ more accurately.
+In addition, at high signal-to-noise ratios ($\eta=0.8$), the performance of LMM-lasso-ggmix was somewhat erratic.
 
-Overall these results show that LMM-lasso-ggmix performs as well as LMM-lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-lasso-Rakitsch in estimating $\eta$ when $n \ll p$.
+Overall these results show that LMM-lasso-ggmix performs similarly to LMM-lasso-Rakitsch in estimating $\bbeta$ and outperforms LMM-lasso-Rakitsch in estimating $\eta$ when $n \ll p$.
 However, these are also the settings in which LMM-lasso-ggmix is least computationally efficient.
 The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n \ll p$, LMM-lasso-ggmix is up to 60 times slower than LMM-lasso-Rakitsch.
-Conversely, in settings where $n \ge p$, LMM-lasso-ggmix is up to 10 times faster than LMM-lasso-Rakitsch.
-However, these are also settings in which LMM-lasso Rakitsch substantially outperforms LMM-lasso-ggmix in estimating $\bbeta$ and $\eta$ accurately and consistently.
+In settings where $n \ge p$, LMM-lasso-ggmix is ``faster'' than LMM-lasso-Rakitsch, but this seems to be due to the algorithm getting stuck in a local minimum and prematurely exiting, a phenomenon that also explains the extremely poor estimation accuracy of LMM-lasso-ggmix in the $\eta=0.8$, $n=1400$ setting.
 
-LMM-lasso-ggmix's poor performance in these scenarios are likely attributable to a number of a factors such as an ill-conditioned RRM  in $n \ge p$ settings \citep{ledoit2004well}.
-In addition, because LMM-lasso-ggmix attempts to estimate more parameters than LMM-lasso-Rakitsch, its optimization problem is not necessarily convex, making it prone to getting stuck in local minima before it is able to extensively explore the coefficient path.
-An examination of the number of iterations required for convergence for each $\lambda$ in the coefficient path suggests that this is the case for our various simulation settings (Supplementary Figure \ref{fig:niter}).
-Converging to a local minimum in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, which is what we observe in Figure \ref{fig:eta_beta_mse}.
+The underlying reason for this is that the RRM becomes ill-conditioned when $n > p$ \citep{ledoit2004well}.
+Although both approaches utilize the RRM, because LMM-lasso-ggmix attempts to estimate more parameters than LMM-lasso-Rakitsch, its optimization problem is no longer convex, making it prone to becoming stuck in local minima before it is able to sufficiently explore the coefficient path.
+This phenomenon is illustrated in greater detail in Supplementary Figure \ref{fig:niter}.
+%Converging to a local minimum in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, which is what we observe in Figure \ref{fig:eta_beta_mse}.
 
@@
     \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
+     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
+\pb{Can we make this figure less tall?}}
     \label{fig:eta_beta_mse}
@@
 
-In this work, we have reviewed the concept of population structure in a framework which formally differentiates between effects of population structure and those of the environment.
+In this work, we reviewed the concept of population structure in a framework which formally differentiates between effects of population structure and those of the environment.
 We have provided a broad outline of existing  methods for ameliorating the effects of these unobserved confounding influences.
-In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data, now applied within a novel penalized framework: PC-lasso and LMM-lasso.
+In addition, we have provided a thorough review of two penalized regression approaches that correct for population structure in genetic data: PC-lasso and LMM-lasso.
 
 We have detailed the assumptions made by PC-lasso and LMM-lasso, which model confounding influences as fixed and random effects, respectively.
-We have also described the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity.
-In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated.
-We then compare the estimation and variable selection accuracy of the ordinary lasso, PC-lasso, and LMM-lasso in various confounding scenarios.
+We have also described the ways in which prior analyses of these methods may fail to accurately capture the effects of environmental effects in terms of contributing to selection and estimation bias.
+We hope that our result partitioning environmental effects into confounding and non-confounded components will be particularly useful in terms of understanding the performance of these approaches.
 
-Unsurprisingly, we find that PC-lasso and LMM-lasso outperform the ordinary lasso whenever environmental confounding effects are present.
-However, in the absence of such effects, PC-lasso may hinder estimation accuracy and perform more poorly than even the ordinary lasso.
-This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability.
-Additionally, we find that LMM-lasso consistently outperforms PC-lasso in the presence of fine subpopulation structure, while PC-lasso appears to its best advantage in the presence of coarse subpopulation structure.
-LMM-lasso may be a more robust method for practitioners seeking to obtain unbiased estimates when the nature of the population structure is unknown.
+Our simulations experiments reveal, unsurprisingly, that PC-lasso and LMM-lasso outperform the ordinary lasso whenever environmental confounding effects are present.
+However, in the absence of such effects, PC-lasso may hinder estimation accuracy.
+Additionally, we find that LMM-lasso consistently outperforms PC-lasso, especially when the population structure is more complex.
+These results indicate that LMM-lasso is a more robust method in practice when the nature of the population structure is unknown.
 
 We considered two implementations of the LMM-lasso method: LMM-lasso-Rakitsch and LMM-lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
-Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-lasso-ggmix has several drawbacks.
-Firstly, it suffers from estimation instability in $n \ge p$ settings.
-Secondly, although LMM-lasso-Rakitsch and LMM-lasso-ggmix perform comparably when $n \ll p$, LMM-lasso-ggmix requires up to 60 times the computing time of LMM-lasso-Rakitsch in these settings due to its iterative estimation procedure.
-The additional time required of LMM-lasso-ggmix may be worthwhile if the objective of an analysis is to accurately estimate the narrow-sense heritability in $n \ll p$ settings.
-However, the LMM-lasso-Rakitsch method is more robust to varying levels of environmental confounding and sample sizes and thus more broadly applicable.
+Although both methods perform similarly in terms of estimating $\bbeta$, the much greater computational efficiency and numerical stability of LMM-lasso-Rakitsch make it a more attractive choice in most scenarios, although LMM-lasso-ggmix may be useful if the objective of the analysis is to accurately estimate narrow-sense heritability.
 
@@
 
-As an additional point of interest, recall that we provide qualitative and quantitative explanations of how the degree to which differential environmental exposures may confound the genotype-phenotype relationship is dependent upon its similarity to population structure.
-While the objective of these explanations, and this work as a whole, is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved and unmodeled environmental effects could lead to substantial bias in narrow-sense heritability estimates.
-A more detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper.
-However, we hope this emphasizes the broad implications and importance of the subtle distinction between the effect of population structure and unobserved environmental heterogeneity.
+Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.
+In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{tau}$ unless steps are taken to address population structure.
+A detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper, but we hope this emphasizes the broad implications and importance of distinguishing between population structure and environmental effects.
 
-Finally, it is worth nothing that even in scenarios where LMM-lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods.
-Because of this, bias-reduction techniques for penalized LMMs are a potentially attractive avenue for future research.
+Finally, we note that this article has focused entirely on the lasso as a penalization scheme.
+This is reasonable, as it is the most widely used method.
+However, many other penalized regression approaches have been developed that aim to reduce bias, better account for correlation between features, and to account for the grouping of SNPs into genes.
+All of these approaches are potentially attractive avenues for future research into GWAS methodology.

### 2020-12-11 main.tex
Updates from Overleaf
Edits to 1, 3, 4, 5, bibliography, eps figure format

@@
 
-\pb{work this citation in; maybe others like it: https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1009141}
-
 The objective of many genetic studies is to accurately identify variants associated with complex traits or diseases, and to estimate the effects of those variants.
@@
 However, this effort may be hindered by several factors such as non-independent samples, large numbers of variants, and linkage disequilibrium (LD).
-
-% \anna{keep this for later use when talking about genetic background?} Genetic variants are often correlated with each other such that univariate testing may be prone to false positives and biased estimates.
-When a genetic variant of interest is highly correlated with others, the effects of these will be attributed to the variant in question.
-This may result in causal SNPs being removed from analysis, so that only SNPs in LD with the causal loci, as opposed to the causal loci themselves, can be identified and their effects estimated.
 
@@
 These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
+The lasso is widely used in this context and its application to ultra high-dimensional genetic data is the subject of recent research \citep{qian2020fast, prive2018efficient}.
+%The lasso has been increasingly used in this context; of particular note is its use in the analysis of the UK Biobank data \citep{qian2020fast, prive2018efficient}.
 
@@
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2019simultaneous, Sillanpaeae2011}.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2020simultaneous, Sillanpaeae2011}.
 While many standard correction techniques may in principal be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and non-experimental conditions \citep{lawson2019population, barton2019population}.
@@
 The remainder of this paper is organized as follows.
-In Section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this paper.
+In Section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this work.
 We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures \citep{Sillanpaeae2011, sul2018population, vilhjalmsson2012nature, barton2019population}.
@@
 
-Population structure is defined by the existence of allele frequency differences that define sub-populations and is driven by the combined effects of evolutionary processes such as genetic drift, migration, demographic history, and natural selection \citep{gibson2015primer, tibayrenc2017genetics}.
+Population structure is defined by the existence of allele frequency differences that define subpopulations and is driven by the combined effects of evolutionary processes such as genetic drift, migration, demographic history, and natural selection \citep{gibson2015primer, tibayrenc2017genetics}.
 Population structure is broadly categorized based on whether it describes recent or ancient relatedness.
@@
 Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
-An important concept to note, and one that we will return to in Section \ref{sec:partition}, is that population structure only leads to confounding if it is correlated with the environment.
+An important concept to note, and one that we will return to in Section \ref{sec:partition}, is that population structure only leads to confounding if it is correlated with distinct environmental effects.
 In the presence of environmental heterogeneity which affects the phenotype of interest in a manner that is orthogonal to population structure, SNP-phenotype associations will not be confounded.
@@
 Assuming for now that population structure is correlated with environmental exposures and can be accurately inferred from available data, we return to our consideration of methods that attempt to control for confounding via the mechanism depicted in diagram \eqref{diag:ps_env_block}.
-Such methods include LMMs, principal-component-based approaches, and stratified analyses.
-Stratified analyses refer to a general class of methods which attempt to infer subject membership within a discrete number of nonoverlapping subpopulations.
+Such methods include LMMs, principal-component-based approaches, and structured association.
+Structured association refers to a general class of methods which attempt to infer subject membership within a discrete number of nonoverlapping subpopulations.
 Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.
@@
 
-% rely on a pair-wise similarity matrix
 Like PC-adjustment methods, LMMs utilize the left singular vectors of a similarity matrix to estimate and subsequently correct for relatedness among subjects.
@@
 
-Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependencies among subjects \citep{Rakitsch2012, bhatnagar2019simultaneous}.
+Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependencies among subjects \citep{Rakitsch2012, bhatnagar2020simultaneous}.
 This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes.
@@
 Specifically, PC-adjustment and LMM methods, which we will refer to as PC-lasso and LMM-lasso, respectively.
-We consider two similar implementations of the LMM-lasso method, one developed by \citet{Rakitsch2012} and a more recent implementation from \citet{bhatnagar2019simultaneous}.
+We consider two similar implementations of the LMM-lasso method, one developed by \citet{Rakitsch2012} and a more recent implementation from \citet{bhatnagar2020simultaneous}.
 Subsequent references to LMM-lasso apply to both implementations, unless otherwise stated.
@@
 Using the singular value decomposition (SVD) of the standardized genotype matrix, we can derive the spectral decomposition of the RRM:
-
 \begin{equation}
@@
 A lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
-LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2019simultaneous}.
+LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}.
 This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
 We further discuss differences between these methods in Section \ref{sec:results}.
-
-%------------------------------%
-% \subsubsection{\anna{Intuition from GLS (?) p-values and SEs in response to rotation?}}
 
@@
 
-%We argue that this is somewhat misleading from the perspective of understanding the statistical properties of the method in the presence of environmental heterogeneity.
-Averaging over situations in which the environmental confounding is constantly changing direction and magnitude has the effect of ``washing out'' bias in $\bbetaHat$.
-
 To illustrate this concept mathematically, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for the sake of simplicity:
@@
 
-Finally, it is worth notiong that the proportion of environmental effects in the column space of $\P_{\bX}$ (i.e., the portion that will cause confounding) is given by
+Finally, it is worth noting that the proportion of environmental effects in the column space of $\P_{\bX}$ (i.e., the portion that will cause confounding) is given by
 \begin{equation}
@@
 For all data types $p = 1000$ with 50 causal SNPs.
-For the simulated 1D Linear Admixture and Independent Subpopulations data $n = \{200, 400, 600, 800 \}$, and $n = 588$ for the Empirical data.
+For the simulated 1D Linear Admixture and Independent Subpopulations data $n = 200$ unless otherwise specified, and $n = 588$ for the Empirical data.
 
@@
 In our experiments, however, neither the heterogeneity nor the concordance of the environmental effects had any meaningful impact on the performance of the methods under consideration.
-Thus, throughout Section~\ref{sec:results}, we present results only for disconcordant, dichotomous environmental effects, as the other results are very similar.
-
-%Firstly, we considered two different patterns of effect magnitudes, which we refer to as exponential and dichotomous.
-We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations, while exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes.
-Secondly, we considered whether individuals in genetically similar subpopulations would experience similar environmental effects.
-We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.
-\anna{We examined all combinations of these patterns of environmental effects, but found they did not result in appreciable differences in performance.
-We thus focus our attention elsewhere.}
+Thus, throughout Section~\ref{sec:results}, we present results only for discordant, dichotomous environmental effects, as the other results are very similar.
 
@@
 We begin by comparing the ordinary lasso, PC-lasso, and LMM-lasso as the level of environmental confounding is varied.
-Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE) \pb{include a formula}.
+Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error of the coefficients (MSE).
+Suppose $\bbetaHat(\lambda)$ is the vector of coefficients estimated by the lasso with a penalty of $\lambda$, and $\bbeta$ the coefficient values used to simulate the data.
+Mean squared error is then defined as
+\begin{equation}
+    \text{MSE} = \mathbb{E}||\bbeta - \bbetaHat(\lambda)||^2_2.
+\end{equation}
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/beta_mse.png}
+    \includegraphics[width = \textwidth]{figures/beta_mse.eps}
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
@@
 In order to assess the behavior of these methods in this context, we conducted a simulation based on the Empirical genotype type, but restricted to just two subpopulations, Caucasian and African-American.
-\pb{DESCRIBE HOW SNPS CHOSEN HERE}
+1000 SNPs were chosen from the genome-wide Empirical data such that the selected SNPs' absolute difference in MAF between the two subpopulations ranged uniformly from 0 to 0.5.
 As in the earlier simulations, the SNR was set to 1 ($\eta = 0.5$).
@@
 \centering
-    \includegraphics[width = \textwidth]{figures/pop_inf_snps.png}
-    \caption{\anna{Association of population informative nature of SNPs and false selection across methods in the presence and absence of environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}}
+\includegraphics[width = \textwidth]{figures/pop_inf_snps_box.eps}
+\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
 \label{fig:pop_inf}
 \end{figure}
-
-%\begin{figure}[H]
-%\begin{subfigure}{\textwidth}
-%  \centering
-%  % include first image
-%  \includegraphics[scale = 0.9]{figures/pop_inf_snps_loess.png}  
-%  \caption{FPR in the presence of varying levels of population informative SNPs.}
-%  \label{fig:pop_inf_scatter}
-%\end{subfigure}
-%\begin{subfigure}{\textwidth}
-%  \centering
-%  % include second image
-%  \includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}  
-%  \caption{Absolute allele frequency differences of falsely selected, non-causal SNPs across methods.}
-%  \label{fig:pop_inf_box}
-%\end{subfigure}
-%\caption{Association of population informative nature of SNPs and false selection across methods in the presence and absence of environmental confounding for for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
-%\label{fig:pop_inf}
-%\end{figure}
-
-%% \anna{maybe no table ? If table, make more like table 1}
-%% \begin{table}[H]
-%% \centering
-%% \begin{tabular}{rrlr}
-%%   \hline
-%% eta & xi & Method & $r_s$ \\ 
-%%   \hline
-%% 0.5 & 0.0 & lasso & 0.42 \\ 
-%%   0.5 & 0.0 & PC-lasso & -0.77 \\ 
-%%   0.5 & 0.0 & LMM-lasso & -0.71 \\ 
-%%   0.5 & 0.8 & lasso & 0.67 \\ 
-%%   0.5 & 0.8 & PC-lasso & -0.61 \\ 
-%%   0.5 & 0.8 & LMM-lasso & -0.34 \\ 
-%%    \hline
-%% \end{tabular}
-%% \caption{\anna{Spearman's rank correlation coefficient between SNP FPR and absolute difference in standardized allele frequencies across methods.}}
-%% \label{tab:pop_inf}
-%% \end{table}
-
-%% \anna{Table \ref{tab:pop_inf} displays the Spearman's rank correlation coefficient for each of these scenarios.} In the absence of environmental confounding, the lasso demonstrates a moderately positive correlation with allele frequency differences ($r_s = 0.42$), while the PC-lasso and LMM-lasso methods are negatively correlated with allele frequency differences.
-Overall, however, the FPR does not exceed 0.39, 0.19, and 0.22 for the lasso, PC-lasso, and LMM-lasso methods, respectively.
 
@@
 
-%Nearly all of its false positive selections are lasso's correlation with allele frequency increases ($r_s = 0.67$), and particularly struggles to correctly select the most population-informative SNPs.
-PC-lasso and LMM-lasso are able to remain more robust to allele frequency differences, yet remain somewhat negatively correlated with allele frequency differences.
-This is advantageous in preventing dramatically high FPRs in the presence of highly population-informative SNPs.
-From a marginal perspective, LMM-lasso seems to exhibit even stronger control of this than PC-lasso.
-This echoes the superior TPRs of LMM-lasso compared to those of the PC-lasso for the Empirical data in Table \ref{tab:var_sel}.
-Conditional on selection, however, LMM-lasso may be more likely to confuse MAF differences with true signal.
-
 %------------------------------%
@@
 
-\pb{Need to explicitly say how many subpopulations in each case, both in text and legend.}
-
 Although PC-lasso and LMM-lasso outperform the ordinary lasso in the presence of environmental confounding, they are not equally effective across all population structures.
-Here, we compare the performance of these two methods in two settings, as described in Section~\ref{sec:subpopulation_structure}: ``coarse'' (\pb{XXX subpopulations}) and ``fine'' (\pb{XXX subpopulations}).
+Here, we compare the performance of these two methods in two settings, as described in Section~\ref{sec:subpopulation_structure}: ``coarse'' (4 subpopulations) and ``fine'' (100 subpopulations).
 Synthetic phenotypes were simulated using a 1:1 SNR ($\eta = 0.5$) and varying levels of environmental confounding, $\xi \in \{0.2, 0.5,0.8\}$.
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
+    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.eps}
     \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
@@
 This phenomenon is illustrated in greater detail in Supplementary Figure \ref{fig:niter}.
-%Converging to a local minimum in this manner would lead to few iterations required for each $\lambda$ value before convergence, allowing for an advantage in terms of computational speed, but poor estimation, which is what we observe in Figure \ref{fig:eta_beta_mse}.
 
@@
     \centering
-    \includegraphics[scale = 1]{figures/eta_beta_hat.png}
-     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
-\pb{Can we make this figure less tall?}}
+    \includegraphics[width = \textwidth]{figures/eta_beta_hat.eps}
+     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
     \label{fig:eta_beta_mse}

### 2020-12-11 supplement.tex
Edits to 1, 3, 4, 5, bibliography, eps figure format

--- 2020-09-29 supplement.tex
@@
 
+% 1D LINEAR ADMIXTURE KINSHIP
 \begin{figure}[H]
     \centering
-    \includegraphics[scale = 0.45]{figures/niter_perLambda_point.png}
+    \includegraphics[scale = 1]{figures/admixed_kinship.eps}
+    \caption{1D Linear Admixture with coarse subpopulation structure (4 subpopulations, 50 subjects per subpopulation), RRM.}
+    \label{fig:admixed}
+\end{figure}
+
+% INDEPENDENT SUBPOPULATIONS, COARSE, KINSHIP
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/indep_coarse_kinship.eps}
+    \caption{Independent Subpopulations with coarse subpopulation structure (4 subpopulations, 50 subjects per subpopulation), RRM.}
+    \label{fig:indep_coarse}
+\end{figure}
+
+% INDEPENDENT SUBPOPULATIONS, FINE, KINSHIP
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/indep_fine_kinship.eps}
+    \caption{Independent Subpopulations with fine subpopulation structure (100 subpopulations, 2 subjects per subpopulation), RRM.}
+    \label{fig:indep_fine}
+\end{figure}
+
+% EMPIRICAL KINSHIP
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale = 1]{figures/empirical_kinship.eps}
+    \caption{Empirical RRM.}
+    \label{fig:empirical}
+\end{figure}
+
+\begin{sidewaysfigure}
+    \centering
+    \includegraphics[width=\textwidth, height=\textheight, keepaspectratio]{figures/niter_perLambda_point.png}
     \caption{The number of iterations per $\lambda$ value for LMM-Lasso-Rakitsch and LMM-Lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
 Note that 100 is the maximum number of iterations for both implementations of LMM-Lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
-    \label{fig:s4}
-\end{figure}
+    \label{fig:niter}
+\end{sidewaysfigure}

### 2020-12-11 supplement.tex
Supplement

@@
     \includegraphics[width=\textwidth, height=\textheight, keepaspectratio]{figures/niter_perLambda_point.png}
-    \caption{The number of iterations per $\lambda$ value for LMM-Lasso-Rakitsch and LMM-Lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
-Note that 100 is the maximum number of iterations for both implementations of LMM-Lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
+    \caption{The number of iterations per $\lambda$ value for LMM-lasso-Rakitsch and LMM-lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
+Note that 100 is the maximum number of iterations for both implementations of LMM-lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}
     \label{fig:niter}

### 2020-12-11 main.tex
3.6.2, 4.1, 5 edits

@@
 For all data types $p = 1000$ with 50 causal SNPs.
-For the simulated 1D Linear Admixture and Independent Subpopulations data $n = 200$ unless otherwise specified, and $n = 588$ for the Empirical data.
+For the simulated 1D Linear Admixture and Independent Subpopulations data \anna{$n = 200$ unless otherwise specified}, and $n = 588$ for the Empirical data.
 
@@
 \caption{True positive rates, mean (SD), for different data types in the presence ($\xi = 0$) and absence ($\xi = 0.8$) of environmental effects.
-For the data sets with simulated genotypes, \pb{???} subpopulations are present.}
+For the data sets with simulated genotypes, 4 subpopulations are present.}
 \label{tab:var_sel}
@@
 
-LMM-lasso-ggmix does tend to provide improved estimates of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$ (\ref{fig:eta_beta_mse}, middle panel), although these improvements did not help it to estimate $\bbeta$ more accurately.
+LMM-lasso-ggmix does tend to provide improved estimates of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$ (Figure \ref{fig:eta_beta_mse}, middle panel), although these improvements did not help it to estimate $\bbeta$ more accurately.
 In addition, at high signal-to-noise ratios ($\eta=0.8$), the performance of LMM-lasso-ggmix was somewhat erratic.
@@
 Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.
-In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{tau}$ unless steps are taken to address population structure.
+In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{\tau}$ unless steps are taken to address population structure.
 A detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper, but we hope this emphasizes the broad implications and importance of distinguishing between population structure and environmental effects.

### 2020-12-16 main.tex
Minor word tweaks throughout

@@
 An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
-However, classical approaches rely on collecting data with more observations than unknown parameters.
+However, classical multivariate approaches rely on collecting data with more observations than unknown parameters.
 This is typically infeasible in genome-wide studies where the number of features greatly exceeds the number of observations.
@@
 These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
-The lasso is widely used in this context and its application to ultra high-dimensional genetic data is the subject of recent research \citep{qian2020fast, prive2018efficient}.
-%The lasso has been increasingly used in this context; of particular note is its use in the analysis of the UK Biobank data \citep{qian2020fast, prive2018efficient}.
+The lasso is widely used in this context and its application to ultra high-dimensional genetic data is the subject of recent computaional and methodological research \citep{qian2020fast, prive2018efficient}.
 
@@
 
-Population structure is defined by the existence of allele frequency differences that define subpopulations and is driven by the combined effects of evolutionary processes such as genetic drift, migration, demographic history, and natural selection \citep{gibson2015primer, tibayrenc2017genetics}.
+Population structure is defined by the existence of allele frequency differences that distinguish subpopulations and is driven by the combined effects of evolutionary processes such as genetic drift, migration, demographic history, and natural selection \citep{gibson2015primer, tibayrenc2017genetics}.
 Population structure is broadly categorized based on whether it describes recent or ancient relatedness.
@@
 Population structure is a common phenomenon in genetic data.
-Varying levels of relatedness are almost always present among genetic samples, even in samples of unrelated individuals and seemingly homogeneous populations.
+Varying levels of relatedness are almost always present among genetic samples, even in those of unrelated individuals and seemingly homogeneous populations.
 For example, European American, Han Chinese, and most recently, cohorts within the UK Biobank data have been shown to exhibit patterns of population and geographic structure despite their seemingly similar subjects \citep{campbell2005demonstrating, xu2009genomic, chen2009genetic, haworth2019apparent}.
@@
 Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are present.
-Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations when population structure is associated with differences in both allele frequency and the trait or disease \citep{gibson2015primer}.
+Population stratification in particular has been of great concern in genetic studies due to its potential to lead to spurious associations when population structure is associated with differences in both allele frequency and the trait or disease of interest \citep{gibson2015primer}.
 This phenomenon is commonly described as \textbf{confounding due to population stratification}.
@@
 Also suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
-A GWAS of data from these subpopulations would find SNP $X$ to be significantly associated with lung cancer.
+A GWAS of data from these subpopulations would likely find SNP $X$ to be significantly associated with lung cancer.
 This spurious association is not identified because of the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
 Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be found to be associated with the phenotype.
-
-% \anna{description of DAG below, in reference to previous one}
 
@@
 \end{tikzpicture}
-%\caption{Directed acyclic graph (DAG) of confounding due to population structure.}
 \label{diag:ps_env}
 \end{equation}
+
+%\begin{equation}
+%\centering
+%\includegraphics{figures/tikz1.eps}
+%\label{diag:ps_env}
+%\end{equation}
 
@@
 One of the first methods to address the problem of confounding due to population structure is the genomic control (GC) method.
-Because the effect of this confounding is to inflate the distribution of test statistics, GC introduces a factor to counteract this effect and restore the appropriate null distribution.
+Because the effect of this confounding is to inflate the distribution of association test statistics, GC introduces a factor to counteract this effect and restore the appropriate null distribution.
 This deflation factor is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
@@
 
-A different approach is to create an approximation of population structure using surrogate variables, and to adjust for these as an additional model covariates.
+A different approach is to create an approximation of population structure using surrogate variables, and to adjust for these as additional model covariates.
 Principal component-adjustment (PC-adjustment) methods use the first several left singular vectors of a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
@@
 $$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI), $$
-where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ their corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \eqref{eqn:model}.
+where $\bU_{1:k}$ is the $n \times k$ matrix of principal component vectors, $\boldsymbol{\alpha}$ a corresponding $k \times 1$ coefficient vector, and all other parameters are as defined in equation \eqref{eqn:model}.
 In a penalized regression framework, this model yields the objective function
@@
 
-We generated these synthetic phenotypes using three types of high-dimensional ($p > n$) $\bG$ matrices, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
+We generated these synthetic phenotypes using three types of high-dimensional ($p > n$) $\bG$ matrices, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) \anna{a subset of} real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
 We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
@@
 \end{tabularx}
-\caption{True positive rates, mean (SD), for different data types in the presence ($\xi = 0$) and absence ($\xi = 0.8$) of environmental effects.
+\caption{True positive rates, mean (SD), for different data types in the absence ($\xi = 0$) and presence ($\xi = 0.8$) of environmental effects.
 For the data sets with simulated genotypes, 4 subpopulations are present.}
@@
 1000 SNPs were chosen from the genome-wide Empirical data such that the selected SNPs' absolute difference in MAF between the two subpopulations ranged uniformly from 0 to 0.5.
-As in the earlier simulations, the SNR was set to 1 ($\eta = 0.5$).
+As in the earlier simulations, a 1:1 SNR ($\eta = 0.5$) was used.
 
@@
 LMM-lasso-ggmix does tend to provide improved estimates of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$ (Figure \ref{fig:eta_beta_mse}, middle panel), although these improvements did not help it to estimate $\bbeta$ more accurately.
-In addition, at high signal-to-noise ratios ($\eta=0.8$), the performance of LMM-lasso-ggmix was somewhat erratic.
+In addition, at high signal-to-noise ratios ($\eta=0.8$), the performance of LMM-lasso-ggmix was \anna{highly variable}.
 
@@
 
-\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-lasso-Rakitsch method using the equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
+\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
 

### 2020-12-17 main.tex
Getting one tikz.eps to work (?)
final edits, figures formatted for pdf submission

@@
 %\centering
-%\includegraphics{figures/tikz1.eps}
+%\includegraphics{figures/dag1.png}
 %\label{diag:ps_env}
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/beta_mse.eps}
+    \includegraphics[width = \textwidth]{figures/beta_mse.png}
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
@@
 \centering
-\includegraphics[width = \textwidth]{figures/pop_inf_snps_box.eps}
+\includegraphics[width = \textwidth]{figures/pop_inf_snps_box.png}
 \caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.eps}
+    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
     \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/eta_beta_hat.eps}
+    \includegraphics[width = \textwidth]{figures/eta_beta_hat.png}
      \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}

### 2020-12-17 supplement.tex
Supplement
eps figures
Minor word tweaks throughout
final edits, figures formatted for pdf submission

@@
     \centering
-    \includegraphics[scale = 1]{figures/admixed_kinship.eps}
+    \includegraphics[scale = 1]{figures/admixed_kinship.png}
     \caption{1D Linear Admixture with coarse subpopulation structure (4 subpopulations, 50 subjects per subpopulation), RRM.}
@@
     \centering
-    \includegraphics[scale = 1]{figures/indep_coarse_kinship.eps}
+    \includegraphics[scale = 1]{figures/indep_coarse_kinship.png}
     \caption{Independent Subpopulations with coarse subpopulation structure (4 subpopulations, 50 subjects per subpopulation), RRM.}
@@
     \centering
-    \includegraphics[scale = 1]{figures/indep_fine_kinship.eps}
+    \includegraphics[scale = 1]{figures/indep_fine_kinship.png}
     \caption{Independent Subpopulations with fine subpopulation structure (100 subpopulations, 2 subjects per subpopulation), RRM.}
@@
     \centering
-    \includegraphics[scale = 1]{figures/empirical_kinship.eps}
-    \caption{Empirical RRM.}
+    \includegraphics[scale = 1]{figures/empirical_kinship.png}
+    \caption{Empirical data with three subpopulations ($n_1=417, n_2 = 134, n = 37$), RRM.}
     \label{fig:empirical}
@@
 
-\begin{sidewaysfigure}
+\begin{figure}[H]
     \centering
-    \includegraphics[width=\textwidth, height=\textheight, keepaspectratio]{figures/niter_perLambda_point.png}
+    \includegraphics[width=\textwidth]{figures/niter_perLambda_point.png}
     \caption{The number of iterations per $\lambda$ value for LMM-lasso-Rakitsch and LMM-lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
@@
     \label{fig:niter}
-\end{sidewaysfigure}
+\end{figure}

### 2020-12-17 lmm-genetic-review.tex
Section 2: putting DAGs in eq environment
Figure 2 boxplot and related edits.
First attempt at including eps version of file - required for genetic epi submission
eps figs not working... trying to avoid merge conflicts
Edits to 3.5, 4, and 5
Edits to 1, 3, 4, 5, bibliography, eps figure format
Supplement
Edits to abstract

@@
 \usepackage{tikz}
+\usepackage{tikz-cd}
 \usepackage{tkz-euclide}
@@
 \newcommand{\bepsT}[1][]{\widetilde{\boldsymbol{\varepsilon}{#1}}}
-\newcommand{\bX}[1][]{\textbf{X}#1}
-\newcommand{\bXT}[1][]{\widetilde{\textbf{X}{#1}}}
-\newcommand{\bZ}{\textbf{Z}}
-\newcommand{\bZT}{\widetilde{\textbf{Z}}}
-\newcommand{\by}{\textbf{y}}
-\newcommand{\byT}{\widetilde{\textbf{y}}}
-\newcommand{\bD}{\textbf{D}}
-\newcommand{\bG}{\textbf{G}}
-\newcommand{\bK}{\textbf{K}}
-\newcommand{\bI}{\textbf{I}}
-\newcommand{\bV}{\textbf{V}}
-\newcommand{\bu}{\textbf{u}}
-\newcommand{\bU}{\textbf{U}}
-\newcommand{\bR}{\textbf{R}}
-\newcommand{\bW}{\textbf{W}} 
+\newcommand{\bD}{\mathbf{D}}
+\newcommand{\bG}{\mathbf{G}}
+\newcommand{\bK}{\mathbf{K}}
+\newcommand{\bI}{\mathbf{I}}
+\renewcommand{\P}{\mathbf{P}}
+\newcommand{\bR}{\mathbf{R}}
+\newcommand{\bu}{\mathbf{u}}
+\newcommand{\bU}{\mathbf{U}}
+\newcommand{\bV}{\mathbf{V}}
+\newcommand{\bW}{\mathbf{W}}
+\newcommand{\bX}{\mathbf{X}}
+\newcommand{\by}{\mathbf{y}}
+\newcommand{\bZ}{\mathbf{Z}}
+\newcommand{\bXT}{\widetilde{\mathbf{X}}}
+\newcommand{\byT}{\widetilde{\mathbf{y}}}
+\newcommand{\bZT}{\widetilde{\mathbf{Z}}}
 \newcommand{\sigmaee}{\sigma^2_e}
@@
 \begin{abstract}
-Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to unobserved confounding factors, such as environmental heterogeneity.
+Many genetic studies that aim to identify genetic variants associated with complex phenotypes are subject to unobserved confounding factors arising from environmental heterogeneity.
 This poses a challenge to detecting associations of interest and is known to induce spurious associations when left unaccounted for.
@@
 Despite an extensive literature on penalized regression and LMMs separately, the two are rarely discussed together.
-The aim of this review is to do so while rigorously examining the statistical properties of penalized LMMs.
-Specifically, the ability of penalized LMMs to accurately estimate fixed, sparse genetic effects in the presence of environmental heterogeneity, for which population structure may serve as a proxy, has not been well studied.
-In order to clarify the importance yet subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
+The aim of this review is to do so while examining the statistical properties of penalized LMMs in the genetic association setting.
+Specifically, the ability of penalized LMMs to accurately estimate genetic effects in the presence of environmental confounding has not been well studied.
+In order to clarify the important yet subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
 In addition, we evaluate the performance of penalized LMMs and competing methods in terms of estimation and selection accuracy in the presence of a number of confounding structures.

### 2020-12-17 main.tex
Figure 2
Edits to Section 1

@@
 
-One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs, so that only those correlated with each other below some desired threshold are used in modeling and testing procedures.
-LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
-An alternative solution to the problem of LD is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
-However, classical multivariate approaches rely on collecting data with more observations than unknown parameters.
-This is typically infeasible in genome-wide studies where the number of features greatly exceeds the number of observations.
+One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.
+%LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
+An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
+However, classical multivariate approaches rely on data with more observations than unknown parameters.
+This is typically infeasible in genome-wide studies where the number of variants greatly exceeds the number of individuals.
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
@@
 It also provides considerable computational advantages that scale up well to very large data sets.
-These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying a list of variants most highly associated with a particular trait is often of interest.
-The lasso is widely used in this context and its application to ultra high-dimensional genetic data is the subject of recent computaional and methodological research \citep{qian2020fast, prive2018efficient}.
+These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest.
+Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.
 
@@
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, until recently the majority of this work has focused on plant and animal breeding data, much of which is experimentally controlled \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2020simultaneous, Sillanpaeae2011}.
-While many standard correction techniques may in principal be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and non-experimental conditions \citep{lawson2019population, barton2019population}.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, the majority of this work has focused on plant and animal breeding data, a setting which affords much greater experimental control than studies of human genetics \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2020simultaneous, Sillanpaeae2011}.
+While many standard correction techniques may in principle be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and confounding arising from observational study \citep{lawson2019population, barton2019population}.
 
@@
 We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
-In Section \ref{sec:results}, we illustrate the concepts described in Section \ref{sec:methods} via simulation studies, where the data-generating mechanism formally differentiates between population structure and environmental heterogeneity.
+In Section \ref{sec:results}, we illustrate these concepts via simulation studies, where the data-generating mechanism allows us to distinguish between population structure and environmental heterogeneity.
 In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
@@
 \centering
-\includegraphics[width = \textwidth]{figures/pop_inf_snps_box.png}
+\includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}
 \caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}

### 2020-12-17 main.tex
Edits to Section 2.1 and 2.2

@@
 The presence of distinct ancestry groups with different allele frequencies in a sample is known as \textbf{population stratification}.
-Recent relatedness describes the sharing of a common ancestor only several generations previously.
+Recent relatedness describes the sharing of a common ancestor only a few generations previous.
 Pedigree-based methods may be used to explicitly model recent relatedness if familial relationships are known.
@@
 
-In this review we focus on methods to account for unobserved relatedness such as population stratification and cryptic relatedness.
+In this review we focus on methods to account for unobserved dependence between samples, whether arising from population stratification or cryptic relatedness.
 Throughout this paper, the terms \textbf{population structure} and \textbf{structured population} will be used to describe a sample of individuals in which population stratification, cryptic relatedness, or both are present.
@@
 The mechanism of this phenomenon warrants some discussion as it is often overlooked that population structure in and of itself does not confound the genotype-phenotype relationship.
-While population structure may directly affect observed allele frequencies, there must also exist some non-genetic mechanism by which population stratification affects the phenotype, namely, the environment \citep{barton2019population, vilhjalmsson2012nature}.
+Rather, there must also exist some non-genetic mechanism by which population stratification affects the phenotype, namely, the environment \citep{barton2019population, vilhjalmsson2012nature}.
 
@@
 As an example, consider a genome wide association study (GWAS) to assess genetic variants associated with lung cancer in a sample comprised of subjects from two distinct subpopulations, A and B.
-Assume the minor allele of SNP $X$ is present with higher frequency in subpopulation A compared to subpopulation B, but has no direct effect on lung cancer.
+Assume the minor allele of SNP $X$ is more common in subpopulation B compared to subpopulation A, but has no direct effect on lung cancer.
 Also suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
 A GWAS of data from these subpopulations would likely find SNP $X$ to be significantly associated with lung cancer.
-This spurious association is not identified because of the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
-Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be found to be associated with the phenotype.
+This spurious association is does not arise from the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
+Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be associated with the phenotype.
 
@@
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
-Otherwise stated, we do not assume that environmental factors directly influence population-specific allele frequencies, or vice versa, only that the two may be related.
-In the context of our air quality example, this means that we assume air quality is not directly altering individuals' allele frequencies, nor are allele frequencies causing individuals to move to regions of better or worse air quality.
+%Otherwise stated, we do not assume that environmental factors directly influence population-specific allele frequencies, or vice versa, only that the two may be related.
+In the context of our air quality example, this means that we assume air quality is not directly altering individuals' allele frequencies, nor are allele frequencies causing individuals to move to regions of better or worse air quality, only that they are now correlated.
 Genetic relatedness and environmental effects may become associated over time for any number of complex reasons beyond the scope of this paper.
@@
 LD pruning is commonly employed to reduce confounding due to genetic background in univariate association testing.
-However, we will focus on methods which address this via genomewide multivariate analysis.
+Here, we focus on methods which address this via genomewide multivariate analysis.
 

### 2020-12-17 submissions/genepi-cover/cover.tex
Adding cover letter (in progress)

\documentclass{uiletter}

\begin{document}

\StartLetter{}{Dear editors of {\em Genetic Epidemiology},}

I am submitting the manuscript ``XXX'' for your consideration as a Research Paper for publication in {\em Genetic Epidemiology}.

Penalized regression methods are an attractive approach for simultaneously selecting features and fitting a model based on those features, but have long been hampered by the difficulty of applying inferential tools.  In particular, it is of tremendous practical interest to be able to calculate a false discovery rate for the features that have been selected.  There have been several recent papers in this area proposing estimators of various kinds, but the focus has been almost entirely on least-squares regression; very little work has been done on estimating FDRs for penalized logistic regression or Cox regression models.  This article derives a general method for calculating the marginal false discovery rate for any likelihood-based regression model that satisfies the general regularity conditions for classical likelihood-based inference.  We have implemented this method in the R package {\tt ncvreg} and carefully investigate this approach for logistic and Cox regression via simulation and apply it to two high-dimensional studies, one involving binary outcomes and the other involving a time-to-event outcome.  I believe this manuscript would be of interest to the readers of {\em Biometrical Journal}, with the potential to become widely used in the field.

This work has not been previously published, is not under review elsewhere, and all authors concur with this submission.  Thank you very much for your consideration.

\EndLetter

\end{document}

### 2020-12-17 main.tex
Data availability statement
Edits to 2.3

@@
 
-An additional source of confounding in genetic studies that we will briefly mention is that of confounding due to genetic background.
+A different type of confounding in genetic studies that we will briefly mention is due to genetic background.
 This occurs in univariate association methods when a non-causal SNP is tested for association with a trait of interest, and that SNP is correlated with a causal SNP.
@@
 Given the importance of both genetic and environmental contributions to many complex phenotypes, both genetic background and environmental confounding are important factors in our ability to identify true causal genetic relationships between SNPs and phenotype.
-LD pruning is commonly employed to reduce confounding due to genetic background in univariate association testing.
+As mentioned earlier, LD pruning is a common approach to reduce confounding due to genetic background in univariate association testing.
 Here, we focus on methods which address this via genomewide multivariate analysis.
@@
 
-One of the first methods to address the problem of confounding due to population structure is the genomic control (GC) method.
-Because the effect of this confounding is to inflate the distribution of association test statistics, GC introduces a factor to counteract this effect and restore the appropriate null distribution.
+One of the earliest approaches to address the problem of confounding due to population structure was the method of genomic control (GC).
+Because confounding inflates the distribution of association test statistics, GC introduces a factor to counteract this effect and restore the appropriate null distribution.
 This deflation factor is estimated using loci independent of the phenotype of interest and the loci being tested for association \citep{devlin1999genomic, bacanu2000power, wang2009testing}.
@@
 
-Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
-An important concept to note, and one that we will return to in Section \ref{sec:partition}, is that population structure only leads to confounding if it is correlated with distinct environmental effects.
+%Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
+As mentioned earlier, population structure only leads to confounding if it is correlated with a causal effect in the environment.
 In the presence of environmental heterogeneity which affects the phenotype of interest in a manner that is orthogonal to population structure, SNP-phenotype associations will not be confounded.
@@
 In such a scenario, failing to account for unobserved environmental effects is likely to result in greater variability, but SNP effect estimates will remain unbiased \citep{greenland1999causal}.
-We will return to this idea and its mathematical justification in Section \ref{sec:methods}.
+We will discuss this fact in greater detail in Section \ref{sec:partition}.
 
@@
 Such methods include LMMs, principal-component-based approaches, and structured association.
-Structured association refers to a general class of methods which attempt to infer subject membership within a discrete number of nonoverlapping subpopulations.
-Subsequent analyses are conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.
+Structured, or stratified, association refers to a general class of methods which attempt to infer subject membership within a discrete number of nonoverlapping subpopulations.
+The analysis is conducted within each subpopulation, and the subpopulation-specific results may then be combined via meta-analysis \citep{pritchard1999use, pritchard2000association}.
 By partitioning the overall sample size in this manner, the structured association method is vulnerable to a loss of power.
@@
 This assumption may not accurately reflect the characteristics of a particular sample, such as one consisting of admixed populations, or recent relatedness.
-Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, there may also be individuals who fall between these two extremes and are characterized by continuously varying relatedness and a gradient of air quality.
-By clustering those individuals within in one subpopulation or another, we lose information about relatedness by oversimplifying the population structure.
+Considered in the context of our air quality example, rather than the existence of two subpopulations with exposure to either good or bad air quality, we may have a continuous gradient of relatedness and air quality.
+By clustering individuals into one subpopulation or another, we oversimplify the population structure.
 
-A different approach is to create an approximation of population structure using surrogate variables, and to adjust for these as additional model covariates.
-Principal component-adjustment (PC-adjustment) methods use the first several left singular vectors of a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
+A different approach is to use create surrogate variables to represent population structure, and to adjust for these as additional model covariates.
+Principal component-adjustment (PC-adjustment) methods use the first several eigenvectors a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
 This approach may be used in univariate or multivariate frameworks, and allows for analysis on the entirety of a structured sample.
 
-Like PC-adjustment methods, LMMs utilize the left singular vectors of a similarity matrix to estimate and subsequently correct for relatedness among subjects.
-In an LMM, however, these left singular vectors are treated as random, rather than fixed effects.
 The term LMM has been used to describe a number of distinct but related procedures for analyzing structured genetic data which differ in their objectives and methods.
-There are two main objectives of LMMs used in a structured genetic context: estimating narrow-sense heritability, and estimating individual SNP effects.
+There are two main objectives of LMMs in the genetics context: estimating narrow-sense heritability, and estimating individual SNP effects.
 
-Among LMM methods that aim to estimate heritability, the most well-known example is genome-wide complex trait analysis (GCTA) \citep{yang2011gcta}.
+Among LMM methods that aim to estimate heritability, the most well-known is genome-wide complex trait analysis (GCTA) \citep{yang2011gcta}.
 In this framework, all observed SNPs are treated as random effects.
@@
 LMMs can also be used to estimate SNP effects, and this can be done in either a univariate or multivariate manner.
-Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure by modeling it as a random effect using a genetic similarity matrix \citep{yu2006unified, kang2010variance, kang2008efficient}.
-However, as with all univariate testing approaches, univariate LMM implementations are prone to spurious associations and biased effect estimates in the presence of LD.
+Like PC-adjustment methods, LMMs utilize a similarity matrix to estimate and subsequently correct for relatedness among subjects.
+In an LMM, however, the surrogate variables are treated as random, rather than fixed effects
+%Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure by modeling it as a random effect using a genetic similarity matrix
+\citep{yu2006unified, kang2010variance, kang2008efficient}.
+However, as with all univariate testing approaches, univariate LMM implementations are subject to spurious associations and biased effect estimates in the presence of LD.
 
@@
 Several studies have shown that LMMs outperform PC-adjustment methods in the univariate framework \citep{wang2013analytical, kang2010variance, zhao2007arabidopsis}.
-However, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetic relatedness and the environmental mechanism of confounding due to population structure.\\
+However, to the best of our knowledge this is the only comprehensive review and comparison of these methods in a penalized multivariate framework and using a simulation scheme motivated by human genetics and environmental confounding.\\
 

### 2020-12-17 main.tex
Citing ggmix paper, not package

--- 2020-12-17 main.tex
@@
 LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}.
-This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
+This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{bhatnagar2020simultaneous}.
 We further discuss differences between these methods in Section \ref{sec:results}.
@@
 All of these approaches are potentially attractive avenues for future research into GWAS methodology.
+
+\subsection*{Data Availability}
+
+Data sharing is not applicable to this article as no new data were collected or analyzed in this study.

### 2020-12-17 main.tex
Edits to Section 3

--- 2020-12-17 main.tex
@@
 The lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
-It relies on an assumption of sparsity, which allows coefficient estimation and variable selection to be accomplished simultaneously.
-It also provides considerable computational advantages that scale up well to very large data sets.
+It relies on an assumption of sparsity; in other words, that many features will have no effect.
+Thus, although the total number of features may be large, the number that actually affect the outcome will be small.
+This allows coefficient estimation and variable selection to be accomplished simultaneously, and provides considerable computational advantages that scale up well to very large data sets.
 These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest.
@@
 To describe and compare PC-lasso and LMM-lasso, we consider an $n \times p$ matrix of genotype data $\bG$ for $n$ subjects and $p$ SNPs, where each $G_{ij} \in \{ 0, 1, 2 \}$ enumerates the number of allele copies subject $i$ has for SNP $j$.
-In order to equitably compare SNP effects, regardless of allele frequency, in practice $\bG$ is standardized.
+In order to equitably compare SNP effects across differing allele frequencies, in practice $\bG$ is standardized; not doing so introduces a bias in which common SNPs are more likely to be selected.
 We denote the standardized version of this genotype matrix as $\bX$, where each $X_{ij} = (G_{ij} - 2 \pi_j) / \sqrt{2\pi_j (1 - \pi_j)}$ and $\pi_j$ is the minor allele frequency (MAF) of SNP $j$ \citep{zhang2015principal, price2006principal}.
 Additionally, $\bX_0$ represents an $n \times p_0$ matrix of unpenalized covariates and the intercept.
-Thus, at a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
+At a minimum, $\bX_0$ will be an $n \times 1$ vector of 1s to represent the intercept, in which case $\bbeta_0 \equiv \beta_0$ is the scalar intercept.
 However, $\bX_0$ may additionally include observed covariates such as age, sex, or smoking history, which are included in the model without penalty.
@@
 Often it is assumed that subjects belong to $q \le p$ subpopulations, and $\bZ$ is a corresponding incidence matrix of 0s and 1s.
-However, as we will discuss, this is not the only possibility.
+However, this is not the only possibility.
 
@@
 \end{equation}
-where $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ on the top-left diagonal, followed by $n - k$ zeroes on the bottom-right diagonal.
-$\bU$ is the corresponding $n \times n$ orthonormal matrix of left eigenvectors, or principal components, of $\bK$.
+where $\bR = \bU\bD$ and $\bD^2$ is an $n \times n$ diagonal matrix containing the $k$ nonzero eigenvalues of $\bK$ on the top-left diagonal, followed by $n - k$ zeroes.
+$\bU$ is the corresponding $n \times n$ orthonormal matrix of eigenvectors of $\bK$.
 The first $k$ columns of $\bU$ correspond to the nonzero eigenvalues of $\bK$, and the columns of $\bR$ are thus the principal components of $\bK$ weighted by their corresponding singular values.
@@
 \begin{equation}
+  \begin{aligned}
     \label{eqn:lmm_reg}
-    \by = \bX_0\bbeta_0 + \bX\bbeta + \bX\bgamma+ \beps,
+    \by &= \bX_0\bbeta_0 + \bX\bbeta + \bX\bgamma+ \beps, \\
+    \beps &\sim N(\boldsymbol{0}, \sigmaee \bI) \\
+ \bgamma &\sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bX \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK),
+  \end{aligned}
 \end{equation}
-$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI) $$
-$$ \bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bX \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK), $$
 where $\sigmagg$ is the additive genetic variance, and we assume $\beps \independent \bgamma$.
@@
 \end{equation}
-where $\bXT{^\prime} = \bU^\T \bX{^\prime}$, so that all of the data are transformed, but only the elements of $\bbeta{^\prime}$ corresponing to SNPs are penalized.
+where $\bXT{^\prime} = \bU^\T \bX{^\prime}$, so that all of the data are transformed, but only the elements of $\bbeta{^\prime}$ corresponding to SNPs are penalized.
 
@@
 
-Different variance component estimation methods lead to different forms of $\widehat{\bW}$ and two distinct LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
+Different variance component estimation methods lead to different estimates of $\bW$ and two distinct LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
 LMM-lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
@@
 Using the result of the decomposition from \eqref{eqn:k}, $\bK = p^{-1} \bR \bR^\T$ , the LMM model \eqref{eqn:lmm_reg} can be written equivalently as
-
 \begin{equation}
+  \begin{aligned}
     \label{eqn:pc_lmm_equiv}
-    \by = \bX_0 \bbeta_0 + \bX \bbeta + \bR \bgamma + \beps
+    \by &= \bX_0 \bbeta_0 + \bX \bbeta + \bR \bgamma + \beps \\
+    \beps &\sim N(\boldsymbol{0}, \sigmaee \bI) \\
+    \bgamma &\sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bR \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK).
+  \end{aligned}
 \end{equation}
-
-$$ \beps \sim N(\boldsymbol{0}, \sigmaee \bI) $$
-$$ \bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bR \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK).$$
 Based on the relationship between equations \eqref{eqn:pca_reg} and \eqref{eqn:pc_lmm_equiv}, we can see that modeling the principal components of $\bK$ as fixed or random effects relies on the same underlying regression model.
@@
 
-Given the nature of environmental confounding in our setting of interest, and in order to assess methodological performance in terms of estimation accuracy and bias, we simulate from the data-generating model
+Given the nature of environmental confounding in our setting of interest, we simulate from
 \begin{equation}
@@
 \end{equation}
-where $\by$ represents the synthetic phenotype, $\bgamma$ is a fixed quantity (see Section~\ref{sec:partition}), and $a, b, c$ are scaling factors determined such that the variance of $\by$ is equal to 1 and is partitioned into genetic signal, environmental confounding, and random noise components,
+where $\by$ represents the synthetic phenotype, $\bgamma$ is a fixed quantity (see Section~\ref{sec:partition}), and $a, b, c$ are scaling factors.
+These factors are determined such that the variance of $\by$ is equal to 1 and can be partitioned into genetic signal, environmental confounding, and random noise components, respectively, as follows:
 \begin{equation}
@@
 \end{equation}
-$\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$.
+In \eqref{eqn:parition_y}, $\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability.
+%, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$.
 The parameter $\eta$ also describes the data-generating model's signal to noise ratio (SNR), while $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
-So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, and 40\% due to unobserved environmental effects.
+So, for example $\eta = 0.5$ and $\xi = 0.8$ corresponds to a 1:1 SNR, and means that 50\% of the variability in $\by$ is due to additive genetic effects, 40\% due to unobserved environmental effects, and the remaining 10\% due to other factors.
 
@@
 
-We generated these synthetic phenotypes using three types of high-dimensional ($p > n$) $\bG$ matrices, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) \anna{a subset of} real genome-wide SNP data from known caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
-We refer to these as the 1D Linear Admixture, Independent Subpopulations, and Empirical data, respectively.
-For all data types $p = 1000$ with 50 causal SNPs.
-For the simulated 1D Linear Admixture and Independent Subpopulations data \anna{$n = 200$ unless otherwise specified}, and $n = 588$ for the Empirical data.
+We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
+We refer to these as 1D Linear Admixture, Independent Subpopulations, and Empirical, respectively.
+For all data types, there are $p = 1000$ total SNPs, 50 of which have causal effects.
+For the simulated 1D Linear Admixture and Independent Subpopulations data $n = 200$ unless otherwise specified, and $n = 588$ for the Empirical data.
+\pb{NEED A BNPSD REFERENCE HERE??}
 
@@
 We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example.
-Conversely, we define a fine subpopulation structure to be one characterized by many small subpopulations, representing a continuum of environmental heterogeneity.
+Conversely, we define a fine structure to be one characterized by many small subpopulations, representing a continuum of environmental heterogeneity.
 Whether fine or coarse, the simulated data were generated such that subpopulations (also known as founding populations) are of equal size.
@@
 
-Although cross-validation and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons across methods.
-Cross-validation may be inadvisable for LMMs, and in the presence of samples with underlying structure, as is the case here \citep{roberts2017cross}.
-While information criterion methods such as BIC present another option for selecting $\lambda$, they may be unstable in high dimensions.
-In addition, because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach would not lend itself to comparable results across methods.
+Although cross-validation and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons.
+Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods.
+Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples is an open question \citep{roberts2017cross}.
 
-1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
+A total of 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
 

### 2020-12-17 main.tex
3.3 how to describe sigma_g^2 / sigma_s^2?

--- 2020-12-17 main.tex
@@
 $$ \bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bX \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK), $$
-where $\sigmagg$ is the additive genetic variance, and we assume $\beps \independent \bgamma$.
+where $\sigmagg$ is the \anna{additive genetic variance}, and we assume $\beps \independent \bgamma$.
 Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields an equivalent model.

### 2020-12-17 main.tex

--- 2020-12-17 main.tex
@@
 LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}.
-This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix} \citep{ggmix}.
+This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix}.
 We further discuss differences between these methods in Section \ref{sec:results}.

### 2020-12-17 main.tex
Changing sigmagg to sigmass

@@
     \beps &\sim N(\boldsymbol{0}, \sigmaee \bI) \\
- \bgamma &\sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bX \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK),
+ \bgamma &\sim N(\boldsymbol{0}, p^{-1} \sigmass \bI) \Rightarrow \bX \bgamma \sim N(\boldsymbol{0}, \sigmass \bK),
   \end{aligned}
 \end{equation}
-where $\sigmagg$ is the additive genetic variance, and we assume $\beps \independent \bgamma$.
+where $\sigmass$ determines the size of the structured component to the variance, and we assume $\beps \independent \bgamma$.
 Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields an equivalent model.
@@
 We will first derive the LMM-lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
-Following equation \eqref{eqn:lmm_reg}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$.
+Following equation \eqref{eqn:lmm_reg}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmass \textbf{K} + \sigmaee \textbf{I}$.
 LMM-lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
@@
 \begin{align*}
-    \bV^{-1} &= (\sigmagg \bK + \sigmaee \bI)^{-1}\\
-    &=(\sigmagg \bU \bD^2 \bU^\T + \sigmaee \bU \bU^\T)^{-1}\\
-    &= (\bU (\sigmagg \bD^2 + \sigmaee \bI) \bU^\T)^{-1}\\
-    &=\bU (\sigmagg \bD^2 + \sigmaee \bI)^{-1} \bU^\T\\
+    \bV^{-1} &= (\sigmass \bK + \sigmaee \bI)^{-1}\\
+    &=(\sigmass \bU \bD^2 \bU^\T + \sigmaee \bU \bU^\T)^{-1}\\
+    &= (\bU (\sigmass \bD^2 + \sigmaee \bI) \bU^\T)^{-1}\\
+    &=\bU (\sigmass \bD^2 + \sigmaee \bI)^{-1} \bU^\T\\
     &=\bU \bW^2 \bU^\T.
@@
 By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^\T \bX$, $\byT = \bU^\T \by$, $\bW = \text{diag}\{w_1, ...
-w_n\}$, and $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$.
+w_n\}$, and $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmass \bD^2 + \sigmaee \bI)^{-1}$.
 Together, this yields the objective function
@@
 Different variance component estimation methods lead to different estimates of $\bW$ and two distinct LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
-LMM-lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$ \citep{Rakitsch2012}.
+LMM-lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmass$ \citep{Rakitsch2012}.
 These variance components and the corresponding weights are estimated once under the assumption of a null model.
@@
     \beps &\sim N(\boldsymbol{0}, \sigmaee \bI) \\
-    \bgamma &\sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI) \Rightarrow \bR \bgamma \sim N(\boldsymbol{0}, \sigmagg \bK).
+    \bgamma &\sim N(\boldsymbol{0}, p^{-1} \sigmass \bI) \Rightarrow \bR \bgamma \sim N(\boldsymbol{0}, \sigmass \bK).
   \end{aligned}
@@
 \end{align}
-Treating the environmental effects as random with $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmagg \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmagg\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
+Treating the environmental effects as random with $\bgamma \sim N(\boldsymbol{0}, p^{-1} \sigmass \bI)$, equations \eqref{eqn:expectation} and \eqref{eqn:variance} yield $\mathbb{E}(\bbetaHat) = \bbeta$, and $\mathbb{V}(\bbetaHat) = p^{-1}\sigmass\bI + \sigmaee (\bX^\T \bX)^{-1}$, respectively.
 Treating them as fixed, on the other hand, we have $\mathbb{E}(\bbetaHat) = \bbeta + \bgamma$ and $\mathbb{V}(\bbetaHat) = \sigmaee (\bX^\T \bX)^{-1}$.
@@
 In \eqref{eqn:parition_y}, $\eta$ defines the proportion of the variance of $\by$ attributable to additive genetic factors, commonly referred to as the narrow-sense heritability.
-%, $h^2 = \sigmagg/(\sigmagg + \sigmaee)$.
+%, $h^2 = \sigmass/(\sigmass + \sigmaee)$.
 The parameter $\eta$ also describes the data-generating model's signal to noise ratio (SNR), while $\xi$ partitions the remaining variance of $\by$ into latent environmental heterogeneity and random noise.
@@
 
-\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmagg$ and $\delta := \sigmaee / \sigmagg$ as previously discussed, however, we have implemented the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
+\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$ as previously discussed, however, we have implemented the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
 

### 2020-12-18 main.tex
Edits to 4 & 5; cover letter; changed title

@@
 We begin by comparing the ordinary lasso, PC-lasso, and LMM-lasso as the level of environmental confounding is varied.
-Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error of the coefficients (MSE).
-Suppose $\bbetaHat(\lambda)$ is the vector of coefficients estimated by the lasso with a penalty of $\lambda$, and $\bbeta$ the coefficient values used to simulate the data.
-Mean squared error is then defined as
-\begin{equation}
-    \text{MSE} = \mathbb{E}||\bbeta - \bbetaHat(\lambda)||^2_2.
-\end{equation}
+Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE) of the coefficients, MSE = $\mathbb{E}||\bbeta - \bbetaHat(\lambda)||^2_2$, where $\bbetaHat(\lambda)$ is the vector of coefficients estimated by the lasso with a penalty of $\lambda$, and $\bbeta$ the true coefficient values.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
@@
     \includegraphics[width = \textwidth]{figures/beta_mse.png}
-    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-discordant environmental effects.}
+    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.}
     \label{fig:mse}
@@
 This improvement is more substantial in the Empirical and Independent Subpopulations data than in the 1D Linear Admixture data.
-This is due to the fact that the Independent Subpopulations and Empirical data sets are characterized by nonoverlapping genetic subpopulations, meaning that a greater proportion of the environmental effect is in the column space of $\P_{\bX}$.
-We can quantify the proportion of the environmental component in the column space of $\P_{\bX}$ for each data type using the CP metric defined in \eqref{eqn:ratio} of Section \ref{sec:partition}.
+This is because there is less overlap between subpopulations than in the admixture data, meaning that a greater proportion of the environmental effect is in the column space of $\P_{\bX}$.
+We can quantify this proportion using the CP metric defined in \eqref{eqn:ratio}.
 Here, we have CP = 0.47, 0.87, and 0.99 for the 1D Linear Admixture, Empirical, and Independent Subpopulations, respectively.
@@
 Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
-The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, while the addition of the 10 unpenalized PCs leads to a slight decrease in TPR in this setting for the 1D Linear Admixture and Independent Subpopulations data sets.
+The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, with PC-lasso slightly less accurate at identifying causal SNPs.
 In the presence of environmental confounding, both LMM-lasso and PC-lasso achieve substantial improvements in TPR over the ordinary lasso.
@@
 
-In order to assess the behavior of these methods in this context, we conducted a simulation based on the Empirical genotype type, but restricted to just two subpopulations, Caucasian and African-American.
-1000 SNPs were chosen from the genome-wide Empirical data such that the selected SNPs' absolute difference in MAF between the two subpopulations ranged uniformly from 0 to 0.5.
-As in the earlier simulations, a 1:1 SNR ($\eta = 0.5$) was used.
+In order to assess the behavior of these methods in this context, we conducted a simulation using Empirical genotypes, but restricted to just two subpopulations, Caucasian and African-American.
+A total of 1000 SNPs were chosen such that the selected SNPs' absolute difference in MAF between the two subpopulations ranged uniformly from 0 to 0.5.
+As in the earlier simulations, a 1:1 SNR ($\eta = 0.5$) was used to simulate the phenotype.
 
-\begin{figure}[H]
+\begin{figure}[hbt]
 \centering
 \includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}
-\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, dichotomous-discordant effects, $\eta = 0.5$.}
+\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$.}
 \label{fig:pop_inf}
@@
 Even in the absence of environmental effects, the ordinary lasso shows a preference for selecting population-informative SNPs, while PC-lasso and LMM-lasso are reluctant to select such SNPs (recall that with this design, the median MAF difference between populations is 0.25).
-
 In the presence of confounding, however, the lasso shows an overwhelming tendency to select population-informative SNPs.
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations data, $\eta = 0.5$, and dichotomous-discordant environmental effects.
-}
+    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$. }
     \label{fig:big_vs_small}
@@
 LMM-lasso-ggmix does tend to provide improved estimates of $\eta$ compared to LMM-lasso-Rakitsch, particularly in the presence of lower $\eta$ values and where $n \ll p$ (Figure \ref{fig:eta_beta_mse}, middle panel), although these improvements did not help it to estimate $\bbeta$ more accurately.
-In addition, at high signal-to-noise ratios ($\eta=0.8$), the performance of LMM-lasso-ggmix was \anna{highly variable}.
+In addition, at high signal-to-noise ratios ($\eta=0.8$), the performance of LMM-lasso-ggmix was highly variable.
 
@@
 The bottom panel of Figure \ref{fig:eta_beta_mse} shows the computing time required of each method in the various simulation settings, and shows that when $n \ll p$, LMM-lasso-ggmix is up to 60 times slower than LMM-lasso-Rakitsch.
-In settings where $n \ge p$, LMM-lasso-ggmix is ``faster'' than LMM-lasso-Rakitsch, but this seems to be due to the algorithm getting stuck in a local minimum and prematurely exiting, a phenomenon that also explains the extremely poor estimation accuracy of LMM-lasso-ggmix in the $\eta=0.8$, $n=1400$ setting.
+In settings where $n \ge p$, LMM-lasso-ggmix is ``faster'' than LMM-lasso-Rakitsch, but this seems to be due to the algorithm getting stuck in a local minimum and prematurely terminating, a phenomenon that also explains the extremely poor estimation accuracy of LMM-lasso-ggmix in the $\eta=0.8$, $n=1400$ setting.
 
@@
     \includegraphics[width = \textwidth]{figures/eta_beta_hat.png}
-     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-lasso-Rakitsch and LMM-lasso-ggmix for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-discordant environmental effect structure.}
+     \caption{Estimation accurate of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$.}
     \label{fig:eta_beta_mse}
@@
 In addition, we have provided a thorough review of two penalized regression approaches that correct for population structure in genetic data: PC-lasso and LMM-lasso.
-
 We have detailed the assumptions made by PC-lasso and LMM-lasso, which model confounding influences as fixed and random effects, respectively.
@@
 We considered two implementations of the LMM-lasso method: LMM-lasso-Rakitsch and LMM-lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
-Although both methods perform similarly in terms of estimating $\bbeta$, the much greater computational efficiency and numerical stability of LMM-lasso-Rakitsch make it a more attractive choice in most scenarios, although LMM-lasso-ggmix may be useful if the objective of the analysis is to accurately estimate narrow-sense heritability.
+Although both methods perform similarly in terms of estimating $\bbeta$, the much greater computational efficiency and numerical stability of LMM-lasso-Rakitsch make it a more attractive choice in most scenarios, although LMM-lasso-ggmix may be useful if the objective of the analysis is to accurately estimate variance components.
 

### 2020-12-18 main.tex
Fixing typos

@@
 where $\sigmass$ determines the size of the structured component to the variance, and we assume $\beps \independent \bgamma$.
-Rather than assuming anything about $\bgamma$ directly, a common representation paramaterizes the model in terms of $\bu=\bX\bgamma$, which yields an equivalent model.
+Rather than assuming anything about $\bgamma$ directly, a common representation parameterizes the model in terms of $\bu=\bX\bgamma$, which yields an equivalent model.
 
-We will first derive the LMM-lasso objective function in the absence of unpenalized covariates for the sake of notational simplicity.
+We will first derive the LMM-lasso objective function in the absence of unpenalized covariates for the sake of simplicity.
 Following equation \eqref{eqn:lmm_reg}, in the absence of unpenalized covariates the LMM assumes $\by \sim N(\bX \bbeta, \bV)$, where $\bV = \sigmass \textbf{K} + \sigmaee \textbf{I}$.
-LMM-lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generlized least squares (GLS).
+LMM-lasso uses the covariance structure of the errors and properties of the multivariate normal distribution to transform the original data and diagonalize the errors in a manner akin to generalized least squares (GLS).
 Specifically, left-multiplying the original data by $\bV^{-1/2}$ yields
@@
 
-To illustrate this concept mathematically, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generatlity, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for the sake of simplicity:
+To illustrate this concept mathematically, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generality, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for the sake of simplicity:
 \begin{align}
@@
 For the fully simulated data, two subpopulation structures were considered, which we refer to as ``fine'' and ``coarse''.
-We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quaility example.
+We define a coarse structure to be one characterized by a few large subpopulations intended to model large-scale environmental heterogeneity and population stratification, such as that described in our motivating air quality example.
 Conversely, we define a fine structure to be one characterized by many small subpopulations, representing a continuum of environmental heterogeneity.
@@
 For example, both the Independent Subpopulation and 1D Linear Admixture simulated data may be comprised of fine or coarse structure.
-However, the subpopulations will be more distinct in the unadmixed Independent Subpopulation data than in the 1D Linear Admixed data.
+However, the subpopulations will be more distinct in the Independent Subpopulation data than in the 1D Linear Admixed data.
 The Empirical data is comprised of three racial subpopulations; one relatively large, one small, and one in the middle.
@@
 
-LMM-lasso-Rakitsh and LMM-lasso-ggmix achieve nearly identical MSE for $\bbeta$ in almost all scenarios, particularly when $n \ll p$ (Figure \ref{fig:eta_beta_mse}, top panel).
+LMM-lasso-Rakitsch and LMM-lasso-ggmix achieve nearly identical MSE for $\bbeta$ in almost all scenarios, particularly when $n \ll p$ (Figure \ref{fig:eta_beta_mse}, top panel).
 However, when $n \ge p$ the estimation performance of LMM-ggmix becomes erratic, occasionally far worse than that of LMM-lasso-Rakitsch.

### 2020-12-18 submissions/genetic_epi.tex
Formatting for genetic epi submission; using endfloat
eps figs not working... trying to avoid merge conflicts
final edits, figures formatted for pdf submission

--- 2020-10-08 lmm-genetic-review.tex
@@
 \newcommand{\bepsT}[1][]{\widetilde{\boldsymbol{\varepsilon}{#1}}}
-\newcommand{\bX}[1][]{\textbf{X}#1}
-\newcommand{\bXT}[1][]{\widetilde{\textbf{X}{#1}}}
-\newcommand{\bZ}{\textbf{Z}}
-\newcommand{\bZT}{\widetilde{\textbf{Z}}}
-\newcommand{\by}{\textbf{y}}
-\newcommand{\byT}{\widetilde{\textbf{y}}}
-\newcommand{\bD}{\textbf{D}}
-\newcommand{\bG}{\textbf{G}}
-\newcommand{\bK}{\textbf{K}}
-\newcommand{\bI}{\textbf{I}}
-\newcommand{\bV}{\textbf{V}}
-\newcommand{\bu}{\textbf{u}}
-\newcommand{\bU}{\textbf{U}}
-\newcommand{\bR}{\textbf{R}}
-\newcommand{\bW}{\textbf{W}} 
+\newcommand{\bD}{\mathbf{D}}
+\newcommand{\bG}{\mathbf{G}}
+\newcommand{\bK}{\mathbf{K}}
+\newcommand{\bI}{\mathbf{I}}
+\renewcommand{\P}{\mathbf{P}}
+\newcommand{\bR}{\mathbf{R}}
+\newcommand{\bu}{\mathbf{u}}
+\newcommand{\bU}{\mathbf{U}}
+\newcommand{\bV}{\mathbf{V}}
+\newcommand{\bW}{\mathbf{W}}
+\newcommand{\bX}{\mathbf{X}}
+\newcommand{\by}{\mathbf{y}}
+\newcommand{\bZ}{\mathbf{Z}}
+\newcommand{\bXT}{\widetilde{\mathbf{X}}}
+\newcommand{\byT}{\widetilde{\mathbf{y}}}
+\newcommand{\bZT}{\widetilde{\mathbf{Z}}}
 \newcommand{\sigmaee}{\sigma^2_e}
 \newcommand{\sigmagg}{\sigma^2_g}
+\newcommand{\sigmass}{\sigma^2_s}
 \newcommand{\T}{\text{T}}

### 2020-12-18 main.tex
Adding BNPSD reference

@@
 
-We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations, (2) simulated SNP data from independent subpopulations, and (3) real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
+We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
 We refer to these as 1D Linear Admixture, Independent Subpopulations, and Empirical, respectively.
@@
 For the simulated 1D Linear Admixture and Independent Subpopulations data $n = 200$ unless otherwise specified, and $n = 588$ for the Empirical data.
-\pb{NEED A BNPSD REFERENCE HERE??}
 

### 2020-12-18 main.tex
Fixing figure caption typos

@@
     \includegraphics[width = \textwidth]{figures/eta_beta_hat.png}
-     \caption{Estimation accurate of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$.}
+     \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$.}
     \label{fig:eta_beta_mse}

### 2020-12-18 supplement.tex
Adding BNPSD reference
Using endfloat for genetic epi submission, un-numbering supplement
Fixing figure caption typos

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Supplementary materials}
+\section*{Supplementary materials}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \includegraphics[scale = 1]{figures/empirical_kinship.png}
-    \caption{Empirical data with three subpopulations ($n_1=417, n_2 = 134, n = 37$), RRM.}
+    \caption{Empirical data with three subpopulations, Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$), RRM.}
     \label{fig:empirical}
@@
     \includegraphics[width=\textwidth]{figures/niter_perLambda_point.png}
-    \caption{The number of iterations per $\lambda$ value for LMM-lasso-Rakitsch and LMM-lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
+    \caption{Number of iterations per $\lambda$ value for LMM-lasso-Rakitsch and LMM-lasso-ggmix across varying sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\eta = 0.8$, and $\xi = 0.8$, and dichotomous-discordant environmental effect structure.
 Note that 100 is the maximum number of iterations for both implementations of LMM-lasso, so that numerous incidences of 100 iterations may indicate lack of model convergence.}

### 2021-02-14 main.tex
Fixing Eq 15 typo
Putting in placeholders for where to address reviewer comments

@@
 
-The objective of many genetic studies is to accurately identify variants associated with complex traits or diseases, and to estimate the effects of those variants.
+\anna{More discussion of application in general, and especially up front.} The objective of many genetic studies is to accurately identify variants associated with complex traits or diseases, and to estimate the effects of those variants.
 Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.
@@
 Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.
+\anna{More background on lasso}
 
@@
 
+\anna{More discussion of nonlinear models. In particular, is logistic regression practical? - threshold methods, probit regression by lippert}
+
 The remainder of this review will focus on the performance of such multivariate lasso-penalized methods for the analysis of continuous, normally distributed outcomes.
@@
 
+\anna{"should be clearer descriptions on when the reviewed approaches can/cannot be applied"; be more explicit here?}
+
+%------------------------------%
+\subsection{Software}
+\label{sec:software}
+\anna{More discussion of software. Add a small section? R-based Software for lasso, pc-lasso, ggmix, plmm, PLINK functions, KING, snpRelate}
+
 %------------------------------%
@@
     &\quad \notag\\
-    \mathbb{V}(\bbetaHat) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\by) \notag\\
-    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps) \notag\\
+    \mathbb{V}(\bbetaHat) &= (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\by) \bX  (\bX^\T \bX)^{-1} \notag\\
+    &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bbeta + \bX\bgamma + \beps) \bX  (\bX^\T \bX)^{-1} \notag\\
     &=  (\bX^\T \bX)^{-1} \bX^\T \mathbb{V}(\bX\bgamma + \beps) \bX  (\bX^\T \bX)^{-1} \label{eqn:variance}.
@@
 This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.
+
+\anna{penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.}
 

### 2021-02-14 main.tex
Making DAGs black and white
Sec 5. Responding to: penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.

@@
 The lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
-It relies on an assumption of sparsity; in other words, that many features will have no effect.
+\anna{More background on lasso} It relies on an assumption of sparsity; in other words, that many features will have no effect.
 Thus, although the total number of features may be large, the number that actually affect the outcome will be small.
@@
 Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.
-\anna{More background on lasso}
 
@@
     \path[bidirected, -] (1) edge (2);
-    \path[red, line width = 0.5] (1) edge (3);
+    \path[line width = 0.5] (1) edge (3);
     \path (2) edge (4);
     \path (3) edge (4);
-    \tkzMarkSegment[color=red,pos=.5,mark=x](1,3)
+    \tkzMarkSegment[pos=.5,mark=x](1,3)
 \end{tikzpicture}
@@
     \node (4) [below = of 2] {Phenotype};
-    \path[bidirected, -, red] (1) edge (2);
+    \path[bidirected, -] (1) edge (2);
     \path (1) edge (3);
@@
     \path (3) edge (4);
-    \tkzMarkSegment[color=red,pos=.5,mark=x](1,2)
+    \tkzMarkSegment[pos=.5,mark=x](1,2)
 \end{tikzpicture}
@@
 \label{sec:software}
-\anna{More discussion of software. Add a small section? R-based Software for lasso, pc-lasso, ggmix, plmm, PLINK functions, KING, snpRelate}
+\anna{More discussion of software.
+Add a small section?
+R-based Software for lasso, pc-lasso, ggmix, plmm, PLINK functions, KING, snpRelate.
+Remove other mentions of software in other places and put them here}
 
@@
 
-\anna{penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.}
-
-Although cross-validation and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons.
+Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons.
 Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods.
 Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples is an open question \citep{roberts2017cross}.
+\anna{We include a more detailed discussion of methods of penalty parameter selection for penalized LMMs to section \ref{sec:discussion}.}
 
 A total of 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
+\anna{maybe cut this and include in figure/table captions instead since they requested that for fig1/tab1}
+
+%------------------------------%
+\subsection{\anna{Genomewide case study}}
+%------------------------------%
 
@@
 
+\anna{\textbf{penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.}}
+
+The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
+In a real data scenario, the true number of causal SNPs would not be known a priori.
+As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
+As of yet there is no clear consensus on the optimal way to select a penalty parameter for penalized LMMs.
+
+Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \cite{bondell2010joint, schelldorfer2011estimation}.
+However it has been shown that the Akaike information criterion (AIC) and BIC may perform poorly in the high-dimensional setting \citep{fan2013tuning}.
+The high-dimensional BIC introduced by \citet{fan2013tuning} penalizes model parameters more severely in an attempt to perform better in high dimensional settings.
+\citep{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.
+
+Subsampling approaches, such as CV and stability selection, are another common class of variable selection methods often preferred to information criterion-based approaches.
+\anna{cite?
+It seems like you see CV done way more often...} \citet{Rakitsch2012} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
+\citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.
+
+As previously mentioned, however, the advisability of subsampling methods in the presence of correlated observations remains an open question \citep{roberts2017cross, jia2012preconditioning}.
+A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.
+
+\anna{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it's just confusing without a lot of additional explanation.}
+
 Finally, we note that this article has focused entirely on the lasso as a penalization scheme.

### 2021-02-14 main.tex
3.8 Software subsection

@@
 %------------------------------%
-\subsection{Software}
-\label{sec:software}
-\anna{More discussion of software.
-Add a small section?
-R-based Software for lasso, pc-lasso, ggmix, plmm, PLINK functions, KING, snpRelate.
-Remove other mentions of software in other places and put them here}
-
-%------------------------------%
 \subsection{Partitioning the effects of environmental confounding}
@@
 Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples is an open question \citep{roberts2017cross}.
-\anna{We include a more detailed discussion of methods of penalty parameter selection for penalized LMMs to section \ref{sec:discussion}.}
+\anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}
 
@@
 %------------------------------%
+
+%------------------------------%
+\subsection{\anna{Software}}
+%------------------------------%
+\label{sec:software}
+
+All analyses and simulations were performed using \texttt{R} version ???.
+The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}.
+As described previously, LMM-lasso-ggmix was implemented using \texttt{ggmix} \citep{bhatnagar2020simultaneous}.
+The LMM-lasso-Rakitsch method was implemented using the \texttt{R} package, \texttt{penalizedLMM}, available on the github repository \url{https://github.com/areisett/penalizedLMM}.
+Although \citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.
+\texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
 
@@
 
-\citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$ as previously discussed, however, we have implemented the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$ in the \texttt{R} package, \texttt{penalizedLMM}, available on the author's github repository \url{https://github.com/areisett/penalizedLMM}.
-
 Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.

### 2021-02-15 main.tex
Sec 2.3 - Addressing reviewer comment "More discussion of nonlinear models.  In particular, is logistic regression practical?"

@@
 If unaccounted for, this dependency can result in the identification of spurious associations.
-Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, the majority of this work has focused on plant and animal breeding data, a setting which affords much greater experimental control than studies of human genetics \citep{Amin2007, hoffman2013correcting, price2006principal, Rakitsch2012, bhatnagar2020simultaneous, Sillanpaeae2011}.
+Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, the majority of this work has focused on plant and animal breeding data, a setting which affords much greater experimental control than studies of human genetics \citep{Amin2007, hoffman2013correcting, price2006principal, rakitsch2013lasso, bhatnagar2020simultaneous, Sillanpaeae2011}.
 While many standard correction techniques may in principle be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and confounding arising from observational study \citep{lawson2019population, barton2019population}.
@@
 Like PC-adjustment methods, LMMs utilize a similarity matrix to estimate and subsequently correct for relatedness among subjects.
-In an LMM, however, the surrogate variables are treated as random, rather than fixed effects
-%Univariate approaches attempt to estimate the marginal effect of a SNP on the phenotype and assess the statistical significance of SNP-phenotype associations while controlling for population structure by modeling it as a random effect using a genetic similarity matrix
-\citep{yu2006unified, kang2010variance, kang2008efficient}.
+In an LMM, however, the surrogate variables are treated as random, rather than fixed effects \citep{yu2006unified, kang2010variance, kang2008efficient}.
 However, as with all univariate testing approaches, univariate LMM implementations are subject to spurious associations and biased effect estimates in the presence of LD.
 
-Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependencies among subjects \citep{Rakitsch2012, bhatnagar2020simultaneous}.
+Multivariate mixed model approaches reduce these problems by modeling the relationship between a phenotype of interest and all SNPs simultaneously via penalized regression, while correcting for dependencies among subjects \citep{rakitsch2013lasso, bhatnagar2020simultaneous}.
 This approach assumes there are a relatively small number of SNPs associated with the phenotype of interest that have large effect sizes.
@@
 
-\anna{More discussion of nonlinear models. In particular, is logistic regression practical? - threshold methods, probit regression by lippert}
+\anna{\textbf{More discussion of nonlinear models.
+In particular, is logistic regression practical?}} \anna{In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
+However, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward.
+LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation.
+This procedure does not generalize beyond the linear regression case where the outcome is continuous.
+Because of this, distinct inferential algorithms are required to extend the LMM modeling methodology to nonlinear models such as logistic regression.
+\citet{mandt2017sparse} generalize the LMM modeling paradigm to binary phenotypes in the interest of classification.
+They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference.
+However, to the best of our knowledge, the joint problem of sparse estimation in the presence of dependent observations has been limited to the linear regression case \citep{seeger2011large, vattikuti2014applying, rakitsch2013lasso}.}
 
@@
 Specifically, PC-adjustment and LMM methods, which we will refer to as PC-lasso and LMM-lasso, respectively.
-We consider two similar implementations of the LMM-lasso method, one developed by \citet{Rakitsch2012} and a more recent implementation from \citet{bhatnagar2020simultaneous}.
+We consider two similar implementations of the LMM-lasso method, one developed by \citet{rakitsch2013lasso} and a more recent implementation from \citet{bhatnagar2020simultaneous}.
 Subsequent references to LMM-lasso apply to both implementations, unless otherwise stated.
@@
 Different variance component estimation methods lead to different estimates of $\bW$ and two distinct LMM-lasso variants: LMM-lasso-Rakitsch and LMM-lasso-ggmix.
-LMM-lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmass$ \citep{Rakitsch2012}.
+LMM-lasso-Rakitsch uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmass$ \citep{rakitsch2013lasso}.
 These variance components and the corresponding weights are estimated once under the assumption of a null model.
@@
 The LMM-lasso-Rakitsch method was implemented using the \texttt{R} package, \texttt{penalizedLMM}, available on the github repository \url{https://github.com/areisett/penalizedLMM}.
-Although \citet{Rakitsch2012} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.
+Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.
 \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
@@
 \anna{cite?
-It seems like you see CV done way more often...} \citet{Rakitsch2012} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
+It seems like you see CV done way more often...} \citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
 \citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.

### 2021-02-15 main.tex
Sec 5 - reformatting penalty parameter selection edits

@@
 
-\anna{\textbf{penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.}}
-
-The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
+\anna{\textbf{penalty parameter selection; also a tough one.
+We need to say more, but not a great deal is known.} The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
 In a real data scenario, the true number of causal SNPs would not be known a priori.
 As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
-As of yet there is no clear consensus on the optimal way to select a penalty parameter for penalized LMMs.
+As of yet there is no clear consensus on the optimal way to select a penalty parameter for penalized LMMs.}
 
-Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \cite{bondell2010joint, schelldorfer2011estimation}.
+\anna{Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \cite{bondell2010joint, schelldorfer2011estimation}.
 However it has been shown that the Akaike information criterion (AIC) and BIC may perform poorly in the high-dimensional setting \citep{fan2013tuning}.
 The high-dimensional BIC introduced by \citet{fan2013tuning} penalizes model parameters more severely in an attempt to perform better in high dimensional settings.
-\citep{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.
+\citep{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.}
 
-Subsampling approaches, such as CV and stability selection, are another common class of variable selection methods often preferred to information criterion-based approaches.
+\anna{Subsampling approaches, such as CV and stability selection, are another common class of variable selection methods often preferred to information criterion-based approaches.
 \anna{cite?
@@
 \citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.
+As previously mentioned, however, the advisability of subsampling methods in the presence of correlated observations remains an open question \citep{roberts2017cross, jia2012preconditioning}.
+A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}
 
-As previously mentioned, however, the advisability of subsampling methods in the presence of correlated observations remains an open question \citep{roberts2017cross, jia2012preconditioning}.
-A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.
-
-\anna{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it's just confusing without a lot of additional explanation.}
+\anna{\textbf{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it's just confusing without a lot of additional explanation.}}
 

### 2021-02-15 main.tex
Sec 1 - Responding to reviewer comment "More  background  on  lasso"

@@
 One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.
-%LD pruning can be an important preprocessing step in a GWAS analysis, for which there exist a variety of algorithms and methods.
+
 An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
@@
 This is typically infeasible in genome-wide studies where the number of variants greatly exceeds the number of individuals.
+
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
-The lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
-\anna{More background on lasso} It relies on an assumption of sparsity; in other words, that many features will have no effect.
+\anna{\textbf{More background on lasso}} The lasso is a widely used penalized regression method \citep{tibshirani1996regression}, \anna{which can be used to model continuous, dichotomous, and survival outcomes.}  It relies on an assumption of sparsity; in other words, that many features will have no effect \anna{so that their coefficients will be estimated to be exactly 0}.
 Thus, although the total number of features may be large, the number that actually affect the outcome will be small.
 This allows coefficient estimation and variable selection to be accomplished simultaneously, and provides considerable computational advantages that scale up well to very large data sets.
-These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest.
+\anna{For the small number of features that have an effect on the outcome, large, extreme coefficient magnitudes are assumed to be less realistic and are thus shrunken towards 0.} These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest, \anna{and the effect size of a variant is typically assumed small}.
 Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.
@@
 
-All analyses and simulations were performed using \texttt{R} version ???.
+All analyses and simulations were performed using \texttt{R} version 3.6.3.
 The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}.

### 2021-02-15 main.tex
Sec  3.1 - after re-reading the original reviewer comment, adding a brief subsection to review the essential details of the lasso.

--- 2021-02-15 main.tex
@@
 This is typically infeasible in genome-wide studies where the number of variants greatly exceeds the number of individuals.
+
 Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
 The lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
-\anna{More background on lasso} It relies on an assumption of sparsity; in other words, that many features will have no effect.
+It relies on an assumption of sparsity; in other words, that many features will have no effect.
 Thus, although the total number of features may be large, the number that actually affect the outcome will be small.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%------------------------------%
+\subsection{\anna{lasso}}
+\anna{\textbf{More background on lasso: 'I acknowledge that this article cannot and should not review Lasso and other penalizations.
+But still, it is worthwhile briefly discussing what the key methodological components can do and providing references'}}
+We begin by briefly reviewing the key methodological components of the lasso.
+In the usual linear regression set-up where there is a continuous response $\by \in \real^n$, an $n \times p$ design matrix $\bX$, and a parameter vector $\bbeta \in \real^p$, the ordinary lasso minimizes the objective function
+
+\begin{equation}
+    Q(\bbeta | \lam) = \frac{1}{2n}|| \by -  \beta_0 - \bX \bbeta ||_2^2 + \lam ||\bbeta||_1,
+\end{equation}
+where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm.
+The the first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data.
+The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values.
+The tuning parameter, $\lambda$, controls the tradeoff between these two components.
+For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0.
+In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of ordinary least squares (OLS).
+\anna{\textbf{should be clearer descriptions on when the reviewed approaches can/cannot be applied; be more explicit here?}} The lasso assumes observations are independent.
+However, the $\ell_1$ penalty of the lasso can also be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.
 
@@
     \label{eqn:pca_obj}
-    Q(\bbeta, \boldsymbol{\alpha}|\bX, \by) = \frac{1}{2n} ||\by - \bX_0\bbeta_0 - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1
+    Q(\bbeta, \boldsymbol{\alpha}|\bX, \by) = \frac{1}{2n} ||\by - \bX_0\bbeta_0 - \bX \bbeta- \bU_{1:k}\boldsymbol{\alpha}||_2^2 + \lambda || \bbeta ||_1,
 \end{equation}
-where $\lambda$ is a tuning parameter, and $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm.
-Note that the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
+where the principal component regression coefficients, $\boldsymbol{\alpha}$, are not included in the penalty term.
 
@@
 This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix}.
-We further discuss differences between these methods in Section \ref{sec:results}.
+\anna{Should we still reference glmnet, ncvreg, ggmix here if we now include a software section?} We further discuss differences between these methods in Section \ref{sec:results}.
 
@@
 
+\anna{\textbf{"should be clearer descriptions on when the reviewed approaches can/cannot be applied"; be more explicit here?}}
+
 Additionally, while the fixed effects model considers the principal components of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
@@
 This is the approach we take when implementing PC-lasso in Section \ref{sec:results}.
-
-\anna{"should be clearer descriptions on when the reviewed approaches can/cannot be applied"; be more explicit here?}
 
@@
 
-To illustrate this concept mathematically, consider the ordinary least squares (OLS) estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generality, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for the sake of simplicity:
+To illustrate this concept mathematically, consider the OLS estimates of $\bbeta$ based on model \eqref{eqn:lmm_reg} where, without loss of generality, we assume $\bX_0 \bbeta_0 = \boldsymbol{0}$ for the sake of simplicity:
 \begin{align}

### 2021-02-15 main.tex
Minor tweaks

@@
 
-\anna{\textbf{More discussion of nonlinear models.
-In particular, is logistic regression practical?}} \anna{In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
+\anna{In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
 However, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward.
@@
 \subsection{\anna{lasso}}
-\anna{\textbf{More background on lasso: 'I acknowledge that this article cannot and should not review Lasso and other penalizations.
-But still, it is worthwhile briefly discussing what the key methodological components can do and providing references'}}
 We begin by briefly reviewing the key methodological components of the lasso.
@@
 where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm.
-The the first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data.
+The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data.
 The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values.
@@
 In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of ordinary least squares (OLS).
-\anna{\textbf{should be clearer descriptions on when the reviewed approaches can/cannot be applied; be more explicit here?}} The lasso assumes observations are independent.
+The lasso assumes observations are independent.
 However, the $\ell_1$ penalty of the lasso can also be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.
@@
 %------------------------------%
-\subsection{\anna{Genomewide case study}}
+\subsection{\anna{Genomewide analysis case study}}
 %------------------------------%
@@
 
-\anna{\textbf{penalty parameter selection; also a tough one.
-We need to say more, but not a great deal is known.} The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
+\anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
 In a real data scenario, the true number of causal SNPs would not be known a priori.
@@
 
-\anna{\textbf{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it's just confusing without a lot of additional explanation.}}
+\anna{\textbf{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it would just make things more confusing without a lot of additional explanation.}}
 

### 2021-02-16 main.tex
Adding '1000 replicates' and definition of TPR to relevant figure and table captions.

@@
     \includegraphics[width = \textwidth]{figures/beta_mse.png}
-    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.}
+    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.
+Mean and 1 SE bars displayed for 1000 replications.}
     \label{fig:mse}
@@
 \end{tabularx}
-\caption{True positive rates, mean (SD), for different data types in the absence ($\xi = 0$) and presence ($\xi = 0.8$) of environmental effects.
-For the data sets with simulated genotypes, 4 subpopulations are present.}
+\caption{True positive rates, defined as the proportion of the 50 causal SNPs which are selected with the correct sign, for different data types in the absence ($\xi = 0$) and presence ($\xi = 0.8$) of environmental effects.
+For the data sets with simulated genotypes, 4 subpopulations are present.
+Mean (SD) of 1000 replications.}
 \label{tab:var_sel}
@@
 \includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}
-\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$.}
+\caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$, 1000 replications.}
 \label{fig:pop_inf}
@@
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
-    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$. }
+    \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$, 1000 replications.}
     \label{fig:big_vs_small}
@@
     \includegraphics[width = \textwidth]{figures/eta_beta_hat.png}
-     \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$.}
+     \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, 1000 replications.}
     \label{fig:eta_beta_mse}

### 2021-02-16 main.tex
Removing '1000 simulations' from the text now that it's in all the figure legends.
Sec 1 - addressing reviewer  comments - "directly jumps into a statistical question" More discussion of application in general, and especially up front. - and - "should be clearer descriptions on when the reviewed approaches can/cannot be applied"; be more explicit here?

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
+\section{\anna{Introduction}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Genetic association studies have become increasingly widespread in recent years, due to their success in identifying thousands of genetic variants associated with diseases and complex traits.
+Accurately identifying and estimating the effect sizes of these variants or single nucleotide polymorphisms (SNPs) is of interest for a variety of scientific applications.
+This information can be used to predict an individual's risk of disease, elucidate the genetic architecture underlying a complex trait, and guide drug discovery.
+Genome-wide association studies (GWAS) are the most common method of analyzing genetic data sets, and involve independently testing each variant.
+However, this approach can fail to identify true associations due to the high significance thresholds required to account for multiple comparisons.
 
-\anna{More discussion of application in general, and especially up front.} The objective of many genetic studies is to accurately identify variants associated with complex traits or diseases, and to estimate the effects of those variants.
-Such information is valuable for assessing an individual's risk of developing a particular condition, elucidating the underlying genetic architecture of a trait or disease, and identifying potential drug targets.
-However, this effort may be hindered by several factors such as non-independent samples, large numbers of variants, and linkage disequilibrium (LD).
+An additional hurdle such analyses must contend with is that of population structure such as population stratification or cryptic relatedness.
+The presence of such structure can adversely affect analyses in several ways.
+Firstly, allele frequency differences due to ancestry within a sample can be mistaken for biological signal if steps are not taken to mitigate this effect.
+One common solution, chopping up a sample and performing ancestry-specific analyses, often results in a loss of statistical power from reduced sample sizes.
+Secondly, correlation among subjects within a sample, whether due to common ancestry or more recent relatedness, violates the independent errors assumption that many regression methods rely on, rendering statistical methods such as linear and logistic invalid.
 
-One method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.
-
+Additionally, testing one SNP at a time may not account for linkage disequilibrium (LD), or correlation among loci.
+Consequently, signals of association may be mis-attributed to SNPs in the absence of true biological association.
+One commonly used method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.
 An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
@@
 \subsection{\anna{lasso}}
+\anna{not sure if you/reviewers were thinking more lasso background should go here vs.
+somewhere else.
+I initially tried to add more background to the introduction but didn't think additional information "fit" there given what was already there.
+I'm open to having additional lasso background elsewhere though.}
 We begin by briefly reviewing the key methodological components of the lasso.
@@
 
-\anna{\textbf{"should be clearer descriptions on when the reviewed approaches can/cannot be applied"; be more explicit here?}}
-
 Additionally, while the fixed effects model considers the principal components of $\bK$, the LMM considers the principal components weighted by their corresponding eigenvalues.
@@
 \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}
-
-A total of 1000 simulations were performed for each unique combination of parameters, datatype, subpopulation structure, and environmental effect structure.
-\anna{maybe cut this and include in figure/table captions instead since they requested that for fig1/tab1}
 
@@
 The high-dimensional BIC introduced by \citet{fan2013tuning} penalizes model parameters more severely in an attempt to perform better in high dimensional settings.
-\citep{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.}
+\citet{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.}
 
 \anna{Subsampling approaches, such as CV and stability selection, are another common class of variable selection methods often preferred to information criterion-based approaches.
-\anna{cite?
-It seems like you see CV done way more often...} \citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
+\citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
 \citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.

### 2021-02-21 main.tex
First pass with genome-wide analysis results. Trying different dimensions on cluster.

@@
 %------------------------------%
-\subsection{\anna{Genomewide analysis case study}}
+\subsection{\anna{Genomewide analysis}}
 %------------------------------%
@@
 
+%------------------------------%
+\subsection{\anna{Genome-wide analysis case study}}
+%------------------------------%
+
+\begin{figure}[hbt]
+\centering
+\includegraphics[scale = 0.9]{figures/genomewide_prc.png}
+\caption{Precision recall curves across all $\lambda$ values from the Empirical genome-wide analysis. Single replication, $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_prc}
+\end{figure}
+
+\begin{figure}[hbt]
+\centering
+\includegraphics[scale = 0.9]{figures/genomewide_pop_inf_snps.png}
+\caption{Absolute correlation of selected, non-causal SNPs with the first principal component of the data in Empirical genome-wide analysis.
+Selection is calculated across all $\lambda$ values for a single replication.
+$\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_pop_inf}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-02-25 main.tex
Most recent genome wide figures, full page dag pdfs to get turned into eps files
Updating Figures 5, 6 (empirical genome-wide analysis)

--- 2021-02-16 main.tex
@@
 \end{tikzpicture}
-%\caption{The causal pathway between the environment and observed SNPs may be indirectly blocked by many methods that attempt to reduce confounding due to population structure.}
 \label{diag:ps_env_block}
@@
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.
-Mean and 1 SE bars displayed for 1000 replications.}
+Mean and 1 standard error (SE) bars displayed for 1000 replications.}
     \label{fig:mse}
@@
 
+%------------------------------%
+\subsection{\anna{Genome-wide analysis case study}}
+%------------------------------%
+
+\begin{figure}[hbt]
+\centering
+\includegraphics[scale = 0.9]{figures/genomewide_prc.png}
+\caption{Precision recall curves across $\lambda$ values from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_prc}
+\end{figure}
+
+\begin{figure}[hbt]
+\centering
+\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr.png}
+\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component of the data in Empirical genome-wide analysis.
+Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region for $\pm$ 1 SE.
+Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_pc1_corr}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-02-25 main.tex
Updating figures 5, 6

@@
 
-\begin{figure}[hbt]
+\begin{figure}[H]
 \centering
@@
 \caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component of the data in Empirical genome-wide analysis.
-Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region for $\pm$ 1 SE.
-Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating $\pm$ 1 SE.
+Single replication case study with $\eta = 0.5, \xi = 0.8$.
+\anna{get rid of 'Method' legend title.}}
 \label{fig:gw_pc1_corr}

### 2021-02-25 main.tex
Sec 3.8

@@
 
-We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
+We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) \anna{a subset of} real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
 We refer to these as 1D Linear Admixture, Independent Subpopulations, and Empirical, respectively.
@@
 Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods.
-Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples is an open question \citep{roberts2017cross}.
+Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}.
 \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}
@@
 %------------------------------%
-\subsection{\anna{Genomewide analysis}}
+\subsection{\anna{Genome-wide analysis}}
 %------------------------------%
+
+To illustrate the statistical properties our simulations aim to elucidate on a more realistic scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} with a synthetic phenotype as a pseudo-real data example or case study.
+We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and environmental effects ($\xi = 0.8$).
 
@@
 %------------------------------%
-\subsection{\anna{Genome-wide analysis case study}}
+\subsection{\anna{LMM-lasso outperforms lasso when analyzing genome-wide data}}
 %------------------------------%
@@
 \includegraphics[scale = 0.9]{figures/genomewide_pc1_corr.png}
-\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component of the data in Empirical genome-wide analysis.
+\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the data in Empirical genome-wide analysis.
 Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating $\pm$ 1 SE.

### 2021-03-02 main.tex
Figures 5 and 6

@@
 \includegraphics[scale = 0.9]{figures/genomewide_prc.png}
-\caption{Precision recall curves across $\lambda$ values from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\includegraphics[scale = 0.9]{figures/genomewide_prc2.png}
+\caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
 \label{fig:gw_prc}
 \end{figure}
+\anna{should the y-axes on this and the below fig match? i.e. both be number of selected snps, or both be lambda indices?}\\
+\anna{Try number of snps selected for x-axis, keep y-axis as precision}\\
+\anna{Also why is pc-lasso doing so terribly? \textbf{depends on how you're defining precision... selects a lot of true snps but often they have the wrong sign}}\\
 
@@
 \caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the data in Empirical genome-wide analysis.
-Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating $\pm$ 1 SE.
-Single replication case study with $\eta = 0.5, \xi = 0.8$.
-\anna{get rid of 'Method' legend title.}}
+Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating SE bounds.
+Single replication case study with $\eta = 0.5, \xi = 0.8$.}
 \label{fig:gw_pc1_corr}
 \end{figure}
+\anna{Should y-axis here be \# of selected snps to match above fig?}
 

### 2021-03-03 main.tex
Updating DAG figs, figure 6

@@
 The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction.
-%Otherwise stated, we do not assume that environmental factors directly influence population-specific allele frequencies, or vice versa, only that the two may be related.
+
 In the context of our air quality example, this means that we assume air quality is not directly altering individuals' allele frequencies, nor are allele frequencies causing individuals to move to regions of better or worse air quality, only that they are now correlated.
@@
 \centering
-\begin{tikzpicture}[baseline=(current  bounding  box.center)]
-    \node (1) at (0,0) {Population Stratification};
-    \node (2) [right = of 1] {Environment};
-    \node (3) [below = of 1] {Observed SNPs};
-    \node (4) [below = of 2] {Phenotype};
-    \path[bidirected, -] (1) edge (2);
-    \path (1) edge (3);
-    \path (2) edge (4);
-    \path (3) edge (4);
-\end{tikzpicture}
+\includegraphics{figures/dag1.eps}
 \label{diag:ps_env}
 \end{equation}
-
-%\begin{equation}
-%\centering
-%\includegraphics{figures/dag1.png}
-%\label{diag:ps_env}
-%\end{equation}
 
@@
 \centering
-\begin{tikzpicture}[baseline=(current  bounding  box.center)]
-    \node (1) at (0,0) {Population Stratification};
-    \node (2) [right = of 1] {Environment};
-    \node (3) [below = of 1] {Observed SNPs};
-    \node (4) [below = of 2] {Phenotype};
-    \path[bidirected, -] (1) edge (2);
-    \path[line width = 0.5] (1) edge (3);
-    \path (2) edge (4);
-    \path (3) edge (4);
-    \tkzMarkSegment[pos=.5,mark=x](1,3)
-\end{tikzpicture}
+\includegraphics{figures/dag2.eps}
 \label{diag:ps_env_block}
@@
 
-%Whether or not such methods are effective at reducing confounding effects depends upon whether population structure can be accurately inferred.
 As mentioned earlier, population structure only leads to confounding if it is correlated with a causal effect in the environment.
@@
 \centering
-\begin{tikzpicture}[baseline=(current  bounding  box.center)]
-    \node (1) at (0,0) {Population Stratification};
-    \node (2) [right = of 1] {Environment};
-    \node (3) [below = of 1] {Observed SNPs};
-    \node (4) [below = of 2] {Phenotype};
-    \path[bidirected, -] (1) edge (2);
-    \path (1) edge (3);
-    \path (2) edge (4);
-    \path (3) edge (4);
-    \tkzMarkSegment[pos=.5,mark=x](1,2)
-\end{tikzpicture}
-%\caption{The causal pathway between the environment and observed SNPs may be blocked if genetic information does not serve as an accurate proxy for it.}
+\includegraphics{figures/dag3.eps}
 \label{diag:ps_env_block2}
@@
 \subsection{LMM and PC models}
-\label{Sec:lmm-and-pc}
+\label{sec:lmm-and-pc}
 
@@
 
-In Section~\ref{Sec:correcting-structure}, we remarked that population structure only leads to confounding if it is correlated with the
-environment, and in Section~\ref{Sec:lmm-and-pc}, we remarked that in LMM modeling, a typical albeit somewhat mysterious assumption is that $\bZ=\bX$.
+In Section~\ref{Sec:correcting-structure}, we remarked that population structure only leads to confounding if it is correlated with the environment, and in Section~\ref{sec:lmm-and-pc}, we remarked that in LMM modeling, a typical albeit somewhat mysterious assumption is that $\bZ=\bX$.
 Here, we provide explicit mathematical justifications for those statements.
@@
 
-To illustrate the statistical properties our simulations aim to elucidate on a more realistic scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} with a synthetic phenotype as a pseudo-real data example or case study.
-We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and environmental effects ($\xi = 0.8$).
+To illustrate the statistical properties demonstrated by our simulations on a more realistic scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} with a synthetic phenotype as a pseudo-real data case study.
+We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
 
@@
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.
-Mean and 1 standard error (SE) bars displayed for 1000 replications.}
+Mean and standard error (SE) bars displayed for 1000 replications.}
     \label{fig:mse}

### 2021-03-03 submissions/genetic_epi.tex
Adding BNPSD reference
Using endfloat for genetic epi submission, un-numbering supplement
Sec 5. Responding to: penalty parameter selection; also a tough one. We need to say more, but not a great deal is known.
Sec  3.1 - after re-reading the original reviewer comment, adding a brief subsection to review the essential details of the lasso.
Updating DAG figs, figure 6

@@
 \usepackage{color}
-\usepackage{tikz}
-\usepackage{tikz-cd}
+%\usepackage{tikz}
+%\usepackage{tikz-cd}
 \usepackage{tkz-euclide}
 \usepackage{lscape}
+%%% UNCOMMENT THIS WHEN DONE EDITING TO PUT FIGS AT END
+%\usepackage{endfloat}
 
@@
 
-% Tikz settings optimized for causal graphs.
-% Just copy-paste this part
-\usetikzlibrary{shapes,decorations,decorations.markings,arrows,calc,arrows.meta,fit,positioning}
-\tikzset{
-    -Latex,auto,node distance =1 cm and 3 cm,semithick,
-    state/.style ={ellipse, draw, minimum width = 0.7 cm},
-    point/.style = {circle, draw, inner sep=0.04cm,fill,node contents={}},
-    bidirected/.style={Latex-Latex,dashed},
-    el/.style = {inner sep=2pt, align=left, sloped},
-    strike through/.append style={
-    decoration={markings, mark=at position 0.5 with {
-    \draw[-] ++ (-4pt,-4pt) -- (4pt,4pt);}
-  },postaction={decorate}}
-}
-
 \usepackage{subcaption}
@@
 \usepackage{natbib}
-%\usepackage{biblatex}
-%\AtEveryCitekey{%
-%  \clearfield{note}%
-%}
+
 \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
 \def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
-\usetikzlibrary{arrows}
 \providecommand{\pb}[1]{\textcolor{red}{#1}}
@@
 \newcommand{\lam}{\lambda}
+\providecommand{\real}{\mathbb{R}}
 
-\title{A comparative analysis of penalized linear mixed models in structured genetic data}
+\title{Penalized linear mixed models for structured genetic data}
 \author{Anna Reisetter \qquad Patrick Breheny\\
@@
 In addition, we evaluate the performance of penalized LMMs and competing methods in terms of estimation and selection accuracy in the presence of a number of confounding structures.
-% in the presence of confounding due to varying levels of environmental heterogeneity and relatedness.
 \end{abstract}

### 2021-03-03 main.tex
5.2 case study figure 6 results

@@
 %------------------------------%
-\subsection{\anna{Genome-wide analysis}}
-%------------------------------%
-
-To illustrate the statistical properties demonstrated by our simulations on a more realistic scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} with a synthetic phenotype as a pseudo-real data case study.
-We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
-
-%------------------------------%
 \subsection{\anna{Software}}
@@
 \label{sec:software}
+\anna{Should this be here as a methods subsection or should it get its own section after discussion, before data availability?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results} \label{sec:results}
+\section{\anna{Simulation results}} \label{sec:results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %------------------------------%
-\subsection{PC-lasso and LMM-lasso increase accuracy}
+\subsection{PC-lasso and LMM-lasso increase estimation accuracy}
 \label{sec:sim-mse}
@@
 In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
-Throughout, we use ``true positive'' to mean both that the SNP was selected and that its direction was estimated correctly.
-Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
+\anna{Throughout, we use ``true positive'' to mean both that the SNP was selected and that its direction was estimated correctly.} Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
 The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, with PC-lasso slightly less accurate at identifying causal SNPs.
@@
 %------------------------------%
-\subsection{PC-lasso and LMM-lasso reduce selection bias}
+\subsection{PC-lasso and LMM-lasso reduce selection bias} \label{sec:sel_bias}
 %------------------------------%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %------------------------------%
-\subsection{\anna{LMM-lasso outperforms lasso when analyzing genome-wide data}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{\anna{Genome-wide case study}} \label{sec:gw_cs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %------------------------------%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To illustrate the statistical properties demonstrated by our simulations on a more realistic genome-wide scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} and a synthetic phenotype as a pseudo-real data case study.
+We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
+In order to compare all methods' performance fairly, we present results for the entirety of each method's $\lambda$ path.
+
+%------------------------------%
+\subsection{\anna{LMM-lasso identifies causal SNPs with greater precision than lasso}}
+%------------------------------%
+\begin{figure}[H]
+\centering
+\includegraphics[scale = 0.9]{figures/genomewide_prc.png}
+\caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_prc}
+\end{figure}
+
+$$\text{Precision }= 1 - \frac{\#\{ \beta = 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}} = \frac{\#\{ \beta \ne 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
 
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_prc.png}
 \includegraphics[scale = 0.9]{figures/genomewide_prc2.png}
 \caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-\label{fig:gw_prc}
+\label{fig:gw_prc2}
 \end{figure}
-\anna{should the y-axes on this and the below fig match? i.e. both be number of selected snps, or both be lambda indices?}\\
-\anna{Try number of snps selected for x-axis, keep y-axis as precision}\\
-\anna{Also why is pc-lasso doing so terribly? \textbf{depends on how you're defining precision... selects a lot of true snps but often they have the wrong sign}}\\
+
+$$\text{Precision }= \frac{\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
+
+\anna{Why is pc-lasso doing so terribly? \textbf{original bad seed + depends on how you're defining precision... selects a lot of true snps but often they have the wrong sign}}\\
+
+\noindent\anna{should the y-axes on this and the below fig match? i.e. both be number of selected snps, or both be lambda indices?}\\
+
+%------------------------------%
+\subsection{\anna{PC-lasso and LMM-lasso reduce selection bias in genome-wide data}}
+%------------------------------%
+
+A central issue explored in this work has been the potential for population stratification and environmental confounding to lead to the identification of spurious associations.
+In Section~\ref{sec:sel_bias} we characterized the three lasso methods (ordinary, PC and LMM) in terms of their susceptibility to this type of selection bias in a simplified setting of two subpopulations.
+We now seek to examine these methods' performance in a genome-wide, pathwise manner.
+
+Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a particular $\lambda$ value with the first principal component (PC1) of the Empirical data for each method across the full $\lambda$ path.
+The corresponding shaded regions indicate the SE about the mean; overall these bounds are largest early on in each $\lambda$ path, where only a few SNPs are be selected, but narrow as $\lambda$ increases and more variables are allowed to enter the model.\anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.
 
@@
 \end{figure}
-\anna{Should y-axis here be \# of selected snps to match above fig?}
+
+The ordinary lasso is clearly most susceptible to selecting non-causal SNPs due to their association with population structure.
+This corresponds to the simulation results shown in Section~\ref{sec:sel_bias}, though in Figure \ref{fig:gw_pc1_corr} we more plainly see that this bias persists across the $\lambda$ path.
+Likewise, PC-lasso most effectively controls this type of selection bias.
+This is to be expected given that we are assessing the population-informative nature of incorrectly selected SNPs via correlation with PC1, which is explicitly adjusted for in the PC-lasso method.
+LMM-lasso yields a substantial improvement compared to the ordinary lasso in controlling this selection bias.
+It should be noted that the Empirical data's RRM in Figure~\ref{fig:empirical} shows that its structure is predominantly coarse.
+Given our results in Section~\ref{sec:sim-fine-coarse}, we might expect a still more substantial improvement over the ordinary lasso in the analysis of more finely structured genome-wide data.
 
@@
 
+\anna{Add discussion of genome-wide results}
+
 We considered two implementations of the LMM-lasso method: LMM-lasso-Rakitsch and LMM-lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
@@
 
+\textbf{\anna{Add reference to that paper that mentions this challenge explicitly}}
+
 \anna{Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \cite{bondell2010joint, schelldorfer2011estimation}.
@@
 
-\anna{\textbf{I don't think we should bring up the idea of CV/subsampling on original vs transformed data in this paper at all - I think it would just make things more confusing without a lot of additional explanation.}}
-
 Finally, we note that this article has focused entirely on the lasso as a penalization scheme.

### 2021-03-04 main.tex
Adding pathwise MSE figure

@@
 We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
-In order to compare all methods' performance fairly, we present results for the entirety of each method's $\lambda$ path.
+In order to compare all methods' performance fairly and completely, we present results for the entirety of each method's $\lambda$ path.
 
@@
 \anna{Why is pc-lasso doing so terribly? \textbf{original bad seed + depends on how you're defining precision... selects a lot of true snps but often they have the wrong sign}}\\
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale = 0.9]{figures/genomewide_mse.png}
+\caption{Pathwise MSE from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_mse}
+\end{figure}
 
@@
 LMM-lasso yields a substantial improvement compared to the ordinary lasso in controlling this selection bias.
-It should be noted that the Empirical data's RRM in Figure~\ref{fig:empirical} shows that its structure is predominantly coarse.
-Given our results in Section~\ref{sec:sim-fine-coarse}, we might expect a still more substantial improvement over the ordinary lasso in the analysis of more finely structured genome-wide data.
 
@@
 
-Our simulations experiments reveal, unsurprisingly, that PC-lasso and LMM-lasso outperform the ordinary lasso whenever environmental confounding effects are present.
+Our simulation experiments reveal, unsurprisingly, that PC-lasso and LMM-lasso outperform the ordinary lasso whenever environmental confounding effects are present.
 However, in the absence of such effects, PC-lasso may hinder estimation accuracy.
@@
 
-\anna{Add discussion of genome-wide results}
-
 We considered two implementations of the LMM-lasso method: LMM-lasso-Rakitsch and LMM-lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
@@
 
-Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.
-In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{\tau}$ unless steps are taken to address population structure.
-A detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper, but we hope this emphasizes the broad implications and importance of distinguishing between population structure and environmental effects.
+\anna{The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
+Although PC-lasso was found to control this selection bias most effectively, it should be noted that because the Empirical data consists of three distinct ethnic groups, its RRM (Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
+Given our results in Section~\ref{sec:sim-fine-coarse}, we might expect LMM-lasso to provide greater improvement over the ordinary lasso in the analysis of more finely structured genome-wide data.}
 
@@
 
+Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.
+In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{\tau}$ unless steps are taken to address population structure.
+A detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper, but we hope this emphasizes the broad implications and importance of distinguishing between population structure and environmental effects.
+
 Finally, we note that this article has focused entirely on the lasso as a penalization scheme.

### 2021-03-04 main.tex
Additional edits throughout - reorganizing, making more 'friendly'

@@
 However, this approach can fail to identify true associations due to the high significance thresholds required to account for multiple comparisons.
-
-An additional hurdle such analyses must contend with is that of population structure such as population stratification or cryptic relatedness.
-The presence of such structure can adversely affect analyses in several ways.
-Firstly, allele frequency differences due to ancestry within a sample can be mistaken for biological signal if steps are not taken to mitigate this effect.
-One common solution, chopping up a sample and performing ancestry-specific analyses, often results in a loss of statistical power from reduced sample sizes.
-Secondly, correlation among subjects within a sample, whether due to common ancestry or more recent relatedness, violates the independent errors assumption that many regression methods rely on, rendering statistical methods such as linear and logistic invalid.
 
@@
 
-As with multiple linear regression models, lasso penalized regression assumes independent observations and uncorrelated errors.
+% As with multiple linear regression models, lasso penalized regression assume subjects are independent of each other with uncorrelated errors.
 In the presence of population structure or cryptic relatedness, however, samples are correlated, violating this assumption.
 If unaccounted for, this dependency can result in the identification of spurious associations.
+
+As with multiple linear regression models, lasso penalized regression assume subjects are independent of each other with uncorrelated errors.
+In the presence of population structure or cryptic relatedness, however, samples are correlated, violating this assumption.
+The presence of structure can adversely affect analyses in several ways.
+Firstly, allele frequency differences due to ancestry within a sample can be mistaken for biological signal if steps are not taken to mitigate this effect.
+One common solution, dividing a sample and performing ancestry-specific analyses, often results in a loss of statistical power from reduced sample sizes.
+Secondly, correlation among subjects within a sample, whether due to common ancestry or more recent relatedness, violates an assumption that many regression methods rely on, rendering these statistical methods invalid.
+
 Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, the majority of this work has focused on plant and animal breeding data, a setting which affords much greater experimental control than studies of human genetics \citep{Amin2007, hoffman2013correcting, price2006principal, rakitsch2013lasso, bhatnagar2020simultaneous, Sillanpaeae2011}.
@@
 The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data.
+As in ordinary least squares (OLS) regression, the lasso assumes observations are independent.
 The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values.
@@
 For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0.
-In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of ordinary least squares (OLS).
-The lasso assumes observations are independent.
-However, the $\ell_1$ penalty of the lasso can also be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.
+In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of OLS.
+The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.
 
@@
 In Section~\ref{sec:sel_bias} we characterized the three lasso methods (ordinary, PC and LMM) in terms of their susceptibility to this type of selection bias in a simplified setting of two subpopulations.
-We now seek to examine these methods' performance in a genome-wide, pathwise manner.
+We now examine these methods' performance in a genome-wide data set with greater structural complexity.
 
 Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a particular $\lambda$ value with the first principal component (PC1) of the Empirical data for each method across the full $\lambda$ path.
-The corresponding shaded regions indicate the SE about the mean; overall these bounds are largest early on in each $\lambda$ path, where only a few SNPs are be selected, but narrow as $\lambda$ increases and more variables are allowed to enter the model.\anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.
+The corresponding shaded regions indicate the SE about the mean.
+Overall, these bounds are largest early on in each $\lambda$ path, where only a few SNPs are be selected, but narrow as $\lambda$ increases and more variables are allowed to enter the model.\anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.
 
@@
 
-\anna{The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
+\anna{\textbf{Additional genome-wide results based on other figures we decide to include.} The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
 Although PC-lasso was found to control this selection bias most effectively, it should be noted that because the Empirical data consists of three distinct ethnic groups, its RRM (Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
 Given our results in Section~\ref{sec:sim-fine-coarse}, we might expect LMM-lasso to provide greater improvement over the ordinary lasso in the analysis of more finely structured genome-wide data.}
+
+Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.
+In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{\tau}$ unless steps are taken to address population structure.
+A detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper, but we hope this emphasizes the broad implications and importance of distinguishing between population structure and environmental effects.
 
@@
 As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
-As of yet there is no clear consensus on the optimal way to select a penalty parameter for penalized LMMs.}
+Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm, which refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso.
+At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no clear consensus on the optimal way to select a penalty parameter for such algorithms.}
 
-\textbf{\anna{Add reference to that paper that mentions this challenge explicitly}}
-
-\anna{Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \cite{bondell2010joint, schelldorfer2011estimation}.
+\anna{Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \citep{bondell2010joint, schelldorfer2011estimation}.
 However it has been shown that the Akaike information criterion (AIC) and BIC may perform poorly in the high-dimensional setting \citep{fan2013tuning}.
@@
 
-\anna{Subsampling approaches, such as CV and stability selection, are another common class of variable selection methods often preferred to information criterion-based approaches.
+\anna{Subsampling approaches, such as CV and stability selection are another common class of variable selection methods often preferred to information criterion-based approaches in practice.
 \citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
@@
 A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}
-
-Although the purpose of this article is to elucidate the manner in which failing to correct for unobserved environmental effects can bias SNP effect estimates, equation \eqref{eqn:partition_e} also highlights the manner in which such unobserved environmental effects can lead to substantial bias in narrow-sense heritability estimates using SNP data.
-In particular, these estimates are likely to be inflated by the quantity $\boldsymbol{\tau}$ unless steps are taken to address population structure.
-A detailed treatment of narrow-sense heritability estimation is beyond the scope of this paper, but we hope this emphasizes the broad implications and importance of distinguishing between population structure and environmental effects.
 

### 2021-03-04 main.tex
Removing reference to 'CV is bad for dependent data' paper (Roberts 2017)

@@
 Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods.
-Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}.
+\anna{ I want to cut the following sentence but I think we may need some other reason why we didn't do CV...} Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}.
 \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}
@@
 \citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.
-As previously mentioned, however, the advisability of subsampling methods in the presence of correlated observations remains an open question \citep{roberts2017cross, jia2012preconditioning}.
 A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}

### 2021-03-05 main.tex
Section 5 genome-wide figure options

@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_prc.png}
+\includegraphics[scale = 0.6]{figures/genomewide_prc_p1=100.png}
+\includegraphics[scale = 0.6]{figures/genomewide_prc_p1=500.png}
+\includegraphics[scale = 0.6]{figures/genomewide_prc_p1=1000.png}
 \caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
@@
 
-$$\text{Precision }= 1 - \frac{\#\{ \beta = 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}} = \frac{\#\{ \beta \ne 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
+$$\text{Precision }= \frac{\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
+
+%------------------------------%
+\subsection{\anna{PC-lasso and LMM-lasso increase estimation accuracy}}
+%------------------------------%
 
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_prc2.png}
-\caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-\label{fig:gw_prc2}
-\end{figure}
-
-$$\text{Precision }= \frac{\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
-
-\anna{Why is pc-lasso doing so terribly? \textbf{original bad seed + depends on how you're defining precision... selects a lot of true snps but often they have the wrong sign}}\\
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale = 0.9]{figures/genomewide_mse.png}
+\includegraphics[scale = 0.6]{figures/genomewide_mse_p1=100.png}
+\includegraphics[scale = 0.6]{figures/genomewide_mse_p1=500.png}
+\includegraphics[scale = 0.6]{figures/genomewide_mse_p1=1000.png}
 \caption{Pathwise MSE from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
@@
 
-Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a particular $\lambda$ value with the first principal component (PC1) of the Empirical data for each method across the full $\lambda$ path.
+\anna{rewrite to reflect that the x-axis is number of selected snps, not lambda indices.
+update results / discussion as necessary based on final p1 chosen} Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a particular $\lambda$ value with the first principal component (PC1) of the Empirical data for each method across the full $\lambda$ path.
 The corresponding shaded regions indicate the SE about the mean.
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr.png}
-\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the data in Empirical genome-wide analysis.
-Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating SE bounds.
+\includegraphics[scale = 0.6]{figures/genomewide_pc1_corr_p1=100.png}
+\includegraphics[scale = 0.6]{figures/genomewide_pc1_corr_p1=500.png}
+\includegraphics[scale = 0.6]{figures/genomewide_pc1_corr_p1=1000.png}
+\caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the data in Empirical genome-wide analysis.
+Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular number of incorrectly selected SNPs, with shaded region indicating SE bounds.
 Single replication case study with $\eta = 0.5, \xi = 0.8$.}

### 2021-03-05 main.tex
Finalizing genome-wide figures
Sections 5 and 6

--- 2021-03-04 main.tex
@@
 In addition to MSE, these methods' performance was also assessed in terms of variable selection using the true positive rate (TPR) for the first 50 variables each method allowed into the model.
-\anna{Throughout, we use ``true positive'' to mean both that the SNP was selected and that its direction was estimated correctly.} Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
+Throughout, we use ``true positive'' to mean both that the SNP was selected and that its direction was estimated correctly.
+Table \ref{tab:var_sel} displays the average TPR for the same settings as in Figure \ref{fig:mse}.
 The TPRs echo the MSE results in that lasso and LMM-lasso perform comparably in the absence of environmental confounding across all data types, with PC-lasso slightly less accurate at identifying causal SNPs.
@@
 To illustrate the statistical properties demonstrated by our simulations on a more realistic genome-wide scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} and a synthetic phenotype as a pseudo-real data case study.
-We generate the synthetic phenotype as in \eqref{eqn:sim} using 50,000 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
-In order to compare all methods' performance fairly and completely, we present results for the entirety of each method's $\lambda$ path.
+We generate the synthetic phenotype as in \eqref{eqn:sim} using 100 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
+In order to compare methods fairly and completely, we consider results across the entirety of each methods' $\lambda$ path, indexed by the number of selected variables for interpretatiblity.
 
@@
 %------------------------------%
+
+Selection accuracy was evaluated using selection precision.
+All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.
+Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
 \begin{figure}[H]
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_prc.png}
-\caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\includegraphics[scale = 0.9]{figures/genomewide_prc_p1=100.png}
+\caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
 \label{fig:gw_prc}
@@
 
-$$\text{Precision }= 1 - \frac{\#\{ \beta = 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}} = \frac{\#\{ \beta \ne 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
+PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selection accuracy, particularly for the first $\approx$ 100 SNPs each method allows into the model.
+For the most part, PC-lasso selects SNPs outperforms LMM-lasso in this setting as well.
+All methods struggle as the model becomes less sparse and the number of selected SNPs approaches $n$.
+
+%------------------------------%
+\subsection{\anna{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}}
+%------------------------------%
+
+Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}. Figure~\ref{fig:gw_mse} displays this value for each method and number of selected SNPs.
 
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_prc2.png}
-\caption{Precision values for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-\label{fig:gw_prc2}
-\end{figure}
-
-$$\text{Precision }= \frac{\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}}{\#\{ \widehat{\beta} \ne 0 \}}$$
-
-\anna{Why is pc-lasso doing so terribly? \textbf{original bad seed + depends on how you're defining precision... selects a lot of true snps but often they have the wrong sign}}\\
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale = 0.9]{figures/genomewide_mse.png}
-\caption{Pathwise MSE from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\includegraphics[scale = 0.9]{figures/genomewide_mse_p1=100.png}
+\caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
 \label{fig:gw_mse}
@@
 
-\noindent\anna{should the y-axes on this and the below fig match? i.e. both be number of selected snps, or both be lambda indices?}\\
+In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso.
+This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method.
+Both PC-lasso and LMM-lasso produce more accurate estimates than the ordinary lasso for any number of selected SNPs.
+
+%PC-lasso and LMM-lasso coincide in MSE for the first few selected variables, for the most part, however, PC-lasso outperforms LMM-lasso.
+
+% This is an interesting result in contrast to that observed in the third panel of Figure~\ref{fig:mse}, where we analyzed a subset of the Empirical data using 50 causal SNPs and LMM-lasso slightly outperformed PC-lasso.
+It seems that in the genome-wide setting the coarse, subpopulation-based structure dominates the subtle relatedness present in the sample
 
@@
 
-Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a particular $\lambda$ value with the first principal component (PC1) of the Empirical data for each method across the full $\lambda$ path.
+Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a selected SNPs with the first principal component (PC1) of the Empirical data for each number of incorrectly selected SNPs.
 The corresponding shaded regions indicate the SE about the mean.
-Overall, these bounds are largest early on in each $\lambda$ path, where only a few SNPs are be selected, but narrow as $\lambda$ increases and more variables are allowed to enter the model.\anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.
+\anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.
 
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr.png}
-\caption{Pathwise absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the data in Empirical genome-wide analysis.
-Plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular $\lambda$ value, with shaded region indicating SE bounds.
+\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr_p1=100.png}
+\caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the Empirical genome-wide data for each method and number of selected SNPs.
+The plotted line is the mean absolute correlation of all incorrectly selected SNPs for a particular selection number, with shaded region indicating SE bounds.
 Single replication case study with $\eta = 0.5, \xi = 0.8$.}
@@
 The ordinary lasso is clearly most susceptible to selecting non-causal SNPs due to their association with population structure.
-This corresponds to the simulation results shown in Section~\ref{sec:sel_bias}, though in Figure \ref{fig:gw_pc1_corr} we more plainly see that this bias persists across the $\lambda$ path.
-Likewise, PC-lasso most effectively controls this type of selection bias.
-This is to be expected given that we are assessing the population-informative nature of incorrectly selected SNPs via correlation with PC1, which is explicitly adjusted for in the PC-lasso method.
-LMM-lasso yields a substantial improvement compared to the ordinary lasso in controlling this selection bias.
+This corresponds to the simulation results shown in Section~\ref{sec:sel_bias}, though in Figure \ref{fig:gw_pc1_corr} we more plainly see that this bias persists regardless of the number of variables allowed into the model.
+Likewise, PC-lasso and LMM-lasso both effectively control this type of selection bias.
+PC-lasso's effective control of bias in this setting is to be expected, given that we are assessing the population-informative nature of incorrectly selected SNPs via correlation with PC1, which is explicitly adjusted for in the PC-lasso method.
 
@@
 
+\anna{The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
+Although PC-lasso was found to outperform LMM-lasso in terms of precision and estimation accuracy in this analysis, because the Empirical data consists of three distinct racial groups, its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
+Given our results in Section~\ref{sec:sim-fine-coarse}, we would expect LMM-lasso to outperform PC-lasso in the analysis of more finely structured genome-wide data.}
+
 We considered two implementations of the LMM-lasso method: LMM-lasso-Rakitsch and LMM-lasso-ggmix, which differ in their variance estimation procedures and computational implementations.
 Although both methods perform similarly in terms of estimating $\bbeta$, the much greater computational efficiency and numerical stability of LMM-lasso-Rakitsch make it a more attractive choice in most scenarios, although LMM-lasso-ggmix may be useful if the objective of the analysis is to accurately estimate variance components.
-
-\anna{\textbf{Additional genome-wide results based on other figures we decide to include.} The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
-Although PC-lasso was found to control this selection bias most effectively, it should be noted that because the Empirical data consists of three distinct ethnic groups, its RRM (Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
-Given our results in Section~\ref{sec:sim-fine-coarse}, we might expect LMM-lasso to provide greater improvement over the ordinary lasso in the analysis of more finely structured genome-wide data.}
 
@@
 As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
-Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm, which refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso.
-At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no clear consensus on the optimal way to select a penalty parameter for such algorithms.}
+Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm.
+This refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso.
+At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no consensus on the optimal way to select a penalty parameter for such algorithms.}
 

### 2021-03-05 main.tex
Section 6

@@
 \anna{The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
-Although PC-lasso was found to outperform LMM-lasso in terms of precision and estimation accuracy in this analysis, because the Empirical data consists of three distinct racial groups, its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
-Given our results in Section~\ref{sec:sim-fine-coarse}, we would expect LMM-lasso to outperform PC-lasso in the analysis of more finely structured genome-wide data.}
+We note that because the Empirical data consists of three distinct racial groups, its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
+Given our results in Section~\ref{sec:sim-fine-coarse}, we would expect LMM-lasso to perform even better in the analysis of more finely structured genome-wide data.}
 

### 2021-03-12 submissions/response/response.tex
point-by-point response: Skeleton added

\documentclass{article}
\usepackage[left=0.5in,right=0.5in,bottom=0.75in,top=0.75in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{color}
\providecommand{\note}[1]{\textcolor{red}{#1}}

\begin{document}

\subsubsection*{Response to comments by Reviewer \#1}

Thank you very much for your detailed reading of the paper and your insightful questions. You raise a number of good points and we have modified the manuscript in several places based upon your feedback.  In our opinion, these revisions have nicely improved upon the original submission.

\begin{enumerate}

\item \emph{This is a nicely conducted review/educational article on an important topic. My main ``complaint'' is that it is not ``sufficiently friendly''.  The article directly jumps into a statistical question (and solutions). There is insufficient discussion/description on the application scenarios. There can/should be clearer descriptions on when the reviewed approaches can/cannot be applied.}

  Response.

\item \emph{In the discussion of methods, there are similar concerns. For example, for some or most statisticians, Lasso is not a new concept. But I can bet that not all practitioners know what Lasso does. I acknowledge that this article cannot and should not review Lasso and other penalizations. But still, it is worthwhile briefly discussing what the key methodological components can do and providing references.}

  Response

\item \emph{Simulation is conducted and is informative. However, for practitioners, it is of interest to see what kind of practical impact the approaches may have. Data examples are strongly recommended.}

  Response

\item \emph{Related to the above points: how can someone not savvy in statistics/penalization etc. use these approaches? Any simple software? I can see some scattered information for example in Discussion. One possibility is to have a separate subsection or table for software. Application examples too.}

  Response

\item \emph{I understand that the authors’ focus is on linear model. But, many if not most genetic studies do not have a continuous outcome and linear model. Can the models/approaches be extended to other outcomes/models with reasonable effort? Either way, it is worth knowing.}

  Response

\item \emph{This may seem contradicting the above points: for a few statisticians, it may be of interest to know more about the statistical properties, especially under high-dimensional settings. Brief discussions (maybe a paragraph? Or a subsection?) on the statistical properties may be informative.}

  Response.

\end{enumerate}

\newpage

\subsubsection*{Response to comments by Reviewer \#2}

Thank you very much for your detailed reading of the paper and your insightful questions. You raise a number of good points and we have modified the manuscript in several places based upon your feedback.  In our opinion, these revisions have nicely improved upon the original submission.

\begin{enumerate}

\item \emph{This is a very insightful paper as to the performance of lasso LMM methods in GWAS analyses and provides nice intuition using theoretical derivations on the interplay between  population structure and environmental heterogeneity. The paper could highly benefit from using more realistic simulation scenarios reflecting standard GWAS (increase the number of SNPs in model which is more in the 100,000s in current GWASes) as well as provide some assessment of how robust the results would be to different lasso penalty parameters (as in practice we would not have a priori an optimal value).}

  Response

\item \emph{In the captions of Figs. 1 and Table 1, it is not clear how many replicates are used to estimate these quantities (e.g. all p=1000 SNPs used in simulations or only the 50 SNPs selected by lasso?). For Figure 1, it would be helpful to have error bars included in bar plots. Also, there is a typo in Eqn 15.}

  Response

\end{enumerate}

%\bibliographystyle{ims-nourl}
%\bibliography{articles}

\end{document}

### 2021-03-12 submissions/response/response.tex
Edits to introduction

@@
 
-Thank you very much for your detailed reading of the paper and your insightful questions.
-You raise a number of good points and we have modified the manuscript in several places based upon your feedback.
-In our opinion, these revisions have nicely improved upon the original submission.
+Thank you very much for your detailed reading of the paper and your insightful suggestions.
+You raise a number of good points and we have modified the manuscript considerably based upon your feedback.
+In our opinion, these revisions have very much improved upon the original submission.
 
@@
 
-  Response.
+  You raise a good point.
+We have tried to improve upon this aspect of the paper by rewriting the introduction with the application, rather than the statistical methodology, in mind.
+Furthermore, MENTION DISCUSSION?
+MENTION APPLICATION / SECTION 5?
 

### 2021-03-12 main.tex
Edits to introduction

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\anna{Introduction}}
+\section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Genetic association studies have become increasingly widespread in recent years, due to their success in identifying thousands of genetic variants associated with diseases and complex traits.
-Accurately identifying and estimating the effect sizes of these variants or single nucleotide polymorphisms (SNPs) is of interest for a variety of scientific applications.
-This information can be used to predict an individual's risk of disease, elucidate the genetic architecture underlying a complex trait, and guide drug discovery.
-Genome-wide association studies (GWAS) are the most common method of analyzing genetic data sets, and involve independently testing each variant.
-However, this approach can fail to identify true associations due to the high significance thresholds required to account for multiple comparisons.
 
-Additionally, testing one SNP at a time may not account for linkage disequilibrium (LD), or correlation among loci.
-Consequently, signals of association may be mis-attributed to SNPs in the absence of true biological association.
-One commonly used method for analyzing data in the presence of correlated loci is LD pruning, which involves sequentially thinning SNPs that exceed a correlation threshold, so that only variants that are relatively independent are used in modeling and testing procedures.
-An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of numerous variants simultaneously.
-However, classical multivariate approaches rely on data with more observations than unknown parameters.
-This is typically infeasible in genome-wide studies where the number of variants greatly exceeds the number of individuals.
+\pb{Read this over and let me know what you think}
 
-Penalized regression methods make such problems tractable by constraining the solution space of coefficient estimates.
-The lasso is a widely used penalized regression method \citep{tibshirani1996regression}.
-It relies on an assumption of sparsity; in other words, that many features will have no effect.
-Thus, although the total number of features may be large, the number that actually affect the outcome will be small.
+Genome-wide association studies (GWAS) have become widespread over the past fifteen years, successfully identifying thousands of genetic variants associated with diseases and complex traits.
+Accurately identifying and estimating the effect sizes of these variants or single nucleotide polymorphisms (SNPs) is of interest for a variety of scientific applications including the prediction of an individual's risk of disease, understanding the genetic architecture underlying a complex trait, and drug discovery.
+Genetic association is typically analyzing by testing individual variants one at a time.
+While straightforward, there are inherent limitations to analyzing each variant in isolation including diminished power, an inability to examine epistatic interactions, and bias due to the presence of correlation between variants.
+
+Correlation among genetic loci, or linkage disequilibrium (LD), causes true biological associations to be mis-attributed to nearby, correlated SNPs.
+If one is analyzing each variant in isolation, then additional steps need to be taken to avoid this bias.
+For example, in LD pruning, SNPs that exceed a correlation threshold are sequentially removed until only variants that are relatively independent are used in analysis and testing procedures.
+An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of many variants simultaneously.
+However, classical regression relies on data with more many more observations than variants.
+This is typically infeasible in genome-wide studies where the variants to be analyzed number in the hundreds of thousands or even millions.
+
+Penalized regression methods such as the lasso \citep{tibshirani1996regression} make such problems tractable by introducing sparsity, which means that many variants have no effect on the trait of interest.
+Thus, although the total number of variants may be large, the number that actually affect the outcome will be small.
 This allows coefficient estimation and variable selection to be accomplished simultaneously, and provides considerable computational advantages that scale up well to very large data sets.
-These features make the lasso attractive for the analysis of high-dimensional genetic data where identifying the variants most highly associated with a particular trait is often of interest.
+These features make the lasso attractive for the analysis of large GWAS data sets.
 Although separately testing each SNP is still the most common way of analyzing GWAS data, the use of lasso and related methods is increasing in use due to computational advances that make it practical to apply on a genome-wide scale \citep{qian2020fast, prive2018efficient}.
 
-% As with multiple linear regression models, lasso penalized regression assume subjects are independent of each other with uncorrelated errors.
-In the presence of population structure or cryptic relatedness, however, samples are correlated, violating this assumption.
-If unaccounted for, this dependency can result in the identification of spurious associations.
-
-As with multiple linear regression models, lasso penalized regression assume subjects are independent of each other with uncorrelated errors.
-In the presence of population structure or cryptic relatedness, however, samples are correlated, violating this assumption.
-The presence of structure can adversely affect analyses in several ways.
-Firstly, allele frequency differences due to ancestry within a sample can be mistaken for biological signal if steps are not taken to mitigate this effect.
+Lasso models, however, rely on an assumption that subjects are independent of each other.
+In the presence of population structure or cryptic relatedness, however, this assumption is violated: traits are more likely to be similar within a family or ethnic group.
+The presence of structure can adversely affect analyses in several ways, but the most important concern is that it is likely to introduce false positives.
+Allele frequency differences due to ancestry within a sample are likely to be mistaken for biological signal if steps are not taken to mitigate this effect.
 One common solution, dividing a sample and performing ancestry-specific analyses, often results in a loss of statistical power from reduced sample sizes.
-Secondly, correlation among subjects within a sample, whether due to common ancestry or more recent relatedness, violates an assumption that many regression methods rely on, rendering these statistical methods invalid.
+Furthermore, ancestry-specific analyses cannot easily be applied in the presence of admixture.
 
 Though methods to correct for population stratification and cryptic relatedness have been extensively investigated, the majority of this work has focused on plant and animal breeding data, a setting which affords much greater experimental control than studies of human genetics \citep{Amin2007, hoffman2013correcting, price2006principal, rakitsch2013lasso, bhatnagar2020simultaneous, Sillanpaeae2011}.
-While many standard correction techniques may in principle be applied to human genetic data, their performance has not been well-studied within the context of human ancestry and confounding arising from observational study \citep{lawson2019population, barton2019population}.
+While many standard correction techniques may in principle be applied to human genetic data, their performance has not been well studied within the context of human ancestry and confounding arising from observational study \citep{lawson2019population, barton2019population}.
 
@@
 
-In order to clarify and emphasize the importance of the subtle distinction between population structure and environmental heterogeneity, we present a detailed review of relevant concepts and methods.
+In order to clarify the subtle distinction between population structure and environmental heterogeneity and illustrate its importance, we present a detailed review of relevant concepts and methods.
 The remainder of this paper is organized as follows.
@@
 In addition, we briefly review historical methods of correcting for population stratification and relatedness.
-In Section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness at present: PCA adjusted lasso penalized regression (PC-lasso) and lasso penalized multivariate LMMs (LMM-lasso).
+In Section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness: PCA adjusted lasso penalized regression (PC-lasso) and lasso penalized multivariate LMMs (LMM-lasso).
 We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
-In Section \ref{sec:results}, we illustrate these concepts via simulation studies, where the data-generating mechanism allows us to distinguish between population structure and environmental heterogeneity.
-In addition, these results identify scenarios where particular methods may be most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
+In Section \ref{sec:results}, we illustrate these concepts via simulation studies and identify scenarios where particular methods are most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
+Finally, in Section~\ref{sec:gw_cs} we illustrate the use of these methods in a case study involving genome-wide data involving a sample of diverse racial background.
 

### 2021-03-12 main.tex
Edits to binary paragraph, lasso section, and software

@@
 
-\anna{In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
-However, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward.
-LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation.
-This procedure does not generalize beyond the linear regression case where the outcome is continuous.
-Because of this, distinct inferential algorithms are required to extend the LMM modeling methodology to nonlinear models such as logistic regression.
+\pb{Please read the following paragraph; it has been modified extensively.}
+
+In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
+Unfortunately, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward.
+As we explain in Section~\ref{sec:methods}, LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation.
+This procedure does not directly generalize beyond the linear regression case.
+The situation is certainly not impossible, and many approaches to fitting generalized linear mixed models, or GLMMs, have been proposed in the classical, low-dimensional setting.
+However, very little work has been done to extend this to penalized, high-dimensional models and to the best of our knowledge there is no existing software for doing so at this time.
+We hope that this situation changes in the near future.
+
+\pb{What is the significance of this Mandt paper?
+Isn't there a huge literature on GLMMs?
 \citet{mandt2017sparse} generalize the LMM modeling paradigm to binary phenotypes in the interest of classification.
 They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference.
-However, to the best of our knowledge, the joint problem of sparse estimation in the presence of dependent observations has been limited to the linear regression case \citep{seeger2011large, vattikuti2014applying, rakitsch2013lasso}.}
+}
 
@@
 %------------------------------%
-\subsection{\anna{lasso}}
-\anna{not sure if you/reviewers were thinking more lasso background should go here vs.
-somewhere else.
-I initially tried to add more background to the introduction but didn't think additional information "fit" there given what was already there.
-I'm open to having additional lasso background elsewhere though.}
+\subsection{Lasso}
 We begin by briefly reviewing the key methodological components of the lasso.
-In the usual linear regression set-up where there is a continuous response $\by \in \real^n$, an $n \times p$ design matrix $\bX$, and a parameter vector $\bbeta \in \real^p$, the ordinary lasso minimizes the objective function
+In the usual linear regression setting involving a continuous response $\by \in \real^n$, an $n \times p$ design matrix $\bX$, and a parameter vector $\bbeta \in \real^p$, the ordinary lasso minimizes the objective function
 
@@
 where $|| \boldsymbol{\beta} ||_1 = \sum_{j=1}^p |\beta_j|$ denotes the $\ell_1$ norm of the regression coefficients, and $||\mathbf{x}||_2 = \sqrt{\sum_{i=1}^p x_i^2}$ the $\ell_2$ norm.
-The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data.
-As in ordinary least squares (OLS) regression, the lasso assumes observations are independent.
+The first component of the objective function is the linear regression loss function, which quantifies the overall model fit and captures how well the estimated parameters agree with the observed data (here, assuming that observations are independent).
 The second component is the $\ell_1$ penalty applied to the estimated coefficient values (but not the intercept, $\beta_0$), which discourages large, unrealistic parameter values.
@@
 For sufficiently large values of $\lambda$ all elements of $\bbeta$ will be estimated to be 0.
-In the $p \ll n$ setting, if $\lam = 0$, the lasso solution coincides with that of OLS.
-The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate negative log-likelihood function.
+At the other extreme, if $\lam = 0$ the lasso solution coincides with that of classical ordinary least squares regression.
+The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).
+
+The lasso penalty accomplishes two key things: (1) shrinkage, meaning that coefficients are encouraged to be smaller and less likely to be overestimated and (2) sparsity, meaning that many coefficients will be exactly zero and thus eliminated from the model.
+Both of these components are key to the lasso's ability to work with large, genome-scale data.
+The following sections discuss how to lasso can be modified to account for correlation among individuals.
+
+\pb{additional references?  Elements of statistical learning?}
 
@@
 These variance components and the corresponding weights are estimated once under the assumption of a null model.
-A lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
-LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}.
-This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix}.
-\anna{Should we still reference glmnet, ncvreg, ggmix here if we now include a software section?} We further discuss differences between these methods in Section \ref{sec:results}.
+The resulting optimization problem is thus reduced to that of the ordinary lasso and any software can be used to estimate $\bbeta$.
+LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}, and thus requires special software.
+Section~\ref{sec:software} discusses software packages for fitting these models in greater detail.
 
@@
 \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
+
+\pb{BELOW IS JUST DUMP FROM EARLIER}
+. This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package
+
+A lasso model can then be fit using standard software for
+
+A lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
+
+LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}.
+This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix}.
+We further discuss differences between these methods in Section \ref{sec:results}.
+\pb{END DUMP}
 

### 2021-03-13 main.tex
Edits to tuning parameter selection, software

@@
 
-Selection of the tuning parameter $\lambda$ plays an important role in all penalized regression methods, with many possible approaches.
-Here, for each model, we selected the $\lam$ value such that 50 SNPs were selected (recall that 50 is also the true number of causal SNPs).
-This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.
+Selection of the tuning parameter $\lambda$ plays an important role in all penalized regression methods.
+Many approaches have been suggested in the literature, and can be roughly divided into four categories: (1) selecting a predetermined number of features, (2) using information criteria such as the Bayesian Information Criterion (BIC), (3) cross-validation and other resampling/sample-splitting approaches, and (4) inference-based approaches such as choosing the maximum $\lam$ that controls the false discovery rate (FDR) at a certain threshold.
 
-Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons.
+These approaches have been extensively studied in the case of independent observations, but little work has been done on how they are affected by correlation among individuals.
+We offer a few comments here, but note that our theoretical understanding remains relatively poor and there is no consensus on the optimal way to select $\lam$ in this case \citep{wauthier2013comparative}.
+
+BIC has been used by several authors for variable selection in penalized models with correlated outcomes \citep{bondell2010joint, schelldorfer2011estimation}.
+BIC performs poorly in the high-dimensional setting, although modified high-dimensional variants have been proposed \citep{fan2013tuning,Chen2008a}.
+Cross-validation is also reasonable, with some caveats about the generalizability of prediction error \citep{Rabinowicz2020}.
+In the present context, this means that although cross-validation is a valid way to estimate prediction error of another sample from the same population, we cannot expect for example cross-validation of a GWAS from European samples to estimate prediction error accurately for calculating genetic risk in Chinese individuals.
+Selecting $\lam$ on the basis of FDR is obviously attractive, although there has been little work extending these approaches to correlated outcomes.
+Knockoff filters may be useful for this purpose \citep{Sesia2018}, although they have not been studied in the specific context of penalized LMMs.
+
+Choosing a fixed number of variants, on the other hand, is straightforward and easy to apply to any model regardless of whether observations are correlated or not.
+Its obvious drawback is that it requires an arbitrary decision regarding the number of variants to select, as opposed to the other, more data-driven methods.
+However, these other methods often result in substantially different $\lam$ values being chosen from data set to data set or method to method, making comparisons difficult.
+Fixing the number of variants selected provides an unambiguous way of comparing the accuracy of various methods.
+Since comparing these methods is the goal of the present paper, we use this approach in all of our simulation studies, fixing the number of selected variants at 50 (the true number of causal SNPs).
+
+%% Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons.
 Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods.
@@
 
+%% \anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
+In a real data scenario, the true number of causal SNPs would not be known a priori.
+As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
+Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm.
+This refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso.
+At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no consensus on the optimal way to select a penalty parameter for such algorithms.}
+
+%% \anna{Subsampling approaches, such as CV and stability selection are another common class of variable selection methods often preferred to information criterion-based approaches in practice.
+\citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
+\citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.
+A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}
+
 %------------------------------%
-\subsection{\anna{Software}}
+\subsection{Software}
 %------------------------------%
 \label{sec:software}
-\anna{Should this be here as a methods subsection or should it get its own section after discussion, before data availability?}
 
-All analyses and simulations were performed using \texttt{R} version 3.6.3.
+All analyses and simulations were performed using R version 3.6.3.
 The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}.
-As described previously, LMM-lasso-ggmix was implemented using \texttt{ggmix} \citep{bhatnagar2020simultaneous}.
-The LMM-lasso-Rakitsch method was implemented using the \texttt{R} package, \texttt{penalizedLMM}, available on the github repository \url{https://github.com/areisett/penalizedLMM}.
-Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.
-\texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
-
-\pb{BELOW IS JUST DUMP FROM EARLIER}
-. This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package
-
-A lasso model can then be fit using standard software for
-
-A lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.
-
-LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}.
-This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix}.
-We further discuss differences between these methods in Section \ref{sec:results}.
-\pb{END DUMP}
+As described previously, LMM-lasso-ggmix requires specialized software, but is implemented in the R package \texttt{ggmix} \citep{bhatnagar2020simultaneous}.
+Standard lasso software such as \texttt{glmnet} can be used to fit the LMM-lasso-Rakitsch method as well, although this requires a fair amount of preprocessing and setting up.
+To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files.
+pb{Is this sentence important?
+Why do we care what parameterization Rakitsch used?
+Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
+\pb{Mention imputation capabilities?}
 
@@
 
-\anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
-In a real data scenario, the true number of causal SNPs would not be known a priori.
-As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
-Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm.
-This refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso.
-At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no consensus on the optimal way to select a penalty parameter for such algorithms.}
-
-\anna{Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \citep{bondell2010joint, schelldorfer2011estimation}.
-However it has been shown that the Akaike information criterion (AIC) and BIC may perform poorly in the high-dimensional setting \citep{fan2013tuning}.
-The high-dimensional BIC introduced by \citet{fan2013tuning} penalizes model parameters more severely in an attempt to perform better in high dimensional settings.
-\citet{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.}
-
-\anna{Subsampling approaches, such as CV and stability selection are another common class of variable selection methods often preferred to information criterion-based approaches in practice.
-\citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
-\citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.
-A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}
-
 Finally, we note that this article has focused entirely on the lasso as a penalization scheme.

### 2021-03-13 main.tex
Some minor edits to 5

@@
 
-We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) \anna{a subset of} real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
+We generated three varieties of genotype matrices $\bG$, which were then standardized to yield the corresponding $\bX$ matrix: (1) simulated SNP data from admixed subpopulations \citep{Ochoa2016}, (2) simulated SNP data from independent subpopulations \citep{Ochoa2016}, and (3) a subset of real genome-wide SNP data from known Caucasian ($n = 417$), African American ($n = 134$), and Hispanic ($n = 37$) subpopulations \citep{larkin2015objectives}.
 We refer to these as 1D Linear Admixture, Independent Subpopulations, and Empirical, respectively.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\anna{Simulation results}} \label{sec:results}
+\section{Simulation results} \label{sec:results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\anna{Genome-wide case study}} \label{sec:gw_cs}
+\section{Genome-wide case study} \label{sec:gw_cs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To illustrate the statistical properties demonstrated by our simulations on a more realistic genome-wide scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} and a synthetic phenotype as a pseudo-real data case study.
+To illustrate the statistical properties demonstrated by our simulations on a more realistic genome-wide scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} and a synthetic phenotype as a semi-real data case study.
 We generate the synthetic phenotype as in \eqref{eqn:sim} using 100 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
-In order to compare methods fairly and completely, we consider results across the entirety of each methods' $\lambda$ path, indexed by the number of selected variables for interpretatiblity.
+In order to compare methods fairly and completely, we consider results across the entirety of each methods' $\lambda$ path, indexed by the number of selected variables for interpretability.
 
 %------------------------------%
-\subsection{\anna{LMM-lasso identifies causal SNPs with greater precision than lasso}}
+\subsection{LMM- and PC-lasso identify causal SNPs more accurately than lasso}
 %------------------------------%
+
+\pb{uh, what? This formula does NOT require sign to be correct.  What's right, the formula or the text?}
 

### 2021-03-14 main.tex
Edits to 5 and 6

@@
 
+\pb{How do you feel about combining Sections 5.1 + 5.2, and Figures 5 + 6?  Each section seems really short to me.}
+
 %------------------------------%
@@
 
-\pb{uh, what? This formula does NOT require sign to be correct.  What's right, the formula or the text?}
+\pb{I'm confused -- this formula does NOT require sign to be correct.  What's right, the formula or the text?}
 
@@
 \includegraphics[scale = 0.9]{figures/genomewide_prc_p1=100.png}
-\caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\caption{\pb{I feel like we should stop this at 200.
+The 100-400 range dominates this plot and what I get from this figure is: all the methods are equally bad.} Precision for each method and number of selected SNPs from the Empirical genome-wide analysis.
+Single replication case study with $\eta = 0.5, \xi = 0.8$.}
 \label{fig:gw_prc}
@@
 
-PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selection accuracy, particularly for the first $\approx$ 100 SNPs each method allows into the model.
-For the most part, PC-lasso selects SNPs outperforms LMM-lasso in this setting as well.
-All methods struggle as the model becomes less sparse and the number of selected SNPs approaches $n$.
+PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model.
+For the most part, PC-lasso outperforms LMM-lasso in this setting as well.
+None of the method are able to accurately select more than about 100 SNPs.
+However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \pb{XXX} SNPs selected were all false positives.
 
 %------------------------------%
-\subsection{\anna{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}}
+\subsection{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}
 %------------------------------%
@@
 This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method.
-Both PC-lasso and LMM-lasso produce more accurate estimates than the ordinary lasso for any number of selected SNPs.
+For any number of selected SNPs, however, both PC-lasso and LMM-lasso produce much more accurate estimates than the ordinary lasso.
 
@@
 %------------------------------%
-\subsection{\anna{PC-lasso and LMM-lasso reduce selection bias in genome-wide data}}
+\subsection{PC-lasso and LMM-lasso reduce selection bias in genome-wide data}
 %------------------------------%
@@
 
-Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected for a selected SNPs with the first principal component (PC1) of the Empirical data for each number of incorrectly selected SNPs.
-The corresponding shaded regions indicate the SE about the mean.
-\anna{Do I need to explain why we are using correlation with the first PC as a proxy for MAF difference or is this obvious?} The absolute correlation with PC1 is used as a measure of how population-informative the selected SNPs are given the complex Empirical data structure compared to that in Section~\ref{sec:sel_bias}.
+Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected with the first principal component (PC1).
+The first principal component does not capture the full complexity of the population structure, but it useful in terms of quantifying the extent to which a method is selecting SNPs that merely reflect the population structure and hence are at greater risk of being false positives due to environmental confounding.
+The shaded regions indicate the SE about the mean.
 
@@
 Likewise, PC-lasso and LMM-lasso both effectively control this type of selection bias.
-PC-lasso's effective control of bias in this setting is to be expected, given that we are assessing the population-informative nature of incorrectly selected SNPs via correlation with PC1, which is explicitly adjusted for in the PC-lasso method.
+PC-lasso's effective control of bias in this setting is to be expected, given that we are explicitly adjusting for PC1 in the analysis, but we see here that LMM-lasso accomplishes the same thing in a more indirect manner.
 
@@
 
-\anna{The results of a genome-wide analysis of Empirical data agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
-We note that because the Empirical data consists of three distinct racial groups, its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates predominantly coarse population structure.
-Given our results in Section~\ref{sec:sim-fine-coarse}, we would expect LMM-lasso to perform even better in the analysis of more finely structured genome-wide data.}
+\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.}
+The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
+However, PC-lasso performed somewhat better than LMM-lasso in this case study, in contrast to most of the simulation experiments.
+This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure.
+Thus, its structure is well-suited for PC-lasso.
+This finding suggests that while LMM-lasso may be more robust to complex structures, PC-lasso remains a strong competitor when the population structure is coarse.
 

### 2021-03-15 lmm-genetic-review.tex
Fixing missing \real def for "plain" article
adding renamed figures per journal guidelines

@@
 \usepackage{color}
-\usepackage{tikz}
-\usepackage{tikz-cd}
+%\usepackage{tikz}
+%\usepackage{tikz-cd}
 \usepackage{tkz-euclide}
 \usepackage{lscape}
+%%% UNCOMMENT THIS WHEN DONE EDITING TO PUT FIGS AT END
+%\usepackage{endfloat}
+\usepackage{epstopdf}
 
@@
 \newcommand{\sigmagg}{\sigma^2_g}
+\newcommand{\sigmass}{\sigma^2_s}
 \newcommand{\T}{\text{T}}
 \newcommand{\lam}{\lambda}
+\providecommand{\real}{\mathbb{R}}
 
-\title{A comparative analysis of penalized linear mixed models in structured genetic data}
+\title{Penalized linear mixed models for structured genetic data}
 \author{Anna Reisetter \qquad Patrick Breheny\\

### 2021-03-15 main.tex
adding renamed figures per journal guidelines
Adding ESL ref

@@
 At the other extreme, if $\lam = 0$ the lasso solution coincides with that of classical ordinary least squares regression.
-The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox regression \citep{tibshirani1997lasso} and logistic regression \citep{shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).
+The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox and logistic regression \citep{tibshirani1997lasso, friedman2001elements, shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).
 
@@
 The following sections discuss how to lasso can be modified to account for correlation among individuals.
-
-\pb{additional references?  Elements of statistical learning?}
 
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/beta_mse.png}
+    \includegraphics[width = \textwidth]{figures/figure_01.png}
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.
@@
 \centering
-\includegraphics[scale = 0.9]{figures/pop_inf_snps_box.png}
+\includegraphics[scale = 0.9]{figures/figure_02.png}
 \caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$, 1000 replications.}
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
+    \includegraphics[scale = 0.9]{figures/figure_03.png}
     \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$, 1000 replications.}
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/eta_beta_hat.png}
+    \includegraphics[width = \textwidth]{figures/figure_04.png}
      \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, 1000 replications.}
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_prc_p1=100.png}
+\includegraphics[scale = 0.9]{figures/figure_05.png}
 \caption{\pb{I feel like we should stop this at 200.
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_mse_p1=100.png}
+\includegraphics[scale = 0.9]{figures/figure_06.png}
 \caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
@@
 \centering
-\includegraphics[scale = 0.9]{figures/genomewide_pc1_corr_p1=100.png}
+\includegraphics[scale = 0.9]{figures/figure_07.png}
 \caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the Empirical genome-wide data for each method and number of selected SNPs.

### 2021-03-15 main.tex
Sec 2.3, 3.8, 5 - in progress

@@
 \pb{What is the significance of this Mandt paper?
-Isn't there a huge literature on GLMMs?
-\citet{mandt2017sparse} generalize the LMM modeling paradigm to binary phenotypes in the interest of classification.
+Isn't there a huge literature on GLMMs?}\anna{ \citet{mandt2017sparse} generalize the \textbf{high-dimensional, penalized} LMM modeling paradigm to binary phenotypes in the interest of classification.
 They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference.
@@
 To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files.
-pb{Is this sentence important?
-Why do we care what parameterization Rakitsch used?
-Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
-\pb{Mention imputation capabilities?}
+\anna{\texttt{penalizedLMM} includes preprocessing functions for \anna{SNP and sample-level filtering and} genotype imputation using \textbf{imputation rules - how to describe this} where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
+\anna{refine filtering, imputation functions}
 
@@
 \pb{How do you feel about combining Sections 5.1 + 5.2, and Figures 5 + 6?  Each section seems really short to me.}
+\anna{Fine with combining Sections 5.1 + 5.2 - I'm not sure what you mean by combine Figures 5 + 6. Keep one and not the other? Combine into a panel with parts (a) and (b)?}
 
@@
 
-\pb{I'm confused -- this formula does NOT require sign to be correct.  What's right, the formula or the text?}
-
 Selection accuracy was evaluated using selection precision.
-All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.
-Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
+\anna{All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.} Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
 \begin{figure}[H]
@@
 \includegraphics[scale = 0.9]{figures/figure_05.png}
-\caption{\pb{I feel like we should stop this at 200.
-The 100-400 range dominates this plot and what I get from this figure is: all the methods are equally bad.} Precision for each method and number of selected SNPs from the Empirical genome-wide analysis.
-Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
 \label{fig:gw_prc}
@@
 However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \pb{XXX} SNPs selected were all false positives.
-
-%------------------------------%
-\subsection{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}
-%------------------------------%
 
@@
 
-\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.}
+\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.} \anna{agreed}
 The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.

### 2021-03-17 main.tex
Section 5

@@
 
-\pb{How do you feel about combining Sections 5.1 + 5.2, and Figures 5 + 6?  Each section seems really short to me.}
-\anna{Fine with combining Sections 5.1 + 5.2 - I'm not sure what you mean by combine Figures 5 + 6. Keep one and not the other? Combine into a panel with parts (a) and (b)?}
-
 %------------------------------%
@@
 
-Selection accuracy was evaluated using selection precision.
+Selection accuracy was evaluated using precision.
 \anna{All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.} Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
-\begin{figure}[H]
-\centering
-\includegraphics[scale = 0.9]{figures/figure_05.png}
-\caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-\label{fig:gw_prc}
-\end{figure}
-
 PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model.
@@
 None of the method are able to accurately select more than about 100 SNPs.
-However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \pb{XXX} SNPs selected were all false positives.
+However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \anna{38} SNPs selected were all false positives.
 
-Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}. Figure~\ref{fig:gw_mse} displays this value for each method and number of selected SNPs.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale = 0.9]{figures/figure_06.png}
-\caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-\label{fig:gw_mse}
-\end{figure}
-
+Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}.
+Figure~\ref{fig:gw_mse} displays this value for each method and number of selected SNPs.
 In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso.
@@
 
-%PC-lasso and LMM-lasso coincide in MSE for the first few selected variables, for the most part, however, PC-lasso outperforms LMM-lasso.
-
-% This is an interesting result in contrast to that observed in the third panel of Figure~\ref{fig:mse}, where we analyzed a subset of the Empirical data using 50 causal SNPs and LMM-lasso slightly outperformed PC-lasso.
-It seems that in the genome-wide setting the coarse, subpopulation-based structure dominates the subtle relatedness present in the sample
+\begin{figure}[H]
+     \centering
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{figures/figure_05.png}
+         \caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+        \label{fig:gw_prc}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.475\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{figures/figure_06.png}
+         \caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+        \label{fig:gw_mse}
+     \end{subfigure}
+\end{figure}
 

### 2021-03-17 main.tex
modifying figure 5
Edits to intro, software section, genome-wide section. replacing eps with pdfs

@@
 
-\pb{Read this over and let me know what you think}
+\anna{Overall I think this looks good and definitely more friendly. Fixed some minor typos, repetitive words (e.g. 'however' twice in a sentence)}
 
@@
 Accurately identifying and estimating the effect sizes of these variants or single nucleotide polymorphisms (SNPs) is of interest for a variety of scientific applications including the prediction of an individual's risk of disease, understanding the genetic architecture underlying a complex trait, and drug discovery.
-Genetic association is typically analyzing by testing individual variants one at a time.
+Genetic association is typically assessed by testing individual variants one at a time.
 While straightforward, there are inherent limitations to analyzing each variant in isolation including diminished power, an inability to examine epistatic interactions, and bias due to the presence of correlation between variants.
@@
 An alternative solution is the use of multivariate methods such as multiple linear regression, which adjusts for and estimates the effects of many variants simultaneously.
-However, classical regression relies on data with more many more observations than variants.
+However, classical regression relies on data with many more observations than variants.
 This is typically infeasible in genome-wide studies where the variants to be analyzed number in the hundreds of thousands or even millions.
@@
 Lasso models, however, rely on an assumption that subjects are independent of each other.
-In the presence of population structure or cryptic relatedness, however, this assumption is violated: traits are more likely to be similar within a family or ethnic group.
+In the presence of population structure or cryptic relatedness this assumption is violated: traits are more likely to be similar within a family or ethnic group.
 The presence of structure can adversely affect analyses in several ways, but the most important concern is that it is likely to introduce false positives.
@@
 In Section \ref{sec:results}, we illustrate these concepts via simulation studies and identify scenarios where particular methods are most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
-Finally, in Section~\ref{sec:gw_cs} we illustrate the use of these methods in a case study involving genome-wide data involving a sample of diverse racial background.
+Finally, in Section~\ref{sec:gw_cs} we illustrate the use of these methods in a case study of genome-wide data involving a sample of diverse racial background.
 
@@
 
-\pb{Please read the following paragraph; it has been modified extensively.}
-
 In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
@@
 The situation is certainly not impossible, and many approaches to fitting generalized linear mixed models, or GLMMs, have been proposed in the classical, low-dimensional setting.
-However, very little work has been done to extend this to penalized, high-dimensional models and to the best of our knowledge there is no existing software for doing so at this time.
-We hope that this situation changes in the near future.
+\anna{However, very little work has been done to extend this to penalized, high-dimensional models.
+As far we are aware, only \citet{mandt2017sparse} generalize the high-dimensional penalized LMM modeling paradigm to binary phenotypes.
+They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference.
+However, to the best of our knowledge there is no available software to implement this or any other methods for high-dimensional, penalized methods at this time.} We hope that this situation changes in the near future.
 
 \pb{What is the significance of this Mandt paper?
-Isn't there a huge literature on GLMMs?}\anna{ \citet{mandt2017sparse} generalize the \textbf{high-dimensional, penalized} LMM modeling paradigm to binary phenotypes in the interest of classification.
-They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference.
-}
+Isn't there a huge literature on GLMMs?}\anna{Mandt is specifically for high-dimensional, penalized LMMs that use the data to infer structure of the errors - not just a regular GLMM.
+Includes some of the Rakitsch 2013 authors, extends LMM-lasso to binary outcomes in one way.}
 
@@
 To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files.
-\anna{\texttt{penalizedLMM} includes preprocessing functions for \anna{SNP and sample-level filtering and} genotype imputation using \textbf{imputation rules - how to describe this} where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
-\anna{refine filtering, imputation functions}
+\anna{\texttt{penalizedLMM} includes preprocessing functions for genotype imputation using \textbf{imputation rules - how to describe this} where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
 
@@
      \centering
-     \begin{subfigure}[b]{0.475\textwidth}
+     \begin{subfigure}[b]{0.4\textwidth}
          \centering
-         \includegraphics[width=\textwidth]{figures/figure_05.png}
+         \includegraphics[width=\textwidth]{figures/figure_05a.pdf}
          \caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
@@
     \hfill
-     \begin{subfigure}[b]{0.475\textwidth}
+     \begin{subfigure}[b]{0.4\textwidth}
          \centering
-         \includegraphics[width=\textwidth]{figures/figure_06.png}
+         \includegraphics[width=\textwidth]{figures/figure_05b.pdf}
          \caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
         \label{fig:gw_mse}
+     \end{subfigure}
+     \begin{subfigure}[b]{0.15\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{figures/figure_05c.pdf}
      \end{subfigure}

### 2021-03-18 main.tex
Notes from meeting

@@
 This procedure does not directly generalize beyond the linear regression case.
+
+\anna{other approaches are possible, for example \citet{mandt2017sparse} generalize the penalized lmm modeling paradigm to binary phenotypes by thresholding the outcome variable through a probit likelihood.
+However, there is no work that we are aware of that explicitly addresses the use of such models for GWAS.}
+
+\anna{idea: not impossible, important,b ut not enough has been done for us to include it in the review}
+
 The situation is certainly not impossible, and many approaches to fitting generalized linear mixed models, or GLMMs, have been proposed in the classical, low-dimensional setting.
@@
 However, to the best of our knowledge there is no available software to implement this or any other methods for high-dimensional, penalized methods at this time.} We hope that this situation changes in the near future.
-
-\pb{What is the significance of this Mandt paper?
-Isn't there a huge literature on GLMMs?}\anna{Mandt is specifically for high-dimensional, penalized LMMs that use the data to infer structure of the errors - not just a regular GLMM.
-Includes some of the Rakitsch 2013 authors, extends LMM-lasso to binary outcomes in one way.}
 
@@
 To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files.
-\anna{\texttt{penalizedLMM} includes preprocessing functions for genotype imputation using \textbf{imputation rules - how to describe this} where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
+\anna{\texttt{penalizedLMM} includes preprocessing functions for genotype imputation using functions for imputation based on linkage disequilibrium with tag SNPs where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
+
+\anna{functions for imputation based on linkage disequilibrium with tag SNPs.}
 
@@
 
+\anna{Make sure this gets an overall figure number}
 \begin{figure}[H]

### 2021-03-18 submissions/response/response.tex
Edits to binary paragraph, lasso section, and software
Finishing point-by-point response

@@
 
-Thank you very much for your detailed reading of the paper and your insightful suggestions.
+Thank you very much for your careful reading of the paper and your insightful comments.
 You raise a number of good points and we have modified the manuscript considerably based upon your feedback.
@@
 We have tried to improve upon this aspect of the paper by rewriting the introduction with the application, rather than the statistical methodology, in mind.
-Furthermore, MENTION DISCUSSION?
-MENTION APPLICATION / SECTION 5?
+In addition, we added a new section (more details below) illustrating the method on a case study involving a GWAS with a mixed population (white/black/hispanic).
 
@@
 
-  Response
+  This is an excellent point.  We have now added a new section, 3.1, giving a brief overview of the lasso, as well as references to learn more about it.
 
@@
 
-  Response
+  We have added a new section (Section 5) that contains a ``case study'' GWAS involving real data.
+For a variety of reasons, we chose to use a synthetic phenotype, but all of the genetic data is real.
+This allows us to inspect the accuracy of these methods, as the true causal SNPs are known, and did indeed reveal some additional insights that the smaller-scale simulations did not.
 
@@
 
-  Response
+  We have added a new section (Section 3.8) that specifically addresses the availability of software implementing these methods.
+As far as whether the software is ``simple'', we tried to make our R package, \texttt{penalizedLMM}, as user-friendly as possible, with convenient functions for importing data from PLINK files and preprocessing (e.g., carrying out genomic imputation prior to fitting).
+We hope that this makes these models relatively easy to fit and use for practitioners.
 
@@
 
-  Response
+  This is an excellent point.  We have expanded our discussion of this issue in section 2.3, which now contains the following paragraph:
+
+  \begin{quote}
+    aaa
+  \end{quote}
 
@@
 
-  Response.
+  It's possible that we don't entirely understand what you mean here, as this was the point of all the simulation studies: to better understand the statistical properties of these methods.
+Perhaps you mean the theoretical, mathematical properties?
+If so, unfortunately, not much is known.
+We have expanded Section 3.7.4, which discusses tuning parameter selection (this was suggested by another reviewer).
+In this section, we do discuss some of the work that has been done, including work on inference and estimation of false discovery rates, and provide references.
+We also ``note that our theoretical understanding remains relatively poor and there is no consensus on
+the optimal way to select $\lambda$ in this case (Wauthier et al., 2013).''  We hope that these modifications help to address this issue.
 
@@
 
-Thank you very much for your detailed reading of the paper and your insightful questions.
-You raise a number of good points and we have modified the manuscript in several places based upon your feedback.
+Thank you very much for your detailed reading of the paper and your insightful comments.
+You raise a number of excellent points and we have modified the manuscript in several places based upon your feedback.
 In our opinion, these revisions have nicely improved upon the original submission.
@@
 \item \emph{This is a very insightful paper as to the performance of lasso LMM methods in GWAS analyses and provides nice intuition using theoretical derivations on the interplay between population structure and environmental heterogeneity.
-The paper could highly benefit from using more realistic simulation scenarios reflecting standard GWAS (increase the number of SNPs in model which is more in the 100,000s in current GWASes) as well as provide some assessment of how robust the results would be to different lasso penalty parameters (as in practice we would not have a priori an optimal value).}
+The paper could highly benefit from using more realistic simulation scenarios reflecting standard GWAS (increase the number of SNPs in model which is more in the 100,000s in current GWASes) \ldots}
 
-  Response
+  This is an excellent point, and shares some concerns with a point made by reviewer \#1, who wished to see a real data example.
+We attempted to address both of these concerns with a ``case study'' (Section 5), which uses genotype data from a real GWAS (with 526,266 SNPs) but with a synthetic phenotype, which allows us to assess the accuracy of the various methods in terms of identifying the causal SNPs.
+Indeed, this new setting did reveal some additional insights that the smaller-scale simulations did not.
+
+  \item \emph{\ldots as well as provide some assessment of how robust the results would be to different lasso penalty parameters (as in practice we would not have a priori an optimal value).}
+
+  Another excellent point.
+We have expanded Section 3.7.4 to include a much more thorough discussion of the problem of tuning parameter selection.
+In addition, all of the results in Section 5 (the new GWAS case study) are presented over the entire path, so that readers can see how the statistical properties of the estimates change as $\lambda$ changes.
 
@@
 1 and Table 1, it is not clear how many replicates are used to estimate these quantities (e.g.
-all p=1000 SNPs used in simulations or only the 50 SNPs selected by lasso?).
-For Figure 1, it would be helpful to have error bars included in bar plots.
-Also, there is a typo in Eqn 15.}
+all p=1000 SNPs used in simulations or only the 50 SNPs selected by lasso?).}
 
-  Response
+  Thank you for pointing this out.
+The number of replications is now explicitly stated in all of the figure and table legends.
+Also, we have added this sentence at the beginning of Section 4.1, after defining MSE: ``In particular, note that this calculation includes all $p$ coefficient values, regardless of whether or not they were selected by the model.''
+
+\item \emph{For Figure 1, it would be helpful to have error bars included in bar plots.}
+
+  Figure 1 now included error bars (as does the new Figure 6).
+
+\item \emph{There is a typo in Eqn 15.}
+
+  Thank you very much for pointing this out -- this typo is now fixed.
 

### 2021-03-18 main.tex
Finishing point-by-point response

@@
 At the other extreme, if $\lam = 0$ the lasso solution coincides with that of classical ordinary least squares regression.
-The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox and logistic regression \citep{tibshirani1997lasso, friedman2001elements, shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).
+The $\ell_1$ penalty of the lasso can be applied to nonlinear models such as Cox and logistic regression \citep{tibshirani1997lasso, shevade2003simple} by replacing the residual sum of squares with the appropriate loss function (i.e., the negative log-likelihood).
 
@@
 Both of these components are key to the lasso's ability to work with large, genome-scale data.
+For more information on the lasso, the textbook by \citet{Hastie2009} is a useful reference.
 The following sections discuss how to lasso can be modified to account for correlation among individuals.
@@
 Estimation accuracy for lasso, PC-lasso, and LMM-lasso was evaluated based on mean squared error (MSE) of the coefficients, MSE = $\mathbb{E}||\bbeta - \bbetaHat(\lambda)||^2_2$, where $\bbetaHat(\lambda)$ is the vector of coefficients estimated by the lasso with a penalty of $\lambda$, and $\bbeta$ the true coefficient values.
+In particular, note that this calculation includes all $p$ coefficient values, regardless of whether or not they were selected by the model.
 Figure \ref{fig:mse} presents these results for all three data types in a 1:1 SNR setting, where $\eta = 0.5$.
@@
 However, in the absence of such effects, PC-lasso may hinder estimation accuracy.
-Additionally, we find that LMM-lasso consistently outperforms PC-lasso, especially when the population structure is more complex.
+Additionally, we find that LMM-lasso generally outperforms PC-lasso, especially when the population structure is more complex.
 These results indicate that LMM-lasso is a more robust method in practice when the nature of the population structure is unknown.
@@
 The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.
-However, PC-lasso performed somewhat better than LMM-lasso in this case study, in contrast to most of the simulation experiments.
+However, PC-lasso outperformed LMM-lasso in this case study, in contrast to most of the simulation experiments.
 This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure.

### 2021-03-19 main.tex
Intro, software section edits

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\anna{Overall I think this looks good and definitely more friendly. Fixed some minor typos, repetitive words (e.g. 'however' twice in a sentence)}
 
@@
 This procedure does not directly generalize beyond the linear regression case.
-
-\anna{other approaches are possible, for example \citet{mandt2017sparse} generalize the penalized lmm modeling paradigm to binary phenotypes by thresholding the outcome variable through a probit likelihood.
-However, there is no work that we are aware of that explicitly addresses the use of such models for GWAS.}
-
-\anna{idea: not impossible, important,b ut not enough has been done for us to include it in the review}
-
-The situation is certainly not impossible, and many approaches to fitting generalized linear mixed models, or GLMMs, have been proposed in the classical, low-dimensional setting.
-\anna{However, very little work has been done to extend this to penalized, high-dimensional models.
-As far we are aware, only \citet{mandt2017sparse} generalize the high-dimensional penalized LMM modeling paradigm to binary phenotypes.
-They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference.
-However, to the best of our knowledge there is no available software to implement this or any other methods for high-dimensional, penalized methods at this time.} We hope that this situation changes in the near future.
+Other approaches are possible, for example \citet{mandt2017sparse} generalize the penalized LMM modeling paradigm to binary phenotypes by thresholding the outcome variable through a probit likelihood.
+However, there is no work that we are aware of that explicitly addresses the use of such models for GWAS.
+We hope that this situation changes in the near future.
 
@@
 To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files.
-\anna{\texttt{penalizedLMM} includes preprocessing functions for genotype imputation using functions for imputation based on linkage disequilibrium with tag SNPs where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
-
-\anna{functions for imputation based on linkage disequilibrium with tag SNPs.}
+\texttt{penalizedLMM} includes preprocessing functions for genotype imputation based on linkage disequilibrium with tag SNPs where possible, or else the Hardy-Weinberg Equilibrium values.
+\texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.
 

### 2021-03-19 main.tex
5.1 (figure still to to), 6

@@
 Selection accuracy was evaluated using precision.
-\anna{All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.} Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
+All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.
+\anna{Figure~\ref{fig:gw_prc}} displays each method's precision for the corresponding number of selected SNPs.
 PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model.
@@
 None of the method are able to accurately select more than about 100 SNPs.
-However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \anna{38} SNPs selected were all false positives.
+However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.
 
 Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}.
-Figure~\ref{fig:gw_mse} displays this value for each method and number of selected SNPs.
+\anna{Figure~\ref{fig:gw_mse}} displays this value for each method and number of selected SNPs.
 In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso.
@@
 
-\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.} \anna{agreed}
 The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso.

### 2021-03-19 submissions/response/response.tex
fixing graphics path for submissions/genetic-epi.tex, adding binary paragraph to response

@@
   \begin{quote}
-    aaa
+    In many genetic association studies the phenotype of interest is binary, for example the presence or absence of a heritable trait.
+Unfortunately, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward.
+As we explain in Section 3, LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation.
+This procedure does not directly generalize beyond the linear regression case.
+Other approaches are possible, for example Mandt et al.
+(2017) generalize the penalized LMM modeling paradigm to binary phenotypes by thresholding the outcome variable through a probit likelihood.
+However, there is no work that we are aware of that explicitly addresses the use of such models for GWAS.
+We hope that this situation changes in the near future.
   \end{quote}

### 2021-03-19 main.tex
updated 2-panel figure 5. updating figure refs, additional figure numbers

@@
 \centering
-\includegraphics{figures/dag1.eps}
+\includegraphics{figures/dag1.pdf}
 \label{diag:ps_env}
@@
 \centering
-\includegraphics{figures/dag2.eps}
+\includegraphics{figures/dag2.pdf}
 \label{diag:ps_env_block}
@@
 \centering
-\includegraphics{figures/dag3.eps}
+\includegraphics{figures/dag3.pdf}
 \label{diag:ps_env_block2}
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/figure_01.png}
+    \includegraphics[width = \textwidth]{figures/figure_01.pdf}
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.
@@
 \centering
-\includegraphics[scale = 0.9]{figures/figure_02.png}
+\includegraphics[scale = 0.9]{figures/figure_02.pdf}
 \caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$, 1000 replications.}
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/figure_03.png}
+    \includegraphics[scale = 0.9]{figures/figure_03.pdf}
     \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$, 1000 replications.}
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/figure_04.png}
+    \includegraphics[width = \textwidth]{figures/figure_04.pdf}
      \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, 1000 replications.}
@@
 All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.
-\anna{Figure~\ref{fig:gw_prc}} displays each method's precision for the corresponding number of selected SNPs.
+Figure~\ref{fig:gw_prc_mse}(a) displays each method's precision for the corresponding number of selected SNPs.
 PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model.
@@
 None of the method are able to accurately select more than about 100 SNPs.
-However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.
+However, because the ordinary lasso does not account for environmental confounding, it cannot reliably select \emph{any}: the first 38 SNPs selected were all false positives.
 
-Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}.
-\anna{Figure~\ref{fig:gw_mse}} displays this value for each method and number of selected SNPs.
+In addition to selection, estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}.
+Figure~\ref{fig:gw_prc_mse}(b) displays this value for each method and number of selected SNPs.
 In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso.
@@
 
-\anna{Make sure this gets an overall figure number}
 \begin{figure}[H]
 \centering
-     \begin{subfigure}[b]{0.4\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{figures/figure_05a.pdf}
-         \caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-        \label{fig:gw_prc}
-     \end{subfigure}
-    \hfill
-     \begin{subfigure}[b]{0.4\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{figures/figure_05b.pdf}
-         \caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-        \label{fig:gw_mse}
-     \end{subfigure}
-     \begin{subfigure}[b]{0.15\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{figures/figure_05c.pdf}
-     \end{subfigure}
+\includegraphics[scale = 0.9]{figures/figure_05.pdf}
+\caption{Precision and MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_prc_mse}
 \end{figure}
@@
 \centering
-\includegraphics[scale = 0.9]{figures/figure_07.png}
+\includegraphics[scale = 0.9]{figures/figure_06.pdf}
 \caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the Empirical genome-wide data for each method and number of selected SNPs.

### 2021-03-19 main.tex
Edits to Section 5

--- 2021-03-19 main.tex
@@
 
-To illustrate the statistical properties demonstrated by our simulations on a more realistic genome-wide scale, we present an analysis of the full Empirical genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives} and a synthetic phenotype as a semi-real data case study.
-We generate the synthetic phenotype as in \eqref{eqn:sim} using 100 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
-In order to compare methods fairly and completely, we consider results across the entirety of each methods' $\lambda$ path, indexed by the number of selected variables for interpretability.
+To determine whether the statistical properties demonstrated by our simulations hold in for a genome-wide analysis, we analyzed the full genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives}; a subset of this data was included in the earlier simulations, labeled ``Empirical''.
+A synthetic phenotype was generated as in \eqref{eqn:sim} using 100 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).
+In order to compare methods fairly and completely, we consider results across the entirety of each method's path.
 
@@
 
-Selection accuracy was evaluated using precision.
-All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.
-\anna{Figure~\ref{fig:gw_prc}} displays each method's precision for the corresponding number of selected SNPs.
-PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model.
+We begin by comparing the accuracy of ordinary lasso, LMM-lasso, and PC-lasso in terms of their ability to identify causal variants as well as their ability to estimate the effects of all SNPs.
+
+To quantify selection accuracy, we calculated the precision according to $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but the direction of effect was estimated correctly (in our example, all causal $\beta$ values were positive).
+\anna{Figure~\ref{fig:gw_prc}} displays each method's precision ($1 -$ the false discovery rate) for the corresponding number of selected SNPs.
+PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly over the first $\approx$ 100 SNPs allowed into the model.
 For the most part, PC-lasso outperforms LMM-lasso in this setting as well.
-None of the method are able to accurately select more than about 100 SNPs.
+None of the methods are able to accurately select more than a few causal SNPs with this sample size.
 However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.
 
-Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}.
-\anna{Figure~\ref{fig:gw_mse}} displays this value for each method and number of selected SNPs.
+Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}, and is presented in \anna{Figure~\ref{fig:gw_mse}}.
 In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso.
@@
 \anna{Make sure this gets an overall figure number}
+\pb{I still don't see an overall figure number?}
+
 \begin{figure}[H]

### 2021-03-19 main.tex

@@
 To quantify selection accuracy, we calculated the precision according to $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but the direction of effect was estimated correctly (in our example, all causal $\beta$ values were positive).
-\anna{Figure~\ref{fig:gw_prc}} displays each method's precision ($1 -$ the false discovery rate) for the corresponding number of selected SNPs.
+Figure~\ref{fig:gw_prc_mse}(a) displays each method's precision ($1 -$ the false discovery rate) for the corresponding number of selected SNPs.
 PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly over the first $\approx$ 100 SNPs allowed into the model.
@@
 
-Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}, and is presented in \anna{Figure~\ref{fig:gw_mse}}.
+Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}, and is presented in Figure~\ref{fig:gw_prc_mse}(b).
 In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso.
@@
 
-\anna{Make sure this gets an overall figure number}
-\pb{I still don't see an overall figure number?}
-
 \begin{figure}[H]
 \centering
-     \begin{subfigure}[b]{0.4\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{figures/figure_05a.pdf}
-         \caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-        \label{fig:gw_prc}
-     \end{subfigure}
-    \hfill
-     \begin{subfigure}[b]{0.4\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{figures/figure_05b.pdf}
-         \caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
-        \label{fig:gw_mse}
-     \end{subfigure}
-     \begin{subfigure}[b]{0.15\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{figures/figure_05c.pdf}
-     \end{subfigure}
+\includegraphics[scale = 0.9]{figures/figure_05.pdf}
+\caption{Precision and MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
+\label{fig:gw_prc_mse}
 \end{figure}

### 2021-03-19 main.tex
fixing typo in 5, using enfloat

@@
 None of the methods are able to accurately select more than a few causal SNPs with this sample size.
-However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.
+However, because the ordinary lasso does not account for environmental confounding, it cannot reliably select \emph{any} SNPs: the first 38 SNPs selected were all false positives.
 

### 2021-03-19 submissions/genepi-cover-1/cover.tex
Cover letter for resub

\documentclass{uiletter}

\begin{document}

\StartLetter{}{Dear editors of {\em Genetic Epidemiology},}

We are submitting the manuscript ``Penalized linear mixed models for structured genetic data'' for your consideration as a Review Article for publication in {\em Genetic Epidemiology}.

We feel that this is a review that the readers of {\em Genetic Epidemiology} will find important and timely.  Although one-SNP-a-time analyses are still the most common way of carrying out GWAS, genome-wide penalized regression methods such as the lasso are an increasingly common alternative.  However, little attention has been paid to correcting for population structure in such models.  This review covers this issue in detail, and describes the use of linear mixed models for penalized regression, a topic that has received little attention and been primarily confined to the plant and animal breeding literature.  Our review introduces and compares these methods, describing in detail how they work to reduce the potentially confounding influence of environmental factors in genetic association studies.

This work has not been previously published, is not under review elsewhere, and all authors concur with this submission.  Thank you very much for your consideration.

\EndLetter

\end{document}

### 2021-03-19 submissions/genepi-cover-2/cover.tex
Cover letter for resub

\documentclass{uiletter}

\begin{document}

\StartLetter{}{Dear editors of {\em Genetic Epidemiology},}

We are resubmitting the manuscript ``Penalized linear mixed models for structured genetic data'' for your consideration as a Review Article for publication in {\em Genetic Epidemiology}.

Our revised manuscript addresses all of the issues raised by both reviewers.  In particular, we have added a section including a ``case study'' analysis applying the methods discussed in the paper to actual GWAS data to illustrate the performance of these method at a genome-wide scale.  Our point-by-point response to reviewer comments is included along with the revised manuscript.

Thank you very much for the opportunity to revise and resubmit our manuscript.  We feel that several improvements have been made to the original version and hope that you agree.

\EndLetter

\end{document}

### 2021-03-20 submissions/resubmission/make.sh
creating new directory with script to pull in necessary files and modify paths for easier compilation on journal website

@@ -1,0 +1,11 @@
+#!/bin/bash
+cp ../../main.tex main.tex
+cp ../../supplement.tex supplement.tex 
+cp ../../ims.bst ims.bst
+cp ../../references.bib references.bib
+cp ../../figures/*.eps figures/
+cp ../genetic_epi.tex genetic_epi.tex
+sed -i 's#\\graphicspath{{../}{../logs/}}##g' genetic_epi.tex 
+sed -i 's#../##g' genetic_epi.tex 
+sed -i 's#.pdf}#.eps}#g' main.tex 
+sed -i 's#.pdf}#.eps}#g' supplement.tex 

### 2021-05-15 main.tex
removing commented sections in main
replacing pdf images with pngs for thesis pdf size
Fixing duplicate sec:results refs -> changing to sec:results_paper1 for thesis doc

@@
 We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
-In Section \ref{sec:results}, we illustrate these concepts via simulation studies and identify scenarios where particular methods are most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
+In Section \ref{sec:results_paper1}, we illustrate these concepts via simulation studies and identify scenarios where particular methods are most and least effective in accurately estimating the effect sizes of observed SNPs in the presence of unobserved confounding.
 Finally, in Section~\ref{sec:gw_cs} we illustrate the use of these methods in a case study of genome-wide data involving a sample of diverse racial background.
@@
 Nevertheless, a common practice is to include the ten largest principal components in the analysis \citep{zhao2018practical}.
-This is the approach we take when implementing PC-lasso in Section \ref{sec:results}.
+This is the approach we take when implementing PC-lasso in Section \ref{sec:results_paper1}.
 
@@
 In simulation settings, however, this ratio is a useful metric for quantifying the degree to which $\bZ \bgamma$ reflects population structure and has the potential to bias observed associations.
-We use CP to help interpret the simulation results in Section \ref{sec:results}.
+We use CP to help interpret the simulation results in Section \ref{sec:results_paper1}.
 %In high dimensions, we use a low-rank approximation of $(\bX^\T \bX)^{-1}$ to improve stability.
@@
 In our experiments, however, neither the heterogeneity nor the concordance of the environmental effects had any meaningful impact on the performance of the methods under consideration.
-Thus, throughout Section~\ref{sec:results}, we present results only for discordant, dichotomous environmental effects, as the other results are very similar.
+Thus, throughout Section~\ref{sec:results_paper1}, we present results only for discordant, dichotomous environmental effects, as the other results are very similar.
 
@@
 
-%% Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons.
-Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods.
-\anna{ I want to cut the following sentence but I think we may need some other reason why we didn't do CV...} Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}.
-\anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}
-
-%% \anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods.
-In a real data scenario, the true number of causal SNPs would not be known a priori.
-As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion.
-Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm.
-This refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso.
-At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no consensus on the optimal way to select a penalty parameter for such algorithms.}
-
-%% \anna{Subsampling approaches, such as CV and stability selection are another common class of variable selection methods often preferred to information criterion-based approaches in practice.
-\citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}.
-\citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection.
-A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}
-
 %------------------------------%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulation results} \label{sec:results}
+\section{Simulation results} \label{sec:results_paper1}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/figure_01.pdf}
+    \includegraphics[width = \textwidth]{figures/figure_01.png}
     \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with 4 subpopulations.
@@
 \centering
-\includegraphics[scale = 0.9]{figures/figure_02.pdf}
+\includegraphics[scale = 0.9]{figures/figure_02.png}
 \caption{Absolute allele frequency differences of selected, non-causal SNPs across methods, in the presence and absence of environmental confounding for two Empirical subpopulations, $\eta = 0.5$, 1000 replications.}
@@
     \centering
-    \includegraphics[scale = 0.9]{figures/figure_03.pdf}
+    \includegraphics[scale = 0.9]{figures/figure_03.png}
     \caption{Relative performance of LMM-lasso, PC-lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with Independent Subpopulations, $\eta = 0.5$, 1000 replications.}
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/figure_04.pdf}
+    \includegraphics[width = \textwidth]{figures/figure_04.png}
      \caption{Estimation accuracy and computing time of LMM-lasso-Rakitsch and LMM-lasso-ggmix for $\bbeta$ and $\eta$ for varying $\eta$ levels and sample sizes with Independent Subpopulations data, coarse subpopulation structure, $\xi = 0.8$, 1000 replications.}
@@
 \centering
-\includegraphics[scale = 0.9]{figures/figure_05.pdf}
+\includegraphics[scale = 0.9]{figures/figure_05.png}
 \caption{Precision and MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
@@
 \centering
-\includegraphics[scale = 0.9]{figures/figure_06.pdf}
+\includegraphics[scale = 0.9]{figures/figure_06.png}
 \caption{Absolute correlation of selected, non-causal SNPs with the first principal component (PC1) of the Empirical genome-wide data for each method and number of selected SNPs.

### 2021-05-15 main.tex
Modifying background, methods labels so specific to paper 1 for thesis doc

@@
 The remainder of this paper is organized as follows.
-In Section \ref{sec:background} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this work.
+In Section \ref{sec:background_paper1} we provide a summary of the causes and consequences of structure in the genetic data of seemingly unrelated individuals and formally define terminology used throughout this work.
 We emphasize that confounding due to population structure, frequently cited as a driver of spurious associations, is a function of environmental heterogeneity, where genetic data serves as a proxy for differential environmental exposures \citep{Sillanpaeae2011, sul2018population, vilhjalmsson2012nature, barton2019population}.
@@
 In addition, we briefly review historical methods of correcting for population stratification and relatedness.
-In Section \ref{sec:methods} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness: PCA adjusted lasso penalized regression (PC-lasso) and lasso penalized multivariate LMMs (LMM-lasso).
+In Section \ref{sec:methods_paper1} we provide a detailed review of the two most prevalent multivariate methods of adjusting for population stratification and relatedness: PCA adjusted lasso penalized regression (PC-lasso) and lasso penalized multivariate LMMs (LMM-lasso).
 We review the statistical details of how these methods attempt to correct for population structure and environmental heterogeneity.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background} \label{sec:background}
+\section{Background} \label{sec:background_paper1}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Unfortunately, the extension of LMMs and penalized LMMs to nonlinear models is not straightforward.
-As we explain in Section~\ref{sec:methods}, LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation.
+As we explain in Section~\ref{sec:methods_paper1}, LMMs and penalized LMMs are able to efficiently correct for dependencies among subjects by preprocessing the data by means of a weighted rotation.
 This procedure does not directly generalize beyond the linear regression case.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods} \label{sec:methods}
+\section{Methods} \label{sec:methods_paper1}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In other words, any model specified in terms of $\bZ$ and $\bgamma$ is equivalent to a parameterization using $\bX$ and $\boldsymbol{\tau}$.
-This result provides the explicit mathematical justification for the statement in Section~\ref{sec:background} that population structure only biases the SNP-phenotype association if it is correlated with the environmental effects (in the sense that $\bZ\bgamma$ can be projected onto the column space of $\bX$).
+This result provides the explicit mathematical justification for the statement in Section~\ref{sec:background_paper1} that population structure only biases the SNP-phenotype association if it is correlated with the environmental effects (in the sense that $\bZ\bgamma$ can be projected onto the column space of $\bX$).
 It furthermore illustrates why the assumption $\bZ=\bX$ in the linear mixed model is reasonable, as well as how it works to remove bias in estimation.

### 2021-05-15 main.tex
updating wording on lasso assumptions re Kai Wang comments

@@
 
-Lasso models, however, rely on an assumption that subjects are independent of each other.
-In the presence of population structure or cryptic relatedness this assumption is violated: traits are more likely to be similar within a family or ethnic group.
+Lasso models, however, are based on a loss function motivated by independence among subjects.
+In the presence of population structure or cryptic relatedness this is not the case: traits are more likely to be similar within a family or ethnic group.
 The presence of structure can adversely affect analyses in several ways, but the most important concern is that it is likely to introduce false positives.
@@
 Assume the minor allele of SNP $X$ is more common in subpopulation B compared to subpopulation A, but has no direct effect on lung cancer.
-Also suppose these subpopulations are geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
+Also suppose these subpopulations are sociodemographically or geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
 A GWAS of data from these subpopulations would likely find SNP $X$ to be significantly associated with lung cancer.
-This spurious association is does not arise from the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
+This spurious association does not arise from the presence of population stratification itself, but because that stratification is related to a causal environmental exposure.
 Indeed, if subpopulations A and B were not subject to different air qualities, all else being equal, SNP $X$ would not be associated with the phenotype.
@@
 
-A different approach is to use create surrogate variables to represent population structure, and to adjust for these as additional model covariates.
+A different approach is to create surrogate variables to represent population structure, and to adjust for these as additional model covariates.
 Principal component-adjustment (PC-adjustment) methods use the first several eigenvectors a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
@@
 For more information on the lasso, the textbook by \citet{Hastie2009} is a useful reference.
-The following sections discuss how to lasso can be modified to account for correlation among individuals.
+The following sections discuss how the lasso can be modified to account for correlation among individuals.
 
@@
 BIC has been used by several authors for variable selection in penalized models with correlated outcomes \citep{bondell2010joint, schelldorfer2011estimation}.
-BIC performs poorly in the high-dimensional setting, although modified high-dimensional variants have been proposed \citep{fan2013tuning,Chen2008a}.
+BIC performs poorly in the high-dimensional setting, although modified high-dimensional variants have been proposed \citep{fan2013tuning, chen2008}.
 Cross-validation is also reasonable, with some caveats about the generalizability of prediction error \citep{Rabinowicz2020}.
-In the present context, this means that although cross-validation is a valid way to estimate prediction error of another sample from the same population, we cannot expect for example cross-validation of a GWAS from European samples to estimate prediction error accurately for calculating genetic risk in Chinese individuals.
+In the present context, this means that although cross-validation is a valid way to estimate prediction error of another sample from the same population, we cannot expect for example cross-validation of a GWAS from individuals of European ancestry to estimate prediction error accurately for calculating genetic risk in individuals of Asian ancestry.
 Selecting $\lam$ on the basis of FDR is obviously attractive, although there has been little work extending these approaches to correlated outcomes.

### 2021-05-15 main.tex
edits for KW comments

@@
 Its false positive selections are almost always the ones with the largest differences between populations.
-PC-lasso and LMM-lasso, on the other, show no such such selection bias.
+PC-lasso and LMM-lasso, on the other, show no such selection bias.
 
@@
 
-To determine whether the statistical properties demonstrated by our simulations hold in for a genome-wide analysis, we analyzed the full genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives}; a subset of this data was included in the earlier simulations, labeled ``Empirical''.
+To determine whether the statistical properties demonstrated by our simulations hold in a genome-wide analysis, we analyzed the full genomic data ($n=588, p=526,266$) from \citet{larkin2015objectives}; a subset of this data was included in the earlier simulations, labeled ``Empirical''.
 A synthetic phenotype was generated as in \eqref{eqn:sim} using 100 causal SNPs, a 1:1 SNR ($\eta = 0.5$), and substantial environmental effects ($\xi = 0.8$).

### 2021-05-15 main.tex
Fixing MJ typos

@@
 Assume the minor allele of SNP $X$ is more common in subpopulation B compared to subpopulation A, but has no direct effect on lung cancer.
-Also suppose these subpopulations are sociodemographically or geographically segregated in a such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
+Also suppose these subpopulations are sociodemographically or geographically segregated in such a way that subpopulation A is exposed to good air quality, and subpopulation B to poor air quality, and that air quality does have a direct effect on lung cancer.
 A GWAS of data from these subpopulations would likely find SNP $X$ to be significantly associated with lung cancer.
@@
 %------------------------------%
-\subsection{Simulation study}
+\subsection{Simulation studies}
 %------------------------------%

### 2021-05-19 main.tex
MJ corrections
Updating commands
Updates for KW corrections

@@
 where $\bbeta_0$ is a $p_0 \times 1$ vector which includes the intercept term and the effects of any unpenalized covariates, $\bbeta$ is a $p \times 1$ vector of SNP effects, $\bgamma$ is a $q \times 1$ vector of environmental effects, $\bZ$ is an $n \times q$ matrix that allocates environmental effects among subjects, and $\beps \sim N(\boldsymbol{0}, \sigmaee \bI_n)$.
+While it is important to standardize $\bG$ so that all SNP coefficients are equally penalized by the LMM model, the standardization of unpenalized features is irrelevant.
 Often it is assumed that subjects belong to $q \le p$ subpopulations, and $\bZ$ is a corresponding incidence matrix of 0s and 1s.
@@
 Rather than assuming anything about $\bgamma$ directly, a common representation parameterizes the model in terms of $\bu=\bX\bgamma$, which yields an equivalent model.
+We note that model~\eqref{eqn:lmm_reg} assumes independent, additive SNP effects, with no epistatic or dominance effects.
 
@@
 \begin{equation}
-\bV^{-\frac{1}{2}}\by \sim N\left(\bV^{-\frac{1}{2}} \bX{^\prime} \bbeta{^\prime}, \bI \right),
+\bV^{-\frac{1}{2}}\by \sim N\left(\bV^{-\frac{1}{2}} \bX{^\dagger} \bbeta{^\dagger}, \bI \right),
 \end{equation}
-where $\bX{^\prime} = \begin{bmatrix} \bX_0 & \bX \end{bmatrix}$ and $\bbeta{^\prime} = \begin{bmatrix} \bbeta_0^\T & \bbeta^\T \end{bmatrix}^\T$. This in turn leads the objective function
+where $\bX{^\dagger} = \begin{bmatrix} \bX_0 & \bX \end{bmatrix}$ and $\bbeta{^\dagger} = \begin{bmatrix} \bbeta_0^\T & \bbeta^\T \end{bmatrix}^\T$. This in turn leads the objective function
 \begin{equation}
-Q(\bbeta | \bXT{^\prime}, \byT, \bW) = || \bW (\byT - \bXT{^\prime} \bbeta{^\prime})||_2^2 + \lambda || \bbeta ||_1,
+Q(\bbeta | \bXT{^\dagger}, \byT, \bW) = || \bW (\byT - \bXT{^\dagger} \bbeta{^\dagger})||_2^2 + \lambda || \bbeta ||_1,
 \end{equation}
-where $\bXT{^\prime} = \bU^\T \bX{^\prime}$, so that all of the data are transformed, but only the elements of $\bbeta{^\prime}$ corresponding to SNPs are penalized.
+where $\bXT{^\dagger} = \bU^\T \bX{^\dagger}$, so that all of the data are transformed, but only the elements of $\bbeta{^\dagger}$ corresponding to SNPs are penalized.
 

### 2021-05-19 main.tex
fixing typos

@@
 The presence of distinct ancestry groups with different allele frequencies in a sample is known as \textbf{population stratification}.
-Recent relatedness describes the sharing of a common ancestor only a few generations previous.
+Recent relatedness describes the sharing of a common ancestor only a few generations previously.
 Pedigree-based methods may be used to explicitly model recent relatedness if familial relationships are known.
@@
 A different approach is to create surrogate variables to represent population structure, and to adjust for these as additional model covariates.
-Principal component-adjustment (PC-adjustment) methods use the first several eigenvectors a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
+Principal component-adjustment (PC-adjustment) methods use the first several eigenvectors of a pairwise similarity matrix as these surrogate variables \citep{price2006principal}.
 This approach may be used in univariate or multivariate frameworks, and allows for analysis on the entirety of a structured sample.
@@
 The vertical axis displays the difference in MSE: PC-lasso $-$ LMM-lasso, so that positive numbers correspond to LMM-lasso outperforming PC-lasso and vice versa.
-The dashed horizontal line at 0 indicates no difference in the performance the two methods.
+The dashed horizontal line at 0 indicates no difference in the performance of the two methods.
 
@@
 Figure \ref{fig:gw_pc1_corr} displays the mean absolute correlation of the non-causal SNPs selected with the first principal component (PC1).
-The first principal component does not capture the full complexity of the population structure, but it useful in terms of quantifying the extent to which a method is selecting SNPs that merely reflect the population structure and hence are at greater risk of being false positives due to environmental confounding.
+The first principal component does not capture the full complexity of the population structure, but is useful in terms of quantifying the extent to which a method is selecting SNPs that merely reflect the population structure and hence are at greater risk of being false positives due to environmental confounding.
 The shaded regions indicate the SE about the mean.
